### 2020-09-23 thesis/thesis.tex
feat: Add thesis outline

\documentclass[11pt, a4paper, english]{book}
\usepackage{unir}

\setlength{\headheight}{15pt}

%---------------------------
% Título del trabajo y autor
%---------------------------
\title{Open Clusters Detection in Gaia DR2 Using ML Algorithms}
\author{Carlos David Álvaro Yunta}
\date{13 de Septiembre de 2020}
\director{César Augusto Guzmán Álvarez}
\nombreciudad{Madrid, Spain}

%---------------------------
%marges
%---------------------------
%\usepackage[margin=1.9cm]{geometry}
%---------------------------
%---------------------------
%---------------------------
%---------------------------
\begin{document}
%\renewcommand{\listfigurename}{Índice de Ilustraciones}
%\renewcommand{\listtablename}{Índice de Tablas}
%\renewcommand{\contentsname}{Índice de Contenidos}
%\renewcommand{\figurename}{Figura}
%\renewcommand{\tablename}{Tabla}

\maketitle

\frontmatter
\tableofcontents
\listoffigures
\listoftables

\chapter{Abstract}
{\bf Nota:} En este apartado se introducirá un breve resumen en español del trabajo realizado (extensión máxima: 150 palabras). Este resumen debe incluir el objetivo o propósito de la investigación, la metodología, los resultados y las conclusiones.

{\bf Key words.} open clusters detection --- machine learning --- gaia dr2 --- data analysis

\chapter{Resumen}
{\bf Nota:} En este apartado se introducirá un breve resumen en español del trabajo realizado (extensión máxima: 150 palabras). Este resumen debe incluir el objetivo o propósito de la investigación, la metodología, los resultados y las conclusiones.

{\bf Palabras Clave:} detección de cúmulos abiertos --- inteligencia artificial --- gaia dr2 --- análisis de datos

\mainmatter
\chapter{Introduction}

Hunting for Open Clusters\cite{castro2020hunting}

\chapter{Context and State of the Art}

\chapter{Identificación de Requisitos}

\chapter{Aims}

\chapter{Desarrollo del trabajo}

\chapter{Conclusiones y Trabajo Futuro}

\chapter*{Acknowledgement}
\addcontentsline{toc}{chapter}{Acknowledgement}

This work has made use of data from the European Space Agency (ESA) mission
{\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed by the {\it Gaia}
Data Processing and Analysis Consortium (DPAC,
\url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding for the DPAC
has been provided by national institutions, in particular the institutions
participating in the {\it Gaia} Multilateral Agreement.

\bibliography{references}
\bibliographystyle{plain}
\addcontentsline{toc}{chapter}{Bibliography}

\appendix
\chapter{Appendix}
Atención, deberá generar un pdf con la plantilla de artículo y añadirla como anexo utilizando includepdf.

%\includepdf[pages=-]{anexo.pdf} # TODO: Descomentar cuando el artículo esté hecho
\end{document}

### 2020-09-24 code/scripts/downloader.py
feat: Allow specifying extra_size as a downloader argument

@@ -17,6 +17,13 @@ parser.add_argument('--cluster',
         Cluster name from http://cdsarc.u-strasbg.fr/ftp/cats/B/ocl/clusters.dat list.
         Type ALL to download all of them.
         """)
+parser.add_argument('--extra_size',
+                    type=float,
+                    default=1.5,
+                    help="""
+        Extra size to extend regions to be downloaded.
+        Default to 1.5
+        """)
 parser.add_argument('--verbose', '-v', action='count', default=0)
 
 args = parser.parse_args()
@@ -48,6 +55,6 @@ db_host = os.getenv('DB_HOST', 'localhost')
 db_port = os.getenv('DB_PORT', 5432)
 
 gaia = Gaia(db=DB.instance(host=db_host, port=db_port))
-gaia.download_and_save(regions=clusters, extra_size=1.5)
+gaia.download_and_save(regions=clusters, extra_size=args.extra_size)
 
 logger.info("🚀 Gaia downloader has finished retrieving and saving data")

### 2020-09-25 code/cdalvaro/catalogues/openclust/openclust.py
fix: Change OpenClust Catalogue type

@@ -2,13 +2,13 @@ import astropy.units as u
 from astropy.coordinates import SkyCoord
 import os
 import pandas
-from typing import Set, TypeVar, Union
+from typing import Dict, Set, TypeVar, Union
 
 from ..base_catalogue import BaseCatalogue
 from ...logging import Logger
 from ...models import OpenCluster
 
-Catalogue = TypeVar('Catalogue', bound=Set[OpenCluster])
+Catalogue = TypeVar('Catalogue', bound=Dict[str, OpenCluster])
 
 
 class OpenClust(BaseCatalogue):

### 2020-09-25 code/cdalvaro/catalogues/openclust/openclust.py
doc: Update openclust.py documentation

@@ -45,6 +45,7 @@ class OpenClust(BaseCatalogue):
 
         Args:
             names (Set[str]): The name of the clusters to be recovered.
+            as_dataframe (bool, optional): Flag to recover clusters as a DataFrame. Defaults to False.
 
         Returns:
             Catalogue: A catalogue with the found clusters.

### 2020-09-26 code/cdalvaro/data_base.py
feat: Replace get_region by get_regions method in DB

@@ -13,6 +13,7 @@ from psycopg2.extras import execute_values
 from psycopg2.extras import LoggingConnection, LoggingCursor
 from typing import Dict, List, Set, TypeVar, Union
 
+from .catalogues.base_catalogue import Catalogue
 from .logging import Logger
 from .models.open_cluster import OpenCluster
 from .models.region import Region
@@ -135,39 +136,58 @@ class DB:
 
         return set(map(lambda x: next(iter(x)), result))
 
-    def get_region(self, name: str) -> Union[Region, None]:
+    def get_regions(self, names: Set[str] = {}, as_dataframe: bool = False) -> Union[Catalogue, pandas.DataFrame]:
         """
-        Get the region corresponding to the given name.
+        Get the regions matching the given names.
 
         Args:
-            name (str): The name of the region to be retrieved.
+            names (Set[str]): A set with the names with the regions to retrieve.
+            as_dataframe (bool, optional): Flag to recover regions as a DataFrame. Defaults to False.
 
         Returns:
-            Region: The region recovered.
+            Union[Catalogue, pandas.DataFrame]: A catalogue with found regions.
         """
-        DB._logger.debug(f"Getting region {name} from the DB ...")
+        columns = ('name', 'ra', 'dec', 'diam', 'width', 'height')
+        query = f"SELECT {', '. join(columns)} FROM public.regions"
 
-        query = f"SELECT id, ra, dec, diam, width, height FROM public.regions WHERE name = %s"
+        if len(names) > 0:
+            DB._logger.debug(f"Getting regions: {names} from DB ...")
+            query += " WHERE name = ANY(%s)"
+            params = (list(names), )
+        else:
+            DB._logger.debug(f"Getting all regions from DB ...")
+            params = ()
+
+        if as_dataframe:
+            index_col = 'name'
+            try:
+                return pandas.read_sql_query(query, self.conn, index_col=index_col, params=params)
+            except Exception as error:
+                DB._logger.error(f"Error retrieving regions data. Cause: {error}")
+                df = pandas.DataFrame({}, columns=columns)
+                df.set_index(index_col)
+                return df
 
         cursor = self.conn.cursor(cursor_factory=LoggingCursor)
-        cursor.execute(query, (name, ))
-        (serial, ra, dec, diam, width, height) = cursor.fetchone()
+        cursor.execute(query, params)
+
+        catalogue = dict()
[…]

### 2020-09-26 code/scripts/downloader.py
feat: Add argument to exclude clusters from being downloaded

@@ -17,12 +17,20 @@ parser.add_argument('--cluster',
         Cluster name from http://cdsarc.u-strasbg.fr/ftp/cats/B/ocl/clusters.dat list.
         Type ALL to download all of them.
         """)
+parser.add_argument('--exclude',
+                    '-e',
+                    nargs='*',
+                    type=str,
+                    default={},
+                    help="""
+        Clusters to be ignored from the download. Default to empty.
+        """)
 parser.add_argument('--extra_size',
                     type=float,
                     default=1.5,
                     help="""
         Extra size to extend regions to be downloaded.
-        Default to 1.5
+        Default to 1.5.
         """)
 parser.add_argument('--verbose', '-v', action='count', default=0)
 
@@ -50,6 +58,9 @@ else:
         else:
             logger.error(f"Cluster '{cluster_name}' is not avaiable at the OpenClust catalogue")
             exit(1)
+
+if len(args.exclude) > 0:
+    clusters = set(filter(lambda x: x not in args.exclude, clusters))
 
 db_host = os.getenv('DB_HOST', 'localhost')
 db_port = os.getenv('DB_PORT', 5432)

### 2020-09-26 code/cdalvaro/downloaders/gaia/gaia.py
fix: Change _download_stars method return type

@@ -1,4 +1,5 @@
 import astropy
+from astropy.table import QTable
 import astropy.units as u
 from astroquery import gaia
 import json
@@ -60,7 +61,7 @@ class Gaia:
 
         Gaia._logger.info(f"🏁 Finished downloading stars ...")
 
-    def _download_stars(self, region: Region, extra_size: float, exclude: Set[SourceID] = {}) -> Union[astropy.table, None]:
+    def _download_stars(self, region: Region, extra_size: float, exclude: Set[SourceID] = {}) -> Union[QTable, None]:
         """
         Download data from Gaia DR2 for the given region with an optional extra size
         to extend the given region.
@@ -71,7 +72,7 @@ class Gaia:
             exclude (Set[SourceID]): Source ids to be excluded from the download.
 
         Returns:
-            astropy.table: An astropy table with the downloaded data.
+            Union[QTable, None]: An astropy table with the downloaded data, or None if an error occurs.
         """
         try:
             query = self._compose_query(region=region, extra_size=extra_size, exclude=exclude)
@@ -145,13 +146,13 @@ class Gaia:
 
         return query
 
-    def _save_stars(self, region: Region, stars: astropy.table):
+    def _save_stars(self, region: Region, stars: QTable):
         """
         Method for saving data into cdalvaro database.
 
         Args:
             region (Region): The region associated with to the data.
-            stars (astropy.table): An astropy table with the data to be saved.
+            stars (QTable): An astropy table with the data to be saved.
         """
         Gaia._logger.debug(f"Saving stars into db ...")
 

### 2020-09-26 code/cdalvaro/catalogues/openclust/openclust.py
fix: Change openclust diam columns

@@ -130,7 +130,7 @@ class OpenClust(BaseCatalogue):
             g1_class = None
 
         # Diameter
-        diam = entry[40:47].strip()
+        diam = entry[41:47].strip()
         if len(diam) == 0:
             raise ValueError(f"Cluster '{name}' does not have diameter info")
         diam = u.Quantity(diam, u.arcmin)

### 2020-09-26 code/cdalvaro/catalogues/openclust/openclust.py
feat: Add trumpler field to OpenCluster model

@@ -135,9 +135,12 @@ class OpenClust(BaseCatalogue):
             raise ValueError(f"Cluster '{name}' does not have diameter info")
         diam = u.Quantity(diam, u.arcmin)
 
+        # Trumpler
+        trumpler = entry[144:152].strip()
+
         OpenClust._logger.debug(f"Loaded cluster: {name}")
 
-        return OpenCluster(name=name, coords=coords, diam=diam, g1_class=g1_class)
+        return OpenCluster(name=name, coords=coords, diam=diam, trumpler=trumpler, g1_class=g1_class)
 
     @staticmethod
     def _catalogue_to_dataframe(catalogue: Catalogue) -> pandas.DataFrame:

### 2020-09-30 code/cdalvaro/downloaders/gaia/gaia.py
feat: Improve support for excluding data from download

@@ -1,9 +1,10 @@
 import astropy
-from astropy.table import QTable
+from astropy.table import QTable, Table
 import astropy.units as u
 from astroquery import gaia
 import json
 import logging
+import numpy as np
 import os
 import psycopg2
 from typing import List, Set, TypeVar, Union
@@ -15,7 +16,7 @@ from ...models.open_cluster import OpenCluster
 from ...models.region import Region
 
 Regions = TypeVar('Regions', bound=Set[Region])
-SourceID = TypeVar('SourceID', bound=str)
+SourceID = TypeVar('SourceID', bound=np.int64)
 
 # https://astroquery.readthedocs.io/en/latest/api/astroquery.gaia.Conf.html#astroquery.gaia.Conf
 gaia.Gaia.MAIN_GAIA_TABLE = 'gaiadr2.gaia_source'
@@ -85,8 +86,8 @@ class Gaia:
             Union[QTable, None]: An astropy table with the downloaded data, or None if an error occurs.
         """
         try:
-            query = self._compose_query(region=region, extra_size=extra_size, exclude=exclude)
-            job = gaia.Gaia.launch_job_async(query, verbose=Gaia._logger.level == logging.DEBUG)
+            query, temp_table = self._compose_query(region=region, extra_size=extra_size, exclude=exclude)
+            job = gaia.Gaia.launch_job_async(query)
             result = job.get_results()
             if len(result) > 0:
                 Gaia._logger.info(f"Downloaded {len(result)} stars for {region}")
@@ -97,6 +98,9 @@ class Gaia:
         except Exception as error:
             Gaia._logger.error(f"Error executing job for region {region}. Cause: {error}")
             return None
+        finally:
+            if temp_table is not None:
+                job = gaia.Gaia.delete_user_table(temp_table)
 
         return result
 
@@ -119,16 +123,40 @@ class Gaia:
         dec = region.coords.dec.degree
 
         query = f"""
-            SELECT {', '.join(GaiaMetadata.columns())}
-            FROM {gaia.Gaia.MAIN_GAIA_TABLE}
-            WHERE 1 =
+            SELECT {', '.join(map(lambda x: f"A.{x}", GaiaMetadata.columns()))}
+            FROM {gaia.Gaia.MAIN_GAIA_TABLE} A
             """
+
+        temp_table = None
+        if len(exclude) > 0:
+            if self._logged:
+                try:
+                    temp_table = f"temp_table_{region.name.replace(' ', '')}"
[…]

### 2020-10-02 thesis/thesis.tex
thesis: Add Introduction chapter

@@
 
-Hunting for Open Clusters\cite{castro2020hunting}
+Stellar open clusters (OC) are groups of stars gravitationally bound originated from a single molecular gas cloud.
+Thus they share the same chemical composition and age, and they have similar relative positions and proper motion.
+These astronomical objects are of fundamental importance to understand the spiral structure,
+the dynamics and the chemical evolution of our galaxy.
+
+Although most stars in the Milky Way are presented isolated, it is considered that most of them (or even all)
+are formed in clustered environments and spend a period of time gravitationally bound with their siblings embedded
+in their original molecular cloud.
+\cite[Clarke et al. 2000]{clarke2000theformationofstellarclusters} \cite[Portegies Zwart et al. 2010]{portegies2010young}
+The evolution of these systems tends to sparse them in a few million years by interacting gravitationally with other systems.
+Galactic tidal forces and mechanisms that involves the gas loss driven by stellar feedback are other causes of disruption.
+\cite[Brinkmann et al. 2017]{brinkmann2017bound}
+Nevertheless, a small fraction of these systems will survive in the initial state and persist bound in bigger timescales.
+
+Young OC allow us to research stars formation regions and improve our understanding about the mechanisms that create those stars.
+On the other side, older OC give us information about stellar processes and how the galactic disk evolves.
+Some highly disturbed orbits could also provide evidence of recent merge events and accretion traces from outside the galaxy.
+\cite[Cantat-Gaudin et al. 2016]{cantat2016abundances}
+
+The study of OC has been pushed forward thanks to the huge and precise dataset from the Gaia mission \cite[Gaia DR2]{gaia2018gaia},
+available since 2018. This dataset has helped to review already known open clusters and to find new ones.
 

### 2020-10-05 thesis/thesis.tex
thesis: Add Context and State of the Art

--- 2020-09-23 thesis/thesis.tex
@@
 \documentclass[11pt, a4paper, english]{book}
+
 \usepackage{unir}
+
+\usepackage{caption}
+\usepackage{graphicx}
+\usepackage{natbib}
+\usepackage{subcaption}
+
+% Bibliography style
+\bibliographystyle{abbrvnat}
+\setcitestyle{authoryear,open={(},close={)}}
 
@@
 %---------------------------
-\title{Open Clusters Detection in Gaia DR2 Using ML Algorithms}
+\title{Open Clusters Characterization in Gaia DR2 Using ML Algorithms}
 \author{Carlos David Álvaro Yunta}
@@
 \chapter{Abstract}
-{\bf Nota:} En este apartado se introducirá un breve resumen en español del trabajo realizado (extensión máxima: 150 palabras).
+
+{\bf Nota:} En este apartado se introducirá un breve resumen en inglés del trabajo realizado (extensión máxima: 150 palabras).
 Este resumen debe incluir el objetivo o propósito de la investigación, la metodología, los resultados y las conclusiones.
+
+\medskip
 
@@
 \chapter{Resumen}
+
 {\bf Nota:} En este apartado se introducirá un breve resumen en español del trabajo realizado (extensión máxima: 150 palabras).
 Este resumen debe incluir el objetivo o propósito de la investigación, la metodología, los resultados y las conclusiones.
+
+\medskip
 
@@
 
-Hunting for Open Clusters\cite{castro2020hunting}
+Stellar open clusters (OC) are groups of stars gravitationally bound originated from a single molecular gas cloud.
+Thus they share the same chemical composition and age, and they have similar relative positions and proper motion.
+These astronomical objects are of fundamental importance to understand the spiral structure,
+the dynamics and the chemical evolution of our galaxy.
+
+Although most stars in the Milky Way are presented isolated, it is considered that most of them (or even all)
+are formed in clustered environments and spend a period of time gravitationally bound with their siblings embedded
+in their original molecular cloud
+\cite[]{clarke2000theformationofstellarclusters} \cite[]{portegies2010young}.
+The evolution of these systems tends to sparse them in a few million years by interacting gravitationally with other systems.
+Galactic tidal forces and mechanisms that involves the gas loss driven by stellar feedback are other causes of disruption
+\cite[]{brinkmann2017bound}.
+Nevertheless, a small fraction of these systems will survive in the initial state and persist bound in bigger timescales.
+
+Young OC allow us to research stars formation regions and improve our understanding about the mechanisms that create those stars.
+On the other side, older OC give us information about stellar processes and how the galactic disk evolves.
+Some highly disturbed orbits could also provide evidence of recent merge events and accretion traces from outside the galaxy
+\cite[]{cantat2016abundances}.
+
+The study of OC has been pushed forward thanks to the huge and precise dataset from the Gaia mission
+\cite[]{collaboration2016description} Gaia DR2 \cite[]{gaia2018gaia}, available since 2018.
+This dataset has helped to review already known open clusters and to find new ones.
 
@@
 
-\chapter{Identificación de Requisitos}
+An initial approach to find OC is the search of overdensities along the galactic disk. In general, this is the initial point,
+but even it looks simple, it presents a fundamental problem. The near field around the OC is fulled by two differenciated
+kind of stars populations: the ones which belong to the OC (from a tens or hundreds to a few thousand) and a background composed
+by thousands or millions of stars which don't belong to the OC. Find out which stars belong to the first group is the problem
+treated in this work. This selection is crucial to properly characterize the fundamental properties of the cluster
+(dynamics, total mass, age, chemical composition, etc.).
+
+Sometimes, this problem is easy to solve by looking into astrometric parameters, such as the analysis of proper motion of those
+stars belonging to the overdensity field.
+
+\begin{figure}[htbp]
+  \centering
+  \begin{subfigure}[t]{0.45\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{../figures/pm_melotte_22.pdf}
+    \caption{Proper motions}
+    \label{fig:pm_melotte_22}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[t]{0.45\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{../figures/pm_melotte_22.pdf}
+    \caption{Proper motions with vector lines}
+    \label{fig:pm_vec_melotte_22}
+  \end{subfigure}
+  \caption{Open Cluster Melotte 22 (M45)}
+\end{figure}
+
+However, in general it is not as easy and becomes necessary to take into account other parameters such as distances, or even
+metallicity and age (derived from isochrone curves) and requires photometric data from stars in the studied field.
+
+\begin{figure}[htbp]
+  \centering
+  \begin{subfigure}[t]{0.45\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{../figures/pm_ngc_2353.pdf}
+    \caption{Proper motions}
+    \label{fig:pm_ngc_2353}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[t]{0.45\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{../figures/parallax_ngc_2353.pdf}
+    \caption{Parallax histogram}
+    \label{fig:parallax_ngc_2353}
+  \end{subfigure}
+  \caption{Open Cluster NGC 2353}
+\end{figure}
+
+At this point, it looks reasonable that the study of each individual cluster requires its own parametrization and technique
+to identify it.
+
+With Gaia DR2, which contains great quality astrometrical and photometric data for a huge number of stars,
+a new opportunity to simplify and optimize this processes is presented.
+In this context new approaches are being developed to improve current detection methods and automate the OC characterization.
+This work tries to take the best from \emph{Clusterix 2.0. A Virtual Observatory tool to estimate cluster membership probability}
+\cite[]{balaguer2020clusterix} and \emph{Hunting for open clusters in Gaia DR2: 582 new open clusters
+in the Galactic disc} \cite[]{castro2020hunting}, and merge them into a single \emph{machine learning}
+tool capable of finding open clusters eficiently without requiring powerful machines to perform the computations.
+
+The first one is a non-parametrized method. It determines empirically the frequency functions asociated to astrometrical variables
+without making any assumption about their profiles. However, it heavily does depend on the initial field selection and other
+parameters related to frequency functions. \emph{Clusterix 2.0.} is presented as a web app tool inside the Spanish Virtual Observatory
+(VO) framework and works together with other VO tool such as:
+\emph{TOPCAT} \cite[][\url{http://www.starlink.ac.uk/topcat/}]{taylor2005topcat},
+\emph{Aladin} \cite[][\url{https://aladin.u-strasbg.fr/aladin.gml}]{bonnarel2000aladin},
+\emph{VOSA} \cite[][\url{http://svo2.cab.inta-csic.es/theory/vosa/}]{bayo2008vosa}, etc.
+
+The second tool is based on Big Data techniques. Its purpose is to analyze the whole Gaia DR2 dataset looking for OCs by using
+machine learning algorithms divided in two stages. In the first stage, the algorithm looks for posible OC candidates by searching
+overdensities in the galactic disc.
+The second stage studies candidates found in the first one by generating images with H-R diagrams and passing them
+as sources to an Artificial Neural Network (ANN) which identifies isochrone patterns in photometric data (color and magnitude).
+Before this approach was put into practice there were 1200 OCs known. Now, this number has been increased up to more than 2000 OCs
+available at the Vizier catalogue \cite[][\url{https://vizier.unistra.fr}]{ochsenbein2000vizier}.
+
+The method developed in this project takes as base the approach made by \citeauthor{castro2020hunting} but rapidly diverges.
+In first place, performance has been taken into account since the begining, trying to get a reliable method without compromising the
+needs of supercomputers.
+To achieve this, instead of blindly analyzing the whole galactic disk a fields selection is made focusing in those regions with known OCs.
+Then, these fields are analyzed with an homogeneous methodology based on machine learning algorithms using as sources astrometric and
+photometric variables. The goal is the characterization of each individual star in the cluster, discerning wether the star belongs the OC
+or not.
+
+Thus, the main goal of this project is not the discovery of new open clusters or even candidates, but the development of a model that,
+applied to already knonw open clusters, allows us the characterization of those OCs on an unsupervised way, with a reasonable computing
+power and taking into account the same selection of astrometric and photometric variables for every OC.
+
+\chapter{Requirements Description}
 
@@
 
-\chapter{Desarrollo del trabajo}
+The goal of this work is to \emph{achieve a viable model for open cluster characterization based on machine learning techniques}
+such as clustering algorithms and artificial neural networks.
+This model must be a non-supervised non-parameterized algorithm valid for any selected OC.
 
-\chapter{Conclusiones y Trabajo Futuro}
+The OpenClust catalogue \cite[]{dias2002new} has been used to make a preselection of sky regions for the Gaia DR2 data download.
+This download is not limited to the sized registered in the catalogue. Instead, a wider region is downloaded for each cluster to include
+stars from outside the cluster and take them into account when training the model so the model will be able to take apart those outsiders
+when characterizing a non known cluster.
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=\columnwidth]{../figures/openclust_catalogue.pdf}
+  \caption{OpenClust Catalogue Distribution}
+\end{figure}
+
+A later filter will be applied to select only those OCs that fulfil one or more of the following criteria:
+
+\begin{itemize}
+  \item Cluster diameter above 25.0 arcmin
+  \item Number of stars in the selected region above 40,000 stars
+\end{itemize}
+
+The final downloaded set covers nearly 60 million stars, which is significantly smaller than the whole Gaia DR2 dataset which contains
+information for around 1,600 million stars.
+This restriction responds to a commitment to the project delivery dates and the available computing power,
+and to having a good basis for the correct model training. As an extension of this work, the trained model could be applied to a region
+not covered by this preselection for a later understanding of the results.
+
+The intention is to get a reliable and efficient non-parameterized neither supervised model over those parameterized or supervised ones.
+To measure the reliability of the resulting model comparisons with other VO tools will be made.
+
+The role for the ANN will be validating cluster members based on dynamincs and photometric data.
+
+These are te variables to be considered in this work:
+
+\begin{itemize}
+  \item OCs and stars \emph{right ascension}: $\alpha$ and \emph{declination}: $\delta$
+  \item OCs \emph{estimated sizes}: $r$
+  \item \emph{Proper motion}: $\mu_{\alpha}$, $\mu_{\delta}$ and \emph{parallax}: $\varpi$
+  \item \emph{Photometric}: $E(B-R)$, $G_{mag}$
+  \item \emph{Radial velocity}: $RV$
+  \item \emph{Associated errors} will be taken into account to estimate prediction errors
+\end{itemize}
+
+Initially, no filter will be applied over parallax neither magnitude values to the selected objects from Gaia DR2.
+
+\chapter{Work Development}
+
+\chapter{Conclusions and Future Work}
 
@@
 
+\medskip
+
+This publication makes use of VOSA, developed under the Spanish Virtual Observatory project
+supported by the Spanish MINECO through grant AyA2017-84089.
+VOSA has been partially updated by using funding from the European Union's Horizon 2020 Research
+and Innovation Programme, under Grant Agreement nº 776403 (EXOPLANETS-A)
+
+\medskip
+
+This research has made use of the VizieR catalogue access tool, CDS, Strasbourg, France (DOI : 10.26093/cds/vizier).
+The original description of the VizieR service was published in 2000, \cite[A\&AS 143, 23]{ochsenbein2000vizier}
+
 \bibliography{references}
-\bibliographystyle{plain}
 \addcontentsline{toc}{chapter}{Bibliography}
@@
 \chapter{Appendix}
-Atención, deberá generar un pdf con la plantilla de artículo y añadirla como anexo utilizando includepdf.
 

### 2020-10-07 code/cdalvaro/data_base.py
bug: Ensure closing db cursor

@@ -87,7 +87,7 @@ class DB:
             Dict[str, int]: A dictionary relating region names with their corresponding ids.
         """
         regions_name = list(map(lambda x: x.name, regions))
-        DB._logger.debug(f"Getting regions id for regions: {', '.join(regions_name)} from dab ...")
+        DB._logger.debug(f"Getting regions id for regions: {', '.join(regions_name)} from DB ...")
 
         query = f"""
             SELECT name, id FROM public.regions
@@ -95,10 +95,15 @@ class DB:
             ORDER BY name ASC
             """
 
-        cursor = self.conn.cursor(cursor_factory=LoggingCursor)
-        cursor.execute(query, (regions_name, ))
-        result = cursor.fetchall()
-        cursor.close()
+        try:
+            cursor = self.conn.cursor(cursor_factory=LoggingCursor)
+            cursor.execute(query, (regions_name, ))
+            result = cursor.fetchall()
+        except Exception as error:
+            DB._logger.error(f"An error ocurred recovering regions id from DB. Cause: {error}")
+            raise error
+        finally:
+            cursor.close()
 
         regions_id = dict()
         for (name, serial) in result:
@@ -134,7 +139,7 @@ class DB:
             cursor.execute(query, (regions_name, ))
             result = cursor.fetchall()
         except Exception as error:
-            DB._logger.error(f"An error ocurred while recovering stars source_id from DB. Cause: {error}")
+            DB._logger.error(f"An error ocurred recovering stars source_id from DB. Cause: {error}")
             raise error
         finally:
             cursor.close()
@@ -168,16 +173,21 @@ class DB:
             try:
                 return pandas.read_sql_query(query, self.conn, index_col=index_col, params=params)
             except Exception as error:
-                DB._logger.error(f"Error retrieving regions data. Cause: {error}")
-                df = pandas.DataFrame({}, columns=columns)
-                df.set_index(index_col)
-                return df
+                DB._logger.error(f"An error ocurred recovering regions dataframe from DB. Cause: {error}")
+                raise error
 
-        cursor = self.conn.cursor(cursor_factory=LoggingCursor)
-        cursor.execute(query, params)
+        try:
+            cursor = self.conn.cursor(cursor_factory=LoggingCursor)
+            cursor.execute(query, params)
+            result = cursor.fetchall()
+        except Exception as error:
+            DB._logger.error(f"An error ocurred recovering regions catalogue from DB. Cause: {error}")
[…]

### 2020-10-07 code/scripts/downloader.py
feat: Allow disabling regions updates

@@ -25,13 +25,27 @@ parser.add_argument('--exclude',
                     help="""
         Clusters to be ignored from the download. Default to empty.
         """)
-parser.add_argument('--extra_size',
+parser.add_argument('--extra-size',
+                    '-s',
+                    dest='extra_size',
                     type=float,
                     default=1.5,
                     help="""
         Extra size to extend regions to be downloaded.
         Default to 1.5.
         """)
+
+update_data_parser = parser.add_mutually_exclusive_group(required=False)
+update_data_parser.add_argument('--update-data',
+                                '-u',
+                                dest='update_data',
+                                action='store_true',
+                                help="""
+        Update existing clusters with new data. Default.
+        """)
+update_data_parser.add_argument('--no-update-data', dest='update_data', action='store_false')
+update_data_parser.set_defaults(update_data=True)
+
 parser.add_argument('--verbose', '-v', action='count', default=0)
 
 args = parser.parse_args()
@@ -66,6 +80,15 @@ db_host = os.getenv('DB_HOST', 'localhost')
 db_port = os.getenv('DB_PORT', 5432)
 db = DB.instance(host=db_host, port=db_port)
 
+if not args.update_data:
+    logger.info("Existing regions won't be updated")
+    existing_clusters = db.get_regions().keys()
+    clusters = set(filter(lambda x: x not in existing_clusters, clusters))
+
+if len(clusters) == 0:
+    logger.info("🍻 All regions are already downloaded")
+    exit(0)
+
 gaia_username = os.getenv('GAIA_USER', None)
 gaia_password = os.getenv('GAIA_PASS', None)
 
@@ -73,3 +96,4 @@ gaia = Gaia(db=db, username=gaia_username, password=gaia_password)
 gaia.download_and_save(regions=clusters, extra_size=args.extra_size)
 
 logger.info("🚀 Gaia downloader has finished retrieving and saving data")
+exit(0)

### 2020-10-07 code/cdalvaro/graphics/__init__.py
python: Add plot functions to the graphics package

@@ -1,1 +1,2 @@
 from .color_palette import ColorPalette, set_color_palette
+from .plot import Plot

### 2020-10-08 code/cdalvaro/data_base.py
fix: Add missing rollbacks

@@ -316,6 +316,7 @@ class DB:
             self.conn.commit()
         except Exception as error:
             DB._logger.error(f"An error ocurred saving regions data into DB. Cause: {error}")
+            self.conn.rollback()
             raise error
         finally:
             cursor.close()
@@ -371,6 +372,7 @@ class DB:
             self.conn.commit()
         except Exception as error:
             DB._logger.error(f"An error ocurred saving stars data into DB. Cause: {error}")
+            self.conn.rollback()
             raise error
         finally:
             cursor.close()

### 2020-10-09 code/cdalvaro/downloaders/gaia/gaia.py
fix: Finished jobs are properly removed now

@@ -90,7 +90,6 @@ class Gaia:
             query, temp_table = self._compose_query(region=region, extra_size=extra_size, exclude=exclude)
             job = gaia.Gaia.launch_job_async(query)
             result = job.get_results()
-            gaia.Gaia.remove_jobs(jobs_list=[job.jobid])
             if len(result) > 0:
                 Gaia._logger.info(f"Downloaded {len(result)} stars for {region}")
             elif len(exclude) > 0:
@@ -102,7 +101,13 @@ class Gaia:
             return None
         finally:
             if temp_table is not None:
-                job = gaia.Gaia.delete_user_table(temp_table)
+                gaia.Gaia.delete_user_table(temp_table, force_removal=True, verbose=Gaia._logger.level <= logging.DEBUG)
+
+        try:
+            gaia.Gaia.remove_jobs(jobs_list=[job.jobid], verbose=Gaia._logger.level <= logging.DEBUG)
+            Gaia._logger.debug(f"Job {job.jobid} successfully removed")
+        except Exception as error:
+            Gaia._logger.error(f"Error removing job: {job.jobid} from Gaia server. Cause: {error}")
 
         return result
 
@@ -143,7 +148,9 @@ class Gaia:
                                   names=['source_id'],
                                   dtype=[np.int64],
                                   meta={'meta': f"temporary table for region {region}"})
-                    gaia.Gaia.upload_table(upload_resource=table, table_name=temp_table)
+                    gaia.Gaia.upload_table(upload_resource=table,
+                                           table_name=temp_table,
+                                           verbose=Gaia._logger.level <= logging.DEBUG)
                     query += f"""
                         LEFT JOIN {schema}.{temp_table} B
                         ON A.source_id = B.source_id
@@ -212,7 +219,7 @@ class Gaia:
 
         try:
             Gaia._logger.info("Logging in to Gaia DR2 database...")
-            gaia.Gaia.login(user=self.username, password=self.password)
+            gaia.Gaia.login(user=self.username, password=self.password, verbose=Gaia._logger.level <= logging.DEBUG)
             self._logged = True
         except Exception as error:
             Gaia._logger.error(f"Unable to login to Gaia DR2 database. Cause: {error}")

### 2020-10-09 code/cdalvaro/downloaders/gaia/gaia.py
feat: Ensure job removal

@@ -103,11 +103,11 @@ class Gaia:
             if temp_table is not None:
                 gaia.Gaia.delete_user_table(temp_table, force_removal=True, verbose=Gaia._logger.level <= logging.DEBUG)
 
-        try:
-            gaia.Gaia.remove_jobs(jobs_list=[job.jobid], verbose=Gaia._logger.level <= logging.DEBUG)
-            Gaia._logger.debug(f"Job {job.jobid} successfully removed")
-        except Exception as error:
-            Gaia._logger.error(f"Error removing job: {job.jobid} from Gaia server. Cause: {error}")
+            try:
+                gaia.Gaia.remove_jobs(jobs_list=[job.jobid], verbose=Gaia._logger.level <= logging.DEBUG)
+                Gaia._logger.debug(f"Job {job.jobid} successfully removed")
+            except Exception as error:
+                Gaia._logger.error(f"Error removing job: {job.jobid} from Gaia server. Cause: {error}")
 
         return result
 

### 2020-10-09 code/cdalvaro/downloaders/gaia/gaia.py
feat: Improve temporary tables support

@@ -34,10 +34,11 @@ class Gaia:
 
     _logger = Logger.instance()
 
-    def __init__(self, db: DB, username: str = None, password: str = None):
+    def __init__(self, db: DB, username: str = None, password: str = None, remove_jobs: bool = True):
         self.db = db
         self.username = username
         self.password = password
+        self.remove_jobs = remove_jobs
         self._logged = False
 
     def download_and_save(self, regions: Regions, extra_size: float = 1.0):
@@ -86,9 +87,12 @@ class Gaia:
         Returns:
             Union[QTable, None]: An astropy table with the downloaded data, or None if an error occurs.
         """
+        job = None
         try:
-            query, temp_table = self._compose_query(region=region, extra_size=extra_size, exclude=exclude)
-            job = gaia.Gaia.launch_job_async(query)
+            query, temp_table_name, temp_table = self._compose_query(region=region,
+                                                                     extra_size=extra_size,
+                                                                     exclude=exclude)
+            job = gaia.Gaia.launch_job_async(query, upload_resource=temp_table, upload_table_name=temp_table_name)
             result = job.get_results()
             if len(result) > 0:
                 Gaia._logger.info(f"Downloaded {len(result)} stars for {region}")
@@ -100,14 +104,12 @@ class Gaia:
             Gaia._logger.error(f"Error executing job for region {region}. Cause: {error}")
             return None
         finally:
-            if temp_table is not None:
-                gaia.Gaia.delete_user_table(temp_table, force_removal=True, verbose=Gaia._logger.level <= logging.DEBUG)
-
-            try:
-                gaia.Gaia.remove_jobs(jobs_list=[job.jobid], verbose=Gaia._logger.level <= logging.DEBUG)
-                Gaia._logger.debug(f"Job {job.jobid} successfully removed")
-            except Exception as error:
-                Gaia._logger.error(f"Error removing job: {job.jobid} from Gaia server. Cause: {error}")
+            if job is not None and self._logged and self.remove_jobs:
+                try:
+                    gaia.Gaia.remove_jobs(jobs_list=[job.jobid], verbose=Gaia._logger.level <= logging.DEBUG)
+                    Gaia._logger.debug(f"Job {job.jobid} successfully removed")
+                except Exception as error:
+                    Gaia._logger.error(f"Error removing job: {job.jobid} from Gaia server. Cause: {error}")
 
         return result
 
[…]

### 2020-10-21 code/cdalvaro/data_base.py
python: Filter incomplete entries

@@ -189,7 +189,8 @@ class DB:
                   columns: List[str] = None,
                   limit: int = None,
                   use_region_id: bool = True,
-                  extra_size: float = 1.0) -> pd.DataFrame:
+                  extra_size: float = 1.0,
+                  filter_null_columns: Union[bool, Set[str]] = False) -> pd.DataFrame:
         """
         Returns a Pandas DataFrame containing the stars of the given region.
 
@@ -199,6 +200,7 @@ class DB:
             limit (int, optional): The maximum number of stars to be recovered. Defaults no limit.
             use_region_id (bool, optional): Use the region_id to get stars. If False select starts by position fields. Defaults to True.
             extra_size (float, optional): A positive number to extend the region which contains the stars to be recovered. Defaults to 1.0.
+            filter_null_columns (Union[bool, Set[str]], optional): Filter entries with null values in the given columns. Defaults to False.
 
         Returns:
             DataFrame: A Pandas DataFrame with the recovered stars
@@ -206,7 +208,7 @@ class DB:
         DB._logger.debug(f"Getting stars for region {region}")
 
         index_columns = ['region_id', 'source_id']
-        if columns is not None:
+        if columns is not None and '*' not in columns:
             columns = index_columns + list(filter(lambda x: x not in index_columns, columns))
         else:
             columns = ['*']
@@ -246,6 +248,20 @@ class DB:
                 query += """
                     BOX(POINT(%(ra1)s, %(dec1)s), POINT(%(ra2)s, %(dec2)s)) @> POINT(ra, dec)
                     """
+
+        if isinstance(filter_null_columns, bool):
+            if filter_null_columns:
+                if '*' in columns:
+                    query += "AND gaiadr2_source IS NOT NULL"
+                else:
+                    columns = list(filter(lambda column: column not in index_columns, columns))
+                    query += " ".join(list(map(lambda column: f"AND {column} IS NOT NULL", columns)))
+        elif len(filter_null_columns) > 0:
+            if '*' in filter_null_columns:
+                query += "AND gaiadr2_source IS NOT NULL"
+            else:
+                filter_null_columns = list(filter(lambda column: column not in index_columns, filter_null_columns))
+                query += " ".join(list(map(lambda column: f"AND {column} IS NOT NULL", filter_null_columns)))
 
         query += """
             ORDER BY region_id, source_id ASC

### 2020-10-23 thesis/thesis.tex
thesis: Add GitHub QR code
thesis: Update figures
thesis: First corrections

@@
 \usepackage{caption}
+\usepackage{eso-pic}
+\usepackage{float}
 \usepackage{graphicx}
 \usepackage{natbib}
+\usepackage{picture}
 \usepackage{subcaption}
@@
 
+\chapter{Acknowledgement}
+
+This work has made use of data from the European Space Agency (ESA) mission
+{\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed by the {\it Gaia}
+Data Processing and Analysis Consortium (DPAC,
+\url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding for the DPAC
+has been provided by national institutions, in particular the institutions
+participating in the {\it Gaia} Multilateral Agreement.
+
+\medskip
+
+This publication makes use of VOSA, developed under the Spanish Virtual Observatory project
+supported by the Spanish MINECO through grant AyA2017-84089.
+VOSA has been partially updated by using funding from the European Union's Horizon 2020 Research
+and Innovation Programme, under Grant Agreement nº 776403 (EXOPLANETS-A)
+
+\medskip
+
+This research has made use of the VizieR catalogue access tool, CDS, Strasbourg, France (DOI : 10.26093/cds/vizier).
+The original description of the VizieR service was published in 2000, \cite[A\&AS 143, 23]{ochsenbein2000vizier}
+
 \mainmatter
@@
 
-Stellar open clusters (OC) are groups of stars gravitationally bound originated from a single molecular gas cloud.
+Stellar open clusters (OC) \cite[]{janes1982open} are groups of stars gravitationally bound originated from a single molecular gas cloud.
 Thus they share the same chemical composition and age, and they have similar relative positions and proper motion.
@@
 
-\chapter{Context and State of the Art}
-
-An initial approach to find OC is the search of overdensities along the galactic disk. In general, this is the initial point,
-but even it looks simple, it presents a fundamental problem. The near field around the OC is fulled by two differenciated
-kind of stars populations: the ones which belong to the OC (from a tens or hundreds to a few thousand) and a background composed
-by thousands or millions of stars which don't belong to the OC. Find out which stars belong to the first group is the problem
-treated in this work. This selection is crucial to properly characterize the fundamental properties of the cluster
-(dynamics, total mass, age, chemical composition, etc.).
-
-Sometimes, this problem is easy to solve by looking into astrometric parameters, such as the analysis of proper motion of those
-stars belonging to the overdensity field.
-
-\begin{figure}[htbp]
-  \centering
-  \begin{subfigure}[t]{0.45\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{../figures/pm_melotte_22.pdf}
-    \caption{Proper motions}
-    \label{fig:pm_melotte_22}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[t]{0.45\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{../figures/pm_melotte_22.pdf}
-    \caption{Proper motions with vector lines}
-    \label{fig:pm_vec_melotte_22}
-  \end{subfigure}
-  \caption{Open Cluster Melotte 22 (M45)}
-\end{figure}
-
-However, in general it is not as easy and becomes necessary to take into account other parameters such as distances, or even
-metallicity and age (derived from isochrone curves) and requires photometric data from stars in the studied field.
-
-\begin{figure}[htbp]
-  \centering
-  \begin{subfigure}[t]{0.45\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{../figures/pm_ngc_2353.pdf}
-    \caption{Proper motions}
-    \label{fig:pm_ngc_2353}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[t]{0.45\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{../figures/parallax_ngc_2353.pdf}
-    \caption{Parallax histogram}
-    \label{fig:parallax_ngc_2353}
-  \end{subfigure}
-  \caption{Open Cluster NGC 2353}
-\end{figure}
-
-At this point, it looks reasonable that the study of each individual cluster requires its own parametrization and technique
-to identify it.
-
-With Gaia DR2, which contains great quality astrometrical and photometric data for a huge number of stars,
-a new opportunity to simplify and optimize this processes is presented.
-In this context new approaches are being developed to improve current detection methods and automate the OC characterization.
-This work tries to take the best from \emph{Clusterix 2.0. A Virtual Observatory tool to estimate cluster membership probability}
-\cite[]{balaguer2020clusterix} and \emph{Hunting for open clusters in Gaia DR2: 582 new open clusters
-in the Galactic disc} \cite[]{castro2020hunting}, and merge them into a single \emph{machine learning}
-tool capable of finding open clusters eficiently without requiring powerful machines to perform the computations.
-
-The first one is a non-parametrized method. It determines empirically the frequency functions asociated to astrometrical variables
-without making any assumption about their profiles. However, it heavily does depend on the initial field selection and other
-parameters related to frequency functions. \emph{Clusterix 2.0.} is presented as a web app tool inside the Spanish Virtual Observatory
-(VO) framework and works together with other VO tool such as:
-\emph{TOPCAT} \cite[][\url{http://www.starlink.ac.uk/topcat/}]{taylor2005topcat},
-\emph{Aladin} \cite[][\url{https://aladin.u-strasbg.fr/aladin.gml}]{bonnarel2000aladin},
-\emph{VOSA} \cite[][\url{http://svo2.cab.inta-csic.es/theory/vosa/}]{bayo2008vosa}, etc.
-
-The second tool is based on Big Data techniques. Its purpose is to analyze the whole Gaia DR2 dataset looking for OCs by using
-machine learning algorithms divided in two stages. In the first stage, the algorithm looks for posible OC candidates by searching
-overdensities in the galactic disc.
-The second stage studies candidates found in the first one by generating images with H-R diagrams and passing them
-as sources to an Artificial Neural Network (ANN) which identifies isochrone patterns in photometric data (color and magnitude).
-Before this approach was put into practice there were 1200 OCs known. Now, this number has been increased up to more than 2000 OCs
-available at the Vizier catalogue \cite[][\url{https://vizier.unistra.fr}]{ochsenbein2000vizier}.
-
-The method developed in this project takes as base the approach made by \citeauthor{castro2020hunting} but rapidly diverges.
-In first place, performance has been taken into account since the begining, trying to get a reliable method without compromising the
-needs of supercomputers.
-To achieve this, instead of blindly analyzing the whole galactic disk a fields selection is made focusing in those regions with known OCs.
-Then, these fields are analyzed with an homogeneous methodology based on machine learning algorithms using as sources astrometric and
-photometric variables. The goal is the characterization of each individual star in the cluster, discerning wether the star belongs the OC
-or not.
-
-Thus, the main goal of this project is not the discovery of new open clusters or even candidates, but the development of a model that,
-applied to already knonw open clusters, allows us the characterization of those OCs on an unsupervised way, with a reasonable computing
-power and taking into account the same selection of astrometric and photometric variables for every OC.
-
-\chapter{Requirements Description}
-
-\chapter{Aims}
+\section{Aims}
 
@@
   \item \emph{Photometric}: $E(B-R)$, $G_{mag}$
-  \item \emph{Radial velocity}: $RV$
   \item \emph{Associated errors} will be taken into account to estimate prediction errors
@@
 
+\chapter{State of the Art}
+
+An initial approach to find OC is the search of overdensities along the galactic disk. In general, this is the initial point,
+but even it looks simple, it presents a fundamental problem. The near field around the OC is fulled by two differenciated
+kind of stars populations: the ones which belong to the OC (from a tens or hundreds to a few thousand) and a background composed
+by thousands or millions of stars which don't belong to the OC. Find out which stars belong to the first group is the problem
+treated in this work. This selection is crucial to properly characterize the fundamental properties of the cluster
+(dynamics, total mass, age, chemical composition, etc.).
+
+Sometimes, this problem is easy to solve by looking into astrometric parameters, such as the analysis of proper motion of those
+stars belonging to the overdensity field.
+
+\begin{figure}[htbp]
+  \centering
+  \begin{subfigure}[t]{0.45\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{../figures/pm_melotte_22.pdf}
+    \caption{Proper motions}
+    \label{fig:pm_melotte_22}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[t]{0.45\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{../figures/parallax_melotte_22.pdf}
+    \caption{Parallax}
+    \label{fig:pm_vec_melotte_22}
+  \end{subfigure}
+  \caption{Open Cluster Melotte 22 (M45)}
+\end{figure}
+
+However, in general it is not as easy and becomes necessary to consider other parameters such as distances, or even
+metallicity and age (derived from isochrone curves) and requires photometric data from stars in the studied field.
+
+\begin{figure}[htbp]
+  \centering
+  \begin{subfigure}[t]{0.45\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{../figures/pm_ngc_2682.pdf}
+    \caption{Proper motions}
+    \label{fig:pm_ngc_2682}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[t]{0.45\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{../figures/isochrone_ngc_2682.pdf}
+    \caption{Isochrone curve}
+    \label{fig:parallax_ngc_2682}
+  \end{subfigure}
+  \caption{Open Cluster NGC 2682}
+\end{figure}
+
+At this point, it looks reasonable that the study of each individual cluster requires its own parametrization and technique
+to identify it.
+
+With Gaia DR2, which contains great quality astrometrical and photometric data for a huge number of stars,
+a new opportunity to simplify and optimize this processes is presented.
+In this context new approaches are being developed to improve current detection methods and automate the OC characterization.
+This work tries to take the best from \emph{Clusterix 2.0. A Virtual Observatory tool to estimate cluster membership probability}
+\cite[]{balaguer2020clusterix} and \emph{Hunting for open clusters in Gaia DR2: 582 new open clusters
+in the Galactic disc} \cite[]{castro2020hunting}, and merge them into a single \emph{machine learning}
+tool capable of finding open clusters eficiently without requiring powerful machines to perform the computations.
+
+The first one is a non-parametrized method. It determines empirically the frequency functions asociated to astrometrical variables
+without making any assumption about their profiles. However, it heavily does depend on the initial field selection and other
+parameters related to frequency functions. \emph{Clusterix 2.0.} is a web app tool inside the Spanish Virtual Observatory
+(VO) framework and works together with other VO tool such as:
+\emph{TOPCAT} \cite[][\url{http://www.starlink.ac.uk/topcat/}]{taylor2005topcat},
+\emph{Aladin} \cite[][\url{https://aladin.u-strasbg.fr/aladin.gml}]{bonnarel2000aladin},
+\emph{VOSA} \cite[][\url{http://svo2.cab.inta-csic.es/theory/vosa/}]{bayo2008vosa}, etc.
+
+The second tool is based on Big Data techniques. Its purpose is to analyze the whole Gaia DR2 dataset looking for OCs by using
+machine learning algorithms divided in two stages. In the first stage, the algorithm looks for posible OC candidates by searching
+overdensities in the galactic disc.
+The second stage studies candidates found in the first one by generating images with H-R diagrams and passing them
+as sources to an Artificial Neural Network (ANN) which identifies isochrone patterns in photometric data (color and magnitude).
+Before this approach was put into practice there were 1200 OCs known. Now, this number has been increased up to more than 2000 OCs
+available at the Vizier catalogue \cite[][\url{https://vizier.unistra.fr}]{ochsenbein2000vizier}.
+
+The method developed in this project takes as base the approach made by \citeauthor{castro2020hunting} but rapidly diverges.
+In first place, performance has been taken into account since the begining, trying to get a reliable method without compromising the
+needs of supercomputers.
+To achieve this, instead of blindly analyzing the whole galactic disk a fields selection is made focusing in those regions with known OCs.
+Then, these fields are analyzed with an homogeneous methodology based on machine learning algorithms using as sources astrometric and
+photometric variables. The goal is the characterization of each individual star in the cluster, discerning wether the star belongs the OC
+or not.
+
+Thus, the main goal of this project is not the discovery of new open clusters or even candidates, but the development of a model that,
+applied to already knonw open clusters, allows us the characterization of those OCs on an unsupervised way, with a reasonable computing
+power and taking into account the same selection of astrometric and photometric variables for every OC.
+
+\chapter{Requirements Description}
+
 \chapter{Work Development}
@@
 \chapter{Conclusions and Future Work}
-
-\chapter*{Acknowledgement}
-\addcontentsline{toc}{chapter}{Acknowledgement}
-
-This work has made use of data from the European Space Agency (ESA) mission
-{\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed by the {\it Gaia}
-Data Processing and Analysis Consortium (DPAC,
-\url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding for the DPAC
-has been provided by national institutions, in particular the institutions
-participating in the {\it Gaia} Multilateral Agreement.
-
-\medskip
-
-This publication makes use of VOSA, developed under the Spanish Virtual Observatory project
-supported by the Spanish MINECO through grant AyA2017-84089.
-VOSA has been partially updated by using funding from the European Union's Horizon 2020 Research
-and Innovation Programme, under Grant Agreement nº 776403 (EXOPLANETS-A)
-
-\medskip
-
-This research has made use of the VizieR catalogue access tool, CDS, Strasbourg, France (DOI : 10.26093/cds/vizier).
-The original description of the VizieR service was published in 2000, \cite[A\&AS 143, 23]{ochsenbein2000vizier}
 

### 2020-10-23 code/cdalvaro/data_base.py
fix: Raise error if region is not available in DB

@@ -221,6 +221,9 @@ class DB:
         params = dict()
         if use_region_id:
             region_ids = self.get_regions_id(regions=set([region])).values()
+            if len(region_ids) == 0:
+                raise RuntimeError(f"Region '{region.name}' is not available in DB")
+
             params['region_id'] = next(iter(region_ids))
             query += """
                 region_id = %(region_id)s

### 2020-10-24 code/cdalvaro/catalogues/openclust/__init__.py
python: Rename code directory to src

@@ -1,1 +1,0 @@
-from .openclust import Catalogue, OpenClust

### 2020-10-24 code/cdalvaro/graphics/__init__.py

@@ -1,2 +1,0 @@
-from .color_palette import ColorPalette, set_color_palette
-from .plot import Plot

### 2020-10-24 code/cdalvaro/graphics/plot.py

@@ -1,131 +1,0 @@
-import matplotlib.pyplot as plt
-import numpy as np
-import seaborn as sns
-
-
-class Plot:
-
-    default_ext = 'pdf'
-
-    def __init__(self, save_figs: bool = False, figs_path: str = '.'):
-        self.save_figs = save_figs
-        self.figs_path = figs_path
-
-    def plot_clusters_catalogue_distribution(self,
-                                             df_catalogue,
-                                             img_name: str,
-                                             img_ext: str = default_ext,
-                                             title: str = None,
-                                             xlim: tuple = None,
-                                             ylim: tuple = None):
-        fig, ax = plt.subplots(figsize=(12, 6))
-        if title is not None:
-            ax.set_title(title)
-
-        ax.set_xlabel('Right Ascension (J2000 Degree)')
-        ax.set_ylabel('Declination (J2000 Degree)')
-
-        if xlim is not None:
-            ax.set_xlim(xlim)
-        if ylim is not None:
-            ax.set_ylim(ylim)
-
-        g = sns.scatterplot(data=df_catalogue,
-                            x="ra",
-                            y="dec",
-                            hue="diam",
-                            size="diam",
-                            palette="ch:r=-.5,l=.75",
-                            ax=ax)
-        plt.legend().set_title("Diameter (arcmin)")
-
-        self._save_figure(fig, name=img_name, img_ext=img_ext)
-
-    def plot_cluster_proper_motion(self,
-                                   df_cluster,
-                                   img_name: str,
-                                   img_ext: str = default_ext,
-                                   title: str = None,
-                                   xlim: tuple = None,
-                                   ylim: tuple = None):
-        fig, ax = plt.subplots(figsize=(6, 6))
-        if title is not None:
-            ax.set_title(title)
-
-        ax.set_xlabel(r'Proper Motion in Right Ascension Direction ($mas \cdot year^{-1}$)')
-        ax.set_ylabel(r'Proper Motion in Declination Direction ($mas \cdot year^{-1}$)')
-
-        if xlim is not None:
-            ax.set_xlim(xlim)
-        if ylim is not None:
-            ax.set_ylim(ylim)
-
-        g = sns.scatterplot(data=df_cluster, x="pmra", y="pmdec", hue="is_contained", s=12, ax=ax, legend=False)
-
-        self._save_figure(fig, name=img_name, img_ext=img_ext)
-
-    def plot_cluster_parallax_histogram(self,
-                                        df_cluster,
-                                        img_name: str,
[…]

### 2020-10-24 src/cdalvaro/catalogues/openclust/openclust.py

@@ -1,0 +1,159 @@
+import astropy.units as u
+from astropy.coordinates import SkyCoord
+import os
+import pandas
+from typing import Dict, Set, TypeVar, Union
+
+from ..base_catalogue import BaseCatalogue
+from ...logging import Logger
+from ...models import OpenCluster
+
+Catalogue = TypeVar('Catalogue', bound=Dict[str, OpenCluster])
+
+
+class OpenClust(BaseCatalogue):
+    """
+    Class that contains the OpenClust catalogue.
+
+    More info at: https://heasarc.gsfc.nasa.gov/W3Browse/star-catalog/openclust.html
+    """
+
+    _catalogue_file = os.path.join(os.path.dirname(__file__), "clusters.dat")
+    _catalogue = dict()
+    _logger = Logger.instance()
+
+    @staticmethod
+    def catalogue(as_dataframe: bool = False) -> Union[Catalogue, pandas.DataFrame]:
+        """
+        This method returns the whole OPENCLUST catalogue
+
+        Args:
+            as_dataframe (bool, optional): Flag to recover the catalogue as a DataFrame. Defaults to False.
+
+        Returns:
+            Union[Catalogue, pandas.DataFrame]: The OPENCLUST catalogue.
+        """
+        catalogue = OpenClust._load_catalogue()
+        if as_dataframe:
+            return OpenClust._catalogue_to_dataframe(catalogue)
+        return catalogue
+
+    @staticmethod
+    def get_clusters(names: Set[str], as_dataframe: bool = False) -> Union[Catalogue, pandas.DataFrame]:
+        """
+        Get clusters from the catalogue that matches the given names.
+
+        Args:
+            names (Set[str]): The name of the clusters to be recovered.
+            as_dataframe (bool, optional): Flag to recover clusters as a DataFrame. Defaults to False.
+
+        Returns:
+            Catalogue: A catalogue with the found clusters.
+        """
+        selection = dict()
+        whole_catalogue = OpenClust._load_catalogue()
+        for name in names:
+            if name in whole_catalogue.keys():
+                selection[name] = whole_catalogue[name]
+
+        if as_dataframe:
+            return OpenClust._catalogue_to_dataframe(selection)
+
+        return selection
+
+    @staticmethod
+    def _load_catalogue() -> Catalogue:
+        """
+        Load the OpenClust catalogue.
+
+        Returns:
+            Catalogue: The whole OpenClust catalogue.
+        """
[…]

### 2020-10-24 src/cdalvaro/graphics/__init__.py

@@ -1,0 +1,2 @@
+from .color_palette import ColorPalette, set_color_palette
+from .plot import Plot

### 2020-10-24 src/cdalvaro/logging.py

@@ -1,0 +1,61 @@
+import logging
+import sys
+
+
+class Logger():
+
+    _logger: logging.Logger = None
+
+    @staticmethod
+    def instance() -> logging.Logger:
+        if Logger._logger:
+            return Logger._logger
+
+        Logger._logger = logging.getLogger('Gaia Downloader')
+
+        # Console handler
+        stderr_ch = logging.StreamHandler(stream=sys.stderr)
+        stderr_ch.setFormatter(CustomFormatter())
+        stderr_ch.addFilter(lambda record: record.levelno >= logging.WARNING)
+        Logger._logger.addHandler(stderr_ch)
+
+        stdout_ch = logging.StreamHandler(stream=sys.stdout)
+        stdout_ch.setFormatter(CustomFormatter())
+        stdout_ch.addFilter(lambda record: record.levelno < logging.WARNING)
+        Logger._logger.addHandler(stdout_ch)
+
+        return Logger._logger
+
+
+class ColorCodes:
+    grey = "\033[0;37m"
+    green = "\033[0;32m"
+    yellow = "\033[1;33m"
+    red = "\033[0;31m"
+    bold_red = "\033[31;1m"
+    blue = "\033[0;34m"
+    purple = "\033[0;35m"
+    reset = "\033[0m"
+
+
+def _custom_format(color: str):
+    return f"{color}%(levelname)s{ColorCodes.reset}: %(message)s"
+
+
+class CustomFormatter(logging.Formatter):
+    """Logging Formatter to add colors and count warning / errors"""
+
+    datefmt = "%Y-%m-%d %H:%M:%S"
+
+    FORMATS = {
+        logging.DEBUG: _custom_format(color=ColorCodes.purple),
+        logging.INFO: _custom_format(color=ColorCodes.green),
+        logging.WARNING: _custom_format(color=ColorCodes.yellow),
+        logging.ERROR: _custom_format(color=ColorCodes.red),
+        logging.CRITICAL: _custom_format(color=ColorCodes.bold_red)
+    }
+
+    def format(self, record):
+        log_fmt = self.FORMATS.get(record.levelno)
+        formatter = logging.Formatter(log_fmt)
+        return formatter.format(record)

### 2020-10-24 src/cdalvaro/models/__init__.py

@@ -1,0 +1,2 @@
+from .open_cluster import OpenCluster
+from .region import Region

### 2020-10-24 src/scripts/downloader.py

@@ -1,0 +1,102 @@
+#!/usr/bin/env python3
+
+import argparse
+from cdalvaro import DB, Logger
+from cdalvaro.catalogues import OpenClust
+from cdalvaro.downloaders import Gaia
+import logging
+import numpy as np
+import os
+
+parser = argparse.ArgumentParser(prog='Gaia DR2 Downloader', description='✨ Stars downloader from the GAIA DR2 dataset')
+parser.add_argument('--cluster',
+                    '-c',
+                    nargs='+',
+                    type=str,
+                    default={'ALL'},
+                    help="""
+        Cluster name from http://cdsarc.u-strasbg.fr/ftp/cats/B/ocl/clusters.dat list.
+        Type ALL to download all of them.
+        """)
+parser.add_argument('--exclude',
+                    '-e',
+                    nargs='*',
+                    type=str,
+                    default={},
+                    help="""
+        Clusters to be ignored from the download. Default to empty.
+        """)
+parser.add_argument('--extra-size',
+                    '-s',
+                    dest='extra_size',
+                    type=float,
+                    default=1.5,
+                    help="""
+        Extra size to extend regions to be downloaded.
+        Default to 1.5.
+        """)
+
+update_data_parser = parser.add_mutually_exclusive_group(required=False)
+update_data_parser.add_argument('--update-data',
+                                '-u',
+                                dest='update_data',
+                                action='store_true',
+                                help="""
+        Update existing clusters with new data. Default.
+        """)
+update_data_parser.add_argument('--no-update-data', dest='update_data', action='store_false')
+update_data_parser.set_defaults(update_data=True)
+
+parser.add_argument('--verbose', '-v', action='count', default=0)
+
+args = parser.parse_args()
+
+# Logger settings
+log_level = logging.WARNING
+if args.verbose == 1:
+    log_level = logging.INFO
+elif args.verbose > 1:
+    log_level = logging.DEBUG
+
+logger = Logger.instance()
+logger.setLevel(log_level)
+
+# Select clusters
+catalogue = OpenClust.catalogue()
+if len(args.cluster) == 1 and 'ALL' in args.cluster:
+    clusters = set(catalogue.values())
+else:
+    clusters = set()
+    for cluster_name in args.cluster:
+        if cluster_name in catalogue.keys():
[…]

### 2020-10-24 thesis/thesis.tex
thesis: Set cites APA style

@@
 \usepackage{graphicx}
-\usepackage{natbib}
+\usepackage{apacite}
 \usepackage{picture}
@@
 % Bibliography style
-\bibliographystyle{abbrvnat}
-\setcitestyle{authoryear,open={(},close={)}}
+\bibliographystyle{apacite}
 
@@
 
-Stellar open clusters (OC) \cite[]{janes1982open} are groups of stars gravitationally bound originated from a single molecular gas cloud.
+Stellar open clusters (OC) \cite{janes1982open} are groups of stars gravitationally bound originated from a single molecular gas cloud.
 Thus they share the same chemical composition and age, and they have similar relative positions and proper motion.
@@
 in their original molecular cloud
-\cite[]{clarke2000theformationofstellarclusters} \cite[]{portegies2010young}.
+\cite{clarke2000theformationofstellarclusters} \cite{portegies2010young}.
 The evolution of these systems tends to sparse them in a few million years by interacting gravitationally with other systems.
 Galactic tidal forces and mechanisms that involves the gas loss driven by stellar feedback are other causes of disruption
-\cite[]{brinkmann2017bound}.
+\cite{brinkmann2017bound}.
 Nevertheless, a small fraction of these systems will survive in the initial state and persist bound in bigger timescales.
@@
 Some highly disturbed orbits could also provide evidence of recent merge events and accretion traces from outside the galaxy
-\cite[]{cantat2016abundances}.
+\cite{cantat2016abundances}.
 
 The study of OC has been pushed forward thanks to the huge and precise dataset from the Gaia mission
-\cite[]{collaboration2016description} Gaia DR2 \cite[]{gaia2018gaia}, available since 2018.
+\cite{collaboration2016description} Gaia DR2 \cite{gaia2018gaia}, available since 2018.
 This dataset has helped to review already known open clusters and to find new ones.
@@
 
-The OpenClust catalogue \cite[]{dias2002new} has been used to make a preselection of sky regions for the Gaia DR2 data download.
+The OpenClust catalogue \cite{dias2002new} has been used to make a preselection of sky regions for the Gaia DR2 data download.
 This download is not limited to the sized registered in the catalogue. Instead, a wider region is downloaded for each cluster to include
@@
 This work tries to take the best from \emph{Clusterix 2.0. A Virtual Observatory tool to estimate cluster membership probability}
-\cite[]{balaguer2020clusterix} and \emph{Hunting for open clusters in Gaia DR2: 582 new open clusters
-in the Galactic disc} \cite[]{castro2020hunting}, and merge them into a single \emph{machine learning}
+\cite{balaguer2020clusterix} and \emph{Hunting for open clusters in Gaia DR2: 582 new open clusters
+in the Galactic disc} \cite{castro2020hunting}, and merge them into a single \emph{machine learning}
 tool capable of finding open clusters eficiently without requiring powerful machines to perform the computations.
@@
 (VO) framework and works together with other VO tool such as:
-\emph{TOPCAT} \cite[][\url{http://www.starlink.ac.uk/topcat/}]{taylor2005topcat},
-\emph{Aladin} \cite[][\url{https://aladin.u-strasbg.fr/aladin.gml}]{bonnarel2000aladin},
-\emph{VOSA} \cite[][\url{http://svo2.cab.inta-csic.es/theory/vosa/}]{bayo2008vosa}, etc.
+\emph{TOPCAT} \cite{taylor2005topcat},
+\emph{Aladin} \cite{bonnarel2000aladin},
+\emph{VOSA} \cite{bayo2008vosa}, among other.
 
@@
 Before this approach was put into practice there were 1200 OCs known. Now, this number has been increased up to more than 2000 OCs
-available at the Vizier catalogue \cite[][\url{https://vizier.unistra.fr}]{ochsenbein2000vizier}.
+available at the Vizier catalogue \cite{ochsenbein2000vizier}.
 

### 2020-11-04 src/cdalvaro/graphics/color_palette.py
feat: Update graphics package

@@ -1,22 +1,13 @@
-import seaborn
+import seaborn as sns
 
 
 def set_color_palette():
     """
     Set seaborn color palette
     """
-    seaborn.set_palette(ColorPalette.palette)
+    sns.set_palette(color_palette(as_cmap=False))
 
 
-class ColorPalette:
-    """
-    Color palette extracted from macOS Big Sur wallpaper
-    """
-
-    palette = seaborn.color_palette(
-        ('#5596cb', '#6b3e79', '#dc3759', '#eeb760', '#2c4c7c', '#4b4378', '#69addf', '#eca25c', '#e8765a', '#a83a76',
-         '#2f6cab', '#8cccf7', '#e1bf94', '#b2c2e0', '#cd4676', '#d5e0f1', '#1e406a', '#4484c4', '#873d7a', '#b289af',
-         '#448cc4', '#244c74', '#3778b3', '#c83867', '#e14859', '#7fc1f1', '#c77391', '#a799c6', '#d2c6bc', '#e3635a',
-         '#143c5c', '#96b2db', '#654c7e', '#bac0bc', '#a33c84', '#b63d64', '#ba7c9b', '#ec675c', '#e15d6c', '#374171',
-         '#4e86ba', '#e55456', '#7da5cb', '#3b447c', '#e6624c', '#d8a4bc', '#5e89b9', '#4481b8', '#3c6c9c', '#4c84c4',
-         '#ae638e', '#e8a195', '#54a4d4', '#2c74ac', '#448cbc', '#f46464', '#d86444', '#d4605c', '#3f74a4'))
+def color_palette(as_cmap: bool = True):
+    # https://medium.com/@morganjonesartist/color-guide-to-seaborn-palettes-da849406d44f
+    return sns.color_palette('BrBG_r', as_cmap=as_cmap)

### 2020-11-04 src/cdalvaro/graphics/plot.py

@@ -2,6 +2,8 @@ import matplotlib.pyplot as plt
 import numpy as np
 import seaborn as sns
 
+from .color_palette import color_palette, set_color_palette
+
 
 class Plot:
 
@@ -10,6 +12,7 @@ class Plot:
     def __init__(self, save_figs: bool = False, figs_path: str = '.'):
         self.save_figs = save_figs
         self.figs_path = figs_path
+        set_color_palette()
 
     def plot_clusters_catalogue_distribution(self,
                                              df_catalogue,
@@ -17,7 +20,8 @@ class Plot:
                                              img_ext: str = default_ext,
                                              title: str = None,
                                              xlim: tuple = None,
-                                             ylim: tuple = None):
+                                             ylim: tuple = None,
+                                             hue: str = 'diam'):
         fig, ax = plt.subplots(figsize=(12, 6))
         if title is not None:
             ax.set_title(title)
@@ -30,16 +34,13 @@ class Plot:
         if ylim is not None:
             ax.set_ylim(ylim)
 
-        g = sns.scatterplot(data=df_catalogue,
-                            x="ra",
-                            y="dec",
-                            hue="diam",
-                            size="diam",
-                            palette="ch:r=-.5,l=.75",
-                            ax=ax)
+        g = sns.scatterplot(data=df_catalogue, x="ra", y="dec", hue=hue, size=hue, palette=color_palette(), ax=ax)
         plt.legend().set_title("Diameter (arcmin)")
 
-        self._save_figure(fig, name=img_name, img_ext=img_ext)
+        if self.save_figs:
+            self._save_figure(fig, name=img_name, img_ext=img_ext)
+
+        return fig, ax, g
 
     def plot_cluster_proper_motion(self,
                                    df_cluster,
@@ -47,7 +48,9 @@ class Plot:
                                    img_ext: str = default_ext,
                                    title: str = None,
                                    xlim: tuple = None,
-                                   ylim: tuple = None):
+                                   ylim: tuple = None,
+                                   hue: str = 'clusters',
+                                   legend: bool = True):
         fig, ax = plt.subplots(figsize=(6, 6))
         if title is not None:
             ax.set_title(title)
@@ -60,9 +63,19 @@ class Plot:
         if ylim is not None:
             ax.set_ylim(ylim)
 
-        g = sns.scatterplot(data=df_cluster, x="pmra", y="pmdec", hue="is_contained", s=12, ax=ax, legend=False)
[…]

### 2020-11-04 src/cdalvaro/ml/dec/__init__.py
feat: Update ML package

@@ -1,1 +1,3 @@
+from .autoencoder import autoencoder
+from .clustering_layer import ClusteringLayer
 from .dec import DEC

### 2020-11-04 src/cdalvaro/ml/dec/autoencoder.py

@@ -8,7 +8,7 @@ KernelInitializer = TypeVar('KernelInitializer', bound=Union[str, Initializer])
 
 def autoencoder(dims: List[int],
                 activation: str = 'relu',
-                kernel_initializer: KernelInitializer = 'glorot_uniform') -> Tuple[Model, Model]:
+                initializer: KernelInitializer = 'glorot_uniform') -> Tuple[Model, Model]:
     """
     Fully connected symmetric autoencoder model.
 
@@ -16,7 +16,7 @@ def autoencoder(dims: List[int],
         dims (List[int]): List of the sizes of layers of encoder.
                           dims[0] is input dim, dims[-1] is size of the latent hidden layer.
         activation (str, optional): Activation function. Defaults to 'relu'.
-        kernel_initializer (KernelInitializer, optional): Kernel initializer. Defaults to 'glorot_uniform'.
+        initializer (KernelInitializer, optional): Kernel initializer. Defaults to 'glorot_uniform'.
 
     Returns:
         Tuple[Model, Model]: Autoencoder and encoder models
@@ -30,22 +30,20 @@ def autoencoder(dims: List[int],
     for i_layer in range(n_stacks - 1):
         layer = Dense(dims[i_layer + 1],
                       activation=activation,
-                      kernel_initializer=kernel_initializer,
+                      kernel_initializer=initializer,
                       name=f"encoder_{i_layer}")(layer)
 
     # Latent hidden layer
-    encoded_layer = Dense(dims[-1], kernel_initializer=kernel_initializer, name=f"encoder_{n_stacks - 1}")(layer)
+    encoded_layer = Dense(dims[-1], kernel_initializer=initializer, name=f"encoder_{n_stacks - 1}")(layer)
 
     # Decoder internal layers
     layer = encoded_layer
     for i_layer in range(n_stacks - 1, 0, -1):
-        layer = Dense(dims[i_layer],
-                      activation=activation,
-                      kernel_initializer=kernel_initializer,
+        layer = Dense(dims[i_layer], activation=activation, kernel_initializer=initializer,
                       name=f"decoder_{i_layer}")(layer)
 
     # Decoder output
-    layer = Dense(dims[0], kernel_initializer=kernel_initializer, name='decoder_0')(layer)
+    layer = Dense(dims[0], kernel_initializer=initializer, name='decoder_0')(layer)
     decoded_layer = layer
 
     _autoencoder = Model(inputs=input_layer, outputs=decoded_layer, name='autoencoder')

### 2020-11-04 src/cdalvaro/ml/dec/clustering_layer.py

@@ -3,13 +3,24 @@ from keras.layers import InputSpec, Layer
 
 
 class ClusteringLayer(Layer):
-    """
-    This layer converts the input sample to soft label, i.e. a vector that represents
-    the probability of the sample belonging to each cluster.
-
-    The probability is calculated with student's t-distribution.
-    """
     def __init__(self, n_clusters, weights=None, alpha: float = 1.0, **kwargs):
+        """
+        This layer converts the input sample to soft label, i.e. a vector that represents
+        the probability of the sample belonging to each cluster.
+
+        The probability is calculated with student's t-distribution.
+
+        Args:
+            n_clusters (int): Number of clusters.
+            weights (optional): List of Numpy array with shape `(n_clusters, n_features)` which represents the initial cluster centers.
+            alpha (float, optional): Degrees of freedom of the Student's t-distribution. Defaults to 1.0.
+
+        Input shape:
+            2D tensor with shape: `(n_samples, n_features)`
+
+        Output shape:
+            2D tensor with shape: `(n_samples, n_clusters)`
+        """
         if 'input_shape' not in kwargs and 'input_dim' in kwargs:
             kwargs['input_shape'] = (kwargs.pop('input_dim'), )
         super(ClusteringLayer, self).__init__(**kwargs)

### 2020-11-04 src/cdalvaro/ml/dec/dec.py

@@ -1,37 +1,49 @@
-from enum import auto
-from keras.backend.cntk_backend import update
 from keras.models import Model
-from keras.optimizers import Optimizer, SGD
+from keras.optimizers import Adam, Optimizer, SGD
 import numpy as np
-from scipy.cluster.hierarchy import weighted
+import os
 from sklearn.cluster import KMeans
-from sklearn.metrics import silhouette_score
 from typing import List, Union
 
 from .autoencoder import autoencoder, KernelInitializer
 from .clustering_layer import ClusteringLayer
+from ..utils import estimate_n_clusters
 
 
-class DEC:
-    def __init__(self, data, dims: List[int] = None, n_clusters: int = None, n_jobs: int = None, save_dir: str = None):
-        # TODO: Add documentation
-        if dims is None:
-            self._dims = [data.shape[-1], 500, 500, 2000, 10]
-        elif dims[0] != data.shape[-1]:
-            ValueError(f"First dims value must be equal to data.shape[-1] -> {dims[0]} != {data.shape[-1]}")
-        else:
-            self._dims = dims
+class DEC(object):
+    def __init__(self,
+                 dims: List[int],
+                 n_clusters: int,
+                 activation: str = 'relu',
+                 alpha: float = 1.0,
+                 initializer: KernelInitializer = 'glorot_uniform'):
+        """
+        Unsupervised Deep Embedding for Clustering Analysis (DEC)
+        Link: http://proceedings.mlr.press/v48/xieb16.pdf
 
-        self._data = data
+        DEC learns a mapping from the data space to a lower-dimensional feature space
+        in which it iteratively optimizes a clustering objective.
+
+        Args:
+            dims (List[int]): DEC layers configuration. First element must be the number of features.
+            n_clusters (int): The number of clusters to be found.
+            activation (str, optional): Activation function used in encoder models. Defaults to relu.
+            alpha (float, optional): Degrees of freedom of the Student's t-distribution. Defaults to 1.0.
+            initializer (KernelInitializer, optional): Kernel initializer used in encoder models. Defaults to glorot_uniform.
+            optimizer (Optimizer, optional): Optimizer used in encoder models. Defaults to SGD.
+        """
+        super(DEC, self).__init__()
+
+        self._dims = dims
         self._n_clusters = n_clusters
-        self._n_jobs = n_jobs
-        self.save_dir = save_dir
-        self._kmeans = None
-        self._model = None
-        self._pretrain_autoencoder = None
-        self._encoder = None
[…]

### 2020-11-04 src/cdalvaro/graphics/color_palette.py
python: Update notebooks

@@ -8,6 +8,6 @@ def set_color_palette():
     sns.set_palette(color_palette(as_cmap=False))
 
 
-def color_palette(as_cmap: bool = True):
+def color_palette(as_cmap: bool = True, **kwargs):
     # https://medium.com/@morganjonesartist/color-guide-to-seaborn-palettes-da849406d44f
-    return sns.color_palette('BrBG_r', as_cmap=as_cmap)
+    return sns.color_palette('BrBG_r', as_cmap=as_cmap, **kwargs)

### 2020-11-04 src/cdalvaro/graphics/plot.py

@@ -49,7 +49,7 @@ class Plot:
                                    title: str = None,
                                    xlim: tuple = None,
                                    ylim: tuple = None,
-                                   hue: str = 'clusters',
+                                   hue: str = 'cluster_g',
                                    legend: bool = True):
         fig, ax = plt.subplots(figsize=(6, 6))
         if title is not None:
@@ -86,7 +86,7 @@ class Plot:
                                         xlim: tuple = None,
                                         ylim: tuple = None,
                                         stat: str = 'count',
-                                        hue: str = 'clusters',
+                                        hue: str = 'cluster_g',
                                         legend: bool = True):
         fig, ax = plt.subplots(figsize=(6, 6), tight_layout=True)
         if title is not None:
@@ -122,7 +122,7 @@ class Plot:
                                      title: str = None,
                                      xlim: tuple = None,
                                      ylim: tuple = None,
-                                     hue: str = 'clusters',
+                                     hue: str = 'cluster_g',
                                      legend: bool = True):
         fig, ax = plt.subplots(figsize=(6, 6))
         if title is not None:

### 2020-11-07 src/cdalvaro/graphics/plot.py
python: Fix plot issues

@@ -1,5 +1,5 @@
 import matplotlib.pyplot as plt
-import numpy as np
+import pandas as pd
 import seaborn as sns
 
 from .color_palette import color_palette, set_color_palette
@@ -34,7 +34,13 @@ class Plot:
         if ylim is not None:
             ax.set_ylim(ylim)
 
-        g = sns.scatterplot(data=df_catalogue, x="ra", y="dec", hue=hue, size=hue, palette=color_palette(), ax=ax)
+        g = sns.scatterplot(data=df_catalogue,
+                            x="ra",
+                            y="dec",
+                            hue=hue,
+                            size=hue,
+                            palette=color_palette(n_colors=len(pd.unique(df_catalogue[hue]))),
+                            ax=ax)
         plt.legend().set_title("Diameter (arcmin)")
 
         if self.save_figs:
@@ -67,7 +73,7 @@ class Plot:
                             x="pmra",
                             y="pmdec",
                             hue=hue,
-                            palette=color_palette(),
+                            palette=color_palette(n_colors=len(pd.unique(df_cluster[hue]))),
                             s=12,
                             ax=ax,
                             legend=legend)
@@ -103,7 +109,7 @@ class Plot:
         g = sns.histplot(data=df_cluster,
                          x='parallax',
                          hue=hue,
-                         palette=color_palette(),
+                         palette=color_palette(n_colors=len(pd.unique(df_cluster[hue]))),
                          legend=legend,
                          bins=bins,
                          kde=True,
@@ -140,7 +146,7 @@ class Plot:
                             x="bp_rp",
                             y="phot_g_mean_mag",
                             hue=hue,
-                            palette=color_palette(),
+                            palette=color_palette(n_colors=len(pd.unique(df_cluster[hue]))),
                             size='parallax',
                             sizes=(2, 20),
                             ax=ax,

### 2020-11-07 src/cdalvaro/ml/dec/dec.py
dec: Improve DEC model

@@ -43,7 +43,7 @@ class DEC(object):
 
         # DEC model
         clustering_layer = ClusteringLayer(n_clusters, alpha=alpha, name='clustering')(self._encoder.output)
-        self._model = Model(inputs=self._encoder.input, outputs=[clustering_layer, self._autoencoder.output])
+        self._model = Model(inputs=self._encoder.input, outputs=clustering_layer)
 
     @property
     def dims(self):
@@ -161,11 +161,13 @@ class DEC(object):
         # DEC - Phase 2: Optimization
         # Reference:
         #     Unsupervised Deep Embedding for Clustering Analysis - 3.1 Clustering with KL divergence
-        loss, index = 0, 0
+        loss, index = [0.0], 0
         index_array = np.arange(x.shape[0])
         for it in range(maxiter):
             if it % update_interval == 0:
-                q, _ = self._model.predict(x, verbose=0)
+                if verbose > 0:
+                    print(f"Iteration {it + 1}/{maxiter} - loss: {np.max(loss):.4e}")
+                q = self._model.predict(x, verbose=verbose)
                 p = self._target_distribution(q)
                 y_pred = q.argmax(1)
 
@@ -180,10 +182,8 @@ class DEC(object):
 
             # Train on batch
             idx = index_array[index * batch_size:min((index + 1) * batch_size, x.shape[0])]
-            loss = self._model.train_on_batch(x=x[idx], y=[p[idx], x[idx]])
+            loss = self._model.train_on_batch(x=x[idx], y=p[idx])
             index = index + 1 if (index + 1) * batch_size <= x.shape[0] else 0
-            if verbose > 0:
-                print(f"Iteration {it + 1}/{maxiter} - loss: {np.max(loss):.4e}")
 
         self._trained = True
         if save_dir is not None:
@@ -205,7 +205,7 @@ class DEC(object):
         """
         if not self._trained:
             raise Exception("This model has not been trained yet")
-        q, _ = self._model.predict(x, verbose=verbose)
+        q = self._model.predict(x, verbose=verbose)
         return q.argmax(1)
 
     @staticmethod

### 2020-11-09 thesis/thesis.tex
thesis: Add structure comments
thesis: Update structure

@@
 
+%% TODO: 1.1 Motivation
+
+% ¿Cuál es el problema a tratar?
+% ¿Cuáles crees que son las causas?
+% ¿Por qué es relevante el problema?
+
+%% TODO: 1.2 Planteamiento
+
+% ¿Cómo se podría solucionar el problema?
+% ¿Qué es lo que se propone?
+% Descripción del objetivo del trabajo en términos generales
+
+%% TODO: 1.3 Estructura de la memoria
+
+% Descripción breve de lo que se va a contar en cada uno de los capítulos siguientes.
+
 Stellar open clusters (OC) \cite{janes1982open} are groups of stars gravitationally bound originated from a single molecular gas cloud.
@@
 This dataset has helped to review already known open clusters and to find new ones.
-
-\section{Aims}
-
-The goal of this work is to \emph{achieve a viable model for open cluster characterization based on machine learning techniques}
-such as clustering algorithms and artificial neural networks.
-This model must be a non-supervised non-parameterized algorithm valid for any selected OC.
-
-The OpenClust catalogue \cite{dias2002new} has been used to make a preselection of sky regions for the Gaia DR2 data download.
-This download is not limited to the sized registered in the catalogue. Instead, a wider region is downloaded for each cluster to include
-stars from outside the cluster and take them into account when training the model so the model will be able to take apart those outsiders
-when characterizing a non known cluster.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=\columnwidth]{../figures/openclust_catalogue.pdf}
-  \caption{OpenClust Catalogue Distribution}
-\end{figure}
-
-A later filter will be applied to select only those OCs that fulfil one or more of the following criteria:
-
-\begin{itemize}
-  \item Cluster diameter above 25.0 arcmin
-  \item Number of stars in the selected region above 40,000 stars
-\end{itemize}
-
-The final downloaded set covers nearly 60 million stars, which is significantly smaller than the whole Gaia DR2 dataset which contains
-information for around 1,600 million stars.
-This restriction responds to a commitment to the project delivery dates and the available computing power,
-and to having a good basis for the correct model training. As an extension of this work, the trained model could be applied to a region
-not covered by this preselection for a later understanding of the results.
-
-The intention is to get a reliable and efficient non-parameterized neither supervised model over those parameterized or supervised ones.
-To measure the reliability of the resulting model comparisons with other VO tools will be made.
-
-The role for the ANN will be validating cluster members based on dynamincs and photometric data.
-
-These are te variables to be considered in this work:
-
-\begin{itemize}
-  \item OCs and stars \emph{right ascension}: $\alpha$ and \emph{declination}: $\delta$
-  \item OCs \emph{estimated sizes}: $r$
-  \item \emph{Proper motion}: $\mu_{\alpha}$, $\mu_{\delta}$ and \emph{parallax}: $\varpi$
-  \item \emph{Photometric}: $E(B-R)$, $G_{mag}$
-  \item \emph{Associated errors} will be taken into account to estimate prediction errors
-\end{itemize}
-
-Initially, no filter will be applied over parallax neither magnitude values to the selected objects from Gaia DR2.
 
@@
 
-\chapter{Requirements Description}
+%% TODO: Concluir con una última sección de resumen de conclusiones, resumiendo las principales averiguaciones del estudio del dominio
+% y cómo van a afectar al desarrollo específico del trabajo.
 
-\chapter{Work Development}
+\chapter{Aims}
+
+\section{General}
+
+The main aim of this thesis is to \emph{build an unsupervised clustering model for open cluster characterization}.
+This model must be \emph{non-supervised non-parameterized}, and has to be valid for a wide range of open clusters.
+
+To measure the reliability of the resulting model comparisons with other VO tools will be made.
+
+The role for the ANN will be validating cluster members based on dynamincs data.
+
+These are te variables to be considered in this work:
+
+\begin{itemize}
+  \item OCs and stars \emph{right ascension}: $\alpha$ and \emph{declination}: $\delta$
+  \item OCs \emph{estimated sizes}: $r$
+  \item \emph{Proper motion}: $\mu_{\alpha}$, $\mu_{\delta}$ and \emph{parallax}: $\varpi$
+  \item \emph{Associated errors} will be taken into account to estimate prediction errors
+\end{itemize}
+
+\section{Specific}
+
+To achive this goal, the aim will be divided into the following milestones:
+
+\begin{itemize}
+  \item To recover data from Gaia DR2 database. This data will be taken as source for the machine learning model to characterize open clusters
+  by grouping stars into clusters.
+  \item To research unsupervised clustering algorithms suitable for grouping stars previously recovered.
+  \item To select and implement unsupervised clustering algorithms from the previous study.
+  \item To apply implemented tools to different datasets from first point to find OCs.
+  \item To validate found OCs by determining their accuracy and precission.
+\end{itemize}
+
+\chapter{Method}
+
+In this chapter we address the different steps taken to perform the creation of the unsupervised clustering model for open cluster characterization.
+
+The development has been made with the Python programming language. Other auxiliary tools such as \emph{Docker}, \emph{PostgreSQL}, \emph{Jupyter Notebooks},
+and \emph{Scikit-Learn} and \emph{Keras} frameworks have been used too.
+
+\section{Data Mining}
+
+As mentioned before, Gaia DR2 is the chosen data source for this work. Thus, the first step is to download the required data from Gaia DB
+and store it in a custom database for later access.
+
+Due to the large amount of data inside Gaia database, a complete download is not viable neither useful.
+In order to squize the size of the dataset to be downloaded, the OpenClust catalogue \cite{dias2002new} (Figure \ref{fig:OpenClustComplete})
+has been used to make a preselection of sky regions.
+
+This download is not limited to the size registered in the catalogue. Instead, a wider region is downloaded for each cluster to include
+stars outside the cluster and take them into account when training the model so the model will be able to take apart those outsiders
+when characterizing the cluster.
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=\columnwidth]{../figures/openclust_catalogue.pdf}
+  \caption{OpenClust Catalogue Distribution}
+  \label{fig:OpenClustComplete}
+\end{figure}
+
+The downloaded dataset covers nearly 114 million stars (42GB of compressed data), that is significantly smaller than the whole Gaia DR2 dataset, which contains
+information for around 1,600 million stars, but stils being too large.
+
+Since not all the downloaded clusters are good enough, the following filters are applied to discard some clusters which don't have enough stars or contain too
+many \emph{null} values:
+
+\begin{itemize}
+  \item Cluster diameter above 25.0 arcmin
+  \item Parallax absolute value greater than 0.0
+  \item Number of stars in the selected region above 40,000 stars
+\end{itemize}
+
+There is no upper limit to the number of stars inside a cluster, although for the sake of this work, smaller cluster will be prefered as
+a commitment to the project delivery dates and the available computing power. As an extension of this work, the trained model could be applied to a region
+not covered by this preselection for a later understanding of the results.
+
+%% TODO: Las tecnologías empleadas para el desarrollo de las herramientas,
+% la organización del piloto, cómo ha transcurrido el experimento
+
+%% TODO: Añadir nuevo capítulo - Descripción de los resultados
+
+% Aquí se detallan los resultados obtenidos, con tablas de resumen, gráficas de resultados, identificación de datos relevantes, etc.
+% Es una exposición objetiva, sin valorar los resultados ni justificarlos.
+
+\chapter{Discussion}
+
+% Se aporta la discusión de los resultados presentados en el capítulo anterior. En este capítulo se puede discutir la relevancia de
+% los resultados, presentar posibles explicaciones apra los datos anómalos y resaltar aquellos datos que sean parcitualarmente relevantes
+% para el análisis del experimento.
 
 \chapter{Conclusions and Future Work}
+
+\section{Conclusions}
+
+% Resumen final del trabajo y debe servir para informar del alcance y relevancia de la aportación.
+% Suele estructurarse empezando con un resuemn del problema tratado, de cómo se ha abordado y de por qué la solución sería valída.
+
+% Es recomendable que incluya también un resumen de las contribuciones del trabajo, en el que se relacionen las contribuciones y los
+% resultado obtenidos con ls objetivos panteados para el trabajo, discutiendo hasta qué punto has conseguido resolver los objetivos planteados.
+
+\section{Further Research}
+
+% Debería señalar las perspectivas de futuro que abre el trabajo desarrollado para el campo de estudio definido. En el fondo,
+% debe justificar de qué modo puede emplearse la aportación desarrollada y en qué campos.
 

### 2020-11-09 src/scripts/manage_db.sh
feat: Add script for managing db backups

@@ -1,0 +1,178 @@
+#!/usr/bin/env bash
+
+set -e
+
+DOCKER_COMPOSE_FILE="$(dirname "$0")/../docker-compose.yml"
+VOLUME_NAME="gaia-postgres-data"
+
+BACKUP_DIR="$(pwd)"
+BACKUP_DATE="$(date +%Y-%m-%d_%H%M%S)"
+
+ERRORS_FOUND=false
+STOP_SERVICES=true
+
+BACKUP=false
+RESTORE=false
+
+function help() {
+  echo "manage_db backup|restore|help"
+  echo ""
+  echo "  backup [OPTIONS]                       Create a new backup with the current timestamp"
+  echo "    -d|--dest                            Destination directory for the backup file. Defaults to: ${BACKUP_DIR}"
+  echo "    -v|--volume                          Docker volume to backup. Defaults to: ${VOLUME_NAME}"
+  echo "    -s|--stop                            Stop container before creating backup. Defaults to: ${STOP_SERVICES}"
+  echo "  restore YYYY-mm-dd_HHMMSS [OPTIONS]    Restore the especified backup"
+  echo "    -s|--src                             Source directory of the backup file. Defaults to: ${BACKUP_DIR}"
+  echo "    -v|--volume                          Docker volume to restore. Defaults to: ${VOLUME_NAME}"
+  echo ""
+  echo "  help                                   Show this message and exit"
+  echo ""
+
+  exit 2
+}
+
+function log_error() {
+  (>&2 echo "$@")
+}
+
+case "${1}" in
+backup)
+  BACKUP=true
+  shift
+  while [[ $# -gt 0 ]]; do
+    key="${1}"
+    shift
+    case "${key}" in
+    -d | --dest)
+      BACKUP_DIR="${1}"
+      shift
+      ;;
+    -v | --volume)
+      VOLUME_NAME="${1}"
+      shift
+      ;;
+    -s | --stop)
+      STOP_SERVICES="${2}"
+      shift
+      ;;
+    *)
+      log_error "Unknown parameter ${key}"
+      help
+      ;;
+    esac
+  done
+  ;;
+restore)
+  RESTORE=true
+  shift
+  BACKUP_DATE="${1}"
+  shift
+  while [[ $# -gt 0 ]]; do
+    key="${1}"
+    shift
+    case "${key}" in
+    -s | --src)
+      BACKUP_DIR="${1}"
[…]

### 2020-11-09 src/cdalvaro/graphics/plot.py
feat: Update plots

@@ -2,163 +2,143 @@ import matplotlib.pyplot as plt
 import pandas as pd
 import seaborn as sns
 
-from .color_palette import color_palette, set_color_palette
+from .color_palette import color_palette
 
 
-class Plot:
+def plot_clusters_catalogue_distribution(data: pd.DataFrame,
+                                         title: str = None,
+                                         xlim: tuple = None,
+                                         ylim: tuple = None,
+                                         hue: str = 'diam'):
+    fig, ax = plt.subplots(figsize=(12, 6))
+    if title is not None:
+        ax.set_title(title)
 
-    default_ext = 'pdf'
+    ax.set_xlabel('Right Ascension (J2000 Degree)')
+    ax.set_ylabel('Declination (J2000 Degree)')
 
-    def __init__(self, save_figs: bool = False, figs_path: str = '.'):
-        self.save_figs = save_figs
-        self.figs_path = figs_path
-        set_color_palette()
+    if xlim is not None:
+        ax.set_xlim(xlim)
+    if ylim is not None:
+        ax.set_ylim(ylim)
 
-    def plot_clusters_catalogue_distribution(self,
-                                             df_catalogue,
-                                             img_name: str,
-                                             img_ext: str = default_ext,
-                                             title: str = None,
-                                             xlim: tuple = None,
-                                             ylim: tuple = None,
-                                             hue: str = 'diam'):
-        fig, ax = plt.subplots(figsize=(12, 6))
-        if title is not None:
-            ax.set_title(title)
+    palette = None
+    if hue is not None:
+        n_colors = len(pd.unique(data[hue]))
+        palette = color_palette(n_colors=n_colors)
 
-        ax.set_xlabel('Right Ascension (J2000 Degree)')
-        ax.set_ylabel('Declination (J2000 Degree)')
+    g = sns.scatterplot(data=data, x="ra", y="dec", hue=hue, size=hue, palette=palette, ax=ax)
 
-        if xlim is not None:
-            ax.set_xlim(xlim)
-        if ylim is not None:
-            ax.set_ylim(ylim)
+    plt.legend().set_title("Diameter (arcmin)")
 
-        g = sns.scatterplot(data=df_catalogue,
-                            x="ra",
-                            y="dec",
-                            hue=hue,
-                            size=hue,
-                            palette=color_palette(n_colors=len(pd.unique(df_catalogue[hue]))),
-                            ax=ax)
-        plt.legend().set_title("Diameter (arcmin)")
[…]

### 2020-11-09 thesis/thesis.tex
thesis: Update structure

@@
 
-However, in general it is not as easy and becomes necessary to consider other parameters such as distances, or even
-metallicity and age (derived from isochrone curves) and requires photometric data from stars in the studied field.
+Figure \ref{fig:pm_melotte_22} shows the proper motion distribution in right ascension and declination.
+It is easy to see that a subgroup can be located at [20, -45].
+Figure \ref{fig:pm_vec_melotte_22} shows the parallax density and confirms an overdensity at $\approx 7.5 mas$ corresponding to Mellote 22 OC.
+
+However, as shown in Figures \ref{fig:pm_ngc_2682} and \ref{fig:parallax_ngc_2682}, in general it is not as easy and
+becomes necessary to consider other parameters such as distances, or even metallicity and age (derived from isochrone curves)
+and requires photometric data from stars in the studied field.
 
@@
     \centering
-    \includegraphics[width=\textwidth]{../figures/isochrone_ngc_2682.pdf}
-    \caption{Isochrone curve}
+    \includegraphics[width=\textwidth]{../figures/parallax_ngc_2682.pdf}
+    \caption{Parallax}
     \label{fig:parallax_ngc_2682}
@@
 
+% TODO: Hablar de cómo funciona el descargador y cómo está formada la base de datos
+
+\section{Soft Clustering with K-Means}
+
+% TODO: Describir cómo funciona el clustering con KMedias
+
+\section{Deep Embedded Clustering}
+
+% TODO: Describir cómo funciona el clustering usando dec
+
+\section{Results}
+
 %% TODO: Las tecnologías empleadas para el desarrollo de las herramientas,
@@
 
+\pagebreak
+
 \bibliography{references}

### 2020-11-11 thesis/thesis.tex
thesis: Add Data Mining section

@@
 
-The development has been made with the Python programming language. Other auxiliary tools such as \emph{Docker}, \emph{PostgreSQL}, \emph{Jupyter Notebooks},
-and \emph{Scikit-Learn} and \emph{Keras} frameworks have been used too.
+All code is developed with the Python programming language \cite{Python3}. Other auxiliary tools such as \emph{Docker} \cite{merkel2014docker},
+\emph{PostgreSQL} \cite{postgresql}, \emph{Jupyter Notebooks} \cite{Kluyver2016jupyter}, and \emph{Astropy} \cite{astropy:2013} \cite{astropy:2018},
+\emph{Scikit-Learn} \cite{scikit-learn}, \emph{seaborn} \cite{michael_waskom_2017_883859}, \emph{SQLAlchemy} \cite{sqlalchemy} and
+\emph{Keras} \cite{chollet2015keras} frameworks have been used too.
 
@@
 
-As mentioned before, Gaia DR2 is the chosen data source for this work. Thus, the first step is to download the required data from Gaia DB
+As mentioned before, I have chosen the Gaia DR2 database as the data source for this work. Thus, the first step is to download the required data from Gaia DB
 and store it in a custom database for later access.
@@
 
-There is no upper limit to the number of stars inside a cluster, although for the sake of this work, smaller cluster will be prefered as
-a commitment to the project delivery dates and the available computing power. As an extension of this work, the trained model could be applied to a region
-not covered by this preselection for a later understanding of the results.
+As shown in Figure \ref{fig:OpenClustSelection} this selection stills being a good representation of all clusters in the Milky Way since they are
+equally distributed around the galaxy line.
+
+After this selection, the number of cluster to look up is 169 with nearly 75 million stars.
+
+Since there is no upper limit to the number of stars inside a cluster, for the sake of simplicity and as a commitment to the project delivery dates
+and the available computing power, smaller clusters will be prefered over the greatest ones.
+
+As an extension of this work, the designed model could be applied to a region not covered by this preselection for a later understanding of the results.
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=\columnwidth]{../figures/cluster_selection_tier1.pdf}
+  \caption{OpenClust Catalogue Selection Distribution}
+  \label{fig:OpenClustSelection}
+\end{figure}
+
+The download process has been performed with two Docker containers: a \emph{downloader} and the \emph{database}.
+
+The first one is a container built from a \verb|python:3.8| image that contains the \verb|cdalvaro| package and the \verb|downloader.py| script.
+This script is prepared to load the OpenClust catalogue, connect to the Gaia DR2 database, download all stars contained inside each cluster
+(with a radius multiplier to increse the area to be downloaded) and save the downloaded data inside the custom database hosted in the second container.
+
+The second container is just a \verb|postgres:12.4| image which loads an initial script when the database is not yet initialized for creating the
+database structure. This container is used later as the main database for data analysis.
+
+The database has two main tables, \verb|regions| for storing cluster properties such as location, diameter and other properties, and \verb|gaiadr2_source|
+which contains data for all downloaded stars. This second table is partitioned by region, storing all stars related to one cluster inside an independent
+table for faster access.
+
+These containers are orchestrated with a \verb|docker-compose.yml| file which automatilly launches one instance of each container.
+
+The code developed for this work can be found inside the \verb|src| directory at GitHub:
+\href{https://github.com/cdalvaro/machine-learning-master-thesis}{cdalvaro/machine-learning-master-thesis}
 

### 2020-11-11 src/cdalvaro/graphics/color_palette.py
feat: Improve graphics

@@ -8,6 +8,6 @@ def set_color_palette():
     sns.set_palette(color_palette(as_cmap=False))
 
 
-def color_palette(as_cmap: bool = True, **kwargs):
+def color_palette(as_cmap: bool = False, **kwargs):
     # https://medium.com/@morganjonesartist/color-guide-to-seaborn-palettes-da849406d44f
     return sns.color_palette('BrBG_r', as_cmap=as_cmap, **kwargs)

### 2020-11-11 src/cdalvaro/graphics/plot.py

@@ -1,5 +1,6 @@
 import matplotlib.pyplot as plt
 import pandas as pd
+import numpy as np
 import seaborn as sns
 
 from .color_palette import color_palette
@@ -10,7 +11,7 @@ def plot_clusters_catalogue_distribution(data: pd.DataFrame,
                                          xlim: tuple = None,
                                          ylim: tuple = None,
                                          hue: str = 'diam'):
-    fig, ax = plt.subplots(figsize=(12, 6))
+    fig, ax = plt.subplots(figsize=(12, 6), tight_layout=True)
     if title is not None:
         ax.set_title(title)
 
@@ -23,11 +24,12 @@ def plot_clusters_catalogue_distribution(data: pd.DataFrame,
         ax.set_ylim(ylim)
 
     palette = None
+    hue_order = None
     if hue is not None:
-        n_colors = len(pd.unique(data[hue]))
-        palette = color_palette(n_colors=n_colors)
+        hue_order = np.sort(pd.unique(data[hue]))
+        palette = color_palette(n_colors=len(hue_order))
 
-    g = sns.scatterplot(data=data, x="ra", y="dec", hue=hue, size=hue, palette=palette, ax=ax)
+    g = sns.scatterplot(data=data, x="ra", y="dec", hue=hue, hue_order=hue_order, size=hue, palette=palette, ax=ax)
 
     plt.legend().set_title("Diameter (arcmin)")
 
@@ -40,7 +42,7 @@ def plot_cluster_proper_motion(data: pd.DataFrame,
                                ylim: tuple = None,
                                hue: str = 'cluster_g',
                                legend: bool = True):
-    fig, ax = plt.subplots(figsize=(6, 6))
+    fig, ax = plt.subplots(figsize=(6, 6), tight_layout=True)
     if title is not None:
         ax.set_title(title)
 
@@ -53,16 +55,25 @@ def plot_cluster_proper_motion(data: pd.DataFrame,
         ax.set_ylim(ylim)
 
     palette = None
+    hue_order = None
     if hue is not None:
-        n_colors = len(pd.unique(data[hue]))
-        palette = color_palette(n_colors=n_colors)
+        hue_order = np.sort(pd.unique(data[hue]))
+        palette = color_palette(n_colors=len(hue_order))
 
-    g = sns.scatterplot(data=data, x="pmra", y="pmdec", hue=hue, palette=palette, s=12, ax=ax, legend=legend)
+    g = sns.scatterplot(data=data,
[…]

### 2020-11-11 src/cdalvaro/ml/utils.py
python: Improve estimate_n_clusters method

@@ -1,15 +1,21 @@
 import pandas as pd
 from sklearn.cluster import KMeans
 from sklearn.metrics import silhouette_score
+from typing import Tuple
 
 
-def estimate_n_clusters(x, metric: str = 'euclidean', max_clusters: int = 10, verbose: bool = False) -> int:
+def estimate_n_clusters(x,
+                        metric: str = 'euclidean',
+                        min_clusters: int = 2,
+                        max_clusters: int = 10,
+                        verbose: bool = False) -> Tuple[int, KMeans]:
     """
     Estimate the number of clusters based on silhoutte score.
 
     Args:
         x (array-like): Data to be fit for best number of clusters estimation.
         metric (str, optional): The metric to use when calculating distance between instances in a feature array. Defaults to euclidean.
+        min_clusters (int, optional): The minimum number fo clusters to be tested. Defaults to 2.
         max_clusters (int, optional): The maximum number of clusters to be tested. Defaults to 10.
         verbose (bool, optional): Show algorithm progress. Defaults to False.
 
@@ -20,8 +26,9 @@ def estimate_n_clusters(x, metric: str = 'euclidean', max_clusters: int = 10, ve
         int: The estimated number of clusters
     """
     best_score = -1.0
-    best_n_clusters = None
-    for n_clusters in range(2, max_clusters + 1):
+    best_n_clusters: int = None
+    best_kmeans: KMeans = None
+    for n_clusters in range(min_clusters, max_clusters + 1):
         kmeans = KMeans(n_clusters=n_clusters)
         pred = kmeans.fit_predict(x)
         score = silhouette_score(x, pred, metric=metric)
@@ -30,6 +37,7 @@ def estimate_n_clusters(x, metric: str = 'euclidean', max_clusters: int = 10, ve
         if score > best_score:
             best_score = score
             best_n_clusters = n_clusters
+            best_kmeans = kmeans
 
     if best_n_clusters is None:
         raise ValueError("Unable to estimate the number of clusters to be used")
@@ -37,7 +45,7 @@ def estimate_n_clusters(x, metric: str = 'euclidean', max_clusters: int = 10, ve
     if verbose:
         print(f"Best silhouette score is {best_score} for {best_n_clusters} clusters")
 
-    return best_n_clusters
+    return best_n_clusters, best_kmeans
 
 
 def filter_outliers(df: pd.DataFrame, q: float = 0.05) -> pd.Series:

### 2020-11-11 thesis/thesis.tex
thesis: Add Soft Clustering with K-Means section

@@
 
+\usepackage{apacite}
 \usepackage{caption}
@@
 \usepackage{graphicx}
-\usepackage{apacite}
 \usepackage{picture}
 \usepackage{subcaption}
+
+\hypersetup{
+  colorlinks = true, % Colors links instead of ugly boxes
+  urlcolor   = azulunir, % Color for external hyperlinks
+  linkcolor  = azulunir, % Color of internal links
+  citecolor  = azulunir, % Color of citations
+}
 
@@
 
+% TODO: Hablar sobre los algoritmos de clustering que se van a emplear: KMeans, DEC
+
 %% TODO: Concluir con una última sección de resumen de conclusiones, resumiendo las principales averiguaciones del estudio del dominio
@@
 
+Methods for retrieving information from the database are available through an instance of \verb|cdalvaro.DB|. This class contains methods
+for recovering stars information for a given region as a Pandas \verb|DataFrame| ready for data analysis.
+
 These containers are orchestrated with a \verb|docker-compose.yml| file which automatilly launches one instance of each container.
@@
 
-% TODO: Hablar de cómo funciona el descargador y cómo está formada la base de datos
-
 \section{Soft Clustering with K-Means}
 
-% TODO: Describir cómo funciona el clustering con KMedias
+Our first approach to find the open cluster inside the interest region is to use the K-Means algorithm.
+
+Since we are looking for a cluster, it seems reasonable to use a clustering algorithm and set it to find two clusters.
+One for the desired OC and another cluster which contains stars that don't belong to the open cluster.
+However, this idea is not right.
+
+This is due to the fact that OC's stars are sourounded by other stars with possibly similar properties. So, setting the number of clusters
+to two is too low to separate them properly.
+
+As shown in Figure \ref{fig:kmeans_comparisons_melotte_22}, bigger values for the number of clusters allow us to isolate more accurately
+the resonance in parallax at $\approx 7.5mas$. However, we have the disadvantage that more groups are formed, while our aim is to find just
+one small cluster and another with the rest of stars. This effect complicates our task of finding the desired OC. Thus, we have to find
+the right value to isolate the searched cluster without creating to many groups.
+
+\begin{figure}[htbp]
+  \centering
+  \begin{subfigure}{0.9\textwidth}
+    \centering
+    \begin{subfigure}[t]{0.3\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/kmeans_n2_pm_melotte_22.pdf}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.3\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/kmeans_n5_pm_melotte_22.pdf}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.3\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/kmeans_n8_pm_melotte_22.pdf}
+    \end{subfigure}
+  \end{subfigure}
+  \medskip
+  \begin{subfigure}{0.9\textwidth}
+    \centering
+    \begin{subfigure}[t]{0.3\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/kmeans_n2_parallax_melotte_22.pdf}
+      \caption{N clusters = 2}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.3\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/kmeans_n5_parallax_melotte_22.pdf}
+      \caption{N clusters = 5}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.3\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/kmeans_n8_parallax_melotte_22.pdf}
+      \caption{N clusters = 8}
+    \end{subfigure}
+  \end{subfigure}
+  \caption{K-Means comparisons}
+  \label{fig:kmeans_comparisons_melotte_22}
+\end{figure}
+
+To solve this issue, we will try to estimate the best number of clusters by using the \emph{silhouette score}.
+
+The silhouette score is a metric used to determine how good a cluster is based on intra-cluster distances and nearest-cluster distances.
+The best possible value is 1, and the worst is -1. Negative values mean that some samples have been assigned to the wrong cluster.
 
@@
 
-% TODO: Describir cómo funciona el clustering usando dec
+Our initial approach allows us to find some groups that potentially contain the OC. However we need now a method to merge clusters
+by migrating stars from one group to another. So we can end up with the minimum number of clusters while preserving the open cluster.
+
+For that purpose we have adapted the Unsupervised Deep Embedding for Clustering Analysis or DEC model \cite{xie2016unsupervised} to our work.
+
+DEC is divided in two stages. In the first one and \emph{deep autoencoder} is pretrained to extract information from data and generate the
+the embedding vectors. In the second stage, a parameter optimization is performed, iterating between computing an auxiliary target distribution
+and minimizing the Kullback-Leiber (KL) divergence \cite{kullback1951information} to it.
+
+This model uses K-Means as the starting point, since it is used to create an initial clustering and then the DEC model is used to refined it.
 

### 2020-11-12 thesis/thesis.tex
thesis: Update Deep Embedded Clustering section

@@
 \usepackage{graphicx}
+\usepackage{minted}
 \usepackage{picture}
@@
 \author{Carlos David Álvaro Yunta}
-\date{13 de Septiembre de 2020}
+\date{12th of November, 2020}
 \director{César Augusto Guzmán Álvarez}
@@
   \centering
+  \begin{subfigure}{0.9\textwidth}
+    \centering
     \begin{subfigure}[t]{0.45\textwidth}
@@
     \end{subfigure}
+  \end{subfigure}
   \caption{Open Cluster Melotte 22 (M45)}
@@
 becomes necessary to consider other parameters such as distances, or even metallicity and age (derived from isochrone curves)
-and requires photometric data from stars in the studied field.
+and requires photometric data from stars in the studied field
+\footnote{All figures have been sampled in order to reduce the document size, but are still a faithful representation of the original data}.
 
 \begin{figure}[htbp]
+  \centering
+  \begin{subfigure}{0.9\textwidth}
     \centering
@@
       \label{fig:parallax_ngc_2682}
+    \end{subfigure}
   \end{subfigure}
@@
 
-% TODO: Hablar sobre los algoritmos de clustering que se van a emplear: KMeans, DEC
+% TODO: Hablar sobre los algoritmos de clustering que se van a emplear: K-Means, DEC
 
-%% TODO: Concluir con una última sección de resumen de conclusiones, resumiendo las principales averiguaciones del estudio del dominio
+% TODO: Concluir con una última sección de resumen de conclusiones, resumiendo las principales averiguaciones del estudio del dominio
 % y cómo van a afectar al desarrollo específico del trabajo.
@@
 
+% TODO: Hay que desarrollar más los objetivos
+
 \chapter{Method}
@@
   \centering
-  \includegraphics[width=\columnwidth]{../figures/openclust_catalogue.pdf}
+  \includegraphics[width=0.9\textwidth]{../figures/openclust_catalogue.pdf}
   \caption{OpenClust Catalogue Distribution}
@@
   \centering
-  \includegraphics[width=\columnwidth]{../figures/cluster_selection_tier1.pdf}
+  \includegraphics[width=0.9\textwidth]{../figures/cluster_selection_tier1.pdf}
   \caption{OpenClust Catalogue Selection Distribution}
@@
 
+Downloader image build instructions are available at \href{https://github.com/cdalvaro/machine-learning-master-thesis/blob/main/src/downloader.Dockerfile}{downloader.Dockerfile}
+and latest image is available for pulling from \href{https://github.blog/2020-09-01-introducing-github-container-registry/}{GitHub Container Registry}:
+
+\begin{minted}{sh}
+  docker pull ghcr.io/cdalvaro/gaia-downloader:latest
+\end{minted}
+
+This image is automatically built with \href{https://github.com/features/actions}{GitHub Actions} on every push to the \verb|main| branch.
+
 The second container is just a \verb|postgres:12.4| image which loads an initial script when the database is not yet initialized for creating the
-database structure. This container is used later as the main database for data analysis.
+database schema. This container is used later as the main database for data analysis.
 
@@
 
-These containers are orchestrated with a \verb|docker-compose.yml| file which automatilly launches one instance of each container.
+These containers are orchestrated with a \href{https://github.com/cdalvaro/machine-learning-master-thesis/blob/main/src/docker-compose.yml}{docker-compose.yml}
+file which automatilly launches one instance of each container.
 
@@
   \end{subfigure}
-  \caption{K-Means comparisons}
+  \caption{K-Means comparisons with Melotte 22}
   \label{fig:kmeans_comparisons_melotte_22}
@@
 
-Our initial approach allows us to find some groups that potentially contain the OC. However we need now a method to merge clusters
+Our initial approach allows us to find some groups that potentially contain the desired OC. However we need now a method for merging clusters
 by migrating stars from one group to another. So we can end up with the minimum number of clusters while preserving the open cluster.
+
+Since we don't have a labeled dataset which allows as to train a supervised model, we are required to use an unsupervised self-trained model.
 
@@
 
-DEC is divided in two stages. In the first one and \emph{deep autoencoder} is pretrained to extract information from data and generate the
-the embedding vectors. In the second stage, a parameter optimization is performed, iterating between computing an auxiliary target distribution
-and minimizing the Kullback-Leiber (KL) divergence \cite{kullback1951information} to it.
+As we can see in Figure \ref{fig:dec_melotte_22}, the DEC model is able to merge similar clusters leaving alone the OC. Which is exactly what we are looking for.
 
-This model uses K-Means as the starting point, since it is used to create an initial clustering and then the DEC model is used to refined it.
+\begin{figure}[htbp]
+  \centering
+  \begin{subfigure}{0.9\textwidth}
+    \centering
+    \begin{subfigure}[t]{0.3\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/dec_pm_melotte_22.pdf}
+      \caption{Proper motions}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.3\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/dec_parallax_melotte_22.pdf}
+      \caption{Parallax}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.3\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/dec_isochrone_melotte_22.pdf}
+      \caption{Isochrone curve}
+    \end{subfigure}
+  \end{subfigure}
+  \caption{DEC model applied to Melotte 22}
+  \label{fig:dec_melotte_22}
+\end{figure}
+
+The implementation of this model is available in \verb|cdalvaro.ml.DEC| and it is developed with the Keras framework.
+
+The model is composed by a \emph{deep autoencoder}, a \emph{deep encoder} and a \emph{clustering layer}.
+
+Encoders are used to transform the input data into a latent space using a non-linear mapping function $f_{\theta} : X \rightarrow Z$.
+
+Although in this work we are managing a small number of features, this latent space allows us to reduce the number of features, avoiding
+this way the \emph{"curse of dimensionality"} \cite{bellman1961curse}.
+
+These encoders are pretrained before fitting the model to generate predictions. Then, the encoder is used to transform input data to the
+latent space $Z$. Once the data has been transformed, a K-Means clusterer is used to make an initial clustering of the data.
+
+With that initial configuration, the model iterates alternating between computing an auxiliary target distribution (Soft Assignment)
+and minimizing the Kullback-Leiber (KL) divergence \cite{kullback1951information} to it. This unsupervised algorithm allows us to improve
+the clustering.
+
+In the soft assignment stage, the \emph{Student's t-distribution} is used as a kernel to measure the similarity between the embedded points
+and the cluster centroid. While in the KL divergence minimization the algorithm iteratively refines clusters by learning from their high
+confidence assignments with the help of an auxiliary target distribution. The model is trained by matching the soft assignment to the target
+distribution. The choice of this target distribution is crucial for DEC's performance. In this work we have taken the target distribution
+from DEC's original paper \cite{xie2016unsupervised}, which is defined as follows:
+
+\begin{equation}
+  p_{ij} = \frac{q^{2}_{ij} / f_{j}}{\sum_{j'}q^{2}_{ij'}/f_{j'}}
+\end{equation}
+
+Inside \href{https://github.com/cdalvaro/machine-learning-master-thesis/blob/main/src/notebooks}{src/notebooks},
+there is a notebook called
+\href{https://github.com/cdalvaro/machine-learning-master-thesis/blob/main/src/notebooks/cluster_characterization.ipynb}{cluster\_characterization.ipynb}
+that illustrates how DEC model process a region of stars trying to characterize the OC hidden inside that region.
 
@@
 
-%% TODO: Las tecnologías empleadas para el desarrollo de las herramientas,
+% TODO: Las tecnologías empleadas para el desarrollo de las herramientas,
 % la organización del piloto, cómo ha transcurrido el experimento
 
-%% TODO: Añadir nuevo capítulo - Descripción de los resultados
+% TODO: Añadir nuevo capítulo - Descripción de los resultados
 

### 2020-11-12 thesis/thesis.tex
thesis: Update References chapter

@@
 \usepackage{minted}
+\usepackage[nottoc]{tocbibind}
 \usepackage{picture}
@@
 
-However, as shown in Figures \ref{fig:pm_ngc_2682} and \ref{fig:parallax_ngc_2682}, in general it is not as easy and
+However, as shown in Figures \ref{fig:pm_ngc_6494} and \ref{fig:parallax_ngc_6494}, in general it is not as easy and
 becomes necessary to consider other parameters such as distances, or even metallicity and age (derived from isochrone curves)
@@
       \centering
-      \includegraphics[width=\textwidth]{../figures/pm_ngc_2682.pdf}
+      \includegraphics[width=\textwidth]{../figures/pm_ngc_6494.pdf}
       \caption{Proper motions}
-      \label{fig:pm_ngc_2682}
+      \label{fig:pm_ngc_6494}
     \end{subfigure}
@@
       \centering
-      \includegraphics[width=\textwidth]{../figures/parallax_ngc_2682.pdf}
+      \includegraphics[width=\textwidth]{../figures/parallax_ngc_6494.pdf}
       \caption{Parallax}
-      \label{fig:parallax_ngc_2682}
+      \label{fig:parallax_ngc_6494}
     \end{subfigure}
   \end{subfigure}
-  \caption{Open Cluster NGC 2682}
+  \caption{Open Cluster NGC 6494}
 \end{figure}
@@
 
-\pagebreak
+\newpage
 

### 2020-11-12 src/cdalvaro/graphics/plot.py
graphics: Update plots

@@ -1,9 +1,60 @@
 import matplotlib.pyplot as plt
+from matplotlib.pyplot import xlabel
 import pandas as pd
 import numpy as np
 import seaborn as sns
 
 from .color_palette import color_palette
+
+N_SAMPLES = 4000
+
+
+def _resample_data(data: pd.DataFrame, n_samples: int) -> pd.DataFrame:
+    if n_samples > 0 and data.shape[0] > n_samples:
+        return data.sample(n_samples, random_state=0)
+    return data
+
+
+def _get_color_properties(data: pd.DataFrame, hue: str) -> dict:
+    properties = dict()
+    if hue is not None:
+        hue_order = np.sort(pd.unique(data[hue]))
+        palette = color_palette(n_colors=len(hue_order))
+
+        properties.update({'hue': hue, 'hue_order': hue_order, 'palette': palette})
+
+    return properties
+
+
+def _set_axis_properties(ax,
+                         title: str = None,
+                         xlabel: str = None,
+                         ylabel: str = None,
+                         xlim: tuple = None,
+                         ylim: tuple = None,
+                         invert_xaxis: bool = False,
+                         invert_yaxis: bool = False):
+
+    if title is not None:
+        ax.set_title(title)
+
+    if xlabel is not None:
+        ax.set_xlabel(xlabel)
+
+    if ylabel is not None:
+        ax.set_ylabel(ylabel)
+
+    if xlim is not None:
+        ax.set_xlim(xlim)
+
+    if ylim is not None:
+        ax.set_ylim(ylim)
+
+    if invert_xaxis:
+        ax.invert_xaxis()
+
+    if invert_yaxis:
+        ax.invert_yaxis()
 
 
 def plot_clusters_catalogue_distribution(data: pd.DataFrame,
@@ -11,27 +62,19 @@ def plot_clusters_catalogue_distribution(data: pd.DataFrame,
                                          xlim: tuple = None,
                                          ylim: tuple = None,
                                          hue: str = 'diam'):
+
     fig, ax = plt.subplots(figsize=(12, 6), tight_layout=True)
-    if title is not None:
-        ax.set_title(title)
 
-    ax.set_xlabel('Right Ascension (J2000 Degree)')
-    ax.set_ylabel('Declination (J2000 Degree)')
[…]

### 2020-11-12 thesis/thesis.tex
thesis: Expand Introduction chapter

@@
 
-%% TODO: 1.1 Motivation
+% TODO: 1.1 Motivation
 
@@
 
-%% TODO: 1.2 Planteamiento
+% TODO: 1.2 Planteamiento
 
@@
 
-%% TODO: 1.3 Estructura de la memoria
+% TODO: 1.3 Estructura de la memoria
 
@@
 This dataset has helped to review already known open clusters and to find new ones.
+
+Stars that belong to a same OC share relative positions, so their parallax value is similar among them. They also share similar values of
+proper motion, both in right ascension and declination. Another property, since these stars were born from the same gas cloud and at the same
+temporary stage, their metallicity must be uniform. However, to take into account this last property, we face the inconvinient that this
+parameter is not directly available in Gaia DR2 database. We will avoid this issue by looking at $E(b-r) / G_{mag}$ diagrams with well
+defined isochrone curves \cite{bressan2012parsec}. These isochrone curves are derived from theoretical models that mainly involve metallicity
+and mass/brightness from stars. With these models we can define important properties for the development of stars and their evolution in the
+H-R diagram. Examples of these curves are shown in Figure \ref{fig:examples_of_isochrones}.
+
+Again, since these stars were born close in time, regarding their intrinsic properties (mass/brightness), these stars must present a very
+well defined and sharp profile with low sparsing through the main sequence in the H-R diagram.
+
+We will take advantage of these properties to validate that the open clusters obtained are right.
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=0.9\textwidth]{../figures/theoretical_isochrones_in_hr_diagrams.pdf}
+  \caption{Examples of theoretical isochrones in the HR diagram taken from \protect\citeA[p.~16]{bressan2012parsec}}
+  \label{fig:examples_of_isochrones}
+\end{figure}
 
@@
 The main aim of this thesis is to \emph{build an unsupervised clustering model for open cluster characterization}.
-This model must be \emph{non-supervised non-parameterized}, and has to be valid for a wide range of open clusters.
+
+We want this model to be \emph{non-supervised non-parameterized} in order to fit a wide range of clusters without
+the need of fine tunning hyperparameters regarding the topology or size of the studied clusters.
 

### 2020-11-22 thesis/thesis.tex
thesis: Move and rewrite aims

@@
 
+\chapter{Aims}
+
+\section{General}
+
+The main aim of this thesis is to \emph{build an unsupervised clustering model for open cluster characterization}.
+
+We want this model to be \emph{non-supervised non-parameterized} in order to fit a wide range of clusters without
+the need of fine tunning hyperparameters regarding the topology or size of the studied clusters.
+
+\section{Specific}
+
+To achieve the general goal, we'll follow the following milestones:
+
+\begin{itemize}
+  \item Gather information on the \emph{state of the art}.
+  \item Research unsupervised clustering algorithms suitable for grouping stars previously recovered.
+  \item Recover data from Gaia DR2 database. This data will be taken as source for the machine learning model to characterize open clusters
+  by grouping stars into clusters.
+  \item Select and implement unsupervised clustering algorithms from the previous study.
+  \item Use those algorithms with different datasets to find OCs.
+  \item Look for an independent method to characterize OCs in order to make comparisons.
+  \item Validate found OCs with our custom model by making comparisons with the independent method.
+\end{itemize}
+
+% TODO: Hay que desarrollar más los objetivos
+
 \chapter{State of the Art}
@@
 
-With Gaia DR2, which contains great quality astrometrical and photometric data for a huge number of stars,
+With Gaia DR2, which contains great quality astrometric and photometric data for a huge number of stars,
 a new opportunity to simplify and optimize this processes is presented.
@@
 in the Galactic disc} \cite{castro2020hunting}, and merge them into a single \emph{machine learning}
-tool capable of finding open clusters eficiently without requiring powerful machines to perform the computations.
+tool capable of finding open clusters efficiently without requiring powerful machines to perform the computations.
 
-The first one is a non-parametrized method. It determines empirically the frequency functions asociated to astrometrical variables
+The first one is a non-parametrized method. It determines empirically the frequency functions associated to astrometrical variables
 without making any assumption about their profiles. However, it heavily does depend on the initial field selection and other
@@
 The method developed in this project takes as base the approach made by \citeauthor{castro2020hunting} but rapidly diverges.
-In first place, performance has been taken into account since the begining, trying to get a reliable method without compromising the
+In first place, performance has been taken into account since the beginning, trying to get a reliable method without compromising the
 needs of supercomputers.
@@
 Thus, the main goal of this project is not the discovery of new open clusters or even candidates, but the development of a model that,
-applied to already knonw open clusters, allows us the characterization of those OCs on an unsupervised way, with a reasonable computing
+applied to already known open clusters, allows us the characterization of those OCs on an unsupervised way, with a reasonable computing
 power and taking into account the same selection of astrometric and photometric variables for every OC.
@@
 % y cómo van a afectar al desarrollo específico del trabajo.
-
-\chapter{Aims}
-
-\section{General}
-
-The main aim of this thesis is to \emph{build an unsupervised clustering model for open cluster characterization}.
-
-We want this model to be \emph{non-supervised non-parameterized} in order to fit a wide range of clusters without
-the need of fine tunning hyperparameters regarding the topology or size of the studied clusters.
-
-To measure the reliability of the resulting model comparisons with other VO tools will be made.
-
-The role for the ANN will be validating cluster members based on dynamincs data.
-
-These are te variables to be considered in this work:
-
-\begin{itemize}
-  \item OCs and stars \emph{right ascension}: $\alpha$ and \emph{declination}: $\delta$
-  \item OCs \emph{estimated sizes}: $r$
-  \item \emph{Proper motion}: $\mu_{\alpha}$, $\mu_{\delta}$ and \emph{parallax}: $\varpi$
-  \item \emph{Associated errors} will be taken into account to estimate prediction errors
-\end{itemize}
-
-\section{Specific}
-
-To achive this goal, the aim will be divided into the following milestones:
-
-\begin{itemize}
-  \item To recover data from Gaia DR2 database. This data will be taken as source for the machine learning model to characterize open clusters
-  by grouping stars into clusters.
-  \item To research unsupervised clustering algorithms suitable for grouping stars previously recovered.
-  \item To select and implement unsupervised clustering algorithms from the previous study.
-  \item To apply implemented tools to different datasets from first point to find OCs.
-  \item To validate found OCs by determining their accuracy and precission.
-\end{itemize}
-
-% TODO: Hay que desarrollar más los objetivos
 
@@
 
+As mentioned before, we want to characterize clusters looking at their dynamics properties.
+So, these are the main features we will take into account for clustering:
+
+\begin{itemize}
+  \item OCs and stars \emph{right ascension}: $\alpha$ and \emph{declination}: $\delta$
+  \item OCs \emph{estimated sizes}: $r$
+  \item \emph{Proper motion}: $\mu_{\alpha}$, $\mu_{\delta}$ and \emph{parallax}: $\varpi$
+  \item \emph{Associated errors} will be taken into account to estimate prediction errors
+\end{itemize}
+
 \section{Data Mining}
@@
 Due to the large amount of data inside Gaia database, a complete download is not viable neither useful.
-In order to squize the size of the dataset to be downloaded, the OpenClust catalogue \cite{dias2002new} (Figure \ref{fig:OpenClustComplete})
+In order to reduce the size of the dataset to be downloaded, the OpenClust catalogue \cite{dias2002new} (Figure \ref{fig:OpenClustComplete})
 has been used to make a preselection of sky regions.
@@
 The downloaded dataset covers nearly 114 million stars (42GB of compressed data), that is significantly smaller than the whole Gaia DR2 dataset, which contains
-information for around 1,600 million stars, but stils being too large.
+information for approximately 1,600 million stars, but it is still being too large.
 
@@
 Since there is no upper limit to the number of stars inside a cluster, for the sake of simplicity and as a commitment to the project delivery dates
-and the available computing power, smaller clusters will be prefered over the greatest ones.
+and the available computing power, smaller clusters will be preferred over the greatest ones.
 
@@
 This script is prepared to load the OpenClust catalogue, connect to the Gaia DR2 database, download all stars contained inside each cluster
-(with a radius multiplier to increse the area to be downloaded) and save the downloaded data inside the custom database hosted in the second container.
+(with a radius multiplier to increase the area to be downloaded) and save the downloaded data inside the custom database hosted in the second container.
 
@@
 These containers are orchestrated with a \href{https://github.com/cdalvaro/machine-learning-master-thesis/blob/main/src/docker-compose.yml}{docker-compose.yml}
-file which automatilly launches one instance of each container.
+file which automatically launches one instance of each container.
 
@@
 
-This is due to the fact that OC's stars are sourounded by other stars with possibly similar properties. So, setting the number of clusters
+This is due to the fact that OC's stars are surrounded by other stars with possibly similar properties. So, setting the number of clusters
 to two is too low to separate them properly.
@@
 
-As we can see in Figure \ref{fig:dec_melotte_22}, the DEC model is able to merge similar clusters leaving alone the OC. Which is exactly what we are looking for.
-
-\begin{figure}[htbp]
-  \centering
-  \begin{subfigure}{0.9\textwidth}
-    \centering
-    \begin{subfigure}[t]{0.3\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/dec_pm_melotte_22.pdf}
-      \caption{Proper motions}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.3\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/dec_parallax_melotte_22.pdf}
-      \caption{Parallax}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.3\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/dec_isochrone_melotte_22.pdf}
-      \caption{Isochrone curve}
-    \end{subfigure}
-  \end{subfigure}
-  \caption{DEC model applied to Melotte 22}
-  \label{fig:dec_melotte_22}
-\end{figure}
-
 The implementation of this model is available in \verb|cdalvaro.ml.DEC| and it is developed with the Keras framework.
@@
 \href{https://github.com/cdalvaro/machine-learning-master-thesis/blob/main/src/notebooks/cluster_characterization.ipynb}{cluster\_characterization.ipynb}
-that illustrates how DEC model process a region of stars trying to characterize the OC hidden inside that region.
+that illustrates how DEC model processes a region of stars trying to characterize the OC hidden inside that region.
 
-\section{Results}
+As we can see in Figure \ref{fig:dec_melotte_22}, the DEC model is able to merge similar clusters leaving alone the OC. Which is exactly what we are looking for.
+
+\begin{figure}[htbp]
+  \centering
+  \begin{subfigure}{0.9\textwidth}
+    \centering
+    \begin{subfigure}[t]{0.3\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/dec_pm_melotte_22.pdf}
+      \caption{Proper motions}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.3\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/dec_parallax_melotte_22.pdf}
+      \caption{Parallax}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.3\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/dec_isochrone_melotte_22.pdf}
+      \caption{Isochrone curve}
+    \end{subfigure}
+  \end{subfigure}
+  \caption{DEC model applied to Melotte 22}
+  \label{fig:dec_melotte_22}
+\end{figure}
+
+\chapter{Results}
 
@@
 
-\chapter{Discussion}
+\section{Discussion}
 

### 2020-11-23 thesis/thesis.tex
thesis: Add more content to Method chapter

@@
 \tableofcontents
-\listoffigures
-\listoftables
 
@@
 proper motion, both in right ascension and declination. Another property, since these stars were born from the same gas cloud and at the same
-temporary stage, their metallicity must be uniform. However, to take into account this last property, we face the inconvinient that this
+temporary stage, their metallicity must be uniform. However, to take into account this last property, we face the inconvenient that this
 parameter is not directly available in Gaia DR2 database. We will avoid this issue by looking at $E(b-r) / G_{mag}$ diagrams with well
@@
 and mass/brightness from stars. With these models we can define important properties for the development of stars and their evolution in the
-H-R diagram. Examples of these curves are shown in Figure \ref{fig:examples_of_isochrones}.
+Hertzsprung–Russell (H-R) diagram. Examples of these curves are shown in Figure \ref{fig:examples_of_isochrones}.
 
@@
 
-To achieve the general goal, we'll follow the following milestones:
+To achieve the general goal, we will follow the following milestones:
 
@@
 \chapter{State of the Art}
+
+% TODO: Hablar de la misión Gaia
+% TODO: Hablar de KMedias y DEC
+% TODO: Mencionar trabajos relacionados
 
@@
 
-% TODO: Hablar sobre los algoritmos de clustering que se van a emplear: K-Means, DEC
-
 % TODO: Concluir con una última sección de resumen de conclusiones, resumiendo las principales averiguaciones del estudio del dominio
@@
 
-In this chapter we address the different steps taken to perform the creation of the unsupervised clustering model for open cluster characterization.
+In this chapter we address the different steps followed to perform the creation of the unsupervised clustering model for open cluster characterization.
 
-All code is developed with the Python programming language \cite{Python3}. Other auxiliary tools such as \emph{Docker} \cite{merkel2014docker},
-\emph{PostgreSQL} \cite{postgresql}, \emph{Jupyter Notebooks} \cite{Kluyver2016jupyter}, and \emph{Astropy} \cite{astropy:2013} \cite{astropy:2018},
-\emph{Scikit-Learn} \cite{scikit-learn}, \emph{seaborn} \cite{michael_waskom_2017_883859}, \emph{SQLAlchemy} \cite{sqlalchemy} and
-\emph{Keras} \cite{chollet2015keras} frameworks have been used too.
+All code has been developed with the Python programming language \cite{Python3}. Other auxiliary tools like \emph{Docker} \cite{merkel2014docker},
+\emph{PostgreSQL} \cite{postgresql}, \emph{Jupyter Notebooks} \cite{Kluyver2016jupyter}, and frameworks such as \emph{Astropy} \cite{astropy:2013} \cite{astropy:2018},
+\emph{Scikit-Learn} \cite{scikit-learn}, \emph{Seaborn} \cite{michael_waskom_2017_883859}, \emph{SQLAlchemy} \cite{sqlalchemy} and
+\emph{Keras} \cite{chollet2015keras} have been used too.
 
-As mentioned before, we want to characterize clusters looking at their dynamics properties.
-So, these are the main features we will take into account for clustering:
+\begin{figure}[htbp]
+  \centering
+  \begin{subfigure}{0.5\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{../figures/raw_pm_melotte_22.pdf}
+  \end{subfigure}
+  \caption{Melotte 22 proper motions}
+  \label{fig:raw_pm_melotte_22}
+\end{figure}
 
-\begin{itemize}
-  \item OCs and stars \emph{right ascension}: $\alpha$ and \emph{declination}: $\delta$
-  \item OCs \emph{estimated sizes}: $r$
-  \item \emph{Proper motion}: $\mu_{\alpha}$, $\mu_{\delta}$ and \emph{parallax}: $\varpi$
-  \item \emph{Associated errors} will be taken into account to estimate prediction errors
-\end{itemize}
+For the sake of simplicity, we are taking \emph{Melotte 22 (Messier 45)} \cite{elsanhoury2019ppmxl} to illustrate the whole method. This cluster is well studied
+and their stars are not too mixed with other region stars.
+
+Figure \ref{fig:raw_pm_melotte_22} shows \emph{proper motions in right ascension and declination} for a sample of the downloaded dataset for Melotte 22.
+At first sight, two main clusters can be distinguished, one of them centered nearly at (0, 0) and the second cluster with center at (20, -45). This second cluster
+is the one we are looking for.
+
+However, although this second cluster is almost isolated, there are stars that do not belong to the OC. Thus, we need more information to properly characterize the OC.
+
+\begin{figure}[htbp]
+  \centering
+  \begin{subfigure}{0.9\textwidth}
+    \centering
+    \begin{subfigure}[t]{0.45\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/raw_parallax_melotte_22.pdf}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.45\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/raw_hr_diagram_melotte_22.pdf}
+    \end{subfigure}
+  \end{subfigure}
+  \caption{Melotte 22 parallax and H-R diagram}
+  \label{fig:raw_parallax_hr_diagram_mellote_22}
+\end{figure}
+
+Figure \ref{fig:raw_parallax_hr_diagram_mellote_22} shows the parallax histogram and the H-R diagram for this cluster, respectively.
+
+The first one shows a resonance at $\approx 7.5mas$ belonging the OC. While the second figure would help us to look for isochrone curves and so,
+to determine the age of these stars.
 
@@
 
-As mentioned before, I have chosen the Gaia DR2 database as the data source for this work. Thus, the first step is to download the required data from Gaia DB
+Before being able to develop and test our clustering model, the first step is to download the required data from Gaia repository
 and store it in a custom database for later access.
 
-Due to the large amount of data inside Gaia database, a complete download is not viable neither useful.
+Due to the large amount of data available at Gaia, a complete download is not viable neither useful.
 In order to reduce the size of the dataset to be downloaded, the OpenClust catalogue \cite{dias2002new} (Figure \ref{fig:OpenClustComplete})
-has been used to make a preselection of sky regions.
+has been used to restrict the sky regions to be explored.
 
 This download is not limited to the size registered in the catalogue. Instead, a wider region is downloaded for each cluster to include
-stars outside the cluster and take them into account when training the model so the model will be able to take apart those outsiders
-when characterizing the cluster.
+stars outside the cluster. The idea is that the unsupervised model must be able to clusterize the stars that belongs to the OC and to discard
+those outsiders when characterizing the open cluster.
 
@@
 
-The downloaded dataset covers nearly 114 million stars (42GB of compressed data), that is significantly smaller than the whole Gaia DR2 dataset, which contains
-information for approximately 1,600 million stars, but it is still being too large.
+Taking these considerations into account, the downloaded dataset covers nearly 114 million stars (42GB of compressed data). This dataset is significantly
+smaller than the whole Gaia DR2 dataset, which contains information for approximately 1,600 million stars. However, it is still being too large.
 
-Since not all the downloaded clusters are good enough, the following filters are applied to discard some clusters which don't have enough stars or contain too
-many \emph{null} values:
+Since not all the downloaded clusters are good enough, we will apply a series of filters to discard clusters which do not have enough stars or contain too
+many \emph{null} values.
+
+So, a cluster must fulfil the following filters to be accepted:
 
@@
   \item Parallax absolute value greater than 0.0
-  \item Number of stars in the selected region above 40,000 stars
+  \item Number of stars\footnote{Every star must have all required features completely defined, i.e. without null values} in the selected region above 40,000 stars
 \end{itemize}
 
-As shown in Figure \ref{fig:OpenClustSelection} this selection stills being a good representation of all clusters in the Milky Way since they are
-equally distributed around the galaxy line.
+As shown in Figure \ref{fig:OpenClustSelection}, these constrains give us a smaller dataset but it is still being a good representation of all clusters in the Milky Way
+since they are equally distributed around the galaxy disk.
 
-After this selection, the number of cluster to look up is 169 with nearly 75 million stars.
+After having applied these filters, the number of cluster to analyse is 169 with nearly 75 million stars.
 
-Since there is no upper limit to the number of stars inside a cluster, for the sake of simplicity and as a commitment to the project delivery dates
+Since we have set no upper limit to the number of stars inside a cluster, for the sake of simplicity and as a commitment to the project delivery dates
 and the available computing power, smaller clusters will be preferred over the greatest ones.
 
-As an extension of this work, the designed model could be applied to a region not covered by this preselection for a later understanding of the results.
+As an extension of this work, the designed model could be applied to a region not covered by these constrains for a later understanding of the results.
 
@@
 
+\subsection{Download Process}
+
 The download process has been performed with two Docker containers: a \emph{downloader} and the \emph{database}.
@@
 This script is prepared to load the OpenClust catalogue, connect to the Gaia DR2 database, download all stars contained inside each cluster
-(with a radius multiplier to increase the area to be downloaded) and save the downloaded data inside the custom database hosted in the second container.
+(with a radius multiplier to increase the area to be downloaded) and save the downloaded data inside our custom database hosted in the second container.
 
-Downloader image build instructions are available at \href{https://github.com/cdalvaro/machine-learning-master-thesis/blob/main/src/downloader.Dockerfile}{downloader.Dockerfile}
-and latest image is available for pulling from \href{https://github.blog/2020-09-01-introducing-github-container-registry/}{GitHub Container Registry}:
+\verb|downloader.Dockerfile| contains the build instructions for the downloader image. We use \href{https://github.com/features/actions}{GitHub Actions}
+for automating the process of building a new version of the image with every new push made to the \verb|main| branch.
+
+This image can be pulled from the \href{https://github.blog/2020-09-01-introducing-github-container-registry/}{GitHub Container Registry}:
 
@@
 
-This image is automatically built with \href{https://github.com/features/actions}{GitHub Actions} on every push to the \verb|main| branch.
-
 The second container is just a \verb|postgres:12.4| image which loads an initial script when the database is not yet initialized for creating the
@@
 
-The database has two main tables, \verb|regions| for storing cluster properties such as location, diameter and other properties, and \verb|gaiadr2_source|
-which contains data for all downloaded stars. This second table is partitioned by region, storing all stars related to one cluster inside an independent
-table for faster access.
+The database has two main tables in \verb|public| schema:
+
+\begin{itemize}
+  \item \verb|regions| for storing cluster properties such as \emph{location}, \emph{diameter} and other properties
+  \item \verb|gaiadr2_source| which contains data for all downloaded stars from Gaia
+\end{itemize}
+
+The second table is partitioned by region, storing all stars related to one cluster inside an independent table for optimized access performance.
 
@@
 
-These containers are orchestrated with a \href{https://github.com/cdalvaro/machine-learning-master-thesis/blob/main/src/docker-compose.yml}{docker-compose.yml}
-file which automatically launches one instance of each container.
+These containers are orchestrated with a \verb|docker-compose.yml| file which automatically launches one instance of each container.
 
-The code developed for this work can be found inside the \verb|src| directory at GitHub:
-\href{https://github.com/cdalvaro/machine-learning-master-thesis}{cdalvaro/machine-learning-master-thesis}
+The developed code for this work can be found inside the \verb|src| directory at GitHub:
+\href{https://github.com/cdalvaro/machine-learning-master-thesis}{cdalvaro/machine-learning-master-thesis} repository.
 
@@
 
+% TODO: Hablar sobre la selección de variables
+
 Our first approach to find the open cluster inside the interest region is to use the K-Means algorithm.
+
+As mentioned before, we want to characterize clusters by looking at their dynamics properties.
+So, these are the main features we will take into account for clustering:
+
+\begin{itemize}
+  \item OCs and stars \emph{right ascension}: $\alpha$ and \emph{declination}: $\delta$
+  \item OCs \emph{estimated sizes}: $r$
+  \item \emph{Proper motion}: $\mu_{\alpha}$, $\mu_{\delta}$ and \emph{parallax}: $\varpi$
+  \item \emph{Associated errors} will be taken into account to discard data with excessive uncertainty
+\end{itemize}
 
@@
 
+\listoffigures
+\listoftables
+
 \newpage

### 2020-11-27 thesis/thesis.tex
thesis: Add Feature Selection section

@@
 
+\section{Feature Selection}
+
+As mentioned before, we want our model to characterize open clusters by looking at their dynamics properties.
+
+Also, we want to maintain as simple as possible our clustering model in order to avoid requiring too much computing resources.
+
+Proper motion in right ascension and declination seems like a natural choice since, as we know, stars belonging to the same OC share a common motion vector.
+
+Parallax is another important feature. It lets us know how far stars are from us. And, since all stars within an open cluster were born from the same dust cloud,
+they must all have similar parallax.
+
+However, we are not going to use these raw features. Instead, we are taken a combination of them.
+
+First, we correct proper motion in right ascension and declination by dividing them by the parallax. That way we normalize these quantities and help our clustering
+models to improve their performance.
+
+Another computed property that we are considering is the modulus of the proper motion. That way, we relate both features and therefore force our model to keep them tight.
+
+So, as summary, these are the features we are going to use as sources for our clustering models:
+
+\begin{itemize}
+  \item \emph{Proper motion in right ascension}: $\mu_{\alpha}$
+  \item \emph{Proper motion in declination}: $\mu_{\delta}$
+  \item \emph{Parallax}: $\varpi$
+\end{itemize}
+
+This feature selection has been refined by iterating over the K-Means clustering process with a fixed number of clusters and varying the feature selection in each case.
+The final selection is the one with best silhouette score. This process is explained in more detail in Appendix \ref{chap:feature_selection}.
+
 \section{Soft Clustering with K-Means}
 
-% TODO: Hablar sobre la selección de variables
+Once we have found the set of features that best describes our problem, we can begin to search for the OC in the study region.
 
-Our first approach to find the open cluster inside the interest region is to use the K-Means algorithm.
+Our first approach to find the open cluster within the region of interest is to use the K-Means algorithm.
 
-As mentioned before, we want to characterize clusters by looking at their dynamics properties.
-So, these are the main features we will take into account for clustering:
-
-\begin{itemize}
-  \item OCs and stars \emph{right ascension}: $\alpha$ and \emph{declination}: $\delta$
-  \item OCs \emph{estimated sizes}: $r$
-  \item \emph{Proper motion}: $\mu_{\alpha}$, $\mu_{\delta}$ and \emph{parallax}: $\varpi$
-  \item \emph{Associated errors} will be taken into account to discard data with excessive uncertainty
-\end{itemize}
-
-Since we are looking for a cluster, it seems reasonable to use a clustering algorithm and set it to find two clusters.
+Since we are looking for a single cluster, it seems reasonable to use a clustering algorithm and set it to find two clusters.
 One for the desired OC and another cluster which contains stars that don't belong to the open cluster.
-However, this idea is not right.
+However, this idea is not completely right.
 
@@
 the resonance in parallax at $\approx 7.5mas$. However, we have the disadvantage that more groups are formed, while our aim is to find just
-one small cluster and another with the rest of stars. This effect complicates our task of finding the desired OC. Thus, we have to find
-the right value to isolate the searched cluster without creating to many groups.
+one small cluster and another with the rest of stars. This effect complicates our task of finding the desired OC. Therefore, we have to find
+a way to set the right value for the number of clusters to isolate the searched cluster without creating to many groups.
 
@@
 \appendix
-\chapter{Appendix}
+\chapter{Feature Selection}
+\label{chap:feature_selection}
 

### 2020-11-29 thesis/thesis.tex
thesis: Add Abstract chapter

@@
 
-{\bf Nota:} En este apartado se introducirá un breve resumen en inglés del trabajo realizado (extensión máxima: 150 palabras).
-Este resumen debe incluir el objetivo o propósito de la investigación, la metodología, los resultados y las conclusiones.
+The characterization and knowledge of \emph{Open Clusters} allows us to better understand properties and mechanisms about the Universe
+such as stellar formation and the regions where those events occur. They also provide information about stellar processes and the
+evolution of the galactic disk.
+
+In this work, we develope a method to characterize those clusters by using \emph{artificial intelligence} tools like clustering
+by \emph{K-Means}, or clustering based on \emph{Artificial Neural Networks} by implementing the \emph{Deep Embedded Clustering}
+model. We are using Gaia DR2 database as data source for testing our models.
+
+The developed method aims to improve the existing ones both in terms of \emph{computational efficiency}, with lower computational requirements,
+and \emph{ease of use}, by reducing the number of hyperparameters to configure in order to obtain a good characterization of the analyzed clusters.
+
+As detailed below, our method achieves good results, becoming even better in some cases when the results are compared with other current methods.
 
@@
 
-{\bf Key words.} open clusters detection --- machine learning --- gaia dr2 --- data analysis
+{\bf Key words.} open clusters characterization --- machine learning --- gaia dr2 --- data analysis --- deep embedded clustering
 
@@
 
-{\bf Nota:} En este apartado se introducirá un breve resumen en español del trabajo realizado (extensión máxima: 150 palabras).
-Este resumen debe incluir el objetivo o propósito de la investigación, la metodología, los resultados y las conclusiones.
+La caracterización y conocimiento de \emph{Cúmulos Abiertos} permite conocer mejor propiedades y mecanismos del Universo tales como
+la formación de estrellas y las regiones donde se dan estos procesos. También permiten obtener información sobre procesos estelares
+y la evolución del disco galáctico.
+
+En este trabajo se desarrolla un método para caracterizar estos cúmulos mediante el uso de herramientas de \emph{inteligencia artificial}
+como agrupamiento por \emph{K-Medias} o agrupamiento basado en \emph{Redes Neuronales Artificiales} mediante la implementación del
+modelo \emph{Deep Embedded Clustering} aplicados sobre conjuntos de datos extraídos de la base de datos de \emph{Gaia DR2}.
+
+El método desarrollado pretende mejorar los ya existentes tanto en términos de \emph{eficiencia computacional}, con menores requisitos de cómputo,
+como en \emph{facilidad de uso}, reduciendo el número de hiperparámetros a configurar para obtener una buena caracterización de los cúmulos analizados.
+
+Como se detalla más adelante, nuestro método consigue buenos resultados llegando a ser incluso mejor en algunos casos cuando se comparan
+los resultados con otros métodos actuales.
 
@@
 
-{\bf Palabras Clave:} detección de cúmulos abiertos --- inteligencia artificial --- gaia dr2 --- análisis de datos
+{\bf Palabras Clave:} caracterización de cúmulos abiertos --- inteligencia artificial --- gaia dr2 --- análisis de datos
 

### 2020-11-29 thesis/thesis.tex
thesis: Update Method chapter

@@
 
-The first one shows a resonance at $\approx 7.5mas$ belonging the OC. While the second figure would help us to look for isochrone curves and so,
+The figure on the left shows a resonance at $\approx 7.5mas$ belonging the OC. While the figure on the right would help us to look for isochrone curves and so,
 to determine the age of these stars.
@@
 \section{Feature Selection}
+\label{sec:feature_selection}
 
@@
 
-So, as summary, these are the features we are going to use as sources for our clustering models:
+\begin{figure}[htbp]
+  \centering
+  \begin{subfigure}{0.9\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{../figures/features_melotte_22.pdf}
+  \end{subfigure}
+  \caption{Feature pairplot}
+  \label{fig:features_melotte_22}
+\end{figure}
+
+So, in summary, these are the features we are going to use as sources for our clustering models:
 
 \begin{itemize}
-  \item \emph{Proper motion in right ascension}: $\mu_{\alpha}$
-  \item \emph{Proper motion in declination}: $\mu_{\delta}$
+  \item \emph{Proper motion in right ascension} (corrected by parallax): $\mu_{\alpha}$
+  \item \emph{Proper motion in declination} (corrected by parallax): $\mu_{\delta}$
   \item \emph{Parallax}: $\varpi$
+  \item \emph{Proper motion modulus}: $\left\| \vec{\mu} \right\|$
 \end{itemize}
@@
 This feature selection has been refined by iterating over the K-Means clustering process with a fixed number of clusters and varying the feature selection in each case.
-The final selection is the one with best silhouette score. This process is explained in more detail in Appendix \ref{chap:feature_selection}.
+The final selection is the one with best \emph{silhouette score} \cite{rousseeuw1987silhouettes}.
+
+The silhouette score is a metric used to determine how good a cluster is based on intra-cluster distances and nearest-cluster distances.
+The best possible value is 1, and the worst is -1. Negative values mean that some samples have been assigned to the wrong cluster.
+
+This process is explained in more detail in Appendix \ref{chap:feature_selection}.
 
@@
 
-Once we have found the set of features that best describes our problem, we can begin to search for the OC in the study region.
+Once we have found the set of features that best describes our problem, we can begin searching for the OC in the study region.
 
@@
 Since we are looking for a single cluster, it seems reasonable to use a clustering algorithm and set it to find two clusters.
-One for the desired OC and another cluster which contains stars that don't belong to the open cluster.
+One for the desired OC and another which contains stars that do not belong to the open cluster.
 However, this idea is not completely right.
-
-This is due to the fact that OC's stars are surrounded by other stars with possibly similar properties. So, setting the number of clusters
-to two is too low to separate them properly.
-
-As shown in Figure \ref{fig:kmeans_comparisons_melotte_22}, bigger values for the number of clusters allow us to isolate more accurately
-the resonance in parallax at $\approx 7.5mas$. However, we have the disadvantage that more groups are formed, while our aim is to find just
-one small cluster and another with the rest of stars. This effect complicates our task of finding the desired OC. Therefore, we have to find
-a way to set the right value for the number of clusters to isolate the searched cluster without creating to many groups.
 
@@
 
-To solve this issue, we will try to estimate the best number of clusters by using the \emph{silhouette score}.
+This is due to the fact that OC's stars are surrounded by other stars with possibly similar properties. So, setting the number of clusters
+to two is too low to separate them properly.
 
-The silhouette score is a metric used to determine how good a cluster is based on intra-cluster distances and nearest-cluster distances.
-The best possible value is 1, and the worst is -1. Negative values mean that some samples have been assigned to the wrong cluster.
+As shown in Figure \ref{fig:kmeans_comparisons_melotte_22}, bigger values for the number of clusters allow us to isolate more accurately
+the resonance in parallax at $\approx 7.5mas$. However, we have the disadvantage that more groups are formed, while our aim is to find just
+one small cluster that contains the stars that belong to the OC and another one with the remaining stars.
 
-\section{Deep Embedded Clustering}
+This effect complicates our task of finding the desired OC. Therefore, we have to find
+a way to set the right value for the number of clusters to isolate the searched cluster without creating too many groups.
 
-Our initial approach allows us to find some groups that potentially contain the desired OC. However we need now a method for merging clusters
-by migrating stars from one group to another. So we can end up with the minimum number of clusters while preserving the open cluster.
+To solve this issue, we will try to estimate the best number of clusters by using one more time the \emph{silhouette score}.
 
-Since we don't have a labeled dataset which allows as to train a supervised model, we are required to use an unsupervised self-trained model.
+The following example can be found in the Jupyter notebook: \verb|cluster_characterization.ipynb| (available at \verb|src/notebooks|).
+It shows how to estimate the best number of clusters for Melotte 22 by using \verb|cdalvaro.ml.estimate_n_clusters| method.
 
-For that purpose we have adapted the Unsupervised Deep Embedding for Clustering Analysis or DEC model \cite{xie2016unsupervised} to our work.
+\begin{minted}{python}
+  n_clusters, kmeans = estimate_n_clusters(melotte22_df, min_clusters=3,
+                                           max_clusters=10, verbose=True)
+
+  # Output
+  # Silhouette score for 3 clusters: 0.5420
+  # Silhouette score for 4 clusters: 0.5393
+  # Silhouette score for 5 clusters: 0.5608
+  # Silhouette score for 6 clusters: 0.5336
+  # Silhouette score for 7 clusters: 0.5306
+  # Silhouette score for 8 clusters: 0.4978
+  # Silhouette score for 9 clusters: 0.5214
+  # Silhouette score for 10 clusters: 0.5199
+  # Best silhouette score is 0.5608 for 5 clusters
+\end{minted}
+
+K-Means does a good job making an initial clustering, as shown in Figure \ref{fig:kmeans_melotte_22}. However, too many clusters arise from this characterization
+and the open cluster is still polluted with stars that do not belong to it. Even more, we would like to reduce the amount of clusters too.
+
+\begin{figure}[htbp]
+  \centering
+  \begin{subfigure}{0.9\textwidth}
+    \centering
+    \begin{subfigure}[t]{.45\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/kmeans_pm_melotte_22.pdf}
+      \caption{Proper Motion}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{.45\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/kmeans_parallax_melotte_22.pdf}
+      \caption{Parallax}
+    \end{subfigure}
+  \end{subfigure}
+  \caption{K-Means model applied to Melotte 22}
+  \label{fig:kmeans_melotte_22}
+\end{figure}
+
+If we take a look into the H-R diagram (Figure \ref{fig:kmeans_hr_diagram_melotte_22}), we can identify the isochrone curve for Melotte 22 cluster.
+This curve has a good shape, but again, it contains outsider stars. Therefore, we would like to find a better model that improves this initial characterization
+by reducing the amount of clusters and also that would be able to remove outsiders from the OC.
+
+\begin{figure}[htbp]
+  \centering
+  \begin{subfigure}{0.5\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{../figures/kmeans_isochrone_melotte_22.pdf}
+  \end{subfigure}
+  \caption{Melotte 22 H-R diagram with K-Means characterization}
+  \label{fig:kmeans_hr_diagram_melotte_22}
+\end{figure}
+
+\section{Deep Embedded Clustering (DEC)}
+
+Our initial approach allows us to find some groups that potentially contain the desired OC. However, as we have seen, the result is not enough
+accurate. Therefore, we need a method for merging clusters by migrating stars from one group to another,
+so we can end up with the minimum number of clusters while preserving the open cluster.
+
+Since we do not have a labeled dataset which allows us to train a supervised model, we have no choice but to use an unsupervised self-trained model.
+
+For that reason we have adapted the \emph{Unsupervised Deep Embedding for Clustering Analysis} or \emph{DEC} model \cite{xie2016unsupervised} to our work.
 
@@
 
-Although in this work we are managing a small number of features, this latent space allows us to reduce the number of features, avoiding
-this way the \emph{"curse of dimensionality"} \cite{bellman1961curse}.
+Although, as explained in section \ref{sec:feature_selection}, the number of features we are managing is not too large, this latent space allows us to reduce
+the number of features, avoiding this way the \emph{"curse of dimensionality"} \cite{bellman1961curse}.
 
 These encoders are pretrained before fitting the model to generate predictions. Then, the encoder is used to transform input data to the
-latent space $Z$. Once the data has been transformed, a K-Means clusterer is used to make an initial clustering of the data.
+latent space $Z$. Once the data has been transformed, a K-Means clusterer is used to make an initial clustering.
 
@@
 
-In the soft assignment stage, the \emph{Student's t-distribution} is used as a kernel to measure the similarity between the embedded points
-and the cluster centroid. While in the KL divergence minimization the algorithm iteratively refines clusters by learning from their high
-confidence assignments with the help of an auxiliary target distribution. The model is trained by matching the soft assignment to the target
-distribution. The choice of this target distribution is crucial for DEC's performance. In this work we have taken the target distribution
-from DEC's original paper \cite{xie2016unsupervised}, which is defined as follows:
-
 \begin{equation}
   p_{ij} = \frac{q^{2}_{ij} / f_{j}}{\sum_{j'}q^{2}_{ij'}/f_{j'}}
+  \label{eq:student_tdistribution}
 \end{equation}
 
-Inside \href{https://github.com/cdalvaro/machine-learning-master-thesis/blob/main/src/notebooks}{src/notebooks},
-there is a notebook called
-\href{https://github.com/cdalvaro/machine-learning-master-thesis/blob/main/src/notebooks/cluster_characterization.ipynb}{cluster\_characterization.ipynb}
-that illustrates how DEC model processes a region of stars trying to characterize the OC hidden inside that region.
+In the soft assignment stage, the \emph{Student's t-distribution} is used as a kernel to measure the similarity
+between the embedded points and the cluster centroid. While in the KL divergence minimization the algorithm iteratively refines clusters by learning
+from their high confidence assignments with the help of an auxiliary target distribution. The model is trained by matching the soft assignment to the
+target distribution. The choice of this target distribution is crucial for DEC's performance. In this work we have taken the target distribution
+from DEC's original paper \cite{xie2016unsupervised}, which is defined in Equation \ref{eq:student_tdistribution}.
 
-As we can see in Figure \ref{fig:dec_melotte_22}, the DEC model is able to merge similar clusters leaving alone the OC. Which is exactly what we are looking for.
+Figure \ref{fig:dec_model_setup} shows the layer setup of our DEC model. It is simpler than the one tested on the original paper \cite{xie2016unsupervised},
+since the number of selected features in our work is smaller. Therefore, using the same configuration would result in a model too
+powerful that would incur in overfitting issues unable to make right predictions.
 
@@
     \centering
-    \begin{subfigure}[t]{0.3\textwidth}
+    \includegraphics[width=\textwidth]{../figures/dec_diagram.pdf}
+  \end{subfigure}
+  \caption{DEC model layer setup}
+  \label{fig:dec_model_setup}
+\end{figure}
+
+Once the model is ready, we can test it by applying it to the Melotte 22 dataset. Notebook \verb|cluster_characterization.ipynb| illustrates how DEC model
+processes a region of stars trying to characterize the open cluster hidden within that region.
+
+As shown in Figure \ref{fig:dec_melotte_22}, the DEC model is able to move stars from one group to another removing some clusters and improving the OC characterization.
+This is exactly what we are looking for.
+
+\begin{figure}[htbp]
+  \centering
+  \begin{subfigure}{0.9\textwidth}
+    \centering
+    \begin{subfigure}[t]{0.45\textwidth}
       \centering
       \includegraphics[width=\textwidth]{../figures/dec_pm_melotte_22.pdf}
-      \caption{Proper motions}
+      \caption{Proper Motion}
     \end{subfigure}
     \hfill
-    \begin{subfigure}[t]{0.3\textwidth}
+    \begin{subfigure}[t]{0.45\textwidth}
       \centering
@@
     \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.3\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/dec_isochrone_melotte_22.pdf}
-      \caption{Isochrone curve}
-    \end{subfigure}
   \end{subfigure}
@@
   \label{fig:dec_melotte_22}
+\end{figure}
+
+The H-R diagram has been improved as well, and the isochrone curve of the open cluster is sharper than the one obtained with the K-Means algorithm
+(Figure \ref{fig:kmeans_hr_diagram_melotte_22}).
+
+\begin{figure}[htbp]
+  \centering
+  \begin{subfigure}{0.5\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/dec_isochrone_melotte_22.pdf}
+  \end{subfigure}
+  \caption{Melotte 22 H-R diagram with DEC characterization}
+  \label{fig:dec_hr_diagram_melotte_22}
 \end{figure}

### 2020-11-29 src/cdalvaro/graphics/plot.py
graphics: Add pairplot method

@@ -1,5 +1,4 @@
 import matplotlib.pyplot as plt
-from matplotlib.pyplot import xlabel
 import pandas as pd
 import numpy as np
 import seaborn as sns
@@ -15,13 +14,15 @@ def _resample_data(data: pd.DataFrame, n_samples: int) -> pd.DataFrame:
     return data
 
 
-def _get_color_properties(data: pd.DataFrame, hue: str) -> dict:
+def _get_color_properties(data: pd.DataFrame, hue: str = None, **kwargs) -> dict:
     properties = dict()
     if hue is not None:
         hue_order = np.sort(pd.unique(data[hue]))
-        palette = color_palette(n_colors=len(hue_order))
+        palette = color_palette(n_colors=len(hue_order), **kwargs)
 
         properties.update({'hue': hue, 'hue_order': hue_order, 'palette': palette})
+    else:
+        properties['palette'] = color_palette(**kwargs)
 
     return properties
 
@@ -159,8 +160,23 @@ def plot_cluster_isochrone_curve(data: pd.DataFrame,
 
     return fig, ax, g
 
+def pairplot(data: pd.DataFrame,
+             kind: str = 'scatter',
+             n_samples: int = N_SAMPLES):
+
+    data = _resample_data(data, n_samples).copy()
+    hue = 'cluster_g'
+    data[hue] = 'g0'
+
+    color_props = _get_color_properties(data, hue=hue)
+
+    g = sns.pairplot(data=data, kind=kind, **color_props)
+    g.legend.remove()
+    g.tight_layout()
+
+    return g
+
 
 def save_figure(fig, name: str, img_ext: str = "pdf", save_dir='./'):
-    if save_dir is not None:
-        name = name.replace(' ', '_').lower()
-        fig.savefig(f"{save_dir}/{name}.{img_ext}")
+    name = name.replace(' ', '_').lower()
+    fig.savefig(f"{save_dir}/{name}.{img_ext}")

### 2020-11-29 src/scripts/dec_diagram.py
graphics: Add DEC diagram

@@ -1,0 +1,32 @@
+
+import sys
+sys.path.append('../')
+from pycore.tikzeng import *
+
+# defined your arch
+arch = [
+    to_head( '..' ),
+    to_cor(),
+    to_begin(),
+    to_Conv("input", 4, '', offset="(0,0,0)", to="(0,0,0)", height=2, depth=2, width=2, caption='input' ),
+    to_Conv("encoder_0", 50, '', offset="(2,0,0)", to="(input-east)", height=2, depth=10, width=2, caption='encoder\_0' ),
+    to_connection( "input", "encoder_0"),
+    to_Conv("encoder_1", 50, '', offset="(2,0,0)", to="(encoder_0-east)", height=2, depth=10, width=2, caption='encoder\_1' ),
+    to_connection( "encoder_0", "encoder_1"),
+    to_Conv("encoder_2", 200, '', offset="(2,0,0)", to="(encoder_1-east)", height=2, depth=20, width=2, caption='encoder\_2' ),
+    to_connection( "encoder_1", "encoder_2"),
+    to_Conv("encoder_3", 5, '', offset="(2,0,0)", to="(encoder_2-east)", height=2, depth=2, width=2, caption='encoder\_3' ),
+    to_connection( "encoder_2", "encoder_3"),
+    to_SoftMax("clustering", 5 ,"(2,0,0)", "(encoder_3-east)", height=2, depth=2, width=2, caption='clustering'  ),
+    to_connection("encoder_3", "clustering"),
+    to_Pool("output", offset="(2,0,0)", to="(clustering-east)", height=2, depth=2, width=2, caption='output' ),
+    to_connection("clustering", "output"),
+    to_end()
+    ]
+
+def main():
+    namefile = str(sys.argv[0]).split('.')[0]
+    to_generate(arch, namefile + '.tex' )
+
+if __name__ == '__main__':
+    main()

### 2020-11-29 src/scripts/cluster_characterization.py
thesis: Update graphics with new color palette

@@ -1,0 +1,149 @@
+#!/usr/bin/env python3
+
+from cdalvaro.catalogues import OpenClust
+from cdalvaro.data_base import DB
+from cdalvaro.graphics import plot as cplt
+from cdalvaro.logging import Logger
+from cdalvaro.ml import DEC
+from cdalvaro.ml.utils import estimate_n_clusters, filter_outliers
+from IPython.display import Image
+import logging
+import numpy as np
+import os
+import pandas as pd
+from pathlib import Path
+import re
+from sklearn.preprocessing import MinMaxScaler
+import sys
+
+from keras.initializers import glorot_uniform, VarianceScaling
+from keras.optimizers import SGD
+from keras.utils import plot_model
+
+
+logger = Logger.instance()
+logger.setLevel(logging.ERROR)
+
+db = DB.instance(host='gaia.cdalvaro.io', port=15432)
+
+clusters = OpenClust.catalogue()
+cluster = clusters["Melotte 22"]
+
+# Directories
+project_dir = Path(os.path.dirname(sys.argv[0])).parent.parent
+
+cluster_name = cluster.name.replace(' ', '_').lower()
+figures_path = project_dir / "figures" / cluster_name
+results_path = project_dir / "results" / cluster_name
+
+for dir in [figures_path, results_path]:
+    dir.mkdir(exist_ok=True)
+
+variables = [
+    "ra", "ra_error", "dec", "dec_error", "pmra", "pmra_error",
+    "pmdec", "pmdec_error", "parallax", "parallax_error",
+    "phot_g_mean_mag", "bp_rp"
+]
+
+non_null_columns = list(filter(lambda x: not re.search(r'_error', x), variables))
+
+# Stars selection
+stars_df = db.get_stars(region=cluster, columns=variables, filter_null_columns=non_null_columns)
+
+stars_df = stars_df[(np.abs(stars_df['pmra']) > 0.0)
+                    & (np.abs(stars_df['pmdec']) > 0.0)
+                    & (np.abs(stars_df['parallax']) > 0.0)]
+
+stars_df['pmra_corr'] = stars_df['pmra'] * 1000.0 / stars_df['parallax']
[…]

### 2020-11-29 src/cdalvaro/graphics/plot.py
python: Update notebooks with new graphics

@@ -135,13 +135,13 @@ def plot_cluster_parallax_histogram(data,
     return fig, ax, g
 
 
-def plot_cluster_isochrone_curve(data: pd.DataFrame,
-                                 title: str = None,
-                                 xlim: tuple = None,
-                                 ylim: tuple = None,
-                                 hue: str = 'cluster_g',
-                                 legend: bool = True,
-                                 n_samples: int = N_SAMPLES):
+def plot_cluster_hr_diagram_curve(data: pd.DataFrame,
+                                  title: str = None,
+                                  xlim: tuple = None,
+                                  ylim: tuple = None,
+                                  hue: str = 'cluster_g',
+                                  legend: bool = True,
+                                  n_samples: int = N_SAMPLES):
 
     fig, ax = plt.subplots(figsize=(6, 6), tight_layout=True)
 
@@ -160,9 +160,8 @@ def plot_cluster_isochrone_curve(data: pd.DataFrame,
 
     return fig, ax, g
 
-def pairplot(data: pd.DataFrame,
-             kind: str = 'scatter',
-             n_samples: int = N_SAMPLES):
+
+def pairplot(data: pd.DataFrame, kind: str = 'scatter', n_samples: int = N_SAMPLES):
 
     data = _resample_data(data, n_samples).copy()
     hue = 'cluster_g'

### 2020-11-29 thesis/thesis.tex
thesis: Minor corrections

@@
 \usepackage{minted}
-\usepackage[nottoc]{tocbibind}
 \usepackage{picture}
 \usepackage{subcaption}
+\usepackage[nottoc]{tocbibind}
 
@@
 
-Stellar open clusters (OC) \cite{janes1982open} are groups of stars gravitationally bound originated from a single molecular gas cloud.
+Stellar open clusters (OCs) \cite{janes1982open} are groups of stars gravitationally bound originated from a single molecular gas cloud.
 Thus they share the same chemical composition and age, and they have similar relative positions and proper motion.
@@
 
-Young OC allow us to research stars formation regions and improve our understanding about the mechanisms that create those stars.
-On the other side, older OC give us information about stellar processes and how the galactic disk evolves.
+Young OCs allow us to research stars formation regions and improve our understanding about the mechanisms that create those stars.
+On the other side, older OCs give us information about stellar processes and how the galactic disk evolves.
 Some highly disturbed orbits could also provide evidence of recent merge events and accretion traces from outside the galaxy
@@
 
-The study of OC has been pushed forward thanks to the huge and precise dataset from the Gaia mission
+The study of OCs has been pushed forward thanks to the huge and precise dataset from the Gaia mission
 \cite{collaboration2016description} Gaia DR2 \cite{gaia2018gaia}, available since 2018.
@@
 Stars that belong to a same OC share relative positions, so their parallax value is similar among them. They also share similar values of
-proper motion, both in right ascension and declination. Another property, since these stars were born from the same gas cloud and at the same
-temporary stage, their metallicity must be uniform. However, to take into account this last property, we face the inconvenient that this
+proper motion, both in right ascension and declination. Another property is that their metallicity must be uniform, since these stars were
+born from the same gas cloud and at the same temporary stage. However, to take into account this last property, we face the inconvenient that this
 parameter is not directly available in Gaia DR2 database. We will avoid this issue by looking at $E(b-r) / G_{mag}$ diagrams with well
@@
 
-We will take advantage of these properties to validate that the open clusters obtained are right.
+We will take advantage of these properties to validate our OCs characterization.
 
@@
   \includegraphics[width=0.9\textwidth]{../figures/theoretical_isochrones_in_hr_diagrams.pdf}
-  \caption{Examples of theoretical isochrones in the HR diagram taken from \protect\citeA[p.~16]{bressan2012parsec}}
+  \caption{Examples of theoretical isochrones in the H-R diagram taken from \protect\citeA[p.~16]{bressan2012parsec}}
   \label{fig:examples_of_isochrones}
@@
 
-% TODO: Hay que desarrollar más los objetivos
-
 \chapter{State of the Art}
@@
       \centering
-      \includegraphics[width=\textwidth]{../figures/pm_melotte_22.pdf}
+      \includegraphics[width=\textwidth]{../figures/melotte_22/pm_melotte_22.pdf}
       \caption{Proper motions}
@@
       \centering
-      \includegraphics[width=\textwidth]{../figures/parallax_melotte_22.pdf}
+      \includegraphics[width=\textwidth]{../figures/melotte_22/parallax_melotte_22.pdf}
       \caption{Parallax}
@@
   \end{subfigure}
-  \caption{Open Cluster Melotte 22 (M45)}
+  \caption{Open Cluster Melotte 22 (Messier 45)}
 \end{figure}
@@
       \centering
-      \includegraphics[width=\textwidth]{../figures/pm_ngc_6494.pdf}
+      \includegraphics[width=\textwidth]{../figures/ngc_6494/pm_ngc_6494.pdf}
       \caption{Proper motions}
@@
       \centering
-      \includegraphics[width=\textwidth]{../figures/parallax_ngc_6494.pdf}
+      \includegraphics[width=\textwidth]{../figures/ngc_6494/parallax_ngc_6494.pdf}
       \caption{Parallax}
@@
     \centering
-    \includegraphics[width=\textwidth]{../figures/raw_pm_melotte_22.pdf}
+    \includegraphics[width=\textwidth]{../figures/melotte_22/raw_pm_melotte_22.pdf}
   \end{subfigure}
@@
       \centering
-      \includegraphics[width=\textwidth]{../figures/raw_parallax_melotte_22.pdf}
+      \includegraphics[width=\textwidth]{../figures/melotte_22/raw_parallax_melotte_22.pdf}
     \end{subfigure}
@@
       \centering
-      \includegraphics[width=\textwidth]{../figures/raw_hr_diagram_melotte_22.pdf}
+      \includegraphics[width=\textwidth]{../figures/melotte_22/raw_hr_diagram_melotte_22.pdf}
     \end{subfigure}
@@
     \centering
-    \includegraphics[width=\textwidth]{../figures/features_melotte_22.pdf}
+    \includegraphics[width=\textwidth]{../figures/melotte_22/features_melotte_22.pdf}
   \end{subfigure}
-  \caption{Feature pairplot}
+  \caption{Melotte 22 feature pairplot}
   \label{fig:features_melotte_22}
@@
       \centering
-      \includegraphics[width=\textwidth]{../figures/kmeans_n2_pm_melotte_22.pdf}
+      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n2_pm_melotte_22.pdf}
     \end{subfigure}
@@
       \centering
-      \includegraphics[width=\textwidth]{../figures/kmeans_n5_pm_melotte_22.pdf}
+      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n5_pm_melotte_22.pdf}
     \end{subfigure}
@@
       \centering
-      \includegraphics[width=\textwidth]{../figures/kmeans_n8_pm_melotte_22.pdf}
+      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n8_pm_melotte_22.pdf}
     \end{subfigure}
@@
       \centering
-      \includegraphics[width=\textwidth]{../figures/kmeans_n2_parallax_melotte_22.pdf}
+      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n2_parallax_melotte_22.pdf}
       \caption{N clusters = 2}
@@
       \centering
-      \includegraphics[width=\textwidth]{../figures/kmeans_n5_parallax_melotte_22.pdf}
+      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n5_parallax_melotte_22.pdf}
       \caption{N clusters = 5}
@@
       \centering
-      \includegraphics[width=\textwidth]{../figures/kmeans_n8_parallax_melotte_22.pdf}
+      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n8_parallax_melotte_22.pdf}
       \caption{N clusters = 8}
@@
       \centering
-      \includegraphics[width=\textwidth]{../figures/kmeans_pm_melotte_22.pdf}
+      \includegraphics[width=\textwidth]{../figures/melotte_22/kmeans_pm_melotte_22.pdf}
       \caption{Proper Motion}
@@
       \centering
-      \includegraphics[width=\textwidth]{../figures/kmeans_parallax_melotte_22.pdf}
+      \includegraphics[width=\textwidth]{../figures/melotte_22/kmeans_parallax_melotte_22.pdf}
       \caption{Parallax}
@@
     \centering
-    \includegraphics[width=\textwidth]{../figures/kmeans_isochrone_melotte_22.pdf}
+    \includegraphics[width=\textwidth]{../figures/melotte_22/kmeans_hr_diagram_melotte_22.pdf}
   \end{subfigure}
@@
       \centering
-      \includegraphics[width=\textwidth]{../figures/dec_pm_melotte_22.pdf}
+      \includegraphics[width=\textwidth]{../figures/melotte_22/dec_pm_melotte_22.pdf}
       \caption{Proper Motion}
@@
       \centering
-      \includegraphics[width=\textwidth]{../figures/dec_parallax_melotte_22.pdf}
+      \includegraphics[width=\textwidth]{../figures/melotte_22/dec_parallax_melotte_22.pdf}
       \caption{Parallax}
@@
     \centering
-      \includegraphics[width=\textwidth]{../figures/dec_isochrone_melotte_22.pdf}
+    \includegraphics[width=\textwidth]{../figures/melotte_22/dec_hr_diagram_melotte_22.pdf}
   \end{subfigure}
@@
 
-% TODO: Las tecnologías empleadas para el desarrollo de las herramientas,
-% la organización del piloto, cómo ha transcurrido el experimento
+We have seen in the previous chapter that the DEC model is able to refine the clustering made by K-Means. This allows us to get
+a more accurate characterization of the cluster.
 
-% TODO: Añadir nuevo capítulo - Descripción de los resultados
+In order to validate our model and be able to say how good or bad it is we need a different method to characterize the cluster
+and compare one against each other.
 
-% Aquí se detallan los resultados obtenidos, con tablas de resumen, gráficas de resultados, identificación de datos relevantes, etc.
+For that reason we have used tools from the Virtual Observatory (VO) such as Clusterix and TOPCAT to characterize the studied clusters.
+
+\section{Cluster Characterization with VO Tools}
+
+\section{Comparison}
+
+Until now, we have focus our study in Melotte 22 OC. Since our main aim is to get a model that fits well for a wide range of clusters,
+we present now the results obtained for a selection of clusters with different topologies.
+
+% TODO: Exposición de resultados y comentarios
 % Es una exposición objetiva, sin valorar los resultados ni justificarlos.
@@
 % Se aporta la discusión de los resultados presentados en el capítulo anterior. En este capítulo se puede discutir la relevancia de
-% los resultados, presentar posibles explicaciones apra los datos anómalos y resaltar aquellos datos que sean parcitualarmente relevantes
+% los resultados, presentar posibles explicaciones para los datos anómalos y resaltar aquellos datos que sean particularmente relevantes
 % para el análisis del experimento.
@@
 % Resumen final del trabajo y debe servir para informar del alcance y relevancia de la aportación.
-% Suele estructurarse empezando con un resuemn del problema tratado, de cómo se ha abordado y de por qué la solución sería valída.
+% Suele estructurarse empezando con un resumen del problema tratado, de cómo se ha abordado y de por qué la solución sería válida.
 
 % Es recomendable que incluya también un resumen de las contribuciones del trabajo, en el que se relacionen las contribuciones y los
-% resultado obtenidos con ls objetivos panteados para el trabajo, discutiendo hasta qué punto has conseguido resolver los objetivos planteados.
+% resultado obtenidos con ls objetivos planteados para el trabajo, discutiendo hasta qué punto has conseguido resolver los objetivos planteados.
 

### 2020-12-01 src/cdalvaro/data_base.py
python: Improve cdalvaro.DB package

@@ -1,13 +1,9 @@
-from __future__ import annotations
-
 from astropy.coordinates import SkyCoord
 import astropy.units as u
-import json
 import numpy as np
 import os
 import pandas as pd
 from sqlalchemy import create_engine, MetaData, select, any_
-from sqlalchemy.dialects import postgresql
 from sqlalchemy.dialects.postgresql import insert
 from typing import Dict, List, Set, TypeVar, Union
 
@@ -49,7 +45,7 @@ class DB:
         self.metadata.reflect()
 
     @staticmethod
-    def instance(host: str, port: int) -> DB:
+    def instance(host: str, port: int) -> 'DB':
         """
         Returns a database manager instance pointed to host:port
 

### 2020-12-03 src/cdalvaro/utils/transform.py
python: Add binarize method

@@ -1,0 +1,42 @@
+import pandas as pd
+from typing import List
+
+
+def binarize(data: pd.DataFrame, bins: List[float], by: str, norm_by: str = None) -> pd.DataFrame:
+    bins_it = iter(bins)
+    min_cut = next(bins_it)
+
+    _data = list()
+    grouped_data = list()
+
+    desc = 'desc'
+    keys = pd.unique(data[desc])
+
+    # First bin
+    _data = data[data[by] < min_cut].groupby(by=desc).sum()
+    for key in keys:
+        value = _data.loc[key]['value'] if _data.shape[0] > 0 else 0
+        if norm_by is not None and key != norm_by:
+            value /= _data.loc[norm_by]['value']
+        grouped_data.append({by: f'$<{min_cut}$', 'value': value, 'desc': key})
+
+    # Middle bins
+    for max_cut in bins_it:
+        _data = data[(data[by] >= min_cut) & (data[by] < max_cut)].groupby(by=desc).sum()
+        for key in keys:
+            value = _data.loc[key]['value'] if _data.shape[0] > 0 else 0
+            if norm_by is not None and key != norm_by:
+                value /= _data.loc[norm_by]['value']
+            grouped_data.append({by: f'${min_cut}-{max_cut}$', 'value': value, 'desc': key})
+
+        min_cut = max_cut
+
+    # Last bin
+    _data = data[data[by] >= min_cut].groupby(by=desc).sum()
+    for key in keys:
+        value = _data.loc[key]['value'] if _data.shape[0] > 0 else 0
+        if norm_by is not None and key != norm_by:
+            value /= _data.loc[norm_by]['value']
+        grouped_data.append({by: f'$\geq{min_cut}$', 'value': value, 'desc': key})
+
+    return pd.DataFrame(grouped_data)

### 2020-12-03 thesis/thesis.tex
doc: Update README.md

@@
 by \emph{K-Means}, or clustering based on \emph{Artificial Neural Networks} by implementing the \emph{Deep Embedded Clustering}
-model. We are using Gaia DR2 database as data source for testing our models.
+model. We are using \emph{Gaia DR2 database} as data source for testing our models.
 
@@
       \centering
-      \includegraphics[width=\textwidth]{../figures/melotte_22/pm_melotte_22.pdf}
+      \includegraphics[width=\textwidth]{../figures/melotte_22/pm_melotte_22.png}
       \caption{Proper motions}
@@
       \centering
-      \includegraphics[width=\textwidth]{../figures/melotte_22/parallax_melotte_22.pdf}
+      \includegraphics[width=\textwidth]{../figures/melotte_22/parallax_melotte_22.png}
       \caption{Parallax}
@@
       \centering
-      \includegraphics[width=\textwidth]{../figures/ngc_6494/pm_ngc_6494.pdf}
+      \includegraphics[width=\textwidth]{../figures/ngc_6494/pm_ngc_6494.png}
       \caption{Proper motions}
@@
       \centering
-      \includegraphics[width=\textwidth]{../figures/ngc_6494/parallax_ngc_6494.pdf}
+      \includegraphics[width=\textwidth]{../figures/ngc_6494/parallax_ngc_6494.png}
       \caption{Parallax}
@@
     \centering
-    \includegraphics[width=\textwidth]{../figures/melotte_22/raw_pm_melotte_22.pdf}
+    \includegraphics[width=\textwidth]{../figures/melotte_22/raw_pm_melotte_22.png}
   \end{subfigure}
@@
       \centering
-      \includegraphics[width=\textwidth]{../figures/melotte_22/raw_parallax_melotte_22.pdf}
+      \includegraphics[width=\textwidth]{../figures/melotte_22/raw_parallax_melotte_22.png}
     \end{subfigure}
@@
       \centering
-      \includegraphics[width=\textwidth]{../figures/melotte_22/raw_hr_diagram_melotte_22.pdf}
+      \includegraphics[width=\textwidth]{../figures/melotte_22/raw_hr_diagram_melotte_22.png}
     \end{subfigure}
@@
   \centering
-  \includegraphics[width=0.9\textwidth]{../figures/openclust_catalogue.pdf}
+  \includegraphics[width=0.9\textwidth]{../figures/openclust_catalogue.png}
   \caption{OpenClust Catalogue Distribution}
@@
   \centering
-  \includegraphics[width=0.9\textwidth]{../figures/cluster_selection_tier1.pdf}
+  \includegraphics[width=0.9\textwidth]{../figures/cluster_selection_tier1.png}
   \caption{OpenClust Catalogue Selection Distribution}
@@
     \centering
-    \includegraphics[width=\textwidth]{../figures/melotte_22/features_melotte_22.pdf}
+    \includegraphics[width=\textwidth]{../figures/melotte_22/features_melotte_22.png}
   \end{subfigure}
@@
       \centering
-      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n2_pm_melotte_22.pdf}
+      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n2_pm_melotte_22.png}
     \end{subfigure}
@@
       \centering
-      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n5_pm_melotte_22.pdf}
+      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n5_pm_melotte_22.png}
     \end{subfigure}
@@
       \centering
-      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n8_pm_melotte_22.pdf}
+      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n8_pm_melotte_22.png}
     \end{subfigure}
@@
       \centering
-      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n2_parallax_melotte_22.pdf}
+      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n2_parallax_melotte_22.png}
       \caption{N clusters = 2}
@@
       \centering
-      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n5_parallax_melotte_22.pdf}
+      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n5_parallax_melotte_22.png}
       \caption{N clusters = 5}
@@
       \centering
-      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n8_parallax_melotte_22.pdf}
+      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n8_parallax_melotte_22.png}
       \caption{N clusters = 8}
@@
       \centering
-      \includegraphics[width=\textwidth]{../figures/melotte_22/kmeans_pm_melotte_22.pdf}
+      \includegraphics[width=\textwidth]{../figures/melotte_22/kmeans_pm_melotte_22.png}
       \caption{Proper Motion}
@@
       \centering
-      \includegraphics[width=\textwidth]{../figures/melotte_22/kmeans_parallax_melotte_22.pdf}
+      \includegraphics[width=\textwidth]{../figures/melotte_22/kmeans_parallax_melotte_22.png}
       \caption{Parallax}
@@
     \centering
-    \includegraphics[width=\textwidth]{../figures/melotte_22/kmeans_hr_diagram_melotte_22.pdf}
+    \includegraphics[width=\textwidth]{../figures/melotte_22/kmeans_hr_diagram_melotte_22.png}
   \end{subfigure}
@@
       \centering
-      \includegraphics[width=\textwidth]{../figures/melotte_22/dec_pm_melotte_22.pdf}
+      \includegraphics[width=\textwidth]{../figures/melotte_22/dec_pm_melotte_22.png}
       \caption{Proper Motion}
@@
       \centering
-      \includegraphics[width=\textwidth]{../figures/melotte_22/dec_parallax_melotte_22.pdf}
+      \includegraphics[width=\textwidth]{../figures/melotte_22/dec_parallax_melotte_22.png}
       \caption{Parallax}
@@
     \centering
-    \includegraphics[width=\textwidth]{../figures/melotte_22/dec_hr_diagram_melotte_22.pdf}
+    \includegraphics[width=\textwidth]{../figures/melotte_22/dec_hr_diagram_melotte_22.png}
   \end{subfigure}

### 2020-12-04 src/cdalvaro/utils/__init__.py
python: Add paths package

@@ -1,0 +1,1 @@
+from .paths import safe_path

### 2020-12-04 src/cdalvaro/graphics/plot.py
graphics: Add plot_cluster_position method

@@ -1,4 +1,4 @@
-from typing import Dict
+from typing import Dict, Union
 from matplotlib.axes import Axes
 import matplotlib.pyplot as plt
 import pandas as pd
@@ -10,7 +10,10 @@ from .color_palette import color_palette
 N_SAMPLES = 4000
 
 
-def _resample_data(data: pd.DataFrame, n_samples: int) -> pd.DataFrame:
+def _resample_data(data: pd.DataFrame, n_samples: Union[int, None]) -> pd.DataFrame:
+    if n_samples is None:
+        n_samples = N_SAMPLES
+
     if n_samples > 0 and data.shape[0] > n_samples:
         return data.sample(n_samples, random_state=0)
     return data
@@ -38,7 +41,9 @@ def _set_axis_properties(ax: Axes,
                          invert_xaxis: bool = False,
                          invert_yaxis: bool = False,
                          yscale: str = None,
-                         xscale: str = None):
+                         xscale: str = None,
+                         legend: bool = False,
+                         legend_title: str = None):
 
     if title is not None:
         ax.set_title(title)
@@ -67,6 +72,11 @@ def _set_axis_properties(ax: Axes,
     if xscale is not None:
         ax.set_xscale(xscale)
 
+    if legend and legend_title is not None:
+        _legend = ax.legend()
+        if _legend is not None:
+            _legend.set_title(legend_title)
+
 
 def plot_clusters_catalogue_distribution(data: pd.DataFrame,
                                          title: str = None,
@@ -83,9 +93,9 @@ def plot_clusters_catalogue_distribution(data: pd.DataFrame,
                          xlabel='Right Ascension (J2000 Degree)',
                          ylabel='Declination (J2000 Degree)',
                          xlim=xlim,
-                         ylim=ylim)
-
-    ax.legend().set_title("Diameter (arcmin)")
+                         ylim=ylim,
+                         legend=True,
+                         legend_title=r'Diameter ($arcmin$)')
 
     return fig, ax, g
 
@@ -99,7 +109,7 @@ def plot_downloaded_data_histogram(data: pd.DataFrame,
 
     g = sns.barplot(x='diam', y='value', hue=hue, data=data, palette=color_palette(), ax=ax)
 
-    _set_axis_properties(ax, title=title, xlabel='Diameter (arcmin)', ylabel='', yscale='log')
[…]

### 2020-12-04 thesis/thesis.tex

@@
 
-Stars that belong to a same OC share relative positions, so their parallax value is similar among them. They also share similar values of
-proper motion, both in right ascension and declination. Another property is that their metallicity must be uniform, since these stars were
-born from the same gas cloud and at the same temporary stage. However, to take into account this last property, we face the inconvenient that this
-parameter is not directly available in Gaia DR2 database. We will avoid this issue by looking at $E(b-r) / G_{mag}$ diagrams with well
-defined isochrone curves \cite{bressan2012parsec}. These isochrone curves are derived from theoretical models that mainly involve metallicity
-and mass/brightness from stars. With these models we can define important properties for the development of stars and their evolution in the
-Hertzsprung–Russell (H-R) diagram. Examples of these curves are shown in Figure \ref{fig:examples_of_isochrones}.
+Stars that belong to a same OC share relative positions, inherited from their original gas cloud.
+So their distance from the Earth is similar for all of them and, therefore, they have a narrow dispersion in their parallax value.
+They also share similar values of proper motion, both in right ascension and declination. Another property they share is their chemical composition.
+Thus, their metallicity must be uniform, since these stars were born from the same gas cloud and at the same time stage.
+However, to take this last property into account, we are faced with the drawback that this parameter is poorly reported in Gaia DR2 database.
 
-Again, since these stars were born close in time, regarding their intrinsic properties (mass/brightness), these stars must present a very
-well defined and sharp profile with low sparsing through the main sequence in the H-R diagram.
-
-We will take advantage of these properties to validate our OCs characterization.
+We will avoid this issue by looking at $E(b-r) / G_{mag}$ diagrams that reflect their connection with the isochrone curves, typical of their
+temporal evolution \cite{bressan2012parsec}. The isochrones are derived from theoretical models. These models are mainly based on metallicity
+and mass/brightness ratio of stars. They also have a direct correspondence with that presented by the Hertzsprung–Russell (H-R) diagram of the
+stars belonging to the cluster. Examples of these curves are shown in Figure \ref{fig:examples_of_isochrones}.
 
@@
 \end{figure}
+
+Since these stars were born close in time, they should have a sharp, well-defined profile with little scattering through the main sequence in the H-R diagram.
+Examples for this kind of diagrams are presented in Figure \ref{fig:examples_of_hr_diagrams}.
+
+\begin{figure}[htbp]
+  \centering
+  \begin{subfigure}{0.9\textwidth}
+    \centering
+    \begin{subfigure}[t]{0.45\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/melotte_22/hr_diagram_melotte_22.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.45\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_6494/hr_diagram_ngc_6494.png}
+    \end{subfigure}
+  \end{subfigure}
+  \caption{Examples of typical profiles in H-R diagrams for members of open clusters. On the left Melotte 22, on the right NGC 6494.}
+  \label{fig:examples_of_hr_diagrams}
+\end{figure}
+
+We will take advantage of these properties only to validate our characterization of OCs, but not to determine what stars belong to them. As we will explain later,
+we will only use dynamic properties such as proper motions and parallax to characterize open clusters.
+
+Although the members of a cluster appear in the observation visual field as an overdensity in positions, as shown in Figure \ref{fig:pos_ngc_2682}, these coordinates
+are not useful to separate those stars that belong to the cluster from the other which do not. However, if we look for overdensities in the proper motions configuration space,
+it is possible, at least at first instance, to assume a possible membership delimitation. See Figure \ref{fig:pm_ngc_2682}.
+
+\begin{figure}[htbp]
+  \centering
+  \begin{subfigure}{0.9\textwidth}
+    \centering
+    \begin{subfigure}[t]{0.45\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2682/pos_ngc_2682.png}
+      \caption{Positions map}
+      \label{fig:pos_ngc_2682}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.45\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2682/pm_ngc_2682.png}
+      \caption{Proper motions}
+      \label{fig:pm_ngc_2682}
+    \end{subfigure}
+  \end{subfigure}
+  \caption{NGC 2682 configuration spaces.}
+\end{figure}
+
+Overdensities in proper motion are not always so evident. Additionally, we will consider the parallax distribution to improve our cluster characterization.
+
+\begin{figure}[htbp]
+  \centering
+  \begin{subfigure}{0.9\textwidth}
+    \centering
+    \begin{subfigure}[t]{0.45\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/melotte_22/parallax_zoom_melotte_22.png}
+      \caption{Melotte 22 Parallax}
+      \label{fig:melotte_22_parallax_zoom}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.45\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/melotte_22/pm_melotte_22.png}
+      \caption{Melotte 22 Proper motion}
+    \end{subfigure}
+  \end{subfigure}
+  \caption{Melotte 22 Parallax histogram, clearly differentiated around the value $\approx 7.3mas$ and the corresponding diagram from the photometric magnitudes.}
+  \label{fig:melotte_22_pm_parallax}
+\end{figure}
+
+In many other cases things are not so simple, and the characterization process can turn very difficult and tedious using tools such as the available ones at
+the Virtual Observatory (VO), and frequently it requires parametrizations based on prior knowledge about the studied region. The aim of this work is to test
+an unsupervised and non-parametrized model based on Machine Learning (ML) tools.
 
@@
 
-The figure on the left shows a resonance at $\approx 7.5mas$ belonging the OC. While the figure on the right would help us to look for isochrone curves and so,
+The figure on the left shows a resonance at $\approx 7.3mas$ belonging the OC. While the figure on the right would help us to look for isochrone curves and so,
 to determine the age of these stars.
@@
 As shown in Figure \ref{fig:kmeans_comparisons_melotte_22}, bigger values for the number of clusters allow us to isolate more accurately
-the resonance in parallax at $\approx 7.5mas$. However, we have the disadvantage that more groups are formed, while our aim is to find just
+the resonance in parallax at $\approx 7.3mas$. However, we have the disadvantage that more groups are formed, while our aim is to find just
 one small cluster that contains the stars that belong to the OC and another one with the remaining stars.

### 2020-12-05 thesis/thesis.tex
thesis: Extend State of the Art chapter

@@
 \usepackage{caption}
+\usepackage{csquotes}
 \usepackage{eso-pic}
@@
       \centering
-      \includegraphics[width=\textwidth]{../figures/ngc_6494/hr_diagram_ngc_6494.png}
+      \includegraphics[width=\textwidth]{../figures/ngc_2682/hr_diagram_ngc_2682.png}
     \end{subfigure}
   \end{subfigure}
-  \caption{Examples of typical profiles in H-R diagrams for members of open clusters. On the left Melotte 22, on the right NGC 6494.}
+  \caption{Examples of typical profiles in H-R diagrams for members of open clusters. On the left Melotte 22, on the right NGC 2682.}
   \label{fig:examples_of_hr_diagrams}
@@
       \centering
-      \includegraphics[width=\textwidth]{../figures/melotte_22/pm_melotte_22.png}
-      \caption{Melotte 22 Proper motion}
+      \includegraphics[width=\textwidth]{../figures/melotte_22/hr_diagram_melotte_22.png}
+      \caption{Melotte 22 H-R diagram}
     \end{subfigure}
@@
 In many other cases things are not so simple, and the characterization process can turn very difficult and tedious using tools such as the available ones at
-the Virtual Observatory (VO), and frequently it requires parametrizations based on prior knowledge about the studied region. The aim of this work is to test
+the Virtual Observatory (VO), and frequently it requires parametrization based on prior knowledge about the studied region. The aim of this work is to test
 an unsupervised and non-parametrized model based on Machine Learning (ML) tools.
@@
 
-% TODO: Hablar de la misión Gaia
 % TODO: Hablar de KMedias y DEC
-% TODO: Mencionar trabajos relacionados
 
-An initial approach to find OC is the search of overdensities along the galactic disk. In general, this is the initial point,
-but even it looks simple, it presents a fundamental problem. The near field around the OC is fulled by two differenciated
-kind of stars populations: the ones which belong to the OC (from a tens or hundreds to a few thousand) and a background composed
-by thousands or millions of stars which don't belong to the OC. Find out which stars belong to the first group is the problem
-treated in this work. This selection is crucial to properly characterize the fundamental properties of the cluster
+An initial approach for finding OCs is the search of overdensities along the galactic disk. In general, this is the starting point but,
+although it seems simple, it presents a fundamental problem already discussed. The near field around the OC is filled with two types of
+distinct stars populations: those who belong to the OC (tens or hundreds to a few thousand) and a background made up of thousands or
+millions of stars that are not. Finding out which stars belong to the first group is the problem treated in this work.
+This selection is crucial to properly characterize the fundamental properties of the cluster
 (dynamics, total mass, age, chemical composition, etc.).
 
-Sometimes, this problem is easy to solve by looking into astrometric parameters, such as the analysis of proper motion of those
-stars belonging to the overdensity field.
+Sometimes, the problem is easy to solve, as we have seen, by studying astrometric parameters and looking for overdensities in the proper
+motion configuration space as well as in the parallax space.
 
@@
 It is easy to see that a subgroup can be located at [20, -45].
-Figure \ref{fig:pm_vec_melotte_22} shows the parallax density and confirms an overdensity at $\approx 7.5 mas$ corresponding to Mellote 22 OC.
+Figure \ref{fig:pm_vec_melotte_22} shows the parallax distribution and confirms an overdensity at
+$\approx 7.3 mas$ corresponding to Mellote 22 OC (See also Figure \ref{fig:melotte_22_parallax_zoom}).
 
@@
 becomes necessary to consider other parameters such as distances, or even metallicity and age (derived from isochrone curves)
-and requires photometric data from stars in the studied field
-\footnote{All figures have been sampled in order to reduce the document size, but are still a faithful representation of the original data}.
+and requires photometric data for the stars within the studied field.
 
@@
 
-At this point, it looks reasonable that the study of each individual cluster requires its own parametrization and technique
-to identify it.
+At this point, it seems reasonable that the study of each individual group requires its own parametrization and technique
+to identify it. With Gaia DR2, which contains great high-quality astrometric and photometric data for a huge number of stars,
+a new opportunity presents itself to simplify and optimize these processes. In this context, new approaches are being developed
+to improve current detection methods and automate OC characterization.
 
-With Gaia DR2, which contains great quality astrometric and photometric data for a huge number of stars,
-a new opportunity to simplify and optimize this processes is presented.
-In this context new approaches are being developed to improve current detection methods and automate the OC characterization.
-This work tries to take the best from \emph{Clusterix 2.0. A Virtual Observatory tool to estimate cluster membership probability}
-\cite{balaguer2020clusterix} and \emph{Hunting for open clusters in Gaia DR2: 582 new open clusters
-in the Galactic disc} \cite{castro2020hunting}, and merge them into a single \emph{machine learning}
-tool capable of finding open clusters efficiently without requiring powerful machines to perform the computations.
+\section{Gaia Mission}
 
-The first one is a non-parametrized method. It determines empirically the frequency functions associated to astrometrical variables
-without making any assumption about their profiles. However, it heavily does depend on the initial field selection and other
-parameters related to frequency functions. \emph{Clusterix 2.0.} is a web app tool inside the Spanish Virtual Observatory
-(VO) framework and works together with other VO tool such as:
-\emph{TOPCAT} \cite{taylor2005topcat},
-\emph{Aladin} \cite{bonnarel2000aladin},
-\emph{VOSA} \cite{bayo2008vosa}, among other.
+In this work we make use of Gaia DR2 since, although Release 3 (Gaia DR3) is imminent, it has not been available in time for us.
+Gaia DR2 is a multidimensional dataset obtained by ESA's Gaia mission (located at L2, 1.5 million kilometers from Earth) and
+operational since 2014. The catalogue has high precision and accuracy astrometric data for more than 1.7 billion stellar sources,
+and magnitudes in three photometric filters (G, BP and RP) for more than 1,300 million sources.
+80\% of Gaia DR2 sources are weaker than $G_{mag} \approx 18$. For magnitudes $G \approx 21$ nominal uncertainties reach
+$2 mas$ for parallaxes and $5 mas \cdot years^{-1}$ in proper motions. But for closer sources, (at the bright end, $G < 14$),
+the precision is of the order of $0.02 mas$ and $0.05 mas \cdot year^{-1}$ respectively \cite{cantat2018gaia}. In general,
+open clusters, which is the main subject of this work, are at optimal data thresholds.
 
-The second tool is based on Big Data techniques. Its purpose is to analyze the whole Gaia DR2 dataset looking for OCs by using
-machine learning algorithms divided in two stages. In the first stage, the algorithm looks for posible OC candidates by searching
-overdensities in the galactic disc.
-The second stage studies candidates found in the first one by generating images with H-R diagrams and passing them
-as sources to an Artificial Neural Network (ANN) which identifies isochrone patterns in photometric data (color and magnitude).
-Before this approach was put into practice there were 1200 OCs known. Now, this number has been increased up to more than 2000 OCs
-available at the Vizier catalogue \cite{ochsenbein2000vizier}.
+In this context, and taking advantage of the opportunity provided by Gaia DR2, new approaches are being developed to improve current
+detection methods and automating OC characterization as much as possible. This is the case of tools such as Clusterix 2.0
+\cite{balaguer2020clusterix}, TOPCAT \cite{taylor2005topcat}, and VOSA \cite{bayo2008vosa}, (all of them developed by the Virtual Observatory).
+These tools work together and offer good results, although it is not possible to use them in an unattended process and free of certain
+different parametrization in each case.
 
-The method developed in this project takes as base the approach made by \citeauthor{castro2020hunting} but rapidly diverges.
-In first place, performance has been taken into account since the beginning, trying to get a reliable method without compromising the
-needs of supercomputers.
-To achieve this, instead of blindly analyzing the whole galactic disk a fields selection is made focusing in those regions with known OCs.
-Then, these fields are analyzed with an homogeneous methodology based on machine learning algorithms using as sources astrometric and
-photometric variables. The goal is the characterization of each individual star in the cluster, discerning wether the star belongs the OC
-or not.
+\section{Current Methods}
 
-Thus, the main goal of this project is not the discovery of new open clusters or even candidates, but the development of a model that,
-applied to already known open clusters, allows us the characterization of those OCs on an unsupervised way, with a reasonable computing
-power and taking into account the same selection of astrometric and photometric variables for every OC.
+TOPCAT alone does not offer an independent discrimination mechanisms to researchers. They should try an initial selection, staring from the
+map of proper motions, by delimiting the area where the cluster presumably appears. When this is not possible, it is necessary to have a
+certain knowledge about the cluster profile and to parametrize some approximated values. In both cases, the definition of the selected
+set must then be optimized step by step until reaching a result that can be considered reasonable.
 
-% TODO: Concluir con una última sección de resumen de conclusiones, resumiendo las principales averiguaciones del estudio del dominio
-% y cómo van a afectar al desarrollo específico del trabajo.
+Another tool is Clusterix 2.0, which is an interactive web-based tool \cite{balaguer2020clusterix}. It takes the proper motion diagram
+without assuming any prior profiles assumptions for the stars candidates membership and determines empirically the frequency functions.
+
+Clusterix uses normal Gaussian kernel functions, defined as:
+
+\begin{equation}
+  K(a, b) = \frac{1}{2 \pi h^{2}} \exp{ \left[ -\frac{1}{2}\frac{\left( a - a_{i} \right)^{2} + \left( b - b_{i} \right)^{2}}{ h^{2}} \right]}
+\end{equation}
+
+$(a, b)$ are referred to the proper motion configuration space, while a point located in the center of the cell $(i, j)$ provides the maximum
+contribution for calculating the local density. In this context, the number $h$ is called the \emph{smoothing} parameter and it is measured in
+the same units as proper motion.
+
+With this same purpose of applying non-parametric methods, various clustering algorithms have been applied in the past, UPMASK \cite{krone2014upmask}
+was developed to only use photometry and positions, and then it was adapted to the Gaia data \cite{cantat2018gaia} based only on proper motion and
+parallax. A similar approach has been carried out using DBSCAN and machine learning algorithms to discover new clusters.
+
+In this sense, Clusterix also claims to be a non-parametric method, but it critically depends on the initial selection for the field sizes to be analyzed.
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=0.9\textwidth]{../figures/clusterix/selection.pdf}
+  \caption{Typical Clusterix 2.0 control panel.}
+  \label{fig:clusterix_control_panel}
+\end{figure}
+
+Clusterix relies on several assumptions, among other:
+
+\begin{displayquote}
+The non-field population does not occupy the entire workspace, but is spatially concentrated, which makes it possible to distinguish two regions
+in the workspace: the only field region (label 'f'), dominated by stars field, and the cluster + field region (label 'c+f'), which includes field
+stars and not field stars. \cite{balaguer2020clusterix}
+\end{displayquote}
+
+This fact implies defining three areas or regions with different radius. The first 'c+f' corresponds to the one in which the cluster members are
+presumed to be contained together with other field stars that are not part of the cluster. The second, broader, assumes that it only contains
+stars in an extended visual field without components of the cluster, and a third intermediate region, which is out of analysis (void area), which
+would correspond to a possible transition zone between the other two.
+
+The correct choice of these radii, even having a previous estimation for the 'c+f' region, significantly affects the execution of the algorithm and,
+in general, requires a generously wide field 'f'. There is no general rule that defines relative proportions of these areas.
+
+Finally, when an acceptable result is obtained, what Clusterix provides is a probability field associated with the set of stars under study.
+The dataset, with an extra column for the probability value of each object to belong to the open cluster, can be exported to TOPCAT
+for further optimization and final processing. Figure \ref{fig:clusterix_probability} shows an example of the probability distribution in the proper
+motion space for NGC 2682.
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=0.9\textwidth]{../figures/clusterix/statistics.pdf}
+  \caption{Probability distribution in the proper motion space provided by Clusterix for NGC 2682.}
+  \label{fig:clusterix_probability}
+\end{figure}
+
+Finally, it is necessary to refer to the work of \cite{castro2020hunting}, which presents a method based on machine learning techniques to make
+a systematic search for overdensities in the astrometric space of the galactic disk and a subsequent identification of OC using photometric
+information, also from Gaia DR2.
+
+The method includes two phases: the first one uses an unsupervised clustering algorithm, DBSCAN, to search for overdensities
+$(l, b \pi, \mu_{\alpha} *, \mu_{\delta})$, ans subsequently applies a deep learning ANN trained with magnitude diagrams to identify isochrone
+patterns within the detected overdensities and thus proceed to confirm them as OC.
+
+It should be noted that for the execution of this method, MareNostrum 42 (Bacelona Supercomputing Center) was used. So the neural network could handle
+the image recognition process with isochrone patterns and not applying theoretical models derived from values such as metallicity, masses, among other.
+
+The result of this work is the recognition of 582 new open clusters distributed along the galactic disk for a galactic declination $b < 20^{\circ}$,
+which has meant an increase of 45\% over previously ones.
+
+Our aim in this work, linked to computational limitations compared to MareNostrum, is not to do a blind search for clusters, but to obtain a method,
+in the machine learning environment, that allows the characterization of open clusters through unsupervised and non-parametrized procedures in a
+single step. This method should not require further refinements with other tools.
+
+The result of Clusterix + TOPCAT, applied to the same clusters, will be used only for comparing and validating our results.
 

### 2020-12-05 thesis/thesis.tex
thesis: Extend Method chapter

@@
 For the sake of simplicity, we are taking \emph{Melotte 22 (Messier 45)} \cite{elsanhoury2019ppmxl} to illustrate the whole method. This cluster is well studied
-and their stars are not too mixed with other region stars.
+and their stars are not too mixed with other region stars, so images look clearly.
 
@@
 
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=0.9\textwidth]{../figures/services_diagram.pdf}
+  \caption{Data retrieval and analysis}
+  \label{fig:services_diagram}
+\end{figure}
+
+These containers are orchestrated with a \verb|docker-compose.yml| file which automatically launches one instance of each container.
+Figure \ref{fig:services_diagram} shows a diagram describing the services architecture for this work.
+
 The database has two main tables in \verb|public| schema:
@@
 \begin{itemize}
-  \item \verb|regions| for storing cluster properties such as \emph{location}, \emph{diameter} and other properties
-  \item \verb|gaiadr2_source| which contains data for all downloaded stars from Gaia
+  \item \verb|public.regions| for storing cluster properties such as \emph{location}, \emph{diameter} and other properties
+  \item \verb|public.gaiadr2_source| which contains data for all downloaded stars from Gaia
 \end{itemize}
 
-The second table is partitioned by region, storing all stars related to one cluster inside an independent table for optimized access performance.
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=0.9\textwidth]{../figures/customdb_diagram.pdf}
+  \caption{Custom DB diagram}
+  \label{fig:customdb_diagram}
+\end{figure}
 
-Methods for retrieving information from the database are available through an instance of \verb|cdalvaro.DB|. This class contains methods
-for recovering stars information for a given region as a Pandas \verb|DataFrame| ready for data analysis.
+The second table is partitioned by region, storing all stars related to one region inside an independent table for optimized access performance.
+The partition tables are located inside \verb|gaiadr2| schema. See Figure \ref{fig:customdb_diagram}.
 
-These containers are orchestrated with a \verb|docker-compose.yml| file which automatically launches one instance of each container.
+Although \verb|public.gaiadr2_source| is partitioned, there is no need for us to directly access to one of those tables.
+We only need to know the \verb|id| for the desired region and query for its stars to the main table as follows:
+
+\begin{minted}{sql}
+  -- Get region id
+  SELECT id FROM public.regions WHERE name = 'Melotte 22';
+  -- output: id -> 302
+
+  -- Get stars for the given id
+  -- NOTE: PostgreSQL cannot optimize this query when region_id is
+  --       provided as a subquery. This is why we get first the region id.
+  SELECT * FROM public.gaiadr2_source WHERE region_id = 302;
+\end{minted}
+
+PostgreSQL knows that all stars we want to are located inside \verb|gaiadr2.region_302| so it does not need to query any other tables.
+This allows us to download data for as many regions as we want to without loosing efficiency when accessing for that data.
+
+\verb|gaiadr2_source| and its partitions contain all fields available at the Gaia DR2 database.
+A complete list of these fields and their description can be found at the following link:
+\sloppy\url{https://gea.esac.esa.int/archive/documentation/GDR2/Gaia_archive/chap_datamodel/sec_dm_main_tables/ssec_dm_gaia_source.html}
+
+Table \ref{tab:important_fields} shows a brief description of the fields used in this work.
+
+\begin{table}[h!]
+  \begin{center}
+    \begin{tabular}{r|c|l}
+      \textbf{Name} & \textbf{Units} & \textbf{Description} \\
+      \hline
+      \verb|ra| & Angle [deg] & Right ascension \\
+      \verb|dec| & Angle [deg] & Declination \\
+      \verb|parallax| & Angle [mas] & Parallax \\
+      \verb|pmra| & Angular Velocity [mas/year] & Proper motion in right ascension \\
+      \verb|pmdec| & Angular Velocity [mas/year] & Proper motion in declination \\
+      \verb|phot_g_mean_mag| & Magnitude [mag] & G-band mean magnitude \\
+      \verb|bp_rp| & Magnitude [mag] & BP - RP colour \\
+    \end{tabular}
+    \caption{Properties and descriptions of the fields used in this work.}
+    \label{tab:important_fields}
+  \end{center}
+\end{table}
+
+The downloader takes data from the Gaia DR2 database and save the selection into our custom database without modifying any value,
+so data preserves all its properties and units.
+
+Python methods for retrieving information from the database are available through an instance of \verb|cdalvaro.DB|.
+This class contains methods for recovering stars information for a given region as a Pandas \verb|DataFrame| ready for data analysis.
 
@@
 \href{https://github.com/cdalvaro/machine-learning-master-thesis}{cdalvaro/machine-learning-master-thesis} repository.
+(For the printed version of this work, you can scan the QR code on cover for accessing to it.)
 
@@
 
-Parallax is another important feature. It lets us know how far stars are from us. And, since all stars within an open cluster were born from the same dust cloud,
-they must all have similar parallax.
+Parallax is another important feature. It lets us know how far stars are from us.
+And, since all stars within an open cluster were born from the same dust cloud, they must all have similar parallax.
 
@@
 
-First, we correct proper motion in right ascension and declination by dividing them by the parallax. That way we normalize these quantities and help our clustering
-models to improve their performance.
+First, we correct proper motion in right ascension and declination by dividing them by the parallax.
+That way we normalize these quantities and help our clustering models to improve their performance.
 
-Another computed property that we are considering is the modulus of the proper motion. That way, we relate both features and therefore force our model to keep them tight.
+Another computed property that we are considering is the modulus of the proper motion.
+That way, we relate both features and therefore force our model to keep them tight.
 
@@
   \centering
-  \begin{subfigure}{0.9\textwidth}
+  \begin{subfigure}{\textwidth}
     \centering
@@
 
-So, in summary, these are the features we are going to use as sources for our clustering models:
+\pagebreak
+
+In summary, these are the features we are going to use as sources for our clustering models:
 
@@
 
-The following example can be found in the Jupyter notebook: \verb|cluster_characterization.ipynb| (available at \verb|src/notebooks|).
+The following example is a snippet copied from \verb|cluster_characterization.ipynb| Jupyter notebook (available at \verb|src/notebooks|).
 It shows how to estimate the best number of clusters for Melotte 22 by using \verb|cdalvaro.ml.estimate_n_clusters| method.
@@
 \begin{minted}{python}
-  n_clusters, kmeans = estimate_n_clusters(melotte22_df, min_clusters=3,
-                                           max_clusters=10, verbose=True)
-
+  n_clusters, kmeans = estimate_n_clusters(melotte22_df,
+                                           min_clusters=3, max_clusters=7)
   # Output
@@
   # Silhouette score for 7 clusters: 0.5306
-  # Silhouette score for 8 clusters: 0.4978
-  # Silhouette score for 9 clusters: 0.5214
-  # Silhouette score for 10 clusters: 0.5199
   # Best silhouette score is 0.5608 for 5 clusters
@@
   \centering
-  \begin{subfigure}{0.5\textwidth}
+  \begin{subfigure}{0.9\textwidth}
+    \centering
+    \begin{subfigure}[t]{0.45\textwidth}
       \centering
       \includegraphics[width=\textwidth]{../figures/melotte_22/dec_hr_diagram_melotte_22.png}
+      \caption{DEC H-R diagram}
     \end{subfigure}
-  \caption{Melotte 22 H-R diagram with DEC characterization}
-  \label{fig:dec_hr_diagram_melotte_22}
+    \hfill
+    \begin{subfigure}[t]{0.45\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/melotte_22/hr_diagram_melotte_22.png}
+      \caption{Clusterix+TOPCAT H-R diagram}
+    \end{subfigure}
+  \end{subfigure}
+  \caption{H-R diagram comparison between DEC clustering and Clusterix+TOPCAT method.}
+  \label{fig:hr_diagram_comparison}
 \end{figure}
+
+Finally, Figure \ref{fig:hr_diagram_comparison} shows a comparison between the H-R diagram obtained by DEC clustering and the one obtained
+using the Clusterix + TOPCAT method.
 

### 2020-12-07 src/cdalvaro/ml/utils.py
python: Add ML utils

@@ -1,7 +1,7 @@
 import pandas as pd
 from sklearn.cluster import KMeans
 from sklearn.metrics import silhouette_score
-from typing import Tuple
+from typing import List, Tuple
 
 
 def estimate_n_clusters(x,
@@ -63,3 +63,43 @@ def filter_outliers(df: pd.DataFrame, q: float = 0.05) -> pd.Series:
     lower = df.quantile(q)
     mask = (df < upper) & (df > lower)
     return mask
+
+
+def cluster_centers(df: pd.DataFrame, key: str, columns: List[str] = []) -> pd.DataFrame:
+    """
+    Return a Pandas DataFrame with centers for each cluster.
+
+    Args:
+        df (pd.DataFrame): The DataFrame with data to be analyzed.
+        key (str): The key of the cluster reference column.
+        columns (List[str], optional): List of columns to take into account. Defaults to [].
+
+    Returns:
+        pd.DataFrame: A DataFrame with center for each cluster and column.
+    """
+    if len(columns) > 0 and key not in columns:
+        columns.append(key)
+    __df = df[columns] if len(columns) > 0 else df
+
+    return __df.groupby(by=key).mean()
+
+
+def cluster_stats(df: pd.DataFrame, key: str, columns: List[str] = [], **kwargs) -> pd.DataFrame:
+    """
+    Return a Pandas DataFrame with statistics for each cluster.
+
+    Args:
+        df (pd.DataFrame): The DataFrame with data to be analyzed.
+        key (str): The key of the cluster reference column.
+        columns (List[str], optional): List of columns to take into account. Defaults to [].
+        **kwargs (optional): Extra arguments are passed to Pandas.DataFrame.describe() method.
+        https://pandas.pydata.org/pandas-docs/stable/reference/api/pandas.core.groupby.DataFrameGroupBy.describe.html
+
+    Returns:
+        pd.DataFrame: A DataFrame with statistics for each cluster and column.
+    """
+    if len(columns) > 0 and key not in columns:
+        columns.append(key)
+    __df = df[columns] if len(columns) > 0 else df
+
+    return __df.groupby(by=key).describe(**kwargs)

### 2020-12-07 thesis/thesis.tex
thesis: Update Method chapter

@@
 \usepackage{subcaption}
+\usepackage{tablefootnote}
 \usepackage[nottoc]{tocbibind}
@@
   \includegraphics[width=0.9\textwidth]{../figures/clusterix/selection.pdf}
-  \caption{Typical Clusterix 2.0 control panel.}
+  \caption{Typical Clusterix 2.0 region selection panel.}
   \label{fig:clusterix_control_panel}
@@
 \chapter{Method}
+\label{chap:method}
 
@@
 \section{Data Mining}
+\label{sec:data_mining}
 
@@
 
-This download is not limited to the size registered in the catalogue. Instead, a wider region is downloaded for each cluster to include
-stars outside the cluster. The idea is that the unsupervised model must be able to clusterize the stars that belongs to the OC and to discard
+This download is not limited to the size registered in the catalogue.
+Instead, a wider region (1.5 times the size of the cluster) is downloaded for each cluster to include stars outside the cluster.
+The idea is that the unsupervised model must be able to clusterize the stars that belongs to the OC and to discard
 those outsiders when characterizing the open cluster.
@@
 
-Once the model is ready, we can test it by applying it to the Melotte 22 dataset. Notebook \verb|cluster_characterization.ipynb| illustrates how DEC model
-processes a region of stars trying to characterize the open cluster hidden within that region.
+Once the model is ready, we can test it by applying it to the Melotte 22 dataset.
+Notebook \verb|cluster_characterization.ipynb| illustrates how DEC model processes a region of stars trying to characterize the open cluster hidden within that region.
 
@@
 
+Figure \ref{fig:hr_diagram_comparison} shows a comparison between the H-R diagram obtained by DEC clustering and the one obtained
+using the Clusterix + TOPCAT method.
+
 \begin{figure}[htbp]
@@
 
-Finally, Figure \ref{fig:hr_diagram_comparison} shows a comparison between the H-R diagram obtained by DEC clustering and the one obtained
-using the Clusterix + TOPCAT method.
+Finally, we can refine this selection by filtering those stars which are bellow and above the 0.25 and 0.75 quantiles for each group, respectively.
+That way we remove the most doubtful values from the selection.
+Figure \ref{fig:melotte_22_filtered} shows the result after filtering DEC clustering selection.
+
+\begin{figure}[htbp]
+  \centering
+  \begin{subfigure}{0.9\textwidth}
+    \centering
+    \begin{subfigure}[t]{0.45\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/melotte_22/dec_hr_diagram_filtered_melotte_22.png}
+      \caption{H-R diagram}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.45\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/melotte_22/dec_parallax_filtered_melotte_22.png}
+      \caption{Parallax}
+    \end{subfigure}
+  \end{subfigure}
+  \caption{Melotte 22 H-R diagram and parallax histogram with stars outside 0.25 and 0.75 quantiles filtered.}
+  \label{fig:melotte_22_filtered}
+\end{figure}
+
+Table \ref{tab:results_melotte_22} shows a summary of Melotte 22 with values taken from different sources and methods: SIMBAD Astronomical Database - CDS (Strasbourg),
+Clusterix+TOPCAT method, and the results obtained for each step of this method (KMeans, DEC and DEC filtered).
+
+% TODO: Fix table if figures changes
+\begin{table}[h!]
+  \begin{center}
+    \begin{tabular}{l|c|c|c|c}
+      \textbf{Source / method} & \textbf{$\mu_{\alpha}$ $(mas/yr^{-1})$} & \textbf{$\mu_{\delta}$ $(mas/yr^{-1})$} & \textbf{$\varpi$ $(mas)$} & \emph{\# stars} \\
+      \hline
+      Simbad.u-strasbg.fr\tablefootnote{Results have been taken from \protect\citeA[p.~25. Table A.3. Pleiades]{babusiaux2018gaia}} &
+        19.997 $\pm$ 0.127 & -45.548 $\pm$ 0.101 & 7.364 $\pm$ 0.005 & 1326 \\
+      Clusterix+TOPCAT & 19.98 $\pm$ 1.25 & -45.47 $\pm$ 1.48 & 7.33 $\pm$ 0.21 & 168 \\
+      KMeans & 20.25 $\pm$ 35.16 & -38.01 $\pm$ 40.16 & 7.23 $\pm$ 2.35 & 1378 \\
+      DEC & 23.67 $\pm$ 38.33 & -46.23 $\pm$ 44.54 & 8.04 $\pm$ 2.59 & 878 \\
+      DEC (filtered) & 19.50 $\pm$ 8.59 & -44.23 $\pm$ 8.17 & 7.42 $\pm$ 0.1 & 438 \\
+    \end{tabular}
+    \caption{Melotte 22. The parameters shown are proper motion in right ascension and declination, parallax and number of stars with their respective deviations.}
+    \label{tab:results_melotte_22}
+  \end{center}
+\end{table}
 

### 2020-12-07 thesis/thesis.tex
thesis: Add Cluster Characterization with VO Tools chapter

@@
       \hline
-      Simbad.u-strasbg.fr\tablefootnote{Results have been taken from \protect\citeA[p.~25. Table A.3. Pleiades]{babusiaux2018gaia}} &
+      \textbf{Simbad.u-strasbg.fr}\tablefootnote{Results have been taken from \protect\citeA[p.~25. Table A.3. Pleiades]{babusiaux2018gaia}} &
         19.997 $\pm$ 0.127 & -45.548 $\pm$ 0.101 & 7.364 $\pm$ 0.005 & 1326 \\
-      Clusterix+TOPCAT & 19.98 $\pm$ 1.25 & -45.47 $\pm$ 1.48 & 7.33 $\pm$ 0.21 & 168 \\
+      Clusterix+TOPCAT & 19.98 $\pm$ 1.25 & -45.47 $\pm$ 1.48 & 7.33 $\pm$ 0.21 & 634 \\
       KMeans & 20.25 $\pm$ 35.16 & -38.01 $\pm$ 40.16 & 7.23 $\pm$ 2.35 & 1378 \\
       DEC & 23.67 $\pm$ 38.33 & -46.23 $\pm$ 44.54 & 8.04 $\pm$ 2.59 & 878 \\
-      DEC (filtered) & 19.50 $\pm$ 8.59 & -44.23 $\pm$ 8.17 & 7.42 $\pm$ 0.1 & 438 \\
+      \textbf{DEC (filtered)} & 19.50 $\pm$ 8.59 & -44.23 $\pm$ 8.17 & 7.42 $\pm$ 0.1 & 438 \\
     \end{tabular}
@@
 
-We have seen in the previous chapter that the DEC model is able to refine the clustering made by K-Means. This allows us to get
-a more accurate characterization of the cluster.
-
-In order to validate our model and be able to say how good or bad it is we need a different method to characterize the cluster
-and compare one against each other.
-
-For that reason we have used tools from the Virtual Observatory (VO) such as Clusterix and TOPCAT to characterize the studied clusters.
+We have seen in the previous chapter that the DEC model is capable of refining the clustering selection made by K-Means.
+This allows us to obtain a more precise characterization of the cluster.
+Now, in order to validate our model and be able to estimate how good or bad it is,
+we need a different method to characterize the studied clusters and compare our results with the validation method.
+
+For that reason, we have used an alternative method based on Virtual Observatory (VO) tools such as Clusterix and TOPCAT,
+to characterize the studied clusters.
 
 \section{Cluster Characterization with VO Tools}
+
+The aim of this section is to highlight procedure differences between the our method and this new one,
+as well as using the results given by this method to validate ours results.
+This validation method uses a combination of VO tools: Clusterix 2.0 + TOPCAT.
+We take Melotte 22 again as our study case to make easier taking comparisons between the method presented in Chapter \ref{chap:method}
+and this alternative method.
+
+The initial point, in order to avoid a subjective and manual selection for the cluster members, is taking advantage of Clusterix.
+Figure \ref{fig:clusterix_selection} shows a screenshot of the Clusterix 2.0 selection panel.
+
+\begin{figure}[htbp]
+  \centering
+  \begin{subfigure}{0.9\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{../figures/clusterix/clusterix_selection_panel_melotte_22.png}
+  \end{subfigure}
+  \caption{Clusterix 2.0 selection panel with Melotte 22 settings.
+           The coordinates correspond to the center of Melotte 22 cluster,
+           while the radius is the value registered for this cluster in the
+           OpenClust catalogue multiplied by 1.5 (as explained in \ref{sec:data_mining})}
+  \label{fig:clusterix_selection}
+\end{figure}
+
+This panel allows us to select a region by its coordinates or by searching a cluster by its name in the Gaia DR2 catalogue,
+and specify the size of the region by introducing its radius.
+We can also set limit values in magnitude (for both, minimum and maximum) to discard stars outside those limits.
+Another setting is the flag to enable or disable the \emph{Q-Filter}.
+This filter discards those stars with \emph{null} values in relevant fields.
+
+Clusterix returns a position map for the downloaded stars from Gaia DR2 database using Aladin \cite{bonnarel2000aladin} as an atlas of the sky
+(See Figure \ref{fig:aladin_melotte_22}).
+At this point, we can choose sending the recovered data to TOPCAT or Aladin, or continue to the next step of this assistant.
+
+We select the second choice and proceed with the proper motion frequency space analysis.
+
+\begin{figure}[htbp]
+  \centering
+  \begin{subfigure}{0.9\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{../figures/clusterix/clusterix_aladin_selection_melotte_22.png}
+  \end{subfigure}
+  \caption{Aladin screenshot with equatorial coordinates for Melotte 22 stars referred to J2000.}
+  \label{fig:aladin_melotte_22}
+\end{figure}
+
+In this step, we define three regions that will be used for the frequency analysis:
+
+\begin{itemize}
+  \item \textbf{Inner region} (\emph{c+f}), with stars that belong to the OC and field stars.
+  \item \textbf{External region} or field region (\emph{f}), with stars that do not belong the open cluster.
+  \item \textbf{Void region}. This region is excluded from the analysis since it is supposed to be a transition region with stars that could belong the OC.
+\end{itemize}
+
+It is necessary to notice the importance of this step, since the selection of these regions is critical to get a good result.
+However, despite how crucial is this selection, there is no general rule to define these regions.
+We only have two references: the radius of the downloaded region and the estimated size for the open cluster that sets the \emph{inner radius}.
+The other two radius must be set by trial and error and, in general, they must be generous to be sure we are not taking polluted regions.
+This is a tedious process that requires several tries before achieving good enough probability values.
+
+Furthermore, it is necessary to know in advance some additional profile of the cluster, in addition to the estimated size.
+For example, in the case of Melotte 22, the proper motion in declination direction of the membership stars is around $-45 mas \cdot year^{-1}$,
+so it is necessary to open the $\mu$ range above the estimated values (in absolute value). If we do not do that we loose its detection,
+since Clusterix sets this value to $30 mas \cdot year^{-1}$ by default. But even in this case, it is necessary to set a threshold for
+magnitude values, $G \leq 14$, to help getting the right result.
+Figure \ref{fig:clusterix_step2_melotte_22} shows the region selection panel on Clusterix 2.0.
+
+This makes the selection process an art rather than a science, not to mention the required previous knowledge on certain profiles of the
+objects to characterize. This puts into question the non-parametric character of the method.
+
+\begin{figure}[htbp]
+  \centering
+  \begin{subfigure}{0.9\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{../figures/clusterix/clusterix_step2_melotte_22.png}
+  \end{subfigure}
+  \caption{Clusterix 2.0 Step 2/3: Region selection.}
+  \label{fig:clusterix_step2_melotte_22}
+\end{figure}
+
+After having set all these parameters, Clusterix performs the proper motion frequency analysis and returns the probability field
+associated to each star of the whole region (see Figure \ref{fig:clusterix_probability_melotte_22}).
+
+This step can be repeated over and over again, varying radii and other parameters, until finally reach a valid probability that works.
+The last step presents a table with candidates objects and allows to export this dataset to TOPCAT, Aladin and VOSA to perform a deeper
+filtering and analysis based on the results given by Clusterix.
+
+\begin{figure}[htbp]
+  \centering
+  \begin{subfigure}{0.9\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{../figures/clusterix/clusterix_probability_melotte_22.pdf}
+  \end{subfigure}
+  \caption{Kinematic probability associated to the membership of the open cluster.
+           This probability is derived from the proper motion frequency analysis.}
+  \label{fig:clusterix_probability_melotte_22}
+\end{figure}
+
+Once data is available in TOPCAT, we have to carefully analyze the imported data.
+From there, we can perform a first stars selection by excluding those stars with probability value lower than 0.8.
+This first result can be taken as a good staring point, but we still need to make some fine tunning.
+As we can see in Figure \ref{fig:topcat_prob_selection_melotte_22}, the selection is too restrictive and we have lost too much stars.
+
+Looking at Figure \ref{fig:topcat_prob_parallax_melotte_22}, we can assume that cluster members are around $7.3mas$ in the parallax histogram.
+Also, Figure \ref{fig:topcat_prob_pm_melotte_22} reveals that cluster stars have low scattering rate in the proper motion configuration space.
+
+\begin{figure}[htbp]
+  \centering
+  \begin{subfigure}{0.9\textwidth}
+    \centering
+    \begin{subfigure}[t]{0.45\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/clusterix/topcat_prob_pm_melotte_22.png}
+      \caption{Proper motion}
+      \label{fig:topcat_prob_pm_melotte_22}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.45\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/clusterix/topcat_prob_parallax_melotte_22.png}
+      \caption{Parallax}
+      \label{fig:topcat_prob_parallax_melotte_22}
+    \end{subfigure}
+  \end{subfigure}
+  \caption{Melotte 22 first selection using validation method. Parallax is $7.304mas$ with 0.797 of standard deviation.}
+  \label{fig:topcat_prob_selection_melotte_22}
+\end{figure}
+
+Taken these observations into account, we can proceed to make a new selection based on parallax properties.
+Our second selection uses the values obtained with the previous selection for parallax and its standard deviation and
+selects those stars from the whole set which fall inside the limits $7.304mas \pm 0.797$ (see Figure \ref{fig:topcat_parallax_melotte_22}).
+
+\begin{figure}[htbp]
+  \centering
+  \begin{subfigure}{0.9\textwidth}
+    \centering
+    \begin{subfigure}[t]{0.45\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/clusterix/topcat_2nd_selection_pm_melotte_22.png}
+      \caption{Proper motion}
+      \label{fig:topcat_2nd_selection_pm_melotte_22}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.45\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/clusterix/topcat_2nd_selection_parallax_melotte_22.png}
+      \caption{Parallax}
+      \label{fig:topcat_2nd_selection_parallax_melotte_22}
+    \end{subfigure}
+  \end{subfigure}
+  \caption{Melotte 22 second selection. The estimated number of cluster members is 709.}
+  \label{fig:topcat_parallax_melotte_22}
+\end{figure}
+
+Looking at the H-R diagrams (Figure \ref{fig:clusterix_selection_hr_diagrams_melotte_22}),
+we can compare both selections and see how we have extended the selection without getting too much noise around the isochrone curve.
+
+\begin{figure}[htbp]
+  \centering
+  \begin{subfigure}{0.9\textwidth}
+    \centering
+    \begin{subfigure}[t]{0.45\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/clusterix/topcat_1st_selection_hr_diagram_melotte_22.png}
+      \caption{H-R diagram from first selection.}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.45\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/clusterix/topcat_2nd_selection_hr_diagram_melotte_22.png}
+      \caption{H-R diagram from second selection.}
+    \end{subfigure}
+  \end{subfigure}
+  \caption{Comparison between first and second selection H-R diagrams.}
+  \label{fig:clusterix_selection_hr_diagrams_melotte_22}
+\end{figure}
+
+Finally, we filter the second selection by discarding those stars whose parallax relative error in absolute value is behind 0.05
+(see Figure \ref{fig:clusterix_final_selection_melotte_22}).
+
+\begin{figure}[htbp]
+  \centering
+  \begin{subfigure}{0.9\textwidth}
+    \centering
+    \begin{subfigure}[t]{0.45\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/clusterix/topcat_3rd_selection_parallax_melotte_22.png}
+      \caption{Parallax}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.45\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/clusterix/topcat_3rd_selection_hr_diagram_melotte_22.png}
+      \caption{H-R diagram}
+    \end{subfigure}
+  \end{subfigure}
+  \caption{Comparison between first and second selection H-R diagrams.}
+  \label{fig:clusterix_final_selection_melotte_22}
+\end{figure}
+
+For the sake of completeness, Table \ref{tab:results_melotte_22} is shown again, in order to compare the results obtained with the validation method and our method.
+
+\begin{table}[h!]
+  \begin{center}
+    \begin{tabular}{l|c|c|c|c}
+      \textbf{Source / method} & \textbf{$\mu_{\alpha}$ $(mas/yr^{-1})$} & \textbf{$\mu_{\delta}$ $(mas/yr^{-1})$} & \textbf{$\varpi$ $(mas)$} & \emph{\# stars} \\
+      \hline
+      \textbf{Simbad.u-strasbg.fr} & 19.997 $\pm$ 0.127 & -45.548 $\pm$ 0.101 & 7.364 $\pm$ 0.005 & 1326 \\
+      \textbf{Clusterix+TOPCAT} & 19.98 $\pm$ 1.25 & -45.47 $\pm$ 1.48 & 7.33 $\pm$ 0.21 & 634 \\
+      KMeans & 20.25 $\pm$ 35.16 & -38.01 $\pm$ 40.16 & 7.23 $\pm$ 2.35 & 1378 \\
+      DEC & 23.67 $\pm$ 38.33 & -46.23 $\pm$ 44.54 & 8.04 $\pm$ 2.59 & 878 \\
+      DEC (filtered) & 19.50 $\pm$ 8.59 & -44.23 $\pm$ 8.17 & 7.42 $\pm$ 0.1 & 438 \\
+    \end{tabular}
+  \end{center}
+\end{table}
 

### 2020-12-09 thesis/thesis.tex
thesis: Update title
thesis: Update Introduction chapter

@@
 %---------------------------
-\title{Open Clusters Characterization in Gaia DR2 Using ML Algorithms}
+\title{Using Machine Learning Tools for Open Cluster Characterization with Gaia DR2 Data}
 \author{Carlos David Álvaro Yunta}
@@
 
-% TODO: 1.1 Motivation
-
-% ¿Cuál es el problema a tratar?
-% ¿Cuáles crees que son las causas?
-% ¿Por qué es relevante el problema?
-
-% TODO: 1.2 Planteamiento
-
-% ¿Cómo se podría solucionar el problema?
-% ¿Qué es lo que se propone?
-% Descripción del objetivo del trabajo en términos generales
-
-% TODO: 1.3 Estructura de la memoria
-
-% Descripción breve de lo que se va a contar en cada uno de los capítulos siguientes.
-
 Stellar open clusters (OCs) \cite{janes1982open} are groups of stars gravitationally bound originated from a single molecular gas cloud.
@@
 are formed in clustered environments and spend a period of time gravitationally bound with their siblings embedded
-in their original molecular cloud
-\cite{clarke2000theformationofstellarclusters} \cite{portegies2010young}.
+in their original molecular cloud \cite{clarke2000theformationofstellarclusters} \cite{portegies2010young}.
 The evolution of these systems tends to sparse them in a few million years by interacting gravitationally with other systems.
-Galactic tidal forces and mechanisms that involves the gas loss driven by stellar feedback are other causes of disruption
+Galactic tidal forces and mechanisms that involves gas loss driven by stellar feedback are other causes of disruption
 \cite{brinkmann2017bound}.
@@
 Young OCs allow us to research stars formation regions and improve our understanding about the mechanisms that create those stars.
-On the other side, older OCs give us information about stellar processes and how the galactic disk evolves.
+On the other hand, older open clusters give us information about stellar processes and how the galactic disk evolves.
 Some highly disturbed orbits could also provide evidence of recent merge events and accretion traces from outside the galaxy
@@
 Stars that belong to a same OC share relative positions, inherited from their original gas cloud.
-So their distance from the Earth is similar for all of them and, therefore, they have a narrow dispersion in their parallax value.
-They also share similar values of proper motion, both in right ascension and declination. Another property they share is their chemical composition.
+So their distance to the Earth is similar for all of them and, therefore, they have a narrow dispersion in their parallax value.
+They also share similar values of proper motion, both in right ascension and declination.
+Another property they share is their chemical composition.
 Thus, their metallicity must be uniform, since these stars were born from the same gas cloud and at the same time stage.
@@
 
-We will avoid this issue by looking at $E(b-r) / G_{mag}$ diagrams that reflect their connection with the isochrone curves, typical of their
-temporal evolution \cite{bressan2012parsec}. The isochrones are derived from theoretical models. These models are mainly based on metallicity
-and mass/brightness ratio of stars. They also have a direct correspondence with that presented by the Hertzsprung–Russell (H-R) diagram of the
-stars belonging to the cluster. Examples of these curves are shown in Figure \ref{fig:examples_of_isochrones}.
+We will avoid this issue by looking at $E(b-r) / G_{mag}$ diagrams that reflect their connection with the isochrone curves,
+typical of their temporal evolution \cite{bressan2012parsec}.
+The isochrones are derived from theoretical models.
+These models are mainly based on metallicity and mass/brightness ratio of stars.
+They also have a direct correspondence with that presented by the Hertzsprung–Russell (H-R) diagram of the stars belonging to the cluster.
+Examples of these curves are shown in Figure \ref{fig:examples_of_isochrones}.
 
@@
 
-We will take advantage of these properties only to validate our characterization of OCs, but not to determine what stars belong to them. As we will explain later,
-we will only use dynamic properties such as proper motions and parallax to characterize open clusters.
-
-Although the members of a cluster appear in the observation visual field as an overdensity in positions, as shown in Figure \ref{fig:pos_ngc_2682}, these coordinates
-are not useful to separate those stars that belong to the cluster from the other which do not. However, if we look for overdensities in the proper motions configuration space,
-it is possible, at least at first instance, to assume a possible membership delimitation. See Figure \ref{fig:pm_ngc_2682}.
+We will take advantage of these properties only to validate our characterization of OCs,
+but not to determine what stars belong to them. As we will explain later,
+we will only use dynamic properties such as proper motion and parallax to characterize open clusters.
+
+Although the members of a cluster appear in the observation visual field as an overdensity in positions,
+as shown in Figure \ref{fig:pos_ngc_2682},
+these coordinates are not useful to separate those stars that belong to the cluster from the other which do not.
+However, if we look for overdensities in the proper motions configuration space, it is possible, at least at first instance,
+to assume a possible membership delimitation (see Figure \ref{fig:pm_ngc_2682}.
 
@@
 
-Overdensities in proper motion are not always so evident. Additionally, we will consider the parallax distribution to improve our cluster characterization.
+Overdensities in proper motion are not always so evident.
+Additionally, we will consider the parallax distribution to improve our cluster characterization.
 
@@
 
-In many other cases things are not so simple, and the characterization process can turn very difficult and tedious using tools such as the available ones at
-the Virtual Observatory (VO), and frequently it requires parametrization based on prior knowledge about the studied region. The aim of this work is to test
-an unsupervised and non-parametrized model based on Machine Learning (ML) tools.
+In many other cases things are not so simple,
+and the characterization process can turn very difficult and tedious using tools such as the available ones at the Virtual Observatory (VO),
+and frequently it requires parametrization based on prior knowledge about the studied region.
+The aim of this work is to test an unsupervised and non-parametrized model based on Machine Learning (ML) tools.
+
+We start enumerating the aims we try to achieve with this work.
+Then, we expose the current \emph{state of the art}, mentioning the \emph{Gaia Mission} used as data source in this work and
+describing other related works that try to solve this problem.
+As a last point of the state of the art, we make a brief summary of some of the today available clustering algorithms.
+After the state of the art, we explain our method to improve open cluster characterization.
+Next, we show an alternative method based on current procedures. We have used this alternative method as a validation method to test ours.
+Then, a comparison between a set of results obtained with our method and the results obtained for the same clusters with the validation method is presented.
+Finally, we extract some conclusions based on the obtained results and present further researches that can be done in light of the results.
 
@@
 
-Until now, we have focus our study in Melotte 22 OC. Since our main aim is to get a model that fits well for a wide range of clusters,
+Until now, we have focused our study in Melotte 22 OC. Since our main aim is to get a model that fits well for a wide range of clusters,
 we present now the results obtained for a selection of clusters with different topologies.

### 2020-12-09 thesis/thesis.tex
thesis: Add NGC 2632 results

@@
 This step can be repeated over and over again, varying radii and other parameters, until finally reach a valid probability that works.
-The last step presents a table with candidates objects and allows to export this dataset to TOPCAT, Aladin and VOSA to perform a deeper
-filtering and analysis based on the results given by Clusterix.
+The last step presents a table with candidates objects and allows to export this dataset to TOPCAT,
+Aladin and VOSA to perform a deeper filtering and analysis based on the results given by Clusterix.
 
@@
 This first result can be taken as a good staring point, but we still need to make some fine tunning.
-As we can see in Figure \ref{fig:topcat_prob_selection_melotte_22}, the selection is too restrictive and we have lost too much stars.
+As we can see in Figure \ref{fig:topcat_prob_selection_melotte_22}, the selection is too restrictive and we have lost too many stars.
 
@@
 
-For the sake of completeness, Table \ref{tab:results_melotte_22} is shown again, in order to compare the results obtained with the validation method and our method.
+For the sake of completeness, Table \ref{tab:results_melotte_22} is shown again,
+in order to compare the results obtained with the validation method and our method.
 
@@
 
-Until now, we have focused our study in Melotte 22 OC. Since our main aim is to get a model that fits well for a wide range of clusters,
-we present now the results obtained for a selection of clusters with different topologies.
+Until now, we have focused our study in the open cluster Melotte 22.
+Since our aim is to get a model that fits well for a wide range of clusters,
+we present now some results obtained for a selection of clusters with different topologies.
+
+\pagebreak
+
+\subsection{NGC 2632}
+
+\begin{figure}[H]
+  \centering
+  \begin{subfigure}{0.9\textwidth}
+    \centering
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2632/pm_ngc_2632.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2632/parallax_ngc_2632.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2632/hr_diagram_ngc_2632.png}
+    \end{subfigure}
+  \end{subfigure}
+  \caption{NGC 2632 Clusterix+TOPCAT characterization.}
+  \centering
+  \begin{subfigure}{0.9\textwidth}
+    \centering
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2632/kmeans_pm_ngc_2632.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2632/kmeans_parallax_ngc_2632.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2632/kmeans_hr_diagram_ngc_2632.png}
+    \end{subfigure}
+  \end{subfigure}
+  \caption{NGC 2632 KMeans characterization.}
+  \centering
+  \begin{subfigure}{0.9\textwidth}
+    \centering
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_pm_ngc_2632.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_parallax_ngc_2632.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_hr_diagram_ngc_2632.png}
+    \end{subfigure}
+  \end{subfigure}
+  \caption{NGC 2632 DEC characterization.}
+  \centering
+  \begin{subfigure}{0.9\textwidth}
+    \centering
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_pm_filtered_ngc_2632.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_parallax_filtered_ngc_2632.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_hr_diagram_filtered_ngc_2632.png}
+    \end{subfigure}
+  \end{subfigure}
+  \caption{NGC 2632 DEC (filtered) characterization.}
+\end{figure}
+
+\begin{table}[H]
+  \begin{center}
+    \begin{tabular}{l|c}
+      \textbf{Hyperparameter} & \textbf{Value} \\
+      \hline
+      \# Clusters & 10 \\
+      Clustering Layer & $\left[ 50, 50, 40 \right]$ \\
+      Kernel Initializer Seed & 10 \\
+      Quantil Threshold & 0.1 \\
+    \end{tabular}
+    \caption{NGC 2632 DEC model hyperparameters.}
+    \label{tab:hyperparameters_ngc_2632}
+  \end{center}
+\end{table}
+
+\begin{table}[H]
+  \begin{center}
+    \begin{tabular}{l|c|c|c|c}
+      \textbf{Source / method} & \textbf{$\mu_{\alpha}$ $(mas/yr^{-1})$} & \textbf{$\mu_{\delta}$ $(mas/yr^{-1})$} & \textbf{$\varpi$ $(mas)$} & \emph{\# stars} \\
+      \hline
+      \textbf{Simbad.u-strasbg.fr} & -36.047 $\pm$ 0.110 & -12.917 $\pm$ 0.066 & 5.371 $\pm$ 0.003 & - \\
+      Clusterix+TOPCAT & -36.154 $\pm$ 1.001 & -12.909 $\pm$ 0.806 & 5.327 $\pm$ 0.187 & 371 \\
+      KMeans & -26.352 $\pm$ 20.562 & -15.828 $\pm$ 19.152 & 5.394 $\pm$ 0.696 & 629 \\
+      DEC & -20.012 $\pm$ 20.657 & -14.742 $\pm$ 17.330 & 4.686 $\pm$ 0.855 & 894 \\
+      \textbf{DEC (filtered)} & -21.571 $\pm$ 19.825 & -14.234 $\pm$ 16.388 & 4.719 $\pm$ 0.706 & 714 \\
+    \end{tabular}
+    \caption{NGC 2632 results.}
+    \label{tab:results_ngc_2632}
+  \end{center}
+\end{table}
 

### 2020-12-09 thesis/thesis.tex
thesis: Add NGC 2516 results

@@
 The best possible value is 1, and the worst is -1. Negative values mean that some samples have been assigned to the wrong cluster.
-
-This process is explained in more detail in Appendix \ref{chap:feature_selection}.
 
@@
 
-% TODO: Fix table if figures changes
 \begin{table}[h!]
@@
 
-\pagebreak
-
-\subsection{NGC 2632}
+\newpage
+
+\subsection{NGC 2516}
 
@@
       \centering
-      \includegraphics[width=\textwidth]{../figures/ngc_2632/pm_ngc_2632.png}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/ngc_2632/parallax_ngc_2632.png}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/ngc_2632/hr_diagram_ngc_2632.png}
-    \end{subfigure}
-  \end{subfigure}
-  \caption{NGC 2632 Clusterix+TOPCAT characterization.}
-  \centering
-  \begin{subfigure}{0.9\textwidth}
-    \centering
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/ngc_2632/kmeans_pm_ngc_2632.png}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/ngc_2632/kmeans_parallax_ngc_2632.png}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/ngc_2632/kmeans_hr_diagram_ngc_2632.png}
-    \end{subfigure}
-  \end{subfigure}
-  \caption{NGC 2632 KMeans characterization.}
-  \centering
-  \begin{subfigure}{0.9\textwidth}
-    \centering
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_pm_ngc_2632.png}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_parallax_ngc_2632.png}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_hr_diagram_ngc_2632.png}
-    \end{subfigure}
-  \end{subfigure}
-  \caption{NGC 2632 DEC characterization.}
-  \centering
-  \begin{subfigure}{0.9\textwidth}
-    \centering
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_pm_filtered_ngc_2632.png}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_parallax_filtered_ngc_2632.png}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_hr_diagram_filtered_ngc_2632.png}
-    \end{subfigure}
-  \end{subfigure}
-  \caption{NGC 2632 DEC (filtered) characterization.}
-\end{figure}
-
-\begin{table}[H]
+      \includegraphics[width=\textwidth]{../figures/ngc_2516/pm_ngc_2516.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2516/parallax_ngc_2516.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2516/hr_diagram_ngc_2516.png}
+    \end{subfigure}
+  \end{subfigure}
+  \caption{NGC 2516 Clusterix+TOPCAT characterization.}
+  \centering
+  \begin{subfigure}{0.9\textwidth}
+    \centering
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2516/kmeans_pm_ngc_2516.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2516/kmeans_parallax_ngc_2516.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2516/kmeans_hr_diagram_ngc_2516.png}
+    \end{subfigure}
+  \end{subfigure}
+  \caption{NGC 2516 KMeans characterization.}
+  \centering
+  \begin{subfigure}{0.9\textwidth}
+    \centering
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2516/dec_pm_ngc_2516.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2516/dec_parallax_ngc_2516.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2516/dec_hr_diagram_ngc_2516.png}
+    \end{subfigure}
+  \end{subfigure}
+  \caption{NGC 2516 DEC characterization.}
+  \centering
+  \begin{subfigure}{0.9\textwidth}
+    \centering
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2516/dec_pm_filtered_ngc_2516.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2516/dec_parallax_filtered_ngc_2516.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2516/dec_hr_diagram_filtered_ngc_2516.png}
+    \end{subfigure}
+  \end{subfigure}
+  \caption{NGC 2516 DEC (filtered) characterization.}
+\end{figure}
+
+\newpage
+
+\vfill
+
+\begin{table}[h!]
   \begin{center}
@@
       \hline
-      \# Clusters & 10 \\
+      Number of Clusters & 8 \\
+      Clustering Layer & $\left[ 50, 50, 60 \right]$ \\
+      Kernel Initializer Seed & 2 \\
+      Quantil Threshold & 0.15 \\
+    \end{tabular}
+    \caption{NGC 2516 DEC model hyperparameters.}
+    \label{tab:hyperparameters_ngc_2516}
+  \end{center}
+\end{table}
+
+\begin{table}[h!]
+  \begin{center}
+    \begin{tabular}{l|c|c|c|c}
+      \textbf{Source / method} & \textbf{$\mu_{\alpha}$ $(mas/yr^{-1})$} & \textbf{$\mu_{\delta}$ $(mas/yr^{-1})$} & \textbf{$\varpi$ $(mas)$} & \emph{\# stars} \\
+      \hline
+      \textbf{Simbad.u-strasbg.fr} & -4.6579 $\pm$ 0.0075 & 11.1517 $\pm$ 0.0075 & 2.4118 $\pm$ 0.0006 & 1727 \\
+      Clusterix+TOPCAT & -4.652 $\pm$ 0.523 & 11.203 $\pm$ 0.454 & 2.409 $\pm$ 0.127 & 638 \\
+      KMeans & -4.344 $\pm$ 5.670 & 9.507 $\pm$ 7.546 & 2.268 $\pm$ 0.288 & 1542 \\
+      DEC & -4.426 $\pm$ 6.529 & 9.952 $\pm$ 8.018 & 2.436 $\pm$ 0.358 & 1532 \\
+      \textbf{DEC (filtered)} & -4.502 $\pm$ 4.46 & 10.114 $\pm$ 5.719 & 2.392 $\pm$ 0.135 & 1072 \\
+    \end{tabular}
+    \caption{NGC 2516 results.}
+    \label{tab:results_ngc_2516}
+  \end{center}
+\end{table}
+
+\vfill
+
+\newpage
+
+\subsection{NGC 2632}
+
+\begin{figure}[H]
+  \vfil
+  \centering
+  \begin{subfigure}{0.9\textwidth}
+    \centering
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2632/pm_ngc_2632.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2632/parallax_ngc_2632.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2632/hr_diagram_ngc_2632.png}
+    \end{subfigure}
+  \end{subfigure}
+  \caption{NGC 2632 Clusterix+TOPCAT characterization.}
+  \centering
+  \begin{subfigure}{0.9\textwidth}
+    \centering
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2632/kmeans_pm_ngc_2632.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2632/kmeans_parallax_ngc_2632.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2632/kmeans_hr_diagram_ngc_2632.png}
+    \end{subfigure}
+  \end{subfigure}
+  \caption{NGC 2632 KMeans characterization.}
+  \centering
+  \begin{subfigure}{0.9\textwidth}
+    \centering
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_pm_ngc_2632.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_parallax_ngc_2632.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_hr_diagram_ngc_2632.png}
+    \end{subfigure}
+  \end{subfigure}
+  \caption{NGC 2632 DEC characterization.}
+  \centering
+  \begin{subfigure}{0.9\textwidth}
+    \centering
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_pm_filtered_ngc_2632.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_parallax_filtered_ngc_2632.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_hr_diagram_filtered_ngc_2632.png}
+    \end{subfigure}
+  \end{subfigure}
+  \caption{NGC 2632 DEC (filtered) characterization.}
+\end{figure}
+
+\newpage
+
+\vfill
+
+\begin{table}[h!]
+  \begin{center}
+    \begin{tabular}{l|c}
+      \textbf{Hyperparameter} & \textbf{Value} \\
+      \hline
+      Number of Clusters & 10 \\
       Clustering Layer & $\left[ 50, 50, 40 \right]$ \\
@@
 
-\begin{table}[H]
+\begin{table}[h!]
   \begin{center}
@@
 
+\vfill
+
+\newpage
+
 % TODO: Exposición de resultados y comentarios

### 2020-12-09 thesis/thesis.tex
thesis: Add NGC 2682 results

@@
 
+\subsection{NGC 2682}
+
+\begin{figure}[H]
+  \vfil
+  \centering
+  \begin{subfigure}{0.9\textwidth}
+    \centering
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2682/pm_ngc_2682.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2682/parallax_ngc_2682.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2682/hr_diagram_ngc_2682.png}
+    \end{subfigure}
+  \end{subfigure}
+  \caption{NGC 2682 Clusterix+TOPCAT characterization.}
+  \centering
+  \begin{subfigure}{0.9\textwidth}
+    \centering
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2682/kmeans_pm_ngc_2682.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2682/kmeans_parallax_ngc_2682.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2682/kmeans_hr_diagram_ngc_2682.png}
+    \end{subfigure}
+  \end{subfigure}
+  \caption{NGC 2682 KMeans characterization.}
+  \centering
+  \begin{subfigure}{0.9\textwidth}
+    \centering
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2682/dec_pm_ngc_2682.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2682/dec_parallax_ngc_2682.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2682/dec_hr_diagram_ngc_2682.png}
+    \end{subfigure}
+  \end{subfigure}
+  \caption{NGC 2682 DEC characterization.}
+  \centering
+  \begin{subfigure}{0.9\textwidth}
+    \centering
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2682/dec_pm_filtered_ngc_2682.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2682/dec_parallax_filtered_ngc_2682.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2682/dec_hr_diagram_filtered_ngc_2682.png}
+    \end{subfigure}
+  \end{subfigure}
+  \caption{NGC 2682 DEC (filtered) characterization.}
+\end{figure}
+
+\newpage
+
+\vfill
+
+\begin{table}[h!]
+  \begin{center}
+    \begin{tabular}{l|c}
+      \textbf{Hyperparameter} & \textbf{Value} \\
+      \hline
+      Number of Clusters & 10 \\
+      Clustering Layer & $\left[ 50, 50, 40 \right]$ \\
+      Kernel Initializer Seed & 0 \\
+      Quantil Threshold & 0.1 \\
+    \end{tabular}
+    \caption{NGC 2682 DEC model hyperparameters.}
+    \label{tab:hyperparameters_ngc_2682}
+  \end{center}
+\end{table}
+
+\begin{table}[h!]
+  \begin{center}
+    \begin{tabular}{l|c|c|c|c}
+      \textbf{Source / method} & \textbf{$\mu_{\alpha}$ $(mas/yr^{-1})$} & \textbf{$\mu_{\delta}$ $(mas/yr^{-1})$} & \textbf{$\varpi$ $(mas)$} & \emph{\# stars} \\
+      \hline
+      \textbf{Simbad.u-strasbg.fr} & -10.9737 $\pm$ 0.0064 & -2.9396 $\pm$ 0.0063 & 1.1325 $\pm$ 0.0011 & 1194 \\
+      Clusterix+TOPCAT & -10.970 $\pm$ 0.322 & -2.958 $\pm$ 0.327 & 1.142 $\pm$ 0.080 & 649 \\
+      KMeans & -8.616 $\pm$ 5.591 & -3.710 $\pm$ 6.067 & 1.196 $\pm$ 0.226 & 1374 \\
+      DEC & -8.926 $\pm$ 5.278 & -3.550 $\pm$ 5.425 & 1.144 $\pm$ 0.159 & 1238 \\
+      \textbf{DEC (filtered)} & -9.619 $\pm$ 4.112 & -3.317 $\pm$ 3.976 & 1.140 $\pm$ 0.089 & 990 \\
+    \end{tabular}
+    \caption{NGC 2682 results.}
+    \label{tab:results_ngc_2682}
+  \end{center}
+\end{table}
+
+\vfill
+
+\newpage
+
 % TODO: Exposición de resultados y comentarios

### 2020-12-10 thesis/thesis.tex
thesis: Update State of the Art

@@
 
-% TODO: Hablar de KMedias y DEC
-
 An initial approach for finding OCs is the search of overdensities along the galactic disk. In general, this is the starting point but,
@@
 The result of Clusterix + TOPCAT, applied to the same clusters, will be used only for comparing and validating our results.
+
+\section{Clustering Algorithms}
+
+As we will explain later, what we need to accomplish our aim is an algorithm that manages to make data groups based on the dynamic properties of the stars.
+
+There are several clustering algorithms: KMeans, Mean-Shift Clustering, DBSCAN \cite{ester1996density} among other.
+While each one behaves better according ot the distribution of the objects to clusterize, we have choosen KMeans by its simplicity and good results.
+
+KMeans requires a single parameter, the number of clusters to build (N).
+The algorithm looks for N center points which are vectors of the same dimension as the number of selected features.
+KMeans starts by making an initial groups configuration and then, it reassigns objects to other groups iteratively by minimizing
+the distance among points inside the new group and maximizing the distance with the centers of the other groups.
+
+The algorithm stops once a maximum number of iterations is reached, or when changes between iterations is lower than a minimum.
+
+This algorithm works well and gives good results at first approach.
+However, we need to set a large number of clusters to find the open cluster we are looking for.
+This fact complicates the identification of the OC and many times the OC contains to many outliers.
+For that reason, we have search for a KMeans refinement based on an Artificial Neural Network (ANN).
+
+The \emph{Unsupervised Deep Embedding for Clustering Analysis} model or \emph{DEC} \cite{xie2016unsupervised} takes KMeans as its starting point,
+but then, it trains an autoencoder to reduce the feature space and pass this transformed data through a Clustering Layer which refines the previous selection.
+
+The mechanism of this model is explained in Section \ref{sec:deep_embedding_clustering} and its implementation is available in \verb|cdalvaro.ml.dec| package.
 

### 2020-12-10 thesis/thesis.tex
thesis: Add Melotte 25 results

@@
 
-% TODO: Exposición de resultados y comentarios
-% Es una exposición objetiva, sin valorar los resultados ni justificarlos.
+\subsection{Melotte 25}
+
+\begin{figure}[H]
+  \vfil
+  \centering
+  \begin{subfigure}{0.9\textwidth}
+    \centering
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/melotte_25/pm_melotte_25.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/melotte_25/parallax_melotte_25.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/melotte_25/hr_diagram_melotte_25.png}
+    \end{subfigure}
+  \end{subfigure}
+  \caption{Melotte 25 Clusterix+TOPCAT characterization.}
+  \centering
+  \begin{subfigure}{0.9\textwidth}
+    \centering
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/melotte_25/kmeans_pm_melotte_25.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/melotte_25/kmeans_parallax_melotte_25.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/melotte_25/kmeans_hr_diagram_melotte_25.png}
+    \end{subfigure}
+  \end{subfigure}
+  \caption{Melotte 25 KMeans characterization.}
+  \centering
+  \begin{subfigure}{0.9\textwidth}
+    \centering
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/melotte_25/dec_pm_melotte_25.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/melotte_25/dec_parallax_melotte_25.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/melotte_25/dec_hr_diagram_melotte_25.png}
+    \end{subfigure}
+  \end{subfigure}
+  \caption{Melotte 25 DEC characterization.}
+  \centering
+  \begin{subfigure}{0.9\textwidth}
+    \centering
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/melotte_25/dec_pm_filtered_melotte_25.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/melotte_25/dec_parallax_filtered_melotte_25.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/melotte_25/dec_hr_diagram_filtered_melotte_25.png}
+    \end{subfigure}
+  \end{subfigure}
+  \caption{Melotte 25 DEC (filtered) characterization.}
+\end{figure}
+
+\newpage
+
+\vfill
+
+\begin{table}[h!]
+  \begin{center}
+    \begin{tabular}{l|c}
+      \textbf{Hyperparameter} & \textbf{Value} \\
+      \hline
+      Number of Clusters & 11 \\
+      Clustering Layer & $\left[ 50, 50, 80 \right]$ \\
+      Kernel Initializer Seed & 11 \\
+      Quantil Threshold & 0.1 \\
+    \end{tabular}
+    \caption{Melotte 25 DEC model hyperparameters.}
+    \label{tab:hyperparameters_melotte_25}
+  \end{center}
+\end{table}
+
+\begin{table}[h!]
+  \begin{center}
+    \begin{tabular}{l|c|c|c|c}
+      \textbf{Source / method} & \textbf{$\mu_{\alpha}$ $(mas/yr^{-1})$} & \textbf{$\mu_{\delta}$ $(mas/yr^{-1})$} & \textbf{$\varpi$ $(mas)$} & \emph{\# stars} \\
+      \hline
+      \textbf{Simbad.u-strasbg.fr} & 104.92 $\pm$ 0.12 & -28.00 $\pm$ 0.09 & 21.052 $\pm$ 0.065 & - \\
+      Clusterix+TOPCAT & 106.796 $\pm$ 6.229 & -24.870 $\pm$ 5.417 & 21.210 $\pm$ 1.115 & 109 \\
+      KMeans & 81.410 $\pm$ 71.475 & -45.309 $\pm$ 73.945 & 21.056 $\pm$ 6.018 & 365 \\
+      DEC & x $\pm$ x & x $\pm$ x & x $\pm$ x & x \\
+      \textbf{DEC (filtered)} & x $\pm$ x & x $\pm$ x & x $\pm$ x & x \\
+    \end{tabular}
+    \caption{Melotte 25 results.}
+    \label{tab:results_melotte_25}
+  \end{center}
+\end{table}
+
+\vfill
+
+\newpage
 

### 2020-12-10 thesis/thesis.tex

@@
 \section{Deep Embedded Clustering (DEC)}
+\label{sec:deep_embedding_clustering}
 
@@
 
-For that reason we have adapted the \emph{Unsupervised Deep Embedding for Clustering Analysis} or \emph{DEC} model \cite{xie2016unsupervised} to our work.
+For that reason we have adapted the \emph{Unsupervised Deep Embedding for Clustering Analysis} to our work.
 
@@
 
+All results have been computed with an Apple Mac Pro Late 2013, 2.7 GHz 12-Core Intel Xeon E5 with two AMD FirePro D700 6GB and 64 GB
+of memory, running macOS BigSur 11.0.1, Keras 2.2 and PlaidML 0.7.
+
+The majority of the runnings have taken a few minutes, while big clusters such as Melotte 25, have taken around an hour to perform
+the computations. This contrasts with \citeA{castro2020hunting} work, that used the power of the MareNostrum 42 supercomputer in order
+to perform its calculations.
+
 \newpage
@@
 
-% Se aporta la discusión de los resultados presentados en el capítulo anterior. En este capítulo se puede discutir la relevancia de
-% los resultados, presentar posibles explicaciones para los datos anómalos y resaltar aquellos datos que sean particularmente relevantes
-% para el análisis del experimento.
+We have applied the our proposed method to clusters with different topologies:
+
+\begin{itemize}
+  \item NGC 2516 has it proper motion center deviated from the origin but it is embedded inside a big cloud of stars with similar proper motions.
+  \item NGC 2632 (in addition to Melotte 22) ia s cluster whose proper motion center is not located at the origin and has a well separated parallax center.
+  \item NGC 2682, on the other hand, has its parallax centered within the region's gaussian,
+        which complicates its detection although its proper motion center is deviated from the origin.
+  \item Melotte 25 (as well as Melotte 22) is a cluster closer to us than the other.
+        That makes its memebership stars to be more scattered than previous clusters which are more compact.
+\end{itemize}
+
+All of these clusters have a wide variety of diameters, from 25 to 330 arcmin, as well as the number of stars that belong to them.
+NGC 2682 has 3,000 stars while Melotte 25 is composed by more than 400,000 stars.
+
+% TODO: Meter una tabla con los parámetros de los clusters mencionados
+
+% TODO: Determinar correctamente los errores en las medidas y mencionarlos
+
+In all cases, the model has resolved properly the identity of the cluster and has characterized the membership stars,
+giving results compatible with the ones obtained with classic procedures and VO tools.
+
+Compared with the mentioned tools, our model can be fully calified as a non-parameterized model since it does not depend on parameters referred
+to the clusters itself but only hyperparameters such as the initial number of clusters to find, or the structure of the Clustering Layer used by DEC model.
+
+Our model is also non-supervised since we do not need to tell the model which stars belong to the cluster or assit it while is training.
+
+Only at the end of the characterization, a fine tunning driven by the user, can be applied in orther to improve the selection by discarding those
+stars which fall outside the given quantiles.
+
+In any case we make assumptions about the cluster profile, something that we have to take into account when using VO tools.
+This is another evidence that our model is non-parameterized.
 
@@
 
-% Resumen final del trabajo y debe servir para informar del alcance y relevancia de la aportación.
-% Suele estructurarse empezando con un resumen del problema tratado, de cómo se ha abordado y de por qué la solución sería válida.
-
-% Es recomendable que incluya también un resumen de las contribuciones del trabajo, en el que se relacionen las contribuciones y los
-% resultado obtenidos con ls objetivos planteados para el trabajo, discutiendo hasta qué punto has conseguido resolver los objetivos planteados.
+As shown by the obtained results, we can say that in general our model works fine identifying and characterizing open clusters.
+So we have succedded building a model non-parameterized non-supervised for open cluster characterization.
+However, some improvements should be done in order to improve its accuracy and precision.
+
+We have built a non-parameterized model since the user does not provide any information about the cluster profile or its properties.
+Although some tweaks can be done regarding the cluster, like changing the initial number of clusters or modifying the Clustering Layer structure,
+in order to improve the final characterization.
+
+Another key point in our model is the initializer kernel.
+This kernel prepares data before passing it to the ANN and the results may vary significantly depending on this kernel.
+For example, simply varying the kernel seed can make a huge difference on the result.
+
+Even so, comparing our method with Clusterix, we can still say that the number of parameters (or hyperparameters) in our model is smaller
+and we do not need previous knowledge about the studied cluster.
+That makes easier to test different configurations and automate the process.
+
+With all of this points in mind, we can affirm that this model could be really considered as an alternative model for open cluster characterization and thus,
+it could even be incloud as part of the Virtual Observatory toolset.
 
@@
 
-% Debería señalar las perspectivas de futuro que abre el trabajo desarrollado para el campo de estudio definido. En el fondo,
-% debe justificar de qué modo puede emplearse la aportación desarrollada y en qué campos.
+The first aim after finishing this work is testing the model with a wider range of clusters.
+In general we could think of applying it to the whole VizieR catalogue.
+That way we could have a better idea about the current limits of our model and we could make adjustments to improve it.
+We could determine with better accuracy the topologies that our model works well with, and which ones out model does not resolve properly.
+Or even, we could stablish different sets of hyperparameters regarding the topology of the cluster.
+
+With our model, new secondary clusters arise from the data apart of the main ones.
+So, it is also interesting the individual study of these new clusters.
+
+Another possible change in our model would be using DBSCAN instead of KMeans as our initial clustering algorithm.
+Maybe this algorithm gives us a better starting point for the DEC model.
+
+At last, we have used during this work the Gaia DR2 database, but recently the DR3 dataset has been released.
+So it is evident the interest on testing our model with this new data.
 
@@
 \appendix
-\chapter{Feature Selection}
-\label{chap:feature_selection}
 

### 2020-12-30 paper/paper.tex
paper: Add paper template

\documentclass[11pt,a4paper,english,twocolumn]{article}
\usepackage{unir_paper}

%---------------------------
%título del trabajo y autor
%---------------------------
\title{Título}
\author{Nombre y Apellidos del estudiante}
\date{fecha}

%---------------------------
%marges
%---------------------------
%\usepackage[margin=1.9cm]{geometry}
%---------------------------
%---------------------------
%---------------------------
%---------------------------

\resumen{Breve resumen del trabajo realizado (extensión máxima: 150 palabras). Este resumen debe incluir el objetivo o propósito de la investigación, la metodología, los resultados y las conclusiones.}
\palabrasclave{tres a cinco palabras clave por orden alfabético y separadas por comas}

\begin{document}
\twocolumn[
\begin{@twocolumnfalse}
\maketitle
\end{@twocolumnfalse}
]

%\renewcommand{\listfigurename}{Índice de Ilustraciones}
%\renewcommand{\listtablename}{Índice de Tablas}
%\renewcommand{\contentsname}{Índice de Contenidos}
%\renewcommand{\figurename}{Figura}
%\renewcommand{\tablename}{Tabla}
%\twocolumn

%\frontmatter
%\tableofcontents
%\listoffigures
%\listoftables

\section{INTRODUCCIÓN}
Introducción en la que debes resumir de forma esquemática pero suficientemente clara lo esencial de cada una de las partes del trabajo.
La lectura de esta introducción ha de dar una primera idea clara de lo que se pretendía, las conclusiones a las que se ha llegado y del procedimiento seguido.

\section{ESTADO DEL ARTE}
Estudio a fondo el dominio de aplicación, citando numerosas referencias.
Debe aportar un buen resumen del conocimiento que ya existe en el campo de los problemas habituales identificados.
Numerar las citas de forma consecutiva entre corchetes \cite{Chen} y \cite{young}.
\section{OBJETIVOS Y METODOLOGÍA}
Objetivo general, objetivos específicos y metodología de trabajo aplicada.

\section{CONTRIBUCIÓN}
Desarrollar la descripción de tu contribución.

\section{RESULTADOS O EVALUACIÓN }
Descripción de los resultados (Tipo 1. Piloto Experimental o Tipo 4. Comparativa de soluciones) o descripción de la evaluación y los resultados obtenidos (Tipo 2. Desarrollo de Software o Tipo 3. Metodología).
\subsection{Evaluación 1}

\subsection{Evaluación 2 }
En la Tabla \ref{tab_1}
\begin{table}\label{tab_1}
\caption{Unidades de las propiedades magnéticas:}

\begin{tabular}{ccc}\hline\hline
Símbolo & Cantidad & Conversión\\
\hline
$\Phi$ & flujo magnético & $1$Mx $\rightarrow 10^{-8}V\cdot s$\\
... &...&...\\
\hline\hline
\end{tabular}
\end{table}

\section{DISCUSIÓN O ANÁLISIS DE RESULTADOS}
Tras la presentación objetiva de los resultados, querrás aportar una discusión de los mismos.

\section{CONCLUSIONES}
Resumen de las contribuciones del trabajo, en el que relaciones las contribuciones y los resultados obtenidos con los objetivos que habías planteado para el trabajo, discutiendo hasta qué punto has conseguido resolver los objetivos planteados.
Finalmente, hablar de líneas de trabajo futuro que podrían aportar valor añadido al TFM realizado. La sección debería señalar las perspectivas de futuro que abre el trabajo desarrollado para el campo de estudio definido. En el fondo, debes justificar de qué modo puede emplearse la aportación que has desarrollado y en qué campos.

\appendix
\section{APÉNDICES}
Apéndices, en caso de ser necesario.

\bibliographystyle{unsrt}
\bibliography{references}

\end{document}

### 2020-12-31 thesis/thesis.tex
thesis: Apply supervisor suggestions
Co-Authored-By: Cesar Guzmann <15159632+cguz@users.noreply.github.com>

@@
 %---------------------------
-\title{Using Machine Learning Tools for Open Cluster Characterization with Gaia DR2 Data}
+\title{Machine Learning Tools for Open Cluster Characterization with Gaia DR2 Data}
 \author{Carlos David Álvaro Yunta}
-\date{12th of November, 2020}
-\director{César Augusto Guzmán Álvarez}
+\date{31st of December, 2020}
+\director{Dr. César Augusto Guzmán Álvarez}
 \nombreciudad{Madrid, Spain}
@@
 
-The characterization and knowledge of \emph{Open Clusters} allows us to better understand properties and mechanisms about the Universe
-such as stellar formation and the regions where those events occur. They also provide information about stellar processes and the
-evolution of the galactic disk.
-
-In this work, we develope a method to characterize those clusters by using \emph{artificial intelligence} tools like clustering
-by \emph{K-Means}, or clustering based on \emph{Artificial Neural Networks} by implementing the \emph{Deep Embedded Clustering}
-model. We are using \emph{Gaia DR2 database} as data source for testing our models.
-
-The developed method aims to improve the existing ones both in terms of \emph{computational efficiency}, with lower computational requirements,
-and \emph{ease of use}, by reducing the number of hyperparameters to configure in order to obtain a good characterization of the analyzed clusters.
-
-As detailed below, our method achieves good results, becoming even better in some cases when the results are compared with other current methods.
+The characterization and knowledge of \emph{Open Clusters} allow us to better understand properties and mechanisms about the Universe
+such as stellar formation and the regions where those events occur.
+They also provide information about stellar processes and the evolution of the galactic disk.
+
+In this work, we develop a method to characterize open clusters by using \emph{artificial intelligence} tools,
+such as clustering by \emph{K-Means} and clustering based on \emph{Artificial Neural Networks} by implementing the \emph{Deep Embedded Clustering} model.
+We are using \emph{Gaia DR2 database} as data source for testing our models.
+
+The developed method aims to improve current techniques of the state of the art.
+We achieve improvements not only in terms of \emph{computational efficiency}, with lower computational requirements, but in \emph{usability},
+reducing the number of hyperparameters to obtain a good characterization of the analyzed clusters.
+
+Our method achieves good results, becoming even better in some cases when the results are compared with current techniques.
 
@@
 
-{\bf Key words.} open clusters characterization --- machine learning --- gaia dr2 --- data analysis --- deep embedded clustering
+{\bf Key words.} open clusters characterization, machine learning, gaia, data analysis, deep embedded clustering
 
@@
 
-{\bf Palabras Clave:} caracterización de cúmulos abiertos --- inteligencia artificial --- gaia dr2 --- análisis de datos
+{\bf Palabras Clave:} caracterización de cúmulos abiertos, inteligencia artificial, gaia, análisis de datos
 
@@
 Stellar open clusters (OCs) \cite{janes1982open} are groups of stars gravitationally bound originated from a single molecular gas cloud.
-Thus they share the same chemical composition and age, and they have similar relative positions and proper motion.
-These astronomical objects are of fundamental importance to understand the spiral structure,
-the dynamics and the chemical evolution of our galaxy.
+They share the same chemical composition and age. Moreover, they have similar relative positions and proper motion.
+Those astronomical objects are relevant to understand the spiral structure, the dynamics and the chemical evolution of our galaxy.
 
@@
 in their original molecular cloud \cite{clarke2000theformationofstellarclusters} \cite{portegies2010young}.
-The evolution of these systems tends to sparse them in a few million years by interacting gravitationally with other systems.
+The evolution of those systems tends to sparse them in a few million years by interacting gravitationally with other systems.
 Galactic tidal forces and mechanisms that involves gas loss driven by stellar feedback are other causes of disruption
 \cite{brinkmann2017bound}.
-Nevertheless, a small fraction of these systems will survive in the initial state and persist bound in bigger timescales.
+Nevertheless, a small fraction of those systems will survive in the initial state and persist bound in bigger timescales.
 
@@
 
-Stars that belong to a same OC share relative positions, inherited from their original gas cloud.
-So their distance to the Earth is similar for all of them and, therefore, they have a narrow dispersion in their parallax value.
+Stars that belong to the same OC share relative positions, inherited from their original gas cloud.
+Which means that their distances to the Earth are similar for all of them and, therefore, they have a narrow dispersion in their parallax value.
 They also share similar values of proper motion, both in right ascension and declination.
@@
 The main aim of this thesis is to \emph{build an unsupervised clustering model for open cluster characterization}.
-
-We want this model to be \emph{non-supervised non-parameterized} in order to fit a wide range of clusters without
-the need of fine tunning hyperparameters regarding the topology or size of the studied clusters.
+The model will be \emph{non-supervised and non-parameterized} to fit a wide range of clusters without
+the need for fine-tuning a high number of hyperparameters.
 
@@
 
-\section{Comparison}
+\section{Comparing and Validating}
 

### 2021-01-02 thesis/thesis.tex
thesis: Move list of figures and tables
Co-Authored-By: Cesar Guzmann <15159632+cguz@users.noreply.github.com>
thesis: Add some corrections

@@
 
+\listoffigures
+\listoftables
+
 \chapter{Abstract}
@@
 The characterization and knowledge of \emph{Open Clusters} allow us to better understand properties and mechanisms about the Universe
-such as stellar formation and the regions where those events occur.
+such as stellar formation and the regions where these events occur.
 They also provide information about stellar processes and the evolution of the galactic disk.
@@
 
-\chapter{Acknowledgement}
+\chapter{Acknowledgements}
 
@@
 
-Stellar open clusters (OCs) \cite{janes1982open} are groups of stars gravitationally bound originated from a single molecular gas cloud.
+Stellar open clusters (OCs) \cite{janes1982open} are groups of stars gravitationally bounded originated from a single molecular gas cloud.
 They share the same chemical composition and age. Moreover, they have similar relative positions and proper motion.
@@
 Although most stars in the Milky Way are presented isolated, it is considered that most of them (or even all)
-are formed in clustered environments and spend a period of time gravitationally bound with their siblings embedded
+are formed in clustered environments and spend a period of time gravitationally bounded with their siblings embedded
 in their original molecular cloud \cite{clarke2000theformationofstellarclusters} \cite{portegies2010young}.
 The evolution of those systems tends to sparse them in a few million years by interacting gravitationally with other systems.
-Galactic tidal forces and mechanisms that involves gas loss driven by stellar feedback are other causes of disruption
+Galactic tidal forces and mechanisms that involve gas loss driven by stellar feedback are other causes of disruption
 \cite{brinkmann2017bound}.
-Nevertheless, a small fraction of those systems will survive in the initial state and persist bound in bigger timescales.
-
-Young OCs allow us to research stars formation regions and improve our understanding about the mechanisms that create those stars.
+Nevertheless, a small fraction of those systems will survive in the initial state and persist bounded in bigger timescales.
+
+Young OCs allow us to research star formation regions and improve our understanding about the mechanisms that create those stars.
 On the other hand, older open clusters give us information about stellar processes and how the galactic disk evolves.
@@
 
-Although the members of a cluster appear in the observation visual field as an overdensity in positions,
+Although the members of a cluster appear in the observational visual field as an overdensity in positions,
 as shown in Figure \ref{fig:pos_ngc_2682},
-these coordinates are not useful to separate those stars that belong to the cluster from the other which do not.
-However, if we look for overdensities in the proper motions configuration space, it is possible, at least at first instance,
+these coordinates are not useful to separate those stars that belong to the cluster from the other that do not.
+However, if we look for overdensities in the proper motion configuration spaces, it is possible, at least at first instance,
 to assume a possible membership delimitation (see Figure \ref{fig:pm_ngc_2682}.
@@
 
-To achieve the general goal, we will follow the following milestones:
+To achieve the general goal, we will set the following milestones:
 
@@
 although it seems simple, it presents a fundamental problem already discussed. The near field around the OC is filled with two types of
-distinct stars populations: those who belong to the OC (tens or hundreds to a few thousand) and a background made up of thousands or
-millions of stars that are not. Finding out which stars belong to the first group is the problem treated in this work.
+distinct star populations: those who belong to the OC (tens or hundreds to a few thousand) and a background made up of thousands or
+millions of stars that are not. Finding out which stars belong to the first group is the problem faced in this work.
 This selection is crucial to properly characterize the fundamental properties of the cluster
@@
 to identify it. With Gaia DR2, which contains great high-quality astrometric and photometric data for a huge number of stars,
-a new opportunity presents itself to simplify and optimize these processes. In this context, new approaches are being developed
+a new opportunity arises to simplify and optimize these processes. In this context, new approaches are being developed
 to improve current detection methods and automate OC characterization.
@@
 \cite{balaguer2020clusterix}, TOPCAT \cite{taylor2005topcat}, and VOSA \cite{bayo2008vosa}, (all of them developed by the Virtual Observatory).
-These tools work together and offer good results, although it is not possible to use them in an unattended process and free of certain
+These tools work together and offer good results, although it is not possible to use them in an unsupervised process and free of certain
 different parametrization in each case.
@@
 
-TOPCAT alone does not offer an independent discrimination mechanisms to researchers. They should try an initial selection, staring from the
-map of proper motions, by delimiting the area where the cluster presumably appears. When this is not possible, it is necessary to have a
-certain knowledge about the cluster profile and to parametrize some approximated values. In both cases, the definition of the selected
-set must then be optimized step by step until reaching a result that can be considered reasonable.
-
-Another tool is Clusterix 2.0, which is an interactive web-based tool \cite{balaguer2020clusterix}. It takes the proper motion diagram
-without assuming any prior profiles assumptions for the stars candidates membership and determines empirically the frequency functions.
+TOPCAT alone does not offer an independent discrimination mechanisms to researchers.
+They should try an initial selection, starting from the map of proper motions,
+by delimiting the area where the cluster presumably appears. When this is not possible, it is necessary to have a
+certain knowledge about the cluster profile and to parametrize some approximated values.
+In both cases, the definition of the selected set must then be optimized step by step until reaching a result that can be considered reasonable.
+
+Another tool is Clusterix 2.0, which is an interactive web-based tool \cite{balaguer2020clusterix}.
+It takes the proper motion diagram without making any prior assumptions about the membership of the candidate star
+and determines empirically the frequency functions.
 
@@
 The non-field population does not occupy the entire workspace, but is spatially concentrated, which makes it possible to distinguish two regions
-in the workspace: the only field region (label 'f'), dominated by stars field, and the cluster + field region (label 'c+f'), which includes field
-stars and not field stars. \cite{balaguer2020clusterix}
+in the workspace: the only field region (label 'f'), dominated by star fields, and the cluster + field region (label 'c+f'),
+which includes both star fields and not star fields. \cite{balaguer2020clusterix}
 \end{displayquote}
@@
 This fact implies defining three areas or regions with different radius. The first 'c+f' corresponds to the one in which the cluster members are
-presumed to be contained together with other field stars that are not part of the cluster. The second, broader, assumes that it only contains
+presumed to be contained together with other star fields that are not part of the cluster. The second, broader, assumes that it only contains
 stars in an extended visual field without components of the cluster, and a third intermediate region, which is out of analysis (void area), which
@@
 
-It should be noted that for the execution of this method, MareNostrum 42 (Bacelona Supercomputing Center) was used. So the neural network could handle
-the image recognition process with isochrone patterns and not applying theoretical models derived from values such as metallicity, masses, among other.
+It should be noted that for the execution of this method, MareNostrum 42 (Barcelona Supercomputing Center) was used.
+So the neural network could handle the image recognition process with isochrone patterns and not applying theoretical models
+derived from values such as metallicity, masses, among other.
 
@@
 
-Our aim in this work, linked to computational limitations compared to MareNostrum, is not to do a blind search for clusters, but to obtain a method,
-in the machine learning environment, that allows the characterization of open clusters through unsupervised and non-parametrized procedures in a
-single step. This method should not require further refinements with other tools.
+Our aim in this work, linked to computational limitations compared to MareNostrum, is not to do a blind search for clusters,
+but to obtain a method, within the machine learning environment,
+that allows the characterization of open clusters through unsupervised and non-parametrized procedures in a single step.
+This method should not require further refinements with other tools.
 
@@
 
-For the sake of simplicity, we are taking \emph{Melotte 22 (Messier 45)} \cite{elsanhoury2019ppmxl} to illustrate the whole method. This cluster is well studied
-and their stars are not too mixed with other region stars, so images look clearly.
+For the sake of simplicity, we are taking \emph{Melotte 22 (Messier 45)} \cite{elsanhoury2019ppmxl} to illustrate the whole method.
+This cluster is well studied and its stars are not too mixed with other stars that do not belong to the OC although they are contained
+in the same observation field, so images look clear.
 
@@
 
-However, although this second cluster is almost isolated, there are stars that do not belong to the OC. Thus, we need more information to properly characterize the OC.
+However, although this second cluster is almost isolated, there are stars that do not belong to the OC.
+Thus, we need more information to properly characterize the OC.
 
@@
 Instead, a wider region (1.5 times the size of the cluster) is downloaded for each cluster to include stars outside the cluster.
-The idea is that the unsupervised model must be able to clusterize the stars that belongs to the OC and to discard
+The idea is that the unsupervised model must be able to clusterize the stars that belong to the OC and to discard
 those outsiders when characterizing the open cluster.
@@
 
-Taking these considerations into account, the downloaded dataset covers nearly 114 million stars (42GB of compressed data). This dataset is significantly
-smaller than the whole Gaia DR2 dataset, which contains information for approximately 1,600 million stars. However, it is still being too large.
+Taking these considerations into account, the downloaded dataset covers nearly 114 million stars (42GB of compressed data).
+This dataset is significantly smaller than the whole Gaia DR2 dataset, which contains information for approximately 1,600 million stars.
+However, it is still too large.
 
@@
 
-So, a cluster must fulfil the following filters to be accepted:
+Therefore, a cluster must fulfil the following filters in order to be accepted:
 
@@
 
-As shown in Figure \ref{fig:OpenClustSelection}, these constrains give us a smaller dataset but it is still being a good representation of all clusters in the Milky Way
+As shown in Figure \ref{fig:OpenClustSelection}, these constrains give us a smaller dataset but it is still a good representation of all clusters in the Milky Way
 since they are equally distributed around the galaxy disk.
@@
 
-PostgreSQL knows that all stars we want to are located inside \verb|gaiadr2.region_302| so it does not need to query any other tables.
-This allows us to download data for as many regions as we want to without loosing efficiency when accessing for that data.
+PostgreSQL knows that all stars we want are located inside \verb|gaiadr2.region_302| so it does not need to query any other table.
+This allows us to download data for as many regions as we want without loosing efficiency when accessing for that data.
 
@@
 
-The downloader takes data from the Gaia DR2 database and save the selection into our custom database without modifying any value,
+The downloader takes data from the Gaia DR2 database and saves the selection into our custom database without modifying any value,
 so data preserves all its properties and units.
@@
 Python methods for retrieving information from the database are available through an instance of \verb|cdalvaro.DB|.
-This class contains methods for recovering stars information for a given region as a Pandas \verb|DataFrame| ready for data analysis.
+This class contains methods for retrieving star's information for a given region as a Pandas \verb|DataFrame|, ready for data analysis.
 
@@
 
-As mentioned before, we want our model to characterize open clusters by looking at their dynamics properties.
-
-Also, we want to maintain as simple as possible our clustering model in order to avoid requiring too much computing resources.
-
-Proper motion in right ascension and declination seems like a natural choice since, as we know, stars belonging to the same OC share a common motion vector.
+As mentioned before, we want our model to characterize open clusters by looking at their dynamic properties.
+
+Also, we want to maintain as simple as possible our clustering model in order to save computing resources.
+
+Proper motion in right ascension and declination seems like a natural choice since, as we know,
+stars belonging to the same OC share a common motion vector.
 
@@
 
-However, we are not going to use these raw features. Instead, we are taken a combination of them.
+However, we are not going to use these raw features. Instead, we are taking a combination of them.
 
@@
   \caption{Melotte 22 feature pairplot}
+  % TODO: Ampliar la descripción de la figura
   \label{fig:features_melotte_22}
@@
 
-To solve this issue, we will try to estimate the best number of clusters by using one more time the \emph{silhouette score}.
+To solve this issue, we will try to estimate the best number of clusters by using the \emph{silhouette score} one more time.
 
@@
 
-K-Means does a good job making an initial clustering, as shown in Figure \ref{fig:kmeans_melotte_22}. However, too many clusters arise from this characterization
-and the open cluster is still polluted with stars that do not belong to it. Even more, we would like to reduce the amount of clusters too.
+K-Means does a good job making an initial clustering, as shown in Figure \ref{fig:kmeans_melotte_22}.
+However, too many clusters arise from this characterization and the open cluster is still polluted with stars that do not belong to it.
+Moreover, we would like to reduce the amount of clusters too.
 
@@
 
-If we take a look into the H-R diagram (Figure \ref{fig:kmeans_hr_diagram_melotte_22}), we can identify the isochrone curve for Melotte 22 cluster.
-This curve has a good shape, but again, it contains outsider stars. Therefore, we would like to find a better model that improves this initial characterization
-by reducing the amount of clusters and also that would be able to remove outsiders from the OC.
+If we take a look into the H-R diagram (Figure \ref{fig:kmeans_hr_diagram_melotte_22}),
+we can identify the isochrone curve for Melotte 22 cluster.
+This curve has a good shape, but again, it contains outsider stars.
+Therefore, we would like to find a better model that improves this initial characterization
+by reducing the amount of clusters and also that removes outsiders from the OC.
 
@@
 
-Our initial approach allows us to find some groups that potentially contain the desired OC. However, as we have seen, the result is not enough
-accurate. Therefore, we need a method for merging clusters by migrating stars from one group to another,
+Our initial approach allows us to find some groups that potentially contain the desired OC.
+However, as we have seen before, the result is not enough accurate.
+Therefore, we need a method for merging clusters by migrating stars from one group to another,
 so we can end up with the minimum number of clusters while preserving the open cluster.
 
-Since we do not have a labeled dataset which allows us to train a supervised model, we have no choice but to use an unsupervised self-trained model.
+Since we do not have a labeled dataset to train a supervised model,
+we have no choice but to use an unsupervised self-trained model.
 
@@
 
-Although, as explained in section \ref{sec:feature_selection}, the number of features we are managing is not too large, this latent space allows us to reduce
-the number of features, avoiding this way the \emph{"curse of dimensionality"} \cite{bellman1961curse}.
-
-These encoders are pretrained before fitting the model to generate predictions. Then, the encoder is used to transform input data to the
-latent space $Z$. Once the data has been transformed, a K-Means clusterer is used to make an initial clustering.
+Although, as explained in section \ref{sec:feature_selection}, the number of features we are managing is not too large,
+this latent space helps us reduce the number of features and avoids the \emph{"curse of dimensionality"} \cite{bellman1961curse}.
+
+These encoders are pretrained before fitting the model to generate predictions.
+Then, the encoder is used with he aim of transforming input data to the latent space $Z$.
+Once the data has been transformed, a K-Means clusterer is used in order to make an initial clustering.
 
 With that initial configuration, the model iterates alternating between computing an auxiliary target distribution (Soft Assignment)
-and minimizing the Kullback-Leiber (KL) divergence \cite{kullback1951information} to it. This unsupervised algorithm allows us to improve
-the clustering.
+and minimizing the Kullback-Leiber (KL) divergence \cite{kullback1951information} to it.
+This unsupervised algorithm allows us to improve the clustering.
 
@@
 In the soft assignment stage, the \emph{Student's t-distribution} is used as a kernel to measure the similarity
-between the embedded points and the cluster centroid. While in the KL divergence minimization the algorithm iteratively refines clusters by learning
-from their high confidence assignments with the help of an auxiliary target distribution. The model is trained by matching the soft assignment to the
-target distribution. The choice of this target distribution is crucial for DEC's performance. In this work we have taken the target distribution
-from DEC's original paper \cite{xie2016unsupervised}, which is defined in Equation \ref{eq:student_tdistribution}.
-
-Figure \ref{fig:dec_model_setup} shows the layer setup of our DEC model. It is simpler than the one tested on the original paper \cite{xie2016unsupervised},
-since the number of selected features in our work is smaller. Therefore, using the same configuration would result in a model too
+between the embedded points and the cluster centroid.
+While in the KL divergence minimization the algorithm iteratively refines clusters by learning
+from their high confidence assignments with the help of an auxiliary target distribution.
+The model is trained by matching the soft assignment to the target distribution.
+The choice of this target distribution is crucial for DEC's performance.
+In this work we have taken the target distribution from DEC's original paper \cite{xie2016unsupervised},
+which is defined in Equation \ref{eq:student_tdistribution}.
+
+Figure \ref{fig:dec_model_setup} shows the layer setup of our DEC model.
+It is simpler than the one tested on the original paper \cite{xie2016unsupervised},
+since the number of selected features in our work is smaller.
+Therefore, using the same configuration would result in a model so
 powerful that would incur in overfitting issues unable to make right predictions.
@@
 Once the model is ready, we can test it by applying it to the Melotte 22 dataset.
-Notebook \verb|cluster_characterization.ipynb| illustrates how DEC model processes a region of stars trying to characterize the open cluster hidden within that region.
-
-As shown in Figure \ref{fig:dec_melotte_22}, the DEC model is able to move stars from one group to another removing some clusters and improving the OC characterization.
+Notebook \verb|cluster_characterization.ipynb| illustrates how DEC model processes a region of stars
+trying to characterize the open cluster hidden within that region.
+
+As shown in Figure \ref{fig:dec_melotte_22}, the DEC model is able to move stars from one group to another,
+removing some clusters and improving the OC characterization.
 This is exactly what we are looking for.
@@
 
-For that reason, we have used an alternative method based on Virtual Observatory (VO) tools such as Clusterix and TOPCAT,
+For that purpose, we have used an alternative method based on Virtual Observatory (VO) tools such as Clusterix and TOPCAT,
 to characterize the studied clusters.
@@
 
-The aim of this section is to highlight procedure differences between the our method and this new one,
-as well as using the results given by this method to validate ours results.
+The aim of this section is to highlight the procedure differences between our method and this new one,
+as well as using the results provided by this method to validate ours results.
 This validation method uses a combination of VO tools: Clusterix 2.0 + TOPCAT.
-We take Melotte 22 again as our study case to make easier taking comparisons between the method presented in Chapter \ref{chap:method}
-and this alternative method.
-
-The initial point, in order to avoid a subjective and manual selection for the cluster members, is taking advantage of Clusterix.
+We take Melotte 22 again as our study case to make easier comparisons between the method presented in Chapter \ref{chap:method}
+and this alternative one.
+
+The initial point is taking advantage of Clusterix in order to avoid a subjective and manual selection for the cluster members.
 Figure \ref{fig:clusterix_selection} shows a screenshot of the Clusterix 2.0 selection panel.
@@
 This panel allows us to select a region by its coordinates or by searching a cluster by its name in the Gaia DR2 catalogue,
-and specify the size of the region by introducing its radius.
-We can also set limit values in magnitude (for both, minimum and maximum) to discard stars outside those limits.
+as well as specify the size of the region by introducing its radius.
+We can also set limit values in magnitude (for both minimum and maximum) to discard stars outside these limits.
 Another setting is the flag to enable or disable the \emph{Q-Filter}.
@@
 
-It is necessary to notice the importance of this step, since the selection of these regions is critical to get a good result.
+It is necessary to remark the importance of this step, since the selection of these regions is critical to get a good result.
 However, despite how crucial is this selection, there is no general rule to define these regions.
 We only have two references: the radius of the downloaded region and the estimated size for the open cluster that sets the \emph{inner radius}.
-The other two radius must be set by trial and error and, in general, they must be generous to be sure we are not taking polluted regions.
+The other two radius must be set by trial and error. In general, they should be large enough to be sure we are not taking polluted regions.
 This is a tedious process that requires several tries before achieving good enough probability values.
 
-Furthermore, it is necessary to know in advance some additional profile of the cluster, in addition to the estimated size.
-For example, in the case of Melotte 22, the proper motion in declination direction of the membership stars is around $-45 mas \cdot year^{-1}$,
-so it is necessary to open the $\mu$ range above the estimated values (in absolute value). If we do not do that we loose its detection,
-since Clusterix sets this value to $30 mas \cdot year^{-1}$ by default. But even in this case, it is necessary to set a threshold for
-magnitude values, $G \leq 14$, to help getting the right result.
+Furthermore, it is necessary to know in advance some additional profiles of the cluster, in addition to the estimated size.
+For example, in the case of Melotte 22, the proper motion in declination direction of the membership stars around $-45 mas \cdot year^{-1}$,
+so it is necessary to open the $\mu$ range above the estimated values (in absolute value).
+If we do not do that we lose the possibility of detection, since Clusterix sets this value to $30 mas \cdot year^{-1}$ by default.
+But even in this case, it is necessary to set a threshold for magnitude values, $G \leq 14$, to help getting the right result.
 Figure \ref{fig:clusterix_step2_melotte_22} shows the region selection panel on Clusterix 2.0.
@@
 This step can be repeated over and over again, varying radii and other parameters, until finally reach a valid probability that works.
-The last step presents a table with candidates objects and allows to export this dataset to TOPCAT,
-Aladin and VOSA to perform a deeper filtering and analysis based on the results given by Clusterix.
+The last step presents a table with candidate objects and allows to export this dataset to TOPCAT,
+Aladin and VOSA to perform a deeper filtering and analysis based on the results provided by Clusterix.
 
@@
 Once data is available in TOPCAT, we have to carefully analyze the imported data.
-From there, we can perform a first stars selection by excluding those stars with probability value lower than 0.8.
-This first result can be taken as a good staring point, but we still need to make some fine tunning.
+From there, we can perform a first star selection by excluding those stars with probability value lower than 0.8.
+This first result can be taken as a good starting point, but we still need to make some fine tunning.
 As we can see in Figure \ref{fig:topcat_prob_selection_melotte_22}, the selection is too restrictive and we have lost too many stars.
@@
 Taken these observations into account, we can proceed to make a new selection based on parallax properties.
-Our second selection uses the values obtained with the previous selection for parallax and its standard deviation and
+Our second selection makes use of the values obtained with the previous selection for parallax and its standard deviation and
 selects those stars from the whole set which fall inside the limits $7.304mas \pm 0.797$ (see Figure \ref{fig:topcat_parallax_melotte_22}).
@@
 Since our aim is to get a model that fits well for a wide range of clusters,
-we present now some results obtained for a selection of clusters with different topologies.
+we will now proceed to show some results obtained for a selection of clusters with different topologies.
 
 All results have been computed with an Apple Mac Pro Late 2013, 2.7 GHz 12-Core Intel Xeon E5 with two AMD FirePro D700 6GB and 64 GB
-of memory, running macOS BigSur 11.0.1, Keras 2.2 and PlaidML 0.7.
+of memory, running macOS BigSur 11.1, Keras 2.2 and PlaidML 0.7.
 
@@
 
-\listoffigures
-\listoftables
-
 \newpage

### 2021-01-03 thesis/thesis.tex
thesis: Add new corrections

@@
 \begin{equation}
-  K(a, b) = \frac{1}{2 \pi h^{2}} \exp{ \left[ -\frac{1}{2}\frac{\left( a - a_{i} \right)^{2} + \left( b - b_{i} \right)^{2}}{ h^{2}} \right]}
+  K(a, b) = \frac{1}{2 \pi h^{2}} \exp{ \left[ - \frac{1}{2}\frac{\left( a - a_{i} \right)^{2} + \left( b - b_{i} \right)^{2}}{ h^{2}} \right]}
 \end{equation}
@@
 This allows us to download data for as many regions as we want without loosing efficiency when accessing for that data.
+% TODO: Revisar corrección de Imanol
 
@@
 
-Table \ref{tab:results_melotte_22} shows a summary of Melotte 22 with values taken from different sources and methods: SIMBAD Astronomical Database - CDS (Strasbourg),
+Table \ref{tab:results_melotte_22} shows a summary of Melotte 22 with values taken from different sources and methods:
+SIMBAD Astronomical Database - CDS (Strasbourg),
 Clusterix+TOPCAT method, and the results obtained for each step of this method (KMeans, DEC and DEC filtered).
@@
 
-We have seen in the previous chapter that the DEC model is capable of refining the clustering selection made by K-Means.
-This allows us to obtain a more precise characterization of the cluster.
+We have seen in the previous chapter that the DEC model is capable of refining the clustering selection made by K-Means
+hence helping us have a more precise characterization of the cluster.
 Now, in order to validate our model and be able to estimate how good or bad it is,

### 2021-01-07 thesis/thesis.tex
thesis: Review Abstract

@@
 
-The characterization and knowledge of \emph{Open Clusters} allow us to better understand properties and mechanisms about the Universe
-such as stellar formation and the regions where these events occur.
+The characterization and understanding of \emph{Open Clusters} allow us to better understand properties and mechanisms about the Universe
+such as stellar formation and the kind of regions where these events occur.
 They also provide information about stellar processes and the evolution of the galactic disk.
@@
 La caracterización y conocimiento de \emph{Cúmulos Abiertos} permite conocer mejor propiedades y mecanismos del Universo tales como
-la formación de estrellas y las regiones donde se dan estos procesos. También permiten obtener información sobre procesos estelares
-y la evolución del disco galáctico.
+la formación de estrellas y las regiones donde se dan estos procesos.
+También permiten obtener información sobre procesos estelares y la evolución del disco galáctico.
 
@@
 
-El método desarrollado pretende mejorar los ya existentes tanto en términos de \emph{eficiencia computacional}, con menores requisitos de cómputo,
+El método desarrollado pretende mejorar los ya existentes tanto en términos de \emph{eficiencia computacional}, con menores requisitos de cálculo,
 como en \emph{facilidad de uso}, reduciendo el número de hiperparámetros a configurar para obtener una buena caracterización de los cúmulos analizados.
@@
 Como se detalla más adelante, nuestro método consigue buenos resultados llegando a ser incluso mejor en algunos casos cuando se comparan
-los resultados con otros métodos actuales.
+los resultados con otras técnicas actuales.
 

### 2021-01-07 thesis/thesis.tex
thesis: Review Introduction

@@
 \cite{brinkmann2017bound}.
-Nevertheless, a small fraction of those systems will survive in the initial state and persist bounded in bigger timescales.
+Nevertheless, a small fraction of those systems will survive in the initial state and persist bounded in larger timescales.
 
@@
 Stars that belong to the same OC share relative positions, inherited from their original gas cloud.
-Which means that their distances to the Earth are similar for all of them and, therefore, they have a narrow dispersion in their parallax value.
+This means that their distances to the Earth are similar for all of them and, therefore, they have a narrow dispersion in their parallax value.
 They also share similar values of proper motion, both in right ascension and declination.
@@
 
-We will avoid this issue by looking at $E(b-r) / G_{mag}$ diagrams that reflect their connection with the isochrone curves,
-typical of their temporal evolution \cite{bressan2012parsec}.
+We will avoid this issue by looking at $E(b-r) / G_{mag}$ diagrams that show their connection with the isochrone curves.
+These curves are a tool used to determine how stars evolve in time \cite{bressan2012parsec}.
 The isochrones are derived from theoretical models.
 These models are mainly based on metallicity and mass/brightness ratio of stars.
-They also have a direct correspondence with that presented by the Hertzsprung–Russell (H-R) diagram of the stars belonging to the cluster.
+They also have a direct relation with that presented by the Hertzsprung–Russell (H-R) diagram of the stars belonging to the cluster.
 Examples of these curves are shown in Figure \ref{fig:examples_of_isochrones}.
@@
 
-Since these stars were born close in time, they should have a sharp, well-defined profile with little scattering through the main sequence in the H-R diagram.
+Since these stars were born close in time, they should have a sharp and well-defined profile,
+with low scattering through the main sequence in the H-R diagram.
 Examples for this kind of diagrams are presented in Figure \ref{fig:examples_of_hr_diagrams}.
@@
 However, if we look for overdensities in the proper motion configuration spaces, it is possible, at least at first instance,
-to assume a possible membership delimitation (see Figure \ref{fig:pm_ngc_2682}.
+to assume a possible membership cut (see Figure \ref{fig:pm_ngc_2682}.
 
@@
 Overdensities in proper motion are not always so evident.
-Additionally, we will consider the parallax distribution to improve our cluster characterization.
+In order to improve the characterization capabilities of our model,
+we will consider the parallax distribution as well.
 
@@
   \end{subfigure}
-  \caption{Melotte 22 Parallax histogram, clearly differentiated around the value $\approx 7.3mas$ and the corresponding diagram from the photometric magnitudes.}
+  \caption{Melotte 22 Parallax histogram, clearly differentiated around the value $\approx 7.3mas$
+           and the corresponding diagram from the photometric magnitudes.}
   \label{fig:melotte_22_pm_parallax}
@@
 In many other cases things are not so simple,
-and the characterization process can turn very difficult and tedious using tools such as the available ones at the Virtual Observatory (VO),
-and frequently it requires parametrization based on prior knowledge about the studied region.
+and the characterization process can turn very difficult and tedious using tools such as the available ones at the Virtual Observatory (VO).
+Furthermore, it frequently requires a parametrization based on prior knowledge regarding the studied region.
 The aim of this work is to test an unsupervised and non-parametrized model based on Machine Learning (ML) tools.
 
-We start enumerating the aims we try to achieve with this work.
+We start by enumerating the aims we try to achieve with this work.
 Then, we expose the current \emph{state of the art}, mentioning the \emph{Gaia Mission} used as data source in this work and
 describing other related works that try to solve this problem.
-As a last point of the state of the art, we make a brief summary of some of the today available clustering algorithms.
-After the state of the art, we explain our method to improve open cluster characterization.
-Next, we show an alternative method based on current procedures. We have used this alternative method as a validation method to test ours.
-Then, a comparison between a set of results obtained with our method and the results obtained for the same clusters with the validation method is presented.
-Finally, we extract some conclusions based on the obtained results and present further researches that can be done in light of the results.
+The state of the art ends with a brief summary for some of the clustering algorithms available today.
+In the \emph{Method} chapter, we explain our model to improve open cluster characterization.
+Also, we show an alternative method based on current procedures.
+We have used this alternative technique as a validation method to test ours.
+Then, a comparison between a set of results obtained with our method and the results obtained
+for the same clusters with the validation method is presented.
+Finally, we extract some conclusions based on the obtained results and present
+further researches that can be done in light of the results.
 

### 2021-01-08 thesis/thesis.tex
thesis: Review Aims

@@
 
-The main aim of this thesis is to \emph{build an unsupervised clustering model for open cluster characterization}.
+The primary aim of this thesis is to \emph{build an unsupervised clustering model for open cluster characterization}.
 The model will be \emph{non-supervised and non-parameterized} to fit a wide range of clusters without
@@
   \item Select and implement unsupervised clustering algorithms from the previous study.
-  \item Use those algorithms with different datasets to find OCs.
-  \item Look for an independent method to characterize OCs in order to make comparisons.
-  \item Validate found OCs with our custom model by making comparisons with the independent method.
+  \item Use chosen algorithms with different datasets to find OCs.
+  \item Look for an independent technique to use it as a validation method.
+  \item Validate OCs found with our custom model by making comparisons with the validation method.
 \end{itemize}

### 2021-01-08 thesis/thesis.tex
thesis: Review State of the Art

@@
 
-An initial approach for finding OCs is the search of overdensities along the galactic disk. In general, this is the starting point but,
-although it seems simple, it presents a fundamental problem already discussed. The near field around the OC is filled with two types of
-distinct star populations: those who belong to the OC (tens or hundreds to a few thousand) and a background made up of thousands or
-millions of stars that are not. Finding out which stars belong to the first group is the problem faced in this work.
+An initial approach for finding OCs is the search for overdensities along the galactic disk.
+In general, this is a good starting point but, although it seems simple,
+it presents a fundamental problem already discussed.
+The near field around the OC is filled with two types of distinct star populations:
+those who belong to the OC (tens or hundreds to a few thousand)
+and a background made up of thousands or millions of stars that are not.
+Finding out which stars belong to the first group is the problem faced in this work.
 This selection is crucial to properly characterize the fundamental properties of the cluster
@@
 
-Sometimes, the problem is easy to solve, as we have seen, by studying astrometric parameters and looking for overdensities in the proper
-motion configuration space as well as in the parallax space.
+Sometimes, the problem is easy to solve, as we have seen, by studying astrometric parameters
+and looking for overdensities in the proper motion configuration space as well as in the parallax space.
 
@@
 
-However, as shown in Figures \ref{fig:pm_ngc_6494} and \ref{fig:parallax_ngc_6494}, in general it is not as easy and
-becomes necessary to consider other parameters such as distances, or even metallicity and age (derived from isochrone curves)
-and requires photometric data for the stars within the studied field.
+However, as shown in Figures \ref{fig:pm_ngc_6494} and \ref{fig:parallax_ngc_6494},
+in general it is not as easy and becomes necessary to consider other parameters such as distances,
+or even metallicity and age (derived from isochrone curves).
+Sometimes even, photometric data may be required for the stars within the studied field.
 
@@
 At this point, it seems reasonable that the study of each individual group requires its own parametrization and technique
-to identify it. With Gaia DR2, which contains great high-quality astrometric and photometric data for a huge number of stars,
+to identify it. With Gaia DR2, which contains high-quality astrometric and photometric data for a huge number of stars,
 a new opportunity arises to simplify and optimize these processes. In this context, new approaches are being developed
@@
 
-In this work we make use of Gaia DR2 since, although Release 3 (Gaia DR3) is imminent, it has not been available in time for us.
+In this work we make use of Gaia DR2 since DR3 has not been released in time for us to include it.
 Gaia DR2 is a multidimensional dataset obtained by ESA's Gaia mission (located at L2, 1.5 million kilometers from Earth) and
@@
 and magnitudes in three photometric filters (G, BP and RP) for more than 1,300 million sources.
-80\% of Gaia DR2 sources are weaker than $G_{mag} \approx 18$. For magnitudes $G \approx 21$ nominal uncertainties reach
-$2 mas$ for parallaxes and $5 mas \cdot years^{-1}$ in proper motions. But for closer sources, (at the bright end, $G < 14$),
-the precision is of the order of $0.02 mas$ and $0.05 mas \cdot year^{-1}$ respectively \cite{cantat2018gaia}. In general,
-open clusters, which is the main subject of this work, are at optimal data thresholds.
-
-In this context, and taking advantage of the opportunity provided by Gaia DR2, new approaches are being developed to improve current
-detection methods and automating OC characterization as much as possible. This is the case of tools such as Clusterix 2.0
-\cite{balaguer2020clusterix}, TOPCAT \cite{taylor2005topcat}, and VOSA \cite{bayo2008vosa}, (all of them developed by the Virtual Observatory).
-These tools work together and offer good results, although it is not possible to use them in an unsupervised process and free of certain
-different parametrization in each case.
+80\% of Gaia DR2 sources are weaker than $G_{mag} \approx 18$.
+For magnitudes $G \approx 21$ nominal uncertainties reach $2 mas$ for parallaxes and $5 mas \cdot years^{-1}$ in proper motions.
+But for closer sources, (at the bright end, $G < 14$), the precision is of the order of $0.02 mas$
+and $0.05 mas \cdot year^{-1}$ respectively \cite{cantat2018gaia}.
+In general, open clusters, which are the main subject of this work, are at optimal data thresholds.
+
+In this context, and taking advantage of the opportunity provided by Gaia DR2,
+new approaches are being developed to improve current detection methods and automating OC characterization as much as possible.
+This is the case of tools such as Clusterix 2.0 \cite{balaguer2020clusterix}, TOPCAT \cite{taylor2005topcat}
+and VOSA \cite{bayo2008vosa}, (all of them developed by the Virtual Observatory).
+These tools work together and offer good results,
+although it is not possible to use them in an unsupervised process and they require different parameterizations for each study case.
 
@@
 
-TOPCAT alone does not offer an independent discrimination mechanisms to researchers.
-They should try an initial selection, starting from the map of proper motions,
-by delimiting the area where the cluster presumably appears. When this is not possible, it is necessary to have a
-certain knowledge about the cluster profile and to parametrize some approximated values.
-In both cases, the definition of the selected set must then be optimized step by step until reaching a result that can be considered reasonable.
+TOPCAT by itself does not offer an independent discrimination mechanisms to researchers.
+They have to try an initial selection, starting from the representation of proper motions,
+by delimiting the area where the cluster most probably appears.
+When this is not possible, it is necessary to have a previous knowledge about the cluster profile
+and to parametrize some approximated values.
+In both cases, the obtained selection must then be optimized step by step
+until a result that meets the confidence levels is achieved.
 
 Another tool is Clusterix 2.0, which is an interactive web-based tool \cite{balaguer2020clusterix}.
-It takes the proper motion diagram without making any prior assumptions about the membership of the candidate star
+It takes the proper motion diagram without making any prior assumption about the membership of the candidate star
 and determines empirically the frequency functions.
@@
 
-$(a, b)$ are referred to the proper motion configuration space, while a point located in the center of the cell $(i, j)$ provides the maximum
-contribution for calculating the local density. In this context, the number $h$ is called the \emph{smoothing} parameter and it is measured in
-the same units as proper motion.
-
-With this same purpose of applying non-parametric methods, various clustering algorithms have been applied in the past, UPMASK \cite{krone2014upmask}
-was developed to only use photometry and positions, and then it was adapted to the Gaia data \cite{cantat2018gaia} based only on proper motion and
-parallax. A similar approach has been carried out using DBSCAN and machine learning algorithms to discover new clusters.
-
-In this sense, Clusterix also claims to be a non-parametric method, but it critically depends on the initial selection for the field sizes to be analyzed.
+$\left( a, b \right)$ are referred to the proper motion configuration space,
+while a point located in the center of the cell $\left( i, j \right)$ provides the maximum
+contribution for calculating the local density.
+Number $h$ is called the \emph{smoothing} parameter and it is measured in the same units as proper motion.
+
+Various clustering algorithms have been studied in the past with the purpose of finding a valid non-parameterized method.
+UPMASK \cite{krone2014upmask} was developed to only use photometry and positions,
+and then it was adapted to the Gaia data \cite{cantat2018gaia} based only on proper motion and parallax.
+A similar approach has been carried out using DBSCAN and machine learning algorithms to discover new clusters.
+
+In this sense, Clusterix also claims to be a non-parameterized method,
+but it critically depends on the initial selection for the field sizes to be analyzed.
 
@@
 \begin{displayquote}
-The non-field population does not occupy the entire workspace, but is spatially concentrated, which makes it possible to distinguish two regions
-in the workspace: the only field region (label 'f'), dominated by star fields, and the cluster + field region (label 'c+f'),
-which includes both star fields and not star fields. \cite{balaguer2020clusterix}
+The non-field population does not occupy the entire workspace, but is spatially concentrated,
+which makes it possible to distinguish two regions in the workspace:
+the only field region (label 'f'), dominated by star fields,
+and the cluster + field region (label 'c+f'), which includes both star fields and not star fields.
+\cite{balaguer2020clusterix}
 \end{displayquote}
 
-This fact implies defining three areas or regions with different radius. The first 'c+f' corresponds to the one in which the cluster members are
-presumed to be contained together with other star fields that are not part of the cluster. The second, broader, assumes that it only contains
-stars in an extended visual field without components of the cluster, and a third intermediate region, which is out of analysis (void area), which
-would correspond to a possible transition zone between the other two.
-
-The correct choice of these radii, even having a previous estimation for the 'c+f' region, significantly affects the execution of the algorithm and,
-in general, requires a generously wide field 'f'. There is no general rule that defines relative proportions of these areas.
-
-Finally, when an acceptable result is obtained, what Clusterix provides is a probability field associated with the set of stars under study.
-The dataset, with an extra column for the probability value of each object to belong to the open cluster, can be exported to TOPCAT
-for further optimization and final processing. Figure \ref{fig:clusterix_probability} shows an example of the probability distribution in the proper
-motion space for NGC 2682.
+This fact implies defining three areas or regions with different radius.
+The first 'c+f' corresponds to the one in which the cluster members are
+presumed to be contained together with other star fields that are not part of the cluster.
+The second region is the broadest and assumes that it only contains stars
+in an extended visual field without components of the cluster.
+The third region is the intermediate one and is out of analysis (void area),
+since it would correspond to a possible transition zone between the other two.
+
+The right choice of these radii, even having a previous estimation for the 'c+f' region,
+highly affects the execution of the algorithm and, in general, requires a considerable wide field 'f'.
+There is no rule of thumb that defines relative proportions of these areas.
+
+Finally, when an acceptable result is obtained,
+what Clusterix provides is a probability field associated to the set of stars under study.
+The recovered dataset contains all fields from Gaia database
+and an extra column with the probability value of each object to belong to the open cluster.
+This new dataset can be exported to TOPCAT for further optimization and final processing.
+Figure \ref{fig:clusterix_probability} shows an example of the probability distribution
+in the proper motion space for NGC 2682.
 
@@
 
-Finally, it is necessary to refer to the work of \cite{castro2020hunting}, which presents a method based on machine learning techniques to make
-a systematic search for overdensities in the astrometric space of the galactic disk and a subsequent identification of OC using photometric
-information, also from Gaia DR2.
-
-The method includes two phases: the first one uses an unsupervised clustering algorithm, DBSCAN, to search for overdensities
-$(l, b \pi, \mu_{\alpha} *, \mu_{\delta})$, ans subsequently applies a deep learning ANN trained with magnitude diagrams to identify isochrone
-patterns within the detected overdensities and thus proceed to confirm them as OC.
-
-It should be noted that for the execution of this method, MareNostrum 42 (Barcelona Supercomputing Center) was used.
-So the neural network could handle the image recognition process with isochrone patterns and not applying theoretical models
-derived from values such as metallicity, masses, among other.
-
-The result of this work is the recognition of 582 new open clusters distributed along the galactic disk for a galactic declination $b < 20^{\circ}$,
+Finally, it is necessary to refer to the work developed on \cite{castro2020hunting}.
+It presents a method based on machine learning techniques to make a systematic search
+for overdensities in the astrometric space of the galactic disk
+and a later identification of OC using photometric information, also from Gaia DR2.
+
+The method includes two phases: the first one uses an unsupervised clustering algorithm, DBSCAN,
+to search for overdensities $(l, b \pi, \mu_{\alpha} *, \mu_{\delta})$,
+and then applies a deep learning Artificial Neural Network (ANN),
+previously trained with magnitude diagrams,
+to identify isochrone patterns within the detected overdensities and thus proceed to confirm them as OC.
+
+It should be noted that for the execution of this method,
+MareNostrum 42 (Barcelona Supercomputing Center) was used.
+So the neural network could handle the image recognition process with isochrone patterns
+and not applying theoretical models derived from values such as metallicity or masses, among other.
+
+The result of this work is the recognition of 582 new open clusters
+distributed along the galactic disk for a galactic declination $b < 20^{\circ}$,
 which has meant an increase of 45\% over previously ones.
 
-Our aim in this work, linked to computational limitations compared to MareNostrum, is not to do a blind search for clusters,
-but to obtain a method, within the machine learning environment,
+Our aim in this work, linked to computational limitations compared to MareNostrum,
+is not to do a blind search for clusters, but to obtain a method,
+within the machine learning environment,
 that allows the characterization of open clusters through unsupervised and non-parametrized procedures in a single step.
@@
 
-The result of Clusterix + TOPCAT, applied to the same clusters, will be used only for comparing and validating our results.
+The result of Clusterix + TOPCAT, applied to the same clusters,
+will be used only for comparing and validating our results.
 
@@
 
-As we will explain later, what we need to accomplish our aim is an algorithm that manages to make data groups based on the dynamic properties of the stars.
+As we will explain later, what we need to accomplish our aim is an algorithm
+that manages to make data groups based on the dynamic properties of the stars.
 
 There are several clustering algorithms: KMeans, Mean-Shift Clustering, DBSCAN \cite{ester1996density} among other.
-While each one behaves better according ot the distribution of the objects to clusterize, we have choosen KMeans by its simplicity and good results.
+While each one behaves better according to the distribution of the objects to clusterize,
+we have chosen KMeans by its simplicity and good results.
 
@@
 The algorithm looks for N center points which are vectors of the same dimension as the number of selected features.
-KMeans starts by making an initial groups configuration and then, it reassigns objects to other groups iteratively by minimizing
-the distance among points inside the new group and maximizing the distance with the centers of the other groups.
-
-The algorithm stops once a maximum number of iterations is reached, or when changes between iterations is lower than a minimum.
+KMeans starts by making an initial groups configuration and then,
+it reassigns objects to other groups iteratively by minimizing the distance among points
+inside the new group and maximizing the distance with the centers of the other groups.
+
+The algorithm stops once a maximum number of iterations is reached,
+or when changes between iterations is lower than a minimum.
 
@@
 However, we need to set a large number of clusters to find the open cluster we are looking for.
-This fact complicates the identification of the OC and many times the OC contains to many outliers.
-For that reason, we have search for a KMeans refinement based on an Artificial Neural Network (ANN).
-
-The \emph{Unsupervised Deep Embedding for Clustering Analysis} model or \emph{DEC} \cite{xie2016unsupervised} takes KMeans as its starting point,
-but then, it trains an autoencoder to reduce the feature space and pass this transformed data through a Clustering Layer which refines the previous selection.
-
-The mechanism of this model is explained in Section \ref{sec:deep_embedding_clustering} and its implementation is available in \verb|cdalvaro.ml.dec| package.
+This fact complicates the identification of the OC and many times the found cluster contains to many outliers.
+For that reason, we have search for a KMeans refinement based on an artificial neural network.
+
+The \emph{Unsupervised Deep Embedding for Clustering Analysis} model
+or \emph{DEC} \cite{xie2016unsupervised} takes KMeans as its starting point,
+but then, it trains an autoencoder to reduce the feature space
+and pass this transformed data through a Clustering Layer which refines the previous selection.
+
+The mechanism of this model is explained in Section \ref{sec:deep_embedding_clustering}
+and its implementation is available in \verb|cdalvaro.ml.dec| package.
 

### 2021-01-09 thesis/thesis.tex
thesis: Review Method

@@
 
-In this chapter we address the different steps followed to perform the creation of the unsupervised clustering model for open cluster characterization.
-
-All code has been developed with the Python programming language \cite{Python3}. Other auxiliary tools like \emph{Docker} \cite{merkel2014docker},
-\emph{PostgreSQL} \cite{postgresql}, \emph{Jupyter Notebooks} \cite{Kluyver2016jupyter}, and frameworks such as \emph{Astropy} \cite{astropy:2013} \cite{astropy:2018},
-\emph{Scikit-Learn} \cite{scikit-learn}, \emph{Seaborn} \cite{michael_waskom_2017_883859}, \emph{SQLAlchemy} \cite{sqlalchemy} and
-\emph{Keras} \cite{chollet2015keras} have been used too.
+In this chapter we address the different steps followed to perform the creation of
+the unsupervised clustering model for open cluster characterization.
+
+All code has been developed with the Python programming language \cite{Python3}.
+Other auxiliary tools like \emph{Docker} \cite{merkel2014docker},
+\emph{PostgreSQL} \cite{postgresql}, \emph{Jupyter Notebooks} \cite{Kluyver2016jupyter},
+and frameworks such as \emph{Astropy} \cite{astropy:2013} \cite{astropy:2018},
+\emph{Scikit-Learn} \cite{scikit-learn}, \emph{Seaborn} \cite{michael_waskom_2017_883859},
+\emph{SQLAlchemy} \cite{sqlalchemy} and \emph{Keras} \cite{chollet2015keras} have been used too.
 
@@
 For the sake of simplicity, we are taking \emph{Melotte 22 (Messier 45)} \cite{elsanhoury2019ppmxl} to illustrate the whole method.
-This cluster is well studied and its stars are not too mixed with other stars that do not belong to the OC although they are contained
-in the same observation field, so images look clear.
-
-Figure \ref{fig:raw_pm_melotte_22} shows \emph{proper motions in right ascension and declination} for a sample of the downloaded dataset for Melotte 22.
-At first sight, two main clusters can be distinguished, one of them centered nearly at (0, 0) and the second cluster with center at (20, -45). This second cluster
-is the one we are looking for.
-
-However, although this second cluster is almost isolated, there are stars that do not belong to the OC.
-Thus, we need more information to properly characterize the OC.
+This cluster is well studied and its stars are not too mixed with other stars that do not belong to the OC
+although they are contained in the same observation field, so images look clear.
+
+Figure \ref{fig:raw_pm_melotte_22} shows \emph{proper motions in right ascension and declination}
+for a sample of the downloaded dataset for Melotte 22.
+At first sight, two main clusters can be distinguished, one of them centered nearly at (0, 0)
+and the second one with center at (20, -45). This second cluster is the one we are looking for.
+
+However, although the second cluster is almost isolated, there are stars that do not belong to the OC.
+Thus, we need more information to properly characterize the open cluster.
 
@@
 
-The figure on the left shows a resonance at $\approx 7.3mas$ belonging the OC. While the figure on the right would help us to look for isochrone curves and so,
+The figure on the left shows a resonance at $\approx 7.3mas$ belonging the OC.
+While the figure on the right would help us to look for isochrone curves and so,
 to determine the age of these stars.
@@
 
-Before being able to develop and test our clustering model, the first step is to download the required data from Gaia repository
+Before being able to develop and test our clustering model,
+the first step is to download the required data from Gaia repository
 and store it in a custom database for later access.
@@
 Due to the large amount of data available at Gaia, a complete download is not viable neither useful.
-In order to reduce the size of the dataset to be downloaded, the OpenClust catalogue \cite{dias2002new} (Figure \ref{fig:OpenClustComplete})
+In order to reduce the size of the dataset to be downloaded,
+the OpenClust catalogue \cite{dias2002new} (Figure \ref{fig:OpenClustComplete})
 has been used to restrict the sky regions to be explored.
@@
 Instead, a wider region (1.5 times the size of the cluster) is downloaded for each cluster to include stars outside the cluster.
-The idea is that the unsupervised model must be able to clusterize the stars that belong to the OC and to discard
-those outsiders when characterizing the open cluster.
+The idea is that the unsupervised model must be able to clusterize those stars that belong to the OC
+and to discard outsiders when characterizing the open cluster.
 
@@
 
-Taking these considerations into account, the downloaded dataset covers nearly 114 million stars (42GB of compressed data).
-This dataset is significantly smaller than the whole Gaia DR2 dataset, which contains information for approximately 1,600 million stars.
+Taking these considerations into account, the downloaded dataset covers nearly 114 million stars (around 42GB of compressed data).
+This dataset is significantly smaller than the whole Gaia DR2 dataset,
+which contains information for approximately 1,600 million stars.
 However, it is still too large.
 
-Since not all the downloaded clusters are good enough, we will apply a series of filters to discard clusters which do not have enough stars or contain too
-many \emph{null} values.
+Since not all the downloaded clusters are good enough,
+we will apply a series of filters to discard clusters which do not have enough stars
+or contain too many \emph{null} values.
 
@@
   \item Parallax absolute value greater than 0.0
-  \item Number of stars\footnote{Every star must have all required features completely defined, i.e. without null values} in the selected region above 40,000 stars
+  \item Number of stars\footnote{Every star must have all required features completely defined, i.e. without null values}
+        in the selected region above 40,000 stars
 \end{itemize}
 
-As shown in Figure \ref{fig:OpenClustSelection}, these constrains give us a smaller dataset but it is still a good representation of all clusters in the Milky Way
+As shown in Figure \ref{fig:OpenClustSelection},
+these constrains give us a smaller dataset but it is still a good representation of all clusters in the Milky Way
 since they are equally distributed around the galaxy disk.
 
-After having applied these filters, the number of cluster to analyse is 169 with nearly 75 million stars.
-
-Since we have set no upper limit to the number of stars inside a cluster, for the sake of simplicity and as a commitment to the project delivery dates
+After having applied these filters, the number of cluster to analyze is 169 with nearly 75 million stars.
+
+Since we have set no upper limit to the number of stars inside a cluster,
+for the sake of simplicity and as a commitment to the project delivery dates
 and the available computing power, smaller clusters will be preferred over the greatest ones.
 
-As an extension of this work, the designed model could be applied to a region not covered by these constrains for a later understanding of the results.
+As an extension of this work, the designed model could be applied to a region not covered by these constrains
+for a later understanding of the results.
 
@@
 
-The first one is a container built from a \verb|python:3.8| image that contains the \verb|cdalvaro| package and the \verb|downloader.py| script.
-This script is prepared to load the OpenClust catalogue, connect to the Gaia DR2 database, download all stars contained inside each cluster
-(with a radius multiplier to increase the area to be downloaded) and save the downloaded data inside our custom database hosted in the second container.
-
-\verb|downloader.Dockerfile| contains the build instructions for the downloader image. We use \href{https://github.com/features/actions}{GitHub Actions}
-for automating the process of building a new version of the image with every new push made to the \verb|main| branch.
-
-This image can be pulled from the \href{https://github.blog/2020-09-01-introducing-github-container-registry/}{GitHub Container Registry}:
+The first one is a container built from a \verb|python:3.8| image that contains the \verb|cdalvaro| package
+and the \verb|downloader.py| script.
+This script is prepared to load the OpenClust catalogue, connect to the Gaia DR2 database,
+download all stars contained inside each cluster (with a radius multiplier to increase the area to be downloaded)
+and save the downloaded data inside our custom database hosted in the second container.
+
+\verb|downloader.Dockerfile| contains the build instructions for the downloader image.
+We use \href{https://github.com/features/actions}{GitHub Actions} for automating the process
+for building a new image version with every new push made to the \verb|main| branch.
+
+This image can be pulled from the
+\href{https://github.blog/2020-09-01-introducing-github-container-registry/}{GitHub Container Registry}:
 
@@
 
-The second container is just a \verb|postgres:12.4| image which loads an initial script when the database is not yet initialized for creating the
-database schema. This container is used later as the main database for data analysis.
+The second container is just a \verb|postgres:12.4| image which loads an initial script when
+the database is not yet initialized for creating the database schema.
+This container is used later as the main database for data analysis.
 
@@
 
-The second table is partitioned by region, storing all stars related to one region inside an independent table for optimized access performance.
-The partition tables are located inside \verb|gaiadr2| schema. See Figure \ref{fig:customdb_diagram}.
+The second table is partitioned by region,
+storing all stars related to one region inside an independent table for optimized access performance.
+Partition tables are located inside \verb|gaiadr2| schema. See Figure \ref{fig:customdb_diagram}.
 
@@
 PostgreSQL knows that all stars we want are located inside \verb|gaiadr2.region_302| so it does not need to query any other table.
-This allows us to download data for as many regions as we want without loosing efficiency when accessing for that data.
-% TODO: Revisar corrección de Imanol
+This allows us to download data for as many regions as we want without loosing performance when accessing for that data.
 
@@
 \href{https://github.com/cdalvaro/machine-learning-master-thesis}{cdalvaro/machine-learning-master-thesis} repository.
-(For the printed version of this work, you can scan the QR code on cover for accessing to it.)
+(For the printed version of this work, you can scan the QR code on cover for accessing it.)
 
@@
 Parallax is another important feature. It lets us know how far stars are from us.
-And, since all stars within an open cluster were born from the same dust cloud, they must all have similar parallax.
-
-However, we are not going to use these raw features. Instead, we are taking a combination of them.
+In addition, since all stars within an open cluster were born from the same dust cloud,
+they must all have similar parallax.
+
+However, we are not going to use these raw features.
+Instead, we are taking a combination of them.
 
 First, we correct proper motion in right ascension and declination by dividing them by the parallax.
-That way we normalize these quantities and help our clustering models to improve their performance.
-
-Another computed property that we are considering is the modulus of the proper motion.
-That way, we relate both features and therefore force our model to keep them tight.
+That way, we normalize these quantities and help our clustering models to improve their performance.
+
+The modulus of the proper motion is another computed property that we are considering.
+We use it to relate both features and therefore, to force our model to keep them tight.
 
@@
 
-This feature selection has been refined by iterating over the K-Means clustering process with a fixed number of clusters and varying the feature selection in each case.
+This feature selection has been refined by iterating over the K-Means clustering process
+with a fixed number of clusters and varying the feature selection in each case.
 The final selection is the one with best \emph{silhouette score} \cite{rousseeuw1987silhouettes}.
@@
 The silhouette score is a metric used to determine how good a cluster is based on intra-cluster distances and nearest-cluster distances.
-The best possible value is 1, and the worst is -1. Negative values mean that some samples have been assigned to the wrong cluster.
+The best possible value is 1, and the worst is -1.
+Negative values mean that some samples have been assigned to the wrong cluster.
 
@@
 
-Once we have found the set of features that best describes our problem, we can begin searching for the OC in the study region.
-
-Our first approach to find the open cluster within the region of interest is to use the K-Means algorithm.
-
-Since we are looking for a single cluster, it seems reasonable to use a clustering algorithm and set it to find two clusters.
+Once we have found the set of features that best describes our problem,
+we can begin searching for the OC within the study region.
+
+Our first approach to find the open cluster is using the K-Means algorithm.
+
+Since we are looking for a single cluster,
+it seems reasonable to use a clustering algorithm and set it to find two clusters.
 One for the desired OC and another which contains stars that do not belong to the open cluster.
@@
 
-This is due to the fact that OC's stars are surrounded by other stars with possibly similar properties. So, setting the number of clusters
-to two is too low to separate them properly.
-
-As shown in Figure \ref{fig:kmeans_comparisons_melotte_22}, bigger values for the number of clusters allow us to isolate more accurately
-the resonance in parallax at $\approx 7.3mas$. However, we have the disadvantage that more groups are formed, while our aim is to find just
-one small cluster that contains the stars that belong to the OC and another one with the remaining stars.
-
-This effect complicates our task of finding the desired OC. Therefore, we have to find
-a way to set the right value for the number of clusters to isolate the searched cluster without creating too many groups.
+This is due to the fact that OC's stars are surrounded by other stars with possibly similar properties.
+So, setting the number of clusters to two is too low to separate them properly.
+
+As shown in Figure \ref{fig:kmeans_comparisons_melotte_22},
+larger values for the number of clusters allow us to isolate more accurately the resonance in parallax at $\approx 7.3mas$.
+However, we have the disadvantage that more groups are formed.
+
+This effect complicates our task of finding the desired open cluster,
+since we would like to get just two groups, one for the OC and another with the remaining stars.
+Therefore, we have to find a way to set the right value for the number of clusters
+to isolate the searched cluster without creating too many groups.
 
@@
 
-The following example is a snippet copied from \verb|cluster_characterization.ipynb| Jupyter notebook (available at \verb|src/notebooks|).
+The following example is a snippet copied from \verb|cluster_characterization.ipynb|
+Jupyter notebook (available at \verb|src/notebooks|).
 It shows how to estimate the best number of clusters for Melotte 22 by using \verb|cdalvaro.ml.estimate_n_clusters| method.
@@
 
-K-Means does a good job making an initial clustering, as shown in Figure \ref{fig:kmeans_melotte_22}.
-However, too many clusters arise from this characterization and the open cluster is still polluted with stars that do not belong to it.
+K-Means does a good job making an initial clustering,
+as shown in Figure \ref{fig:kmeans_melotte_22}.
+However, too many clusters arise from this characterization
+and the OC is still polluted with stars that do not belong to it.
 Moreover, we would like to reduce the amount of clusters too.
@@
 Our initial approach allows us to find some groups that potentially contain the desired OC.
-However, as we have seen before, the result is not enough accurate.
+However, as we have seen before, the result is not accurate enough.
 Therefore, we need a method for merging clusters by migrating stars from one group to another,
@@
 
-The implementation of this model is available in \verb|cdalvaro.ml.DEC| and it is developed with the Keras framework.
+The implementation of this model is available in \verb|cdalvaro.ml.DEC|
+and it is developed with the Keras framework.
 
@@
 
-Encoders are used to transform the input data into a latent space using a non-linear mapping function $f_{\theta} : X \rightarrow Z$.
-
-Although, as explained in section \ref{sec:feature_selection}, the number of features we are managing is not too large,
-this latent space helps us reduce the number of features and avoids the \emph{"curse of dimensionality"} \cite{bellman1961curse}.
+Encoders are used to transform the input data into a latent space
+using a non-linear mapping function $f_{\theta} : X \rightarrow Z$.
+
+Although, as explained in section \ref{sec:feature_selection},
+the number of features we are managing is not too large,
+this latent space helps us reduce the number of features
+and avoids the \emph{"curse of dimensionality"} \cite{bellman1961curse}.
 
 These encoders are pretrained before fitting the model to generate predictions.
-Then, the encoder is used with he aim of transforming input data to the latent space $Z$.
+Then, the encoder is used with the aim of transforming input data to the latent space $Z$.
 Once the data has been transformed, a K-Means clusterer is used in order to make an initial clustering.
 
-With that initial configuration, the model iterates alternating between computing an auxiliary target distribution (Soft Assignment)
+With that initial configuration,
+the model iterates alternating between computing an auxiliary target distribution (Soft Assignment)
 and minimizing the Kullback-Leiber (KL) divergence \cite{kullback1951information} to it.
@@
 
-In the soft assignment stage, the \emph{Student's t-distribution} is used as a kernel to measure the similarity
+In the soft assignment stage,
+the \emph{Student's t-distribution} is used as a kernel to measure the similarity
 between the embedded points and the cluster centroid.
@@
 It is simpler than the one tested on the original paper \cite{xie2016unsupervised},
-since the number of selected features in our work is smaller.
+since the number of selected features in our work is smaller than in the original one.
 Therefore, using the same configuration would result in a model so
@@
 
-As shown in Figure \ref{fig:dec_melotte_22}, the DEC model is able to move stars from one group to another,
+As shown in Figure \ref{fig:dec_melotte_22},
+the DEC model is able to move stars from one group to another,
 removing some clusters and improving the OC characterization.
@@
 
-The H-R diagram has been improved as well, and the isochrone curve of the open cluster is sharper than the one obtained with the K-Means algorithm
+The H-R diagram has been improved as well,
+and the isochrone curve of the open cluster is sharper than the one obtained with the K-Means algorithm
 (Figure \ref{fig:kmeans_hr_diagram_melotte_22}).
 
-Figure \ref{fig:hr_diagram_comparison} shows a comparison between the H-R diagram obtained by DEC clustering and the one obtained
-using the Clusterix + TOPCAT method.
+Figure \ref{fig:hr_diagram_comparison} shows a comparison between the H-R diagram obtained
+by DEC clustering and the one obtained using the Clusterix + TOPCAT method.
 
@@
 
-Finally, we can refine this selection by filtering those stars which are bellow and above the 0.25 and 0.75 quantiles for each group, respectively.
+Finally, we can refine this selection by filtering those stars which are bellow and above
+the 0.25 and 0.75 quantiles for each group, respectively.
 That way we remove the most doubtful values from the selection.

### 2021-01-14 thesis/thesis.tex
thesis: Review Results

@@
   \end{subfigure}
-  \caption{Melotte 22 feature pairplot}
-  % TODO: Ampliar la descripción de la figura
+  \caption{Pairwise relationships among variables using Melotte 22 data}
   \label{fig:features_melotte_22}
@@
 \newpage
+
+Figure \ref{fig:features_melotte_22} shows a pairwise relationship among some features available
+in the dataset for Melotte 22 data.
+The diagonal of the grid shows the marginal distribution of the data in each column.
+This plot can be useful to find correlations between pairs of variables or
+to identify resonances in certain variables such as those in the parallax feature.
 
@@
                                            min_clusters=3, max_clusters=7)
-  # Output
   # Silhouette score for 3 clusters: 0.5420
@@
     \begin{tabular}{l|c|c|c|c}
-      \textbf{Source / method} & \textbf{$\mu_{\alpha}$ $(mas/yr^{-1})$} & \textbf{$\mu_{\delta}$ $(mas/yr^{-1})$} & \textbf{$\varpi$ $(mas)$} & \emph{\# stars} \\
+      \textbf{Source / method} & \emph{$\mu_{\alpha}$ $(mas/yr^{-1})$} & \emph{$\mu_{\delta}$ $(mas/yr^{-1})$} & \emph{$\varpi$ $(mas)$} & \emph{\# stars} \\
       \hline
@@
     \end{tabular}
-    \caption{Melotte 22. The parameters shown are proper motion in right ascension and declination, parallax and number of stars with their respective deviations.}
+    \caption{Parameters shown are proper motion in right ascension and declination, parallax
+             with their respective deviations and number of stars corresponding to Melotte 22 data.}
     \label{tab:results_melotte_22}
@@
 
-We have seen in the previous chapter that the DEC model is capable of refining the clustering selection made by K-Means
+We have seen in the previous chapter that DEC model is capable
+of refining the clustering selection made by K-Means
 hence helping us have a more precise characterization of the cluster.
-Now, in order to validate our model and be able to estimate how good or bad it is,
-we need a different method to characterize the studied clusters and compare our results with the validation method.
-
-For that purpose, we have used an alternative method based on Virtual Observatory (VO) tools such as Clusterix and TOPCAT,
-to characterize the studied clusters.
+Now, in order to validate our model and to be able to estimate how good or bad it is,
+we need a different technique to characterize the studied clusters and compare our results
+with those obtained with the validation method.
+
+For that purpose, we have used an alternative method based on Virtual Observatory (VO) tools
+such as Clusterix and TOPCAT, to characterize the studied clusters.
 
@@
 
-The aim of this section is to highlight the procedure differences between our method and this new one,
-as well as using the results provided by this method to validate ours results.
-This validation method uses a combination of VO tools: Clusterix 2.0 + TOPCAT.
-We take Melotte 22 again as our study case to make easier comparisons between the method presented in Chapter \ref{chap:method}
-and this alternative one.
-
-The initial point is taking advantage of Clusterix in order to avoid a subjective and manual selection for the cluster members.
+The aim of this section is to highlight the procedure differences between our method and the validation one,
+as well as using the results provided by the second technique to validate ours results.
+This validation method uses a combination of VO tools: \emph{Clusterix 2.0 + TOPCAT}.
+We take Melotte 22 again as our study case to make easier the comparison process between
+the method presented in Chapter \ref{chap:method} and the alternative one.
+
+The initial point is taking advantage of Clusterix in order to avoid a subjective
+and manual selection for the cluster members.
 Figure \ref{fig:clusterix_selection} shows a screenshot of the Clusterix 2.0 selection panel.
@@
   \caption{Clusterix 2.0 selection panel with Melotte 22 settings.
-           The coordinates correspond to the center of Melotte 22 cluster,
+           Coordinates correspond to the center of Melotte 22 cluster,
            while the radius is the value registered for this cluster in the
@@
 
-This panel allows us to select a region by its coordinates or by searching a cluster by its name in the Gaia DR2 catalogue,
-as well as specify the size of the region by introducing its radius.
-We can also set limit values in magnitude (for both minimum and maximum) to discard stars outside these limits.
+This panel allows us to select a region by its coordinates
+or by searching a cluster by its name in the Gaia DR2 catalogue,
+as well as to specify the size of the region by introducing its radius.
+We can also set limit values in magnitude (for both minimum and maximum)
+to discard stars outside these limits.
 Another setting is the flag to enable or disable the \emph{Q-Filter}.
@@
 
-Clusterix returns a position map for the downloaded stars from Gaia DR2 database using Aladin \cite{bonnarel2000aladin} as an atlas of the sky
-(See Figure \ref{fig:aladin_melotte_22}).
-At this point, we can choose sending the recovered data to TOPCAT or Aladin, or continue to the next step of this assistant.
+Clusterix returns a position map for the downloaded stars from Gaia DR2 database
+using Aladin \cite{bonnarel2000aladin} as an atlas of the sky (See Figure \ref{fig:aladin_melotte_22}).
+At this point, we can choose sending the recovered data to TOPCAT or Aladin,
+or continue to the next step of the assistant.
 
@@
   \item \textbf{External region} or field region (\emph{f}), with stars that do not belong the open cluster.
-  \item \textbf{Void region}. This region is excluded from the analysis since it is supposed to be a transition region with stars that could belong the OC.
+  \item \textbf{Void region}. This region is excluded from the analysis since it is supposed
+                to be a transition region with stars that could belong the OC.
 \end{itemize}
 
-It is necessary to remark the importance of this step, since the selection of these regions is critical to get a good result.
-However, despite how crucial is this selection, there is no general rule to define these regions.
-We only have two references: the radius of the downloaded region and the estimated size for the open cluster that sets the \emph{inner radius}.
-The other two radius must be set by trial and error. In general, they should be large enough to be sure we are not taking polluted regions.
-This is a tedious process that requires several tries before achieving good enough probability values.
-
-Furthermore, it is necessary to know in advance some additional profiles of the cluster, in addition to the estimated size.
-For example, in the case of Melotte 22, the proper motion in declination direction of the membership stars around $-45 mas \cdot year^{-1}$,
+It is necessary to remark the importance of this step,
+since the selection of these regions is critical to get a good result.
+However, despite how crucial this step is, there is no general rule to define these regions.
+We only have two references:
+the radius of the downloaded region and the estimated size for the OC that sets the \emph{inner radius}.
+The other two radius must be set by trial and error.
+In general, they should be large enough to be sure that we are not taking polluted regions.
+This is a tedious process and requires several tries before achieving good enough probability values.
+
+Furthermore, it is necessary to know in advance some additional profiles of the cluster,
+in addition to the estimated size.
+For example, in the case of Melotte 22, the proper motion in declination direction
+of the membership stars around $-45 mas \cdot year^{-1}$,
 so it is necessary to open the $\mu$ range above the estimated values (in absolute value).
-If we do not do that we lose the possibility of detection, since Clusterix sets this value to $30 mas \cdot year^{-1}$ by default.
-But even in this case, it is necessary to set a threshold for magnitude values, $G \leq 14$, to help getting the right result.
+If we do not do that we lose the possibility of detection,
+since Clusterix sets this value to $30 mas \cdot year^{-1}$ by default.
+But even in this case, it is necessary to set a threshold for magnitude values,
+$G \leq 14$, to help getting the right result.
 Figure \ref{fig:clusterix_step2_melotte_22} shows the region selection panel on Clusterix 2.0.
 
-This makes the selection process an art rather than a science, not to mention the required previous knowledge on certain profiles of the
-objects to characterize. This puts into question the non-parametric character of the method.
+This makes the selection process an art rather than a science,
+not to mention the required previous knowledge on certain profiles of the objects to characterize.
+This puts into question the \emph{non-parameterized} character of the validation method.
 
@@
 
-After having set all these parameters, Clusterix performs the proper motion frequency analysis and returns the probability field
+After having set all these parameters,
+Clusterix performs the proper motion frequency analysis and returns the probability field
 associated to each star of the whole region (see Figure \ref{fig:clusterix_probability_melotte_22}).
 
-This step can be repeated over and over again, varying radii and other parameters, until finally reach a valid probability that works.
+This step can be repeated over and over again, varying radii and other parameters,
+until finally reach a valid probability that works.
 The last step presents a table with candidate objects and allows to export this dataset to TOPCAT,
@@
 Once data is available in TOPCAT, we have to carefully analyze the imported data.
-From there, we can perform a first star selection by excluding those stars with probability value lower than 0.8.
+At this point, we can perform a first stars selection by excluding those stars with probability value lower than 0.8.
 This first result can be taken as a good starting point, but we still need to make some fine tunning.
-As we can see in Figure \ref{fig:topcat_prob_selection_melotte_22}, the selection is too restrictive and we have lost too many stars.
-
-Looking at Figure \ref{fig:topcat_prob_parallax_melotte_22}, we can assume that cluster members are around $7.3mas$ in the parallax histogram.
-Also, Figure \ref{fig:topcat_prob_pm_melotte_22} reveals that cluster stars have low scattering rate in the proper motion configuration space.
+As we can see in Figure \ref{fig:topcat_prob_selection_melotte_22},
+the selection is too restrictive and we have lost too many stars.
+
+Looking at Figure \ref{fig:topcat_prob_parallax_melotte_22},
+we can assume that cluster members are around $7.3mas$ in the parallax histogram.
+Also, Figure \ref{fig:topcat_prob_pm_melotte_22} reveals that cluster stars
+have low scattering rate in the proper motion configuration space.
 
@@
 
-Taken these observations into account, we can proceed to make a new selection based on parallax properties.
-Our second selection makes use of the values obtained with the previous selection for parallax and its standard deviation and
-selects those stars from the whole set which fall inside the limits $7.304mas \pm 0.797$ (see Figure \ref{fig:topcat_parallax_melotte_22}).
+Taking these observations into account,
+we can proceed to make a new selection based on parallax properties.
+Our second selection makes use of the values obtained with the previous selection
+for parallax and its standard deviation and selects those stars from the whole set
+which fall inside the limits $7.304mas \pm 0.797$ (see Figure \ref{fig:topcat_parallax_melotte_22}).
 
@@
 Looking at the H-R diagrams (Figure \ref{fig:clusterix_selection_hr_diagrams_melotte_22}),
-we can compare both selections and see how we have extended the selection without getting too much noise around the isochrone curve.
+we can compare both selections and see how we have extended the selection without
+getting too much noise around the isochrone curve.
 
@@
 
-Finally, we filter the second selection by discarding those stars whose parallax relative error in absolute value is behind 0.05
-(see Figure \ref{fig:clusterix_final_selection_melotte_22}).
+Finally, we filter the second selection by discarding those stars whose parallax relative error in absolute value
+is behind 0.05 (see Figure \ref{fig:clusterix_final_selection_melotte_22}).
 
@@
 
-For the sake of completeness, Table \ref{tab:results_melotte_22} is shown again,
-in order to compare the results obtained with the validation method and our method.
+For the sake of completeness, Table \ref{tab:results_melotte_22} is shown again.
+This table summarizes the results obtained with the validation technique and our method.
 
@@
 
-All results have been computed with an Apple Mac Pro Late 2013, 2.7 GHz 12-Core Intel Xeon E5 with two AMD FirePro D700 6GB and 64 GB
+All results have been computed with an Apple Mac Pro Late 2013,
+2.7 GHz 12-Core Intel Xeon E5 with two AMD FirePro D700 6GB and 64 GB
 of memory, running macOS BigSur 11.1, Keras 2.2 and PlaidML 0.7.
 
-The majority of the runnings have taken a few minutes, while big clusters such as Melotte 25, have taken around an hour to perform
-the computations. This contrasts with \citeA{castro2020hunting} work, that used the power of the MareNostrum 42 supercomputer in order
-to perform its calculations.
+The majority of the runnings have taken a few minutes, while big clusters such as Melotte 25,
+have taken around an hour to perform the computations.
+This contrasts with \citeA{castro2020hunting} work,
+that used the power of the MareNostrum 42 supercomputer in order to perform its computations.
 
@@
       \hline
-      Number of Clusters & 11 \\
+      Number of Clusters & 9 \\
       Clustering Layer & $\left[ 50, 50, 80 \right]$ \\
@@
       Clusterix+TOPCAT & 106.796 $\pm$ 6.229 & -24.870 $\pm$ 5.417 & 21.210 $\pm$ 1.115 & 109 \\
-      KMeans & 81.410 $\pm$ 71.475 & -45.309 $\pm$ 73.945 & 21.056 $\pm$ 6.018 & 365 \\
-      DEC & x $\pm$ x & x $\pm$ x & x $\pm$ x & x \\
-      \textbf{DEC (filtered)} & x $\pm$ x & x $\pm$ x & x $\pm$ x & x \\
+      KMeans & 79.380 $\pm$ 71.691 & -45.314 $\pm$ 73.366 & 20.866 $\pm$ 6.029 & 376 \\
+      DEC & 76.463 $\pm$ 72.023 & -43.941 $\pm$ 72.556 & 20.489 $\pm$ 6.138 & 396 \\
+      \textbf{DEC (filtered)} & 81.795 $\pm$ 65.626 & -40.049 $\pm$ 55.112 & 19.888 $\pm$ 2.251 & 316 \\
     \end{tabular}
@@
 \begin{itemize}
-  \item NGC 2516 has it proper motion center deviated from the origin but it is embedded inside a big cloud of stars with similar proper motions.
-  \item NGC 2632 (in addition to Melotte 22) ia s cluster whose proper motion center is not located at the origin and has a well separated parallax center.
+  \item NGC 2516 has it proper motion center deviated from the origin but
+        it is embedded inside a big cloud of stars with similar proper motions.
+  \item NGC 2632 (in addition to Melotte 22) ia s cluster whose proper motion center
+        is not located at the origin and has a well separated parallax center.
   \item NGC 2682, on the other hand, has its parallax centered within the region's gaussian,
@@
 
-All of these clusters have a wide variety of diameters, from 25 to 330 arcmin, as well as the number of stars that belong to them.
+All of these clusters have a wide variety of diameters, from 25 to 330 arcmin,
+as well as the number of stars that belong to them.
 NGC 2682 has 3,000 stars while Melotte 25 is composed by more than 400,000 stars.
-
-% TODO: Meter una tabla con los parámetros de los clusters mencionados
-
-% TODO: Determinar correctamente los errores en las medidas y mencionarlos
-
-In all cases, the model has resolved properly the identity of the cluster and has characterized the membership stars,
-giving results compatible with the ones obtained with classic procedures and VO tools.
-
-Compared with the mentioned tools, our model can be fully calified as a non-parameterized model since it does not depend on parameters referred
-to the clusters itself but only hyperparameters such as the initial number of clusters to find, or the structure of the Clustering Layer used by DEC model.
-
-Our model is also non-supervised since we do not need to tell the model which stars belong to the cluster or assit it while is training.
-
-Only at the end of the characterization, a fine tunning driven by the user, can be applied in orther to improve the selection by discarding those
+More information about the studied clusters is available in Table \ref{tab:clusters_summary}.
+
+\begin{table}[h!]
+  \begin{center}
+    \begin{tabular}{l|c|c|c|c}
+      \textbf{Open Cluster} & \emph{$\alpha$ J2000 $(degrees)$} & \emph{$\delta$ J2000 $(degrees)$} & \emph{Radius $(arcmin)$} & \emph{\# stars} \\
+      \hline
+      NGC 2516 & 119.517 & -60.753 & 15 & 12,869 \\
+      NGC 2632 & 130.1 & 19.667 & 35 & 13,167 \\
+      NGC 2682 & 132.825 & 11.8 & 12.5 & 2,839 \\
+      Melotte 22 & 56.75 & 24.117 & 60 & 61,552 \\
+      Melotte 25 & 66.725 & 15.867 & 165 & 433,996 \\
+    \end{tabular}
+    \caption{Righ ascension, declination, radius and number of stars of studied clusters.
+             The number of stars corresponds to those stars contained within a cone of center
+             $(\alpha, \delta)$ and raius the cluster's radius multiplied by a factor of 1.5.}
+    \label{tab:clusters_summary}
+  \end{center}
+\end{table}
+
+In all cases, the model has resolved properly the identity of the cluster
+and has characterized the membership stars,
+giving compatible results with the ones obtained with classic procedures and VO tools.
+
+Compared with the mentioned tools, our model can be fully calified as a non-parameterized model
+since it does not depend on parameters referred to the cluster itself
+but only hyperparameters such as the initial number of clusters to be found,
+or the structure of the Clustering Layer used by DEC model.
+
+Our model is also non-supervised since we do not need to tell the model
+which stars belong to the cluster or assit it while its training stage.
+
+Only at the end of the characterization, a fine tuning driven by the user,
+can be applied in order to improve the selection by discarding those
 stars which fall outside the given quantiles.
 
-In any case we make assumptions about the cluster profile, something that we have to take into account when using VO tools.
+In any case, we do not make assumptions about the cluster profile,
+something that we have to take into account when using VO tools.
 This is another evidence that our model is non-parameterized.
+
+% TODO: Add references to figures and tables
 

### 2021-01-15 thesis/thesis.tex
thesis: Review Conclusions and Future Work

@@
 \section{Comparing and Validating}
+\label{sec:comparing_and_validating}
 
@@
 
-As shown by the obtained results, we can say that in general our model works fine identifying and characterizing open clusters.
-So we have succedded building a model non-parameterized non-supervised for open cluster characterization.
+From the results shown in Chapter \ref{sec:comparing_and_validating},
+we can say that, in general, our model works fine identifying and characterizing open clusters.
+So we have succedded building a model non-parameterized and non-supervised for open cluster characterization.
 However, some improvements should be done in order to improve its accuracy and precision.
 
-We have built a non-parameterized model since the user does not provide any information about the cluster profile or its properties.
-Although some tweaks can be done regarding the cluster, like changing the initial number of clusters or modifying the Clustering Layer structure,
+We have built a non-parameterized model since the user does not provide any information
+about the cluster profile or its properties.
+Although some tweaks can be done regarding the cluster,
+like changing the initial number of clusters or modifying the Clustering Layer structure,
 in order to improve the final characterization.
@@
 Another key point in our model is the initializer kernel.
-This kernel prepares data before passing it to the ANN and the results may vary significantly depending on this kernel.
+This kernel prepares data before passing it to the ANN
+and the results may vary significantly depending on this kernel.
 For example, simply varying the kernel seed can make a huge difference on the result.
-
-Even so, comparing our method with Clusterix, we can still say that the number of parameters (or hyperparameters) in our model is smaller
+This is something to avoid, so an improvement to solve this issue is necessary in order
+to have a reliable model which does not depend on the dataset order.
+
+Even so, comparing our method with Clusterix,
+we can still say that the number of parameters (or hyperparameters) in our model is smaller
 and we do not need previous knowledge about the studied cluster.
@@
 
-With all of this points in mind, we can affirm that this model could be really considered as an alternative model for open cluster characterization and thus,
-it could even be incloud as part of the Virtual Observatory toolset.
+With all of these points in mind and once the mentioned issues have been solved,
+we can assert this model could be considered as a valid model for open cluster characterization.
+And thus, it could even be included as part of the Virtual Observatory toolset.
 
@@
 That way we could have a better idea about the current limits of our model and we could make adjustments to improve it.
-We could determine with better accuracy the topologies that our model works well with, and which ones out model does not resolve properly.
-Or even, we could stablish different sets of hyperparameters regarding the topology of the cluster.
+We could determine with better accuracy the topologies that our model works well with,
+and which ones the model does not resolve properly.
+Even, we could stablish different sets of hyperparameters regarding the topology of the cluster.
 
@@
 Another possible change in our model would be using DBSCAN instead of KMeans as our initial clustering algorithm.
-Maybe this algorithm gives us a better starting point for the DEC model.
-
-At last, we have used during this work the Gaia DR2 database, but recently the DR3 dataset has been released.
+Maybe this algorithm gives us a better starting point for the DEC model which could improve the results.
+
+At last, we have used the Gaia DR2 database as our data source for this work,
+but recently the DR3 dataset has been released.
 So it is evident the interest on testing our model with this new data.

### 2021-01-15 thesis/thesis.tex
thesis: Fix errors estimation

@@
 Clusterix+TOPCAT method, and the results obtained for each step of this method (KMeans, DEC and DEC filtered).
+The errors associated to KMeans and DEC predictions are calculated as the Standard Error $SE = \sigma / \sqrt{N}$,
+where $N$ is the number of samples in the selected group.
 
@@
       Clusterix+TOPCAT & 19.98 $\pm$ 1.25 & -45.47 $\pm$ 1.48 & 7.33 $\pm$ 0.21 & 634 \\
-      KMeans & 20.25 $\pm$ 35.16 & -38.01 $\pm$ 40.16 & 7.23 $\pm$ 2.35 & 1378 \\
-      DEC & 23.67 $\pm$ 38.33 & -46.23 $\pm$ 44.54 & 8.04 $\pm$ 2.59 & 878 \\
-      \textbf{DEC (filtered)} & 19.50 $\pm$ 8.59 & -44.23 $\pm$ 8.17 & 7.42 $\pm$ 0.1 & 438 \\
+      KMeans & 20.25 $\pm$ 0.95 & -38.01 $\pm$ 1.08 & 7.23 $\pm$ 0.06 & 1378 \\
+      DEC & 23.67 $\pm$ 1.29 & -46.23 $\pm$ 1.50 & 8.04 $\pm$ 0.09 & 878 \\
+      \textbf{DEC (filtered)} & 19.50 $\pm$ 0.41 & -44.23 $\pm$ 0.39 & 7.42 $\pm$ 0.005 & 438 \\
     \end{tabular}
@@
       \textbf{Clusterix+TOPCAT} & 19.98 $\pm$ 1.25 & -45.47 $\pm$ 1.48 & 7.33 $\pm$ 0.21 & 634 \\
-      KMeans & 20.25 $\pm$ 35.16 & -38.01 $\pm$ 40.16 & 7.23 $\pm$ 2.35 & 1378 \\
-      DEC & 23.67 $\pm$ 38.33 & -46.23 $\pm$ 44.54 & 8.04 $\pm$ 2.59 & 878 \\
-      DEC (filtered) & 19.50 $\pm$ 8.59 & -44.23 $\pm$ 8.17 & 7.42 $\pm$ 0.1 & 438 \\
+      KMeans & 20.25 $\pm$ 0.95 & -38.01 $\pm$ 1.08 & 7.23 $\pm$ 0.06 & 1378 \\
+      DEC & 23.67 $\pm$ 1.29 & -46.23 $\pm$ 1.50 & 8.04 $\pm$ 0.09 & 878 \\
+      DEC (filtered) & 19.50 $\pm$ 0.41 & -44.23 $\pm$ 0.39 & 7.42 $\pm$ 0.005 & 438 \\
     \end{tabular}
@@
       Clusterix+TOPCAT & -4.652 $\pm$ 0.523 & 11.203 $\pm$ 0.454 & 2.409 $\pm$ 0.127 & 638 \\
-      KMeans & -4.344 $\pm$ 5.670 & 9.507 $\pm$ 7.546 & 2.268 $\pm$ 0.288 & 1542 \\
-      DEC & -4.426 $\pm$ 6.529 & 9.952 $\pm$ 8.018 & 2.436 $\pm$ 0.358 & 1532 \\
-      \textbf{DEC (filtered)} & -4.502 $\pm$ 4.46 & 10.114 $\pm$ 5.719 & 2.392 $\pm$ 0.135 & 1072 \\
+      KMeans & -4.344 $\pm$ 0.14 & 9.507 $\pm$ 0.19 & 2.268 $\pm$ 0.01 & 1542 \\
+      DEC & -4.426 $\pm$ 0.17 & 9.952 $\pm$ 0.20 & 2.436 $\pm$ 0.01 & 1532 \\
+      \textbf{DEC (filtered)} & -4.502 $\pm$ 0.14 & 10.114 $\pm$ 0.17 & 2.392 $\pm$ 0.004 & 1072 \\
     \end{tabular}
@@
       Clusterix+TOPCAT & -36.154 $\pm$ 1.001 & -12.909 $\pm$ 0.806 & 5.327 $\pm$ 0.187 & 371 \\
-      KMeans & -26.352 $\pm$ 20.562 & -15.828 $\pm$ 19.152 & 5.394 $\pm$ 0.696 & 629 \\
-      DEC & -20.012 $\pm$ 20.657 & -14.742 $\pm$ 17.330 & 4.686 $\pm$ 0.855 & 894 \\
-      \textbf{DEC (filtered)} & -21.571 $\pm$ 19.825 & -14.234 $\pm$ 16.388 & 4.719 $\pm$ 0.706 & 714 \\
+      KMeans & -26.352 $\pm$ 0.82 & -15.828 $\pm$ 0.76 & 5.394 $\pm$ 0.03 & 629 \\
+      DEC & -20.012 $\pm$ 0.69 & -14.742 $\pm$ 0.58 & 4.686 $\pm$ 0.03 & 894 \\
+      \textbf{DEC (filtered)} & -21.571 $\pm$ 0.74 & -14.234 $\pm$ 0.61 & 4.719 $\pm$ 0.03 & 714 \\
     \end{tabular}
@@
       Clusterix+TOPCAT & -10.970 $\pm$ 0.322 & -2.958 $\pm$ 0.327 & 1.142 $\pm$ 0.080 & 649 \\
-      KMeans & -8.616 $\pm$ 5.591 & -3.710 $\pm$ 6.067 & 1.196 $\pm$ 0.226 & 1374 \\
-      DEC & -8.926 $\pm$ 5.278 & -3.550 $\pm$ 5.425 & 1.144 $\pm$ 0.159 & 1238 \\
-      \textbf{DEC (filtered)} & -9.619 $\pm$ 4.112 & -3.317 $\pm$ 3.976 & 1.140 $\pm$ 0.089 & 990 \\
+      KMeans & -8.616 $\pm$ 0.15 & -3.710 $\pm$ 0.16 & 1.196 $\pm$ 0.01 & 1374 \\
+      DEC & -8.926 $\pm$ 0.15 & -3.550 $\pm$ 0.15 & 1.144 $\pm$ 0.005 & 1238 \\
+      \textbf{DEC (filtered)} & -9.619 $\pm$ 0.13 & -3.317 $\pm$ 0.13 & 1.140 $\pm$ 0.003 & 990 \\
     \end{tabular}
@@
       Clusterix+TOPCAT & 106.796 $\pm$ 6.229 & -24.870 $\pm$ 5.417 & 21.210 $\pm$ 1.115 & 109 \\
-      KMeans & 79.380 $\pm$ 71.691 & -45.314 $\pm$ 73.366 & 20.866 $\pm$ 6.029 & 376 \\
-      DEC & 76.463 $\pm$ 72.023 & -43.941 $\pm$ 72.556 & 20.489 $\pm$ 6.138 & 396 \\
-      \textbf{DEC (filtered)} & 81.795 $\pm$ 65.626 & -40.049 $\pm$ 55.112 & 19.888 $\pm$ 2.251 & 316 \\
+      KMeans & 79.380 $\pm$ 3.7 & -45.314 $\pm$ 3.8 & 20.866 $\pm$ 0.31 & 376 \\
+      DEC & 76.463 $\pm$ 3.6 & -43.941 $\pm$ 3.6 & 20.489 $\pm$ 0.31 & 396 \\
+      \textbf{DEC (filtered)} & 81.795 $\pm$ 3.7 & -40.049 $\pm$ 3.1 & 19.888 $\pm$ 0.13 & 316 \\
     \end{tabular}

### 2021-01-17 thesis/thesis.tex
thesis: Review Acknowledgements

@@
 \chapter{Acknowledgements}
+
+I would like to thank my thesis advisor Dr. César Augusto Guzmán Álvarez.
+He has guided me through the process of writing this work and has reviewed
+the whole work suggesting me improvements.
+
+I would also like to acknowledge my father who always gives me its advices and
+helps me by giving me its priceless point of view. I would also like to thank
+my mother and girlfriend, who have supported me through the entire process
+of making and writing this project.
+
+Finally, I would like to mention the acknowledgments that the developers
+of the main tools used in this project politely indicate on their web pages.
+
+\medskip
 

### 2021-01-17 thesis/thesis.tex
thesis: Review Results

@@
 Finally, we can refine this selection by filtering those stars which are bellow and above
-the 0.25 and 0.75 quantiles for each group, respectively.
+the 0.10 and 0.90 quantiles for each group, respectively.
 That way we remove the most doubtful values from the selection.
@@
 Clusterix+TOPCAT method, and the results obtained for each step of this method (KMeans, DEC and DEC filtered).
-The errors associated to KMeans and DEC predictions are calculated as the Standard Error $SE = \sigma / \sqrt{N}$,
-where $N$ is the number of samples in the selected group.
-
-\begin{table}[h!]
+Errors associated to KMeans and DEC predictions have been calculated as the \emph{Standard Error},
+$SE = \sigma / \sqrt{N}$, where $N$ is the number of samples in the selected group.
+
+\vfill
+\begin{table}[h]
   \begin{center}
@@
 \end{table}
+\vfill
 
@@
   \caption{NGC 2516 Clusterix+TOPCAT characterization.}
+  \label{fig:result_ngc_2516_clusterix}
   \centering
@@
   \end{subfigure}
-  \caption{NGC 2516 KMeans characterization.}
+  \caption{NGC 2516 KMeans characterization. Identified as cluster \emph{g3}.}
+  \label{fig:result_ngc_2516_kmeans}
   \centering
@@
   \end{subfigure}
-  \caption{NGC 2516 DEC characterization.}
+  \caption{NGC 2516 DEC characterization. Identified as cluster \emph{g3}.}
+  \label{fig:result_ngc_2516_dec}
   \centering
@@
   \end{subfigure}
-  \caption{NGC 2516 DEC (filtered) characterization.}
+  \caption{NGC 2516 DEC (filtered) characterization. Identified as cluster \emph{g3}.}
+  \label{fig:result_ngc_2516_dec_filtered}
 \end{figure}
@@
 
+Figure \ref{fig:result_ngc_2516_clusterix} shows NGC 2516 characterized using Clusterix+TOPCAT tools.
+Figure \ref{fig:result_ngc_2516_kmeans} shows eigth clusters identified by KMeans.
+In this case, cluster \emph{g3} is the one we are looking for.
+Figures \ref{fig:result_ngc_2516_dec} and \ref{fig:result_ngc_2516_dec_filtered}
+show the groups found using the DEC model and the DEC model filtered, respectively.
+Again, the cluster of interest is the group \emph{g3}.
+Although in general, groups between KMeans model and DEC models may not match.
+
+Table \ref{tab:hyperparameters_ngc_2516} shows the hyperparameters used for characterizing NGC 2516
+with the DEC model. Table \ref{tab:results_ngc_2516} shows a results summary for NGC 2516.
+
 \vfill
 
-\begin{table}[h!]
+\begin{table}[h]
   \begin{center}
@@
 
-\begin{table}[h!]
+\vfill
+
+\begin{table}[h]
   \begin{center}
@@
   \caption{NGC 2632 Clusterix+TOPCAT characterization.}
+  \label{fig:result_ngc_2632_clusterix}
   \centering
@@
   \end{subfigure}
-  \caption{NGC 2632 KMeans characterization.}
+  \caption{NGC 2632 KMeans characterization. Identified as cluster \emph{g1}.}
+  \label{fig:result_ngc_2632_kmeans}
   \centering
@@
   \end{subfigure}
-  \caption{NGC 2632 DEC characterization.}
+  \caption{NGC 2632 DEC characterization. Identified as cluster \emph{g1}.}
+  \label{fig:result_ngc_2632_dec}
   \centering
@@
   \end{subfigure}
-  \caption{NGC 2632 DEC (filtered) characterization.}
+  \caption{NGC 2632 DEC (filtered) characterization. Identified as cluster \emph{g1}.}
+  \label{fig:result_ngc_2632_dec_filtered}
 \end{figure}
@@
 
+Figure \ref{fig:result_ngc_2632_clusterix} shows NGC 2632 characterized by the validation method.
+Figure \ref{fig:result_ngc_2632_kmeans} shows ten clusters identified by KMeans.
+Cluster \emph{g1} is the open cluster we are looking for.
+Figures \ref{fig:result_ngc_2632_dec} and \ref{fig:result_ngc_2632_dec_filtered}
+show the groups found using the DEC model and the DEC model filtered, respectively.
+Open cluster NGC 2632 is labeled as group \emph{g1}.
+
+Table \ref{tab:hyperparameters_ngc_2632} shows the hyperparameters used for characterizing NGC 2632
+with our model. Table \ref{tab:results_ngc_2632} shows a results summary for NGC 2632 analysis.
+
 \vfill
 
-\begin{table}[h!]
+\begin{table}[h]
   \begin{center}
@@
   \caption{NGC 2682 Clusterix+TOPCAT characterization.}
+  \label{fig:result_ngc_2682_clusterix}
   \centering
@@
   \end{subfigure}
-  \caption{NGC 2682 KMeans characterization.}
+  \caption{NGC 2682 KMeans characterization. Identified as cluster \emph{g4}.}
+  \label{fig:result_ngc_2682_kmeans}
   \centering
@@
   \end{subfigure}
-  \caption{NGC 2682 DEC characterization.}
+  \caption{NGC 2682 DEC characterization. Identified as cluster \emph{g2}.}
+  \label{fig:result_ngc_2682_dec}
   \centering
@@
   \end{subfigure}
-  \caption{NGC 2682 DEC (filtered) characterization.}
+  \caption{NGC 2682 DEC (filtered) characterization. Identified as cluster \emph{g2}.}
+  \label{fig:result_ngc_2682_dec_filtered}
 \end{figure}
@@
 
+Figure \ref{fig:result_ngc_2682_clusterix} shows NGC 2682 characterized using Clusterix+TOPCAT tools.
+Figure \ref{fig:result_ngc_2682_kmeans} shows ten clusters identified by KMeans.
+KMeans labels NGC 2682 as group \emph{g4}.
+Figures \ref{fig:result_ngc_2682_dec} and \ref{fig:result_ngc_2682_dec_filtered}
+show the groups found using the DEC model and the DEC model filtered, respectively.
+The cluster of interest is the group \emph{g2}.
+
+Table \ref{tab:hyperparameters_ngc_2682} shows the hyperparameters used for characterizing NGC 2682
+with the DEC model. Table \ref{tab:results_ngc_2682} shows a results summary for NGC 2682.
+
 \vfill
 
-\begin{table}[h!]
+\begin{table}[h]
   \begin{center}
@@
   \caption{Melotte 25 Clusterix+TOPCAT characterization.}
+  \label{fig:result_melotte_25_clusterix}
   \centering
@@
   \end{subfigure}
-  \caption{Melotte 25 KMeans characterization.}
+  \caption{Melotte 25 KMeans characterization. Identified as cluster \emph{g3}.}
+  \label{fig:result_melotte_25_kmeans}
   \centering
@@
   \end{subfigure}
-  \caption{Melotte 25 DEC characterization.}
+  \caption{Melotte 25 DEC characterization. Identified as cluster \emph{g5}.}
+  \label{fig:result_melotte_25_dec}
   \centering
@@
   \end{subfigure}
-  \caption{Melotte 25 DEC (filtered) characterization.}
+  \caption{Melotte 25 DEC (filtered) characterization. Identified as cluster \emph{g5}.}
+  \label{fig:result_melotte_25_dec_filtered}
 \end{figure}
@@
 
+Figure \ref{fig:result_melotte_25_clusterix} shows Melotte 25 characterized using Clusterix+TOPCAT tools.
+Figure \ref{fig:result_melotte_25_kmeans} shows nine clusters identified by KMeans.
+KMeans labels Melotte 25 as group \emph{g3}.
+Figures \ref{fig:result_melotte_25_dec} and \ref{fig:result_melotte_25_dec_filtered}
+show the groups found using the DEC model and the DEC model filtered, respectively.
+Melotte 25 is labeled as group \emph{g5} by DEC model.
+
+Table \ref{tab:hyperparameters_melotte_25} shows the hyperparameters used for characterizing Melotte 25
+with the DEC model. Table \ref{tab:results_melotte_25} shows a results summary for Melotte 25.
+
 \vfill
 
-\begin{table}[h!]
+\begin{table}[h]
   \begin{center}
@@
 
-Compared with the mentioned tools, our model can be fully calified as a non-parameterized model
+Compared with the mentioned tools, our model can be classified as a non-parameterized model
 since it does not depend on parameters referred to the cluster itself
@@
 This is another evidence that our model is non-parameterized.
-
-% TODO: Add references to figures and tables
 

### 2021-01-21 thesis/thesis.tex
thesis: Improve Melotte 25 results

@@
   \end{subfigure}
-  \caption{Melotte 25 KMeans characterization. Identified as cluster \emph{g3}.}
+  \caption{Melotte 25 KMeans characterization. Identified as cluster \emph{g4}.}
   \label{fig:result_melotte_25_kmeans}
@@
       Number of Clusters & 9 \\
-      Clustering Layer & $\left[ 50, 50, 80 \right]$ \\
+      Clustering Layer & $\left[ 50, 50, 40 \right]$ \\
       Kernel Initializer Seed & 11 \\
@@
       Clusterix+TOPCAT & 106.796 $\pm$ 6.229 & -24.870 $\pm$ 5.417 & 21.210 $\pm$ 1.115 & 109 \\
-      KMeans & 79.380 $\pm$ 3.7 & -45.314 $\pm$ 3.8 & 20.866 $\pm$ 0.31 & 376 \\
-      DEC & 76.463 $\pm$ 3.6 & -43.941 $\pm$ 3.6 & 20.489 $\pm$ 0.31 & 396 \\
-      \textbf{DEC (filtered)} & 81.795 $\pm$ 3.7 & -40.049 $\pm$ 3.1 & 19.888 $\pm$ 0.13 & 316 \\
+      KMeans & 79.936 $\pm$ 3.7 & -45.362 $\pm$ 3.8 & 20.901 $\pm$ 0.31 & 374 \\
+      DEC & 104.051 $\pm$ 3.2 & -33.424 $\pm$ 2.5 & 22.072 $\pm$ 0.13 & 219 \\
+      \textbf{DEC (filtered)} & 105.96 $\pm$ 3.5 & -30.00 $\pm$ 2.4 & 21.74 $\pm$ 0.07 & 175 \\
     \end{tabular}

### 2021-01-26 paper/paper.tex
paper: Add title, abstract and key words

@@
-\documentclass[11pt,a4paper,english,twocolumn]{article}
+\documentclass[11pt,a4paper,USenglish,twocolumn]{article}
 \usepackage{unir_paper}
@@
 %---------------------------
-\title{Título}
-\author{Nombre y Apellidos del estudiante}
-\date{fecha}
+\title{Machine Learning Tools for Open Cluster Characterization with Gaia DR2 Data}
+\author{Carlos David Álvaro Yunta}
+\date{24th of January, 2021}
 
@@
 
-\resumen{Breve resumen del trabajo realizado (extensión máxima: 150 palabras).
-Este resumen debe incluir el objetivo o propósito de la investigación, la metodología, los resultados y las conclusiones.}
-\palabrasclave{tres a cinco palabras clave por orden alfabético y separadas por comas}
+\resumen{
+The characterization and understanding of \emph{Open Clusters}
+allow us to better understand properties and mechanisms about the Universe
+such as stellar formation and the kind of regions where these events occur.
+They also provide information about stellar processes and the evolution of the galactic disk.
+
+In this paper, we develop a method to characterize open clusters by using
+\emph{artificial intelligence} tools, such as clustering by \emph{K-Means}
+and clustering based on \emph{Artificial Neural Networks} by implementing
+the \emph{Deep Embedded Clustering} model.
+We are using \emph{Gaia DR2 database} as data source for testing our models.
+
+The developed method aims to improve current techniques of the state of the art.
+We achieve improvements not only in terms of \emph{computational efficiency},
+with lower computational requirements, but in \emph{usability},
+reducing the number of hyperparameters to obtain a good characterization of the analyzed clusters.
+
+Our method achieves good results,
+becoming even better in some cases when the results are compared with current techniques.
+}
+
+\palabrasclave{
+    characterization,
+    data analysis,
+    deep embedded clustering,
+    gaia,
+    machine learning,
+    open cluster
+}
 

### 2021-01-26 paper/paper.tex
paper: Update section titles

@@
 
-\section{INTRODUCCIÓN}
-Introducción en la que debes resumir de forma esquemática pero suficientemente clara lo esencial de cada una de las partes del trabajo.
-La lectura de esta introducción ha de dar una primera idea clara de lo que se pretendía, las conclusiones a las que se ha llegado y del procedimiento seguido.
+\section{INTRODUCTION}
 
-\section{ESTADO DEL ARTE}
+\section{STATE OF THE ART}
+
 Estudio a fondo el dominio de aplicación, citando numerosas referencias.
 Debe aportar un buen resumen del conocimiento que ya existe en el campo de los problemas habituales identificados.
-Numerar las citas de forma consecutiva entre corchetes \cite{Chen} y \cite{young}.
-\section{OBJETIVOS Y METODOLOGÍA}
+\section{AIMS AND METHOD}
+
 Objetivo general, objetivos específicos y metodología de trabajo aplicada.
 
-\section{CONTRIBUCIÓN}
-Desarrollar la descripción de tu contribución.
+\section{CONTRIBUTION}
 
-\section{RESULTADOS O EVALUACIÓN }
-Descripción de los resultados (Tipo 1.
-Piloto Experimental o Tipo 4.
-Comparativa de soluciones) o descripción de la evaluación y los resultados obtenidos (Tipo 2.
-Desarrollo de Software o Tipo 3.
-Metodología).
+\section{RESULTS}
 \subsection{Evaluación 1}
@@
 
-\section{DISCUSIÓN O ANÁLISIS DE RESULTADOS}
+\section{RESULTS DISCUSSION}
+
 Tras la presentación objetiva de los resultados, querrás aportar una discusión de los mismos.
 
-\section{CONCLUSIONES}
+\section{CONCLUSIONS}
+
 Resumen de las contribuciones del trabajo, en el que relaciones las contribuciones y los resultados obtenidos con los objetivos que habías planteado para el trabajo, discutiendo hasta qué punto has conseguido resolver los objetivos planteados.
@@
 \appendix
-\section{APÉNDICES}
+\section{APPENDIX}
 Apéndices, en caso de ser necesario.
 
+\renewcommand{\refname}{REFERENCES}
 \bibliographystyle{unsrt}

### 2021-01-26 paper/paper.tex
paper: Add Introduction section

@@
 
+Stellar open clusters (OCs) \cite{janes1982open} are groups of stars
+gravitationally bounded originated from a single molecular gas cloud.
+They share the same chemical composition and age. Moreover,
+they have similar relative positions and proper motion.
+Those astronomical objects are relevant to understand the spiral structure,
+the dynamics and the chemical evolution of our galaxy.
+
+The study of OCs has been pushed forward thanks to the huge and precise dataset
+from the Gaia mission Gaia DR2 available since 2018
+\cite{collaboration2016description} and \cite{gaia2018gaia}.
+This dataset has helped to review already known open clusters and to find new ones.
+
+Stars that belong to the same OC share relative positions,
+inherited from their original gas cloud.
+This means that their distances to the Earth are similar for all of them and,
+therefore, they have a narrow dispersion in their parallax value.
+They also share similar values of proper motion, both in right ascension and declination.
+Another property they share is their chemical composition.
+Thus, their metallicity must be uniform,
+since these stars were born from the same gas cloud and at the same time stage.
+However, to take this last property into account, we are faced with the drawback
+that this parameter is poorly reported in Gaia DR2 database.
+
+Normally, the characterization process takes into account astrometric features
+such as proper motion in right ascension and declination, and parallax.
+And photometric features like $E(b-r) / G_{mag}$.
+Photometric variables are used for generating the H-R diagram of the OC candidate stars.
+This stars should present a sharp profile corresponding to an isochrone curve derived
+from a theoretical model.
+This model does consider metallicity, mass and brightness of the stars involved.
+At the moment, a set of tools have been used for this purpose, as explained later,
+but in general, these tools require supervised and parameterized models.
+So a previous knowledge of the cluster is necessary or the process must be repeated
+iteratively in order to achieve valid results.
+
+Our aim is to build a machine learning model capable of characterizing open
+clusters inside a stellar region without previous knowledge of that region.
+This model will take advantage of these features to train an Artificial Neural
+Network (ANN) that will clusterize all stars within the given region.
+One of these groups will be the open cluster that interests us.
+With that purpose we have adapted the Unsupervised Deep Embedding for Clustering
+Analysis \cite{xie2016unsupervised} to our problem.
+
+This model fulfills one of our main requirements.
+It has to be \emph{non-supervised} and \emph{non-parameterized} in order to make easier
+the automation process of analyzing a wide range of regions with different topologies.
+Another important requirement for our model is that it must be computationally efficient
+in order to run in common workstations.
+The model has been developed with Python using the Keras framework which takes
+advantage of modern GPUs to make its computations.
+That increases significantly the computational capacity of our model making possible
+to run it on regular workstations with one of those GPUs.
+
+Since the aim of this work is not to find new OCs but to improve the characterization
+of the already known ones, we have taken the approach of analyzing known regions.
+These regions have been taken from the OpenClust catalogue \cite{dias2002new}
+(see Figure \ref{fig:openclust_catalogue}) and downloaded from Gaia DR2 database
+increasing the radius of each region by a factor $1.5$.
+
+\begin{figure}
+    \includegraphics[width=\columnwidth]{../figures/openclust_catalogue.png}
+    \caption{OpenClust Catalogue Distribution}
+    \label{fig:openclust_catalogue}
+\end{figure}
+
+Finally, we have compared the results obtained with the model we purposed with another
+method that involves tools from the Virtual Observatory such as Clusterix \cite{balaguer2020clusterix}.
+
+As shown in Section \ref{sec:results}, the results obtained are promising.
+Although some improvements must be done in order to consider it as valid tool
+that could be part of the standard set of tools for open clusters characterization.
+
 \section{STATE OF THE ART}
@@
 \section{RESULTS}
+\label{sec:results}
+
 \subsection{Evaluación 1}

### 2021-01-26 thesis/thesis.tex
thesis: Sort key words

@@
 
-{\bf Key words.} open clusters characterization, machine learning, gaia, data analysis, deep embedded clustering
+{\bf Key words.} data analysis, deep embedded clustering, gaia, machine learning, open clusters characterization
 
@@
 
-{\bf Palabras Clave:} caracterización de cúmulos abiertos, inteligencia artificial, gaia, análisis de datos
+{\bf Palabras Clave:} análisis de datos, caracterización de cúmulos abiertos, gaia, inteligencia artificial
 

### 2021-01-29 paper/paper.tex
paper: Add State of the Art

@@
 such as proper motion in right ascension and declination, and parallax.
-And photometric features like $E(b-r) / G_{mag}$.
-Photometric variables are used for generating the H-R diagram of the OC candidate stars.
+Photometric features like $E(b-r) / G_{mag}$ are considered too.
+These last variables are used for generating the H-R diagram of the OC candidate stars.
 This stars should present a sharp profile corresponding to an isochrone curve derived
@@
 
-Our aim is to build a machine learning model capable of characterizing open
-clusters inside a stellar region without previous knowledge of that region.
+Our aim is to build a machine learning model capable of characterizing an open
+cluster inside a stellar region with knowing nothing about the region.
 This model will take advantage of these features to train an Artificial Neural
@@
 It has to be \emph{non-supervised} and \emph{non-parameterized} in order to make easier
-the automation process of analyzing a wide range of regions with different topologies.
+the automation process of analyzing a wide range of regions with different typologies.
 Another important requirement for our model is that it must be computationally efficient
@@
 As shown in Section \ref{sec:results}, the results obtained are promising.
-Although some improvements must be done in order to consider it as valid tool
-that could be part of the standard set of tools for open clusters characterization.
+This allows us to consider our model as a valid tool that could be part
+of the standard set of tools for open clusters characterization.
 
@@
 
-Estudio a fondo el dominio de aplicación, citando numerosas referencias.
-Debe aportar un buen resumen del conocimiento que ya existe en el campo de los problemas habituales identificados.
+The initial approach to find open clusters is by searching for overdensities
+in the astrometric space of the galactic disk and a later identification of
+possible OCs using photometric information \cite{castro2020hunting}.
+However, although it looks simple, it is not. Here we face the problem that the
+near field around the OC is composed of two kind of populations.
+The first kind belongs to the membership stars of the OC (from tens or hundreds
+to a few thousand) while the second one is composed by a background of stars
+that do not belong to the OC (from tens to hundreds of thousands).
+To find out which stars do belong to the OC is the problem we try to solve in
+this work as this selection is critical to properly characterize the
+fundamental properties of the cluster (total mass, age, chemical composition, etc.).
+
+Since the components of the open cluster are born from the same gas cloud and they
+share both dynamic properties and relative distance, as well as age and chemical
+composition, the first step is to analyze the proper motion configuration space of
+the region of interest. This approach is the one used by many tools from the Virtual
+Observatory such as TOPCAT \cite{taylor2005topcat}, Clusterix 2.0 \cite{balaguer2020clusterix},
+Aladin \cite{bonnarel2000aladin} or VOSA \cite{bayo2008vosa}.
+
+TOPCAT by itself cannot find members of the open cluster. It requires to parameterize
+some of the properties of the cluster, what requires a previous knowledge of the cluster.
+Or to make a supervised and manual selection of groups based on overdensities
+in the proper motion configuration space. Clusterix 2.0, an interactive web-based tool,
+can help to this process. It takes the proper motion diagram without making any prior
+assumption about the membership of the candidate star and determines empirically
+the frequency functions. Clusterix uses normal Gaussian kernel functions, defined as:
+
+\begin{equation*}
+  K(a, b) = \frac{1}{2 \pi h^{2}} \exp{ \left[ - \frac{1}{2}\frac{\left( a - a_{i} \right)^{2} + \left( b - b_{i} \right)^{2}}{ h^{2}} \right]}
+\end{equation*}
+
+$\left( a, b \right)$ are referred to the proper motion configuration space,
+while a point located in the center of the cell $\left( i, j \right)$
+provides the maximum contribution for calculating the local density.
+Number $h$ is called the \emph{smoothing} parameter and it is measured
+in the same units as proper motion.
+
+Clusterix is defined as an unsupervised and non-parameterized tool, but it
+critically depends on the selection of three radii.
+
+\begin{itemize}
+  \item \textbf{Inner radius} which defines a region with stars that belong to the OC and field stars.
+  \item \textbf{Void region}. This region is excluded from the analysis since it is supposed
+                to be a transition region with stars that could belong the OC.
+  \item \textbf{External region} or field region.
+                This is the outer region with stars that do not belong the open cluster.
+\end{itemize}
+
+If the analysis performed by Clusterix is successful, it returns the probability
+associated to each star in the dataset to belong to the open cluster.
+These results can be imported then into TOPCAT to continue the refining process to
+get a valid characterization.
+
+Another recent approach to solve this problem based on machine learning techniques
+makes a systematic search for overdensities in the astrometric space of the galactic
+disk and a subsequent identification of open clusters using photometric information
+\cite{castro2020hunting}.
+
+This last method includes two phases: the first one uses an unsupervised clustering
+algorithm, DBSCAN, to search for overdensities $(l, b \pi, \mu_{\alpha} *, \mu_{\delta})$,
+and then applies a deep learning ANN, previously trained with magnitude diagrams,
+to identify isochrone patterns within the detected overdensities and thus proceed
+to confirm them as OC.
+
+It should be noted that for the execution of this method, MareNostrum 42
+(Barcelona Supercomputing Center) was used.
+So the neural network could handle the image recognition process with
+isochrone patterns and not applying theoretical models derived from
+values such as metallicity or masses, among other.
+
+This work concludes with the identification of 582 new open clusters
+distributed along the galactic disk for a galactic declination bellow 20 degrees.
+This result increases the number of known open clusters by 45\%.
+
+As told before, our aim in this work, linked to computational limitations compared
+to MareNostrum, is not to do a blind search for clusters, but to obtain a method,
+in the machine learning environment, that allows the characterization of open clusters.
+We require our method to be unsupervised and non-parameterized making it suitable for
+automated processes.
+
+We use the Clusterix+TOPCAT method applied to the same clusters that we have analyzed
+with our method to validate our results.
+
+In this work we make use of Gaia DR2 since DR3 has not been released in time for us to include it.
+Gaia DR2 is a multidimensional dataset obtained by ESA's Gaia mission
+(located at L2, 1.5 million kilometers from Earth) and operational since 2014.
+The catalogue has high precision and accuracy astrometric data for more than 1.7 billion stellar sources,
+and magnitudes in three photometric filters (G, BP and RP) for more than 1,300 million sources.
+
+As we will explain later, what we need to accomplish our aim is an algorithm
+that manages to make data groups based on the dynamic properties of the stars.
+
+There are several clustering algorithms: KMeans, Mean-Shift Clustering, DBSCAN \cite{ester1996density}
+among other. While each one behaves better according to the distribution of the objects
+to clusterize, we have chosen KMeans by its simplicity and good results.
+
+This algorithm works well and gives good results at first approach.
+However, we need to set a large number of clusters to find the open cluster we are looking for.
+This fact complicates the identification of the OC and many times the found cluster contains to many outliers.
+For that reason, we have search for a KMeans refinement based on an artificial neural network.
+
+The \emph{Unsupervised Deep Embedding for Clustering Analysis} model
+or \emph{DEC} \cite{xie2016unsupervised} takes KMeans as its starting point,
+but then, it trains an autoencoder to reduce the feature space
+and pass this transformed data through a Clustering Layer which refines the previous selection.
+
 \section{AIMS AND METHOD}

### 2021-01-29 paper/paper.tex
paper: Update authors
paper: Add Aims and Method

@@
 \usepackage{unir_paper}
+\usepackage{subcaption}
 
@@
 \title{Machine Learning Tools for Open Cluster Characterization with Gaia DR2 Data}
-\author{Carlos David Álvaro Yunta}
+\author{C. D. Álvaro, J. Álvaro, C. A. Guzmán}
 \date{24th of January, 2021}
@@
 
-Objetivo general, objetivos específicos y metodología de trabajo aplicada.
+Our aim is to \emph{build an unsupervised clustering model for open cluster characterization}.
+The model will be \emph{non-supervised and non-parameterized} in order to fit a wide range
+of clusters without the need for fine-tuning a high number of hyperparameters.
+
+\subsection{DOWNLOAD}
+
+We start by selecting a region from the OpenClust catalogue and downloading it from
+Gaia DR2 database into a docker database. This allows us to repeat the analysis without
+having to download the data every time. The radius of the downloaded region from Gaia
+is 1.5 times larger than the one registered in OpenClust. This way we ensure to include
+several stars that do not belong to the open cluster.
+
+\subsection{FEATURE SELECTION}
+\label{sec:feature_selection}
+
+As mentioned before, we want our model to characterize open clusters by looking at
+their dynamic properties. Also, we want to maintain as simple as possible our
+clustering model in order to save computing resources.
+
+Proper motion in right ascension and declination seems like a natural choice since,
+as we know, stars belonging to the same OC share a common motion vector.
+
+Parallax is another important feature. It lets us know how far stars are from us.
+In addition, since all stars within an open cluster were born from the same dust cloud,
+they must all have similar parallax.
+
+However, we are not going to use these raw features.
+Instead, we are taking a combination of them.
+
+First, we correct proper motion in right ascension and declination by dividing them by
+the parallax. That way, we normalize these quantities and help our clustering models
+to improve their performance.
+
+The modulus of the proper motion is another computed property that we are considering.
+We use it to relate both features and therefore, to force our model to keep them tight.
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{../figures/melotte_22/features_melotte_22.png}
+  \caption{Pairwise relationships among variables using Melotte 22 data}
+  \label{fig:features_melotte_22}
+\end{figure}
+
+Figure \ref{fig:features_melotte_22} shows a pairwise relationship among some features
+available in the dataset for Melotte 22 data.
+
+\subsection{SOFT CLUSTERING}
+
+Our first approach to find the open cluster is using the K-Means algorithm.
+Since we are looking for a single cluster, it seems reasonable to use a clustering
+algorithm and set it to find two clusters. One for the desired OC and another which
+contains stars that do not belong to the open cluster.
+However, this idea is not completely right.
+
+\begin{figure}
+  \centering
+  \begin{subfigure}{\columnwidth}
+    \centering
+    \begin{subfigure}[t]{0.3\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n2_pm_melotte_22.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.3\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n5_pm_melotte_22.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.3\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n8_pm_melotte_22.png}
+    \end{subfigure}
+  \end{subfigure}
+  \medskip
+  \begin{subfigure}{\columnwidth}
+    \centering
+    \begin{subfigure}[t]{0.3\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n2_parallax_melotte_22.png}
+      \caption{N = 2}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.3\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n5_parallax_melotte_22.png}
+      \caption{N = 5}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.3\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n8_parallax_melotte_22.png}
+      \caption{N = 8}
+    \end{subfigure}
+  \end{subfigure}
+  \caption{K-Means comparisons with Melotte 22}
+  \label{fig:kmeans_comparisons_melotte_22}
+\end{figure}
+
+This is due to the fact that OC's stars are surrounded by other stars with possibly similar
+properties. So, setting the number of clusters to two is too low to separate them properly.
+
+As shown in Figure \ref{fig:kmeans_comparisons_melotte_22}, larger values for the number of
+clusters allow us to isolate more accurately the resonance in parallax at $\approx 7.3mas$.
+However, we have the disadvantage that more groups are formed.
+
+This effect complicates our task of finding the desired open cluster, since we would like
+to get just two groups, one for the OC and another with the remaining stars.
+Therefore, we have to find a way to set the right value for the number of clusters
+to isolate the searched cluster without creating too many groups.
+
+To solve this issue, we will try to estimate the best number of clusters by using the
+\emph{silhouette score} \cite{rousseeuw1987silhouettes}.
+
+K-Means does a good job making an initial clustering. However, too many clusters arise
+from this characterization and the OC is still polluted with stars that do not belong to it.
+Moreover, we would like to reduce the amount of clusters too.
+
+Therefore, we would like to find a better model that improves this initial characterization
+by reducing the amount of clusters and also that removes outsiders from the OC.
+
+\subsection{DEC MODEL}
+
+Since we do not have a labeled dataset to train a supervised model,
+we have no choice but to use an unsupervised self-trained model.
+
+For that reason we have adapted the \emph{Unsupervised Deep Embedding for Clustering Analysis}
+to our work.
+
+The model is composed by a \emph{deep autoencoder}, a \emph{deep encoder} and
+a \emph{clustering layer}.
+
+Encoders are used to transform the input data into a latent space
+using a non-linear mapping function $f_{\theta} : X \rightarrow Z$.
+
+Although, as explained in section \ref{sec:feature_selection},
+the number of features we are managing is not too large,
+this latent space helps us reduce the number of features
+and avoids the \emph{"curse of dimensionality"} \cite{bellman1961curse}.
+
+These encoders are pretrained before fitting the model to generate predictions. Then,
+the encoder is used with the aim of transforming input data to the latent space $Z$. Once the
+data has been transformed, a K-Means clusterer is used in order to make an initial clustering.
+
+With that initial configuration, the model iterates alternating between computing an auxiliary
+target distribution (Soft Assignment) and minimizing the Kullback-Leiber (KL) divergence
+\cite{kullback1951information} to it. This unsupervised algorithm allows us to improve the clustering.
+
+\begin{equation}
+  p_{ij} = \frac{q^{2}_{ij} / f_{j}}{\sum_{j'}q^{2}_{ij'}/f_{j'}}
+  \label{eq:student_tdistribution}
+\end{equation}
+
+In the soft assignment stage,
+the \emph{Student's t-distribution} is used as a kernel to measure the similarity
+between the embedded points and the cluster centroid.
+While in the KL divergence minimization the algorithm iteratively refines clusters by learning
+from their high confidence assignments with the help of an auxiliary target distribution.
+The model is trained by matching the soft assignment to the target distribution.
+The choice of this target distribution is crucial for DEC's performance.
+In this work we have taken the target distribution from DEC's original paper \cite{xie2016unsupervised},
+which is defined in Equation \ref{eq:student_tdistribution}.
+
+Figure \ref{fig:dec_model_setup} shows the layer setup of our DEC model.
+It is simpler than the one tested on the original paper \cite{xie2016unsupervised},
+since the number of selected features in our work is smaller than in the original one.
+Therefore, using the same configuration would result in a model so
+powerful that would incur in overfitting issues unable to make right predictions.
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{../figures/dec_diagram.pdf}
+  \caption{DEC model layer setup}
+  \label{fig:dec_model_setup}
+\end{figure}
+
+Finally, we can refine this selection by filtering those stars which are bellow and above
+the 0.10 and 0.90 quantiles for each group, respectively.
+That way we remove the most doubtful values from the selection.
+
+All developed resources are available at:
+\href{github.com/cdalvaro/machine-learning-master-thesis}{github.com/cdalvaro/machine-learning-master-thesis}
 

### 2021-01-30 paper/paper.tex
paper: Rearrange pictures

@@
 
+\begin{figure}[htbp]
+  \includegraphics[width=\columnwidth]{../figures/melotte_22/features_melotte_22.png}
+  \caption{Pairwise relationships among variables using Melotte 22 data}
+  \label{fig:features_melotte_22}
+\end{figure}
+
 As mentioned before, we want our model to characterize open clusters by looking at
@@
 
-\begin{figure}
-  \includegraphics[width=\columnwidth]{../figures/melotte_22/features_melotte_22.png}
-  \caption{Pairwise relationships among variables using Melotte 22 data}
-  \label{fig:features_melotte_22}
-\end{figure}
-
 Figure \ref{fig:features_melotte_22} shows a pairwise relationship among some features
@@
 
-Our first approach to find the open cluster is using the K-Means algorithm.
-Since we are looking for a single cluster, it seems reasonable to use a clustering
-algorithm and set it to find two clusters. One for the desired OC and another which
-contains stars that do not belong to the open cluster.
-However, this idea is not completely right.
-
-\begin{figure}
+\begin{figure}[htbp]
   \centering
@@
 \end{figure}
+
+Our first approach to find the open cluster is using the K-Means algorithm.
+Since we are looking for a single cluster, it seems reasonable to use a clustering
+algorithm and set it to find two clusters. One for the desired OC and another which
+contains stars that do not belong to the open cluster.
+However, this idea is not completely right.
 
@@
 
-All developed resources are available at:
-\href{github.com/cdalvaro/machine-learning-master-thesis}{github.com/cdalvaro/machine-learning-master-thesis}
+All resources developed for this project are available at the GitHub repository:
+\href{https://github.com/cdalvaro/machine-learning-master-thesis}{cdalvaro/machine-learning-master-thesis}
+
+It includes the DEC model developed in Python using the Keras framework
+and Jupyter Notebooks \cite{Kluyver2016jupyter} with examples.
 

### 2021-01-30 paper/paper.tex
paper: Add Contribution

@@
 Finally, we have compared the results obtained with the model we purposed with another
-method that involves tools from the Virtual Observatory such as Clusterix \cite{balaguer2020clusterix}.
+method that involves tools from the Virtual Observatory (VO) such as Clusterix \cite{balaguer2020clusterix}.
 
@@
 composition, the first step is to analyze the proper motion configuration space of
-the region of interest. This approach is the one used by many tools from the Virtual
-Observatory such as TOPCAT \cite{taylor2005topcat}, Clusterix 2.0 \cite{balaguer2020clusterix},
+the region of interest. This approach is the one used by many tools from the VO
+such as TOPCAT \cite{taylor2005topcat}, Clusterix 2.0 \cite{balaguer2020clusterix},
 Aladin \cite{bonnarel2000aladin} or VOSA \cite{bayo2008vosa}.
@@
 
+The developed method is successful since it meets all our established requirements:
+
+\begin{itemize}
+  \item Our method gives valid results compatible with current methods as described in Section \ref{sec:results}.
+  \item As required from the beginning, it is non-supervised and non-parameterized.
+  \item The model shows a great performance compared with current methods.
+  \item It is able to deal with large and dense regions. As an example,
+        it has succeeded characterizing Melotte 25 (Hiades) OC. This cluster is
+        really large and dense, and its relative proximity makes it difficult to
+        characterize. Clusterix fails characterizing this cluster. In this case,
+        our method proposes at least two more non-catalogued clusters. The analysis
+        of these new clusters brings out an interesting stellar dynamics in the studied
+        region. It could mean that there was a recent interaction among different molecular clouds.
+\end{itemize}
+
+For all this, we consider that the presented method is good enough to be included
+in the Virtual Observatory toolset. Only a few tweaks must be done in order to make
+our model compatible with other tools of the VO.
+
 \section{RESULTS}

### 2021-01-30 paper/paper.tex
paper: Review Abstract

@@
 allow us to better understand properties and mechanisms about the Universe
-such as stellar formation and the kind of regions where these events occur.
+such as stellar formation and the regions where these events occur.
 They also provide information about stellar processes and the evolution of the galactic disk.
@@
 In this paper, we develop a method to characterize open clusters by using
-\emph{artificial intelligence} tools, such as clustering by \emph{K-Means}
-and clustering based on \emph{Artificial Neural Networks} by implementing
+\emph{machine learning} tools, such as clustering by \emph{K-Means}
+and clustering based on \emph{Artificial Neural Networks} by adapting
 the \emph{Deep Embedded Clustering} model.
-We are using \emph{Gaia DR2 database} as data source for testing our models.
+We are using \emph{Gaia DR2 database} as data source.
 
@@
 We achieve improvements not only in terms of \emph{computational efficiency},
-with lower computational requirements, but in \emph{usability},
-reducing the number of hyperparameters to obtain a good characterization of the analyzed clusters.
+with lower computational requirements, but in \emph{usability}, reducing the number
+of hyperparameters to get a good characterization of the analyzed clusters.
 

### 2021-01-30 paper/paper.tex
paper: Add Conclusions

@@
 This model does consider metallicity, mass and brightness of the stars involved.
-At the moment, a set of tools have been used for this purpose, as explained later,
-but in general, these tools require supervised and parameterized models.
-So a previous knowledge of the cluster is necessary or the process must be repeated
-iteratively in order to achieve valid results.
+At the moment, a set of tools have been used for this purpose, as explained later.
+But in general, these tools require supervised and parameterized models.
+Therefore, a previous knowledge of the cluster is necessary or the process must be
+repeated iteratively in order to achieve valid results.
 
@@
 
-\begin{figure}[htbp]
+Our first approach to find the open cluster is using the K-Means algorithm.
+Since we are looking for a single cluster, it seems reasonable to use a clustering
+algorithm and set it to find two clusters. One for the desired OC and another which
+contains stars that do not belong to the open cluster.
+However, this idea is not completely right.
+
+This is due to the fact that OC's stars are surrounded by other stars with possibly similar
+properties. Therefore, setting the number of clusters to two is too low to separate them properly.
+
+\begin{figure}
   \centering
@@
 \end{figure}
-
-Our first approach to find the open cluster is using the K-Means algorithm.
-Since we are looking for a single cluster, it seems reasonable to use a clustering
-algorithm and set it to find two clusters. One for the desired OC and another which
-contains stars that do not belong to the open cluster.
-However, this idea is not completely right.
-
-This is due to the fact that OC's stars are surrounded by other stars with possibly similar
-properties. So, setting the number of clusters to two is too low to separate them properly.
 
@@
 
-Resumen de las contribuciones del trabajo, en el que relaciones las contribuciones y los resultados obtenidos con los objetivos que habías planteado para el trabajo, discutiendo hasta qué punto has conseguido resolver los objetivos planteados.
-Finalmente, hablar de líneas de trabajo futuro que podrían aportar valor añadido al TFM realizado.
-La sección debería señalar las perspectivas de futuro que abre el trabajo desarrollado para el campo de estudio definido.
-En el fondo, debes justificar de qué modo puede emplearse la aportación que has desarrollado y en qué campos.
+From the results shown in Section \ref{sec:results}, we can say that, in general,
+our model works fine identifying and characterizing open clusters.
+So we have succeeded building a model non-parameterized and non-supervised
+for open cluster characterization. However, some improvements should be done
+in order to improve its accuracy and precision.
 
-\appendix
-\section{APPENDIX}
-Apéndices, en caso de ser necesario.
+We have built a non-parameterized model since the user is not required to provide
+any information about the cluster profile or its properties.
+Nevertheless some tweaks can be done regarding the cluster, like changing the initial
+number of clusters or modifying the clustering layer structure in order to improve
+the final characterization.
+
+Another key point in our model is the initializer kernel.
+This kernel prepares data before passing it to the ANN
+and the results may vary significantly depending on this kernel.
+This is something to avoid, so an improvement to solve this issue is necessary
+in order to have a reliable model which does not depend on the dataset order.
+
+Despite that, comparing our method with Clusterix, we can still say that the
+number of parameters (or hyperparameters) in our model is smaller and that
+we do not need previous knowledge about the studied cluster.
+That makes easier to test different configurations and automate the process.
+
+\subsection{FURTHER RESEARCH}
+
+One open point to do from now is testing the model with a wider range of clusters.
+In general we could think of applying it to the whole VizieR catalogue.
+That way we could have a better idea about the current limitations of our model
+and we could make some adjustments to improve it. We could determine with higher
+accuracy the typologies that our model works well with, and which ones the model
+does not resolve properly. Furthermore, we could stablish different sets of
+hyperparameters regarding the typology of the cluster.
+
+With our model, new uncatalogued clusters arise from the data apart of the main ones.
+Hence, it is also interesting the individual study of these new clusters.
+
+Another possible change in our model would be using DBSCAN instead of KMeans
+as our initial clustering algorithm. Maybe this algorithm gives us a better starting
+point for the DEC model which could improve the results.
+
+Finally, we have used the Gaia DR2 database as our data source for this work,
+but recently the DR3 dataset has been released.
+Therefore, it is evident the interest on testing our model with this new data.
 

### 2021-01-30 thesis/thesis.tex
thesis: Add paper as appendix
thesis: Remove extra toc
thesis: Review spelling and grammar

@@
 
-We have applied the our proposed method to clusters with different topologies:
+We have applied our proposed method to clusters with different topologies:
 
@@
 
-Compared with the mentioned tools, our model can be classified as a non-parameterized model
-since it does not depend on parameters referred to the cluster itself
+Compared with the mentioned tools, our model can be cathegorizeed as a non-parameterized
+model since it does not depend on parameters referred to the cluster itself
 but only hyperparameters such as the initial number of clusters to be found,
@@
 Our model is also non-supervised since we do not need to tell the model
-which stars belong to the cluster or assit it while its training stage.
+which stars belong to the cluster or assist it while its training stage.
 
@@
 we can say that, in general, our model works fine identifying and characterizing open clusters.
-So we have succedded building a model non-parameterized and non-supervised for open cluster characterization.
+So we have succeeded building a model non-parameterized and non-supervised for open cluster characterization.
 However, some improvements should be done in order to improve its accuracy and precision.
@@
 about the cluster profile or its properties.
-Although some tweaks can be done regarding the cluster,
-like changing the initial number of clusters or modifying the Clustering Layer structure,
+Nevertheless some tweaks can be done regarding the cluster,
+like changing the initial number of clusters or modifying the clustering layer structure
 in order to improve the final characterization.
@@
 
-Even so, comparing our method with Clusterix,
+Despite that, comparing our method with Clusterix,
 we can still say that the number of parameters (or hyperparameters) in our model is smaller
-and we do not need previous knowledge about the studied cluster.
+and that we do not need previous knowledge about the studied cluster.
 That makes easier to test different configurations and automate the process.
@@
 That way we could have a better idea about the current limits of our model and we could make adjustments to improve it.
-We could determine with better accuracy the topologies that our model works well with,
+We could determine with higher accuracy the typologies that our model works well with,
 and which ones the model does not resolve properly.
-Even, we could stablish different sets of hyperparameters regarding the topology of the cluster.
+Furthermore, we could stablish different sets of hyperparameters regarding the typology of the cluster.
 
 With our model, new secondary clusters arise from the data apart of the main ones.
-So, it is also interesting the individual study of these new clusters.
+Hence, it is also interesting the individual study of these new clusters.
 
@@
 
-At last, we have used the Gaia DR2 database as our data source for this work,
+Finally, we have used the Gaia DR2 database as our data source for this work,
 but recently the DR3 dataset has been released.
-So it is evident the interest on testing our model with this new data.
+Therefore it is evident the interest on testing our model with this new data.
 
@@
 \bibliography{references}
-\addcontentsline{toc}{chapter}{Bibliography}
-
-\appendix
-
-%\includepdf[pages=-]{anexo.pdf} # TODO: Descomentar cuando el artículo esté hecho
+
+\includepdf[pages={-},fitpaper]{../paper/paper.pdf}
+
 \end{document}

### 2021-01-31 paper/paper.tex
paper: Improve Aims and Method section

@@
-\documentclass[11pt,a4paper,USenglish,twocolumn]{article}
+\documentclass[11pt,a4paper,english,twocolumn]{article}
 \usepackage{unir_paper}
 \usepackage{subcaption}
+\usepackage{tablefootnote}
 
@@
 \title{Machine Learning Tools for Open Cluster Characterization with Gaia DR2 Data}
-\author{C. D. Álvaro, J. Álvaro, C. A. Guzmán}
+\author{CD. Álvaro, J. Álvaro, CAG. Guzmán}
 \date{24th of January, 2021}
@@
 
-The developed method aims to improve current techniques of the state of the art.
+The developed method aims to improve classic techniques of the state of the art.
 We achieve improvements not only in terms of \emph{computational efficiency},
@@
 Clusterix is defined as an unsupervised and non-parameterized tool, but it
-critically depends on the selection of three radii.
-
-\begin{itemize}
-  \item \textbf{Inner radius} which defines a region with stars that belong to the OC and field stars.
-  \item \textbf{Void region}. This region is excluded from the analysis since it is supposed
-                to be a transition region with stars that could belong the OC.
-  \item \textbf{External region} or field region.
-                This is the outer region with stars that do not belong the open cluster.
-\end{itemize}
+critically depends on the selection of three radii in the studied region.
+The inner radius should contain stars belonging or not to the cluster.
+While the outermost radii define a ring which only contains stars that do
+not belong to the cluster. In addition, it is also sensitive to other
+parameters, such as the soften parameter \emph{h}, or other restrictions
+related with the searched proper motions.
 
@@
 
-There are several clustering algorithms: KMeans, Mean-Shift Clustering, DBSCAN \cite{ester1996density}
+There are several clustering algorithms: K-Means, Mean-Shift Clustering, DBSCAN \cite{ester1996density}
 among other. While each one behaves better according to the distribution of the objects
-to clusterize, we have chosen KMeans by its simplicity and good results.
+to clusterize, we have chosen K-Means by its simplicity and good results.
 
@@
 This fact complicates the identification of the OC and many times the found cluster contains to many outliers.
-For that reason, we have search for a KMeans refinement based on an artificial neural network.
+For that reason, we have search for a K-Means refinement based on an artificial neural network.
 
 The \emph{Unsupervised Deep Embedding for Clustering Analysis} model
-or \emph{DEC} \cite{xie2016unsupervised} takes KMeans as its starting point,
+or \emph{DEC} \cite{xie2016unsupervised} takes K-Means as its starting point,
 but then, it trains an autoencoder to reduce the feature space
@@
 
+\begin{figure}[htbp]
+  \centering
+  \begin{subfigure}{\columnwidth}
+    \centering
+    \begin{subfigure}[t]{0.3\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/melotte_22/kmeans_pm_melotte_22.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.3\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/melotte_22/dec_pm_melotte_22.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.3\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/melotte_22/dec_pm_filtered_melotte_22.png}
+    \end{subfigure}
+  \end{subfigure}
+  \medskip
+  \begin{subfigure}{\columnwidth}
+    \centering
+    \begin{subfigure}[t]{0.3\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/melotte_22/kmeans_hr_diagram_melotte_22.png}
+      \caption{K-Means}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.3\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/melotte_22/dec_hr_diagram_melotte_22.png}
+      \caption{DEC}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.3\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/melotte_22/dec_hr_diagram_filtered_melotte_22.png}
+      \caption{DEC (filt.)}
+    \end{subfigure}
+  \end{subfigure}
+  \caption{Evolution of Melotte 22 characterization}
+  \label{fig:melotte_22_characterization_evolution}
+\end{figure}
+
+Figure \ref{fig:melotte_22_characterization_evolution} shows the evolution
+of the characterization of Melotte 22. The first column corresponds to the
+characterization obtained by applying K-Means alone to the dataset. The middle
+column is the characterization made by the DEC model. And the column on the right
+shows the characterization made by the DEC and having filtered the quantiles
+lower than 0.10 and higher than 0.90. The top row shows the proper motion
+configuration space while the bottom one shows the Hertzsprung–Russell diagrams.
+
 All resources developed for this project are available at the GitHub repository:

### 2021-01-31 paper/paper.tex
paper: Add Results and Appendix

@@
   \item The model shows a great performance compared with current methods.
-  \item It is able to deal with large and dense regions. As an example,
-        it has succeeded characterizing Melotte 25 (Hiades) OC. This cluster is
-        really large and dense, and its relative proximity makes it difficult to
+  \item It is able to deal with large and open regions. As an example,
+        it has succeeded characterizing Melotte 25 (Hiades) OC. Also,
+        the relative proximity of this cluster, 45 parsec, makes it difficult to
         characterize. Clusterix fails characterizing this cluster. In this case,
@@
 
-\subsection{Evaluación 1}
+We now present the results obtained after having analyzed Melotte 25.
+This is the largest cluster we have studied and one of the most difficult
+to characterize by its distribution and extension.
 
-\subsection{Evaluación 2 }
-En la Tabla \ref{tab_1}
-\begin{table}\label{tab_1}
-\caption{Unidades de las propiedades magnéticas:}
+More results are available at Appendix \ref{sec:appendix}.
 
-\begin{tabular}{ccc}\hline\hline
-Símbolo & Cantidad & Conversión\\
+Melotte 25 is located at 66.725 degrees in right ascension and 15.867 degrees in
+declination with a radius of 165 arcmin and composed by more than 400,000 stars.
+
+Table \ref{tab:results_melotte_25} shows a results summary for Melotte 25.
+
+\begin{table}[htbp]
+  \begin{center}
+    \resizebox{\columnwidth}{!}{
+      \begin{tabular}{l|c|c|c|c}
+        \textbf{Method} & \emph{$\mu_{\alpha}$ $(mas/yr^{-1})$} & \emph{$\mu_{\delta}$ $(mas/yr^{-1})$} & \emph{$\varpi$ $(mas)$} & \emph{\# stars} \\
 \hline
-$\Phi$ & flujo magnético & $1$Mx $\rightarrow 10^{-8}V\cdot s$\\
-... &...&...\\
-\hline\hline
+        \textbf{Simbad}\tablefootnote{Results have been taken from the SIMBAD astronomical database\cite{wenger2000simbad}} & 104.92 $\pm$ 0.12 & -28.00 $\pm$ 0.09 & 21.052 $\pm$ 0.065 & - \\
+        Clusterix & 106.796 $\pm$ 6.229 & -24.870 $\pm$ 5.417 & 21.210 $\pm$ 1.115 & 109 \\
+        K-Means & 79.936 $\pm$ 3.7 & -45.362 $\pm$ 3.8 & 20.901 $\pm$ 0.31 & 374 \\
+        DEC & 104.051 $\pm$ 3.2 & -33.424 $\pm$ 2.5 & 22.072 $\pm$ 0.13 & 219 \\
+        \textbf{DEC (filt.)} & 105.96 $\pm$ 3.5 & -30.00 $\pm$ 2.4 & 21.74 $\pm$ 0.07 & 175 \\
 \end{tabular}
+    }
+    \caption{Melotte 25 results.}
+    \label{tab:results_melotte_25}
+  \end{center}
+\end{table}
+
+Parameters shown are proper motion in right ascension and declination, parallax
+with their respective deviations and number of stars corresponding to the OC.
+
+Figure \ref{fig:result_melotte_25_dec} shows the groups found using the DEC model.
+This model labels Melotte 25 as group \emph{g5}.
+
+\begin{figure}[htbp]
+  \centering
+  \begin{subfigure}{\columnwidth}
+    \centering
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/melotte_25/dec_pm_melotte_25.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/melotte_25/dec_parallax_melotte_25.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/melotte_25/dec_hr_diagram_melotte_25.png}
+    \end{subfigure}
+  \end{subfigure}
+  \caption{Melotte 25 characterization using the DEC model.
+           From the left to the right: proper motion configuration
+           space, parallax histogram and H-R diagram.}
+  \label{fig:result_melotte_25_dec}
+\end{figure}
+
+Table \ref{tab:hyperparameters_melotte_25} shows the hyperparameters
+used for the characterization.
+
+\begin{table}[h]
+  \begin{center}
+    \begin{tabular}{l|c}
+      \textbf{Hyperparameter} & \textbf{Value} \\
+      \hline
+      Number of Clusters & 9 \\
+      Clustering Layer & $\left[ 50, 50, 40 \right]$ \\
+      Kernel Initializer Seed & 11 \\
+      Quantil Threshold & 0.1 \\
+    \end{tabular}
+    \caption{Melotte 25 DEC hyperparameters.}
+    \label{tab:hyperparameters_melotte_25}
+  \end{center}
 \end{table}
@@
 
-Another possible change in our model would be using DBSCAN instead of KMeans
+Another possible change in our model would be using DBSCAN instead of K-Means
 as our initial clustering algorithm. Maybe this algorithm gives us a better starting
@@
 
+\appendix
+\section{APPENDIX}
+\label{sec:appendix}
+
+\subsection{NGC 2516}
+\label{sec:ngc2516}
+
+\begin{table}[htbp]
+  \begin{center}
+    \resizebox{\columnwidth}{!}{
+      \begin{tabular}{l|c|c|c|c}
+        \textbf{Method} & \emph{$\mu_{\alpha}$ $(mas/yr^{-1})$} & \emph{$\mu_{\delta}$ $(mas/yr^{-1})$} & \emph{$\varpi$ $(mas)$} & \emph{\# stars} \\
+        \hline
+        \textbf{Simbad} & -4.6579 $\pm$ 0.0075 & 11.1517 $\pm$ 0.0075 & 2.4118 $\pm$ 0.0006 & 1727 \\
+        Clusterix & -4.652 $\pm$ 0.523 & 11.203 $\pm$ 0.454 & 2.409 $\pm$ 0.127 & 638 \\
+        K-Means & -4.344 $\pm$ 0.14 & 9.507 $\pm$ 0.19 & 2.268 $\pm$ 0.01 & 1542 \\
+        DEC & -4.426 $\pm$ 0.17 & 9.952 $\pm$ 0.20 & 2.436 $\pm$ 0.01 & 1532 \\
+        \textbf{DEC (filt.)} & -4.502 $\pm$ 0.14 & 10.114 $\pm$ 0.17 & 2.392 $\pm$ 0.004 & 1072 \\
+      \end{tabular}
+    }
+    \caption{NGC 2516 results.}
+    \label{tab:app_results_ngc_2516}
+  \end{center}
+\end{table}
+
+Table \ref{tab:app_results_ngc_2516} shows a results summary for NGC 2516.
+The top row of Figure \ref{fig:app_result_ngc_2516_clusterix_kmeans} shows
+the characterization using Clusterix+TOPCAT tools while the bottom row
+shows eight clusters found by K-Means.
+
+\begin{figure}[htbp]
+  \centering
+  \begin{subfigure}{\columnwidth}
+    \centering
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2516/pm_ngc_2516.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2516/parallax_ngc_2516.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2516/hr_diagram_ngc_2516.png}
+    \end{subfigure}
+  \end{subfigure}
+  \centering
+  \begin{subfigure}{\columnwidth}
+    \centering
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2516/kmeans_pm_ngc_2516.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2516/kmeans_parallax_ngc_2516.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2516/kmeans_hr_diagram_ngc_2516.png}
+    \end{subfigure}
+  \end{subfigure}
+  \caption{NGC 2516 characterization.
+           Top: Clusterix+TOPCAT. Bottom: K-Means (\emph{g3}).}
+  \label{fig:app_result_ngc_2516_clusterix_kmeans}
+\end{figure}
+
+Figure \ref{fig:app_result_ngc_2516_dec} shows the groups found using the
+DEC model (top row) and the DEC model filtered (bottom row).
+K-Means and DEC have labeled NGC 2516 as \emph{g3}. Although in general,
+groups between K-Means and DEC models may not match.
+Hyperparameters used in this characterization are listed in Table
+\ref{tab:app_hyperparameters_ngc_2516}.
+
+\begin{figure}[htbp]
+  \centering
+  \begin{subfigure}{\columnwidth}
+    \centering
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2516/dec_pm_ngc_2516.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2516/dec_parallax_ngc_2516.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2516/dec_hr_diagram_ngc_2516.png}
+    \end{subfigure}
+  \end{subfigure}
+  \centering
+  \begin{subfigure}{\columnwidth}
+    \centering
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2516/dec_pm_filtered_ngc_2516.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2516/dec_parallax_filtered_ngc_2516.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2516/dec_hr_diagram_filtered_ngc_2516.png}
+    \end{subfigure}
+  \end{subfigure}
+  \caption{NGC 2516 characterization (labeled as \emph{g3}).
+           DEC (top). DEC filtered (bottom).}
+  \label{fig:app_result_ngc_2516_dec}
+\end{figure}
+
+\begin{table}[htbp]
+  \begin{center}
+    \begin{tabular}{l|c}
+      \textbf{Hyperparameter} & \textbf{Value} \\
+      \hline
+      Number of Clusters & 8 \\
+      Clustering Layer & $\left[ 50, 50, 60 \right]$ \\
+      Kernel Initializer Seed & 2 \\
+      Quantil Threshold & 0.15 \\
+    \end{tabular}
+    \caption{NGC 2516 DEC hyperparameters.}
+    \label{tab:app_hyperparameters_ngc_2516}
+  \end{center}
+\end{table}
+
+\subsection{NGC 2632}
+\label{sec:ngc2632}
+
+Figure \ref{fig:app_result_ngc_2632_clusterix_kmeans} shows NGC 2632 characterized
+by the validation method (first row) and the characterization made by K-Means
+showing ten clusters (second row). K-Means labels NGC 2632 as \emph{g1}.
+
+\begin{figure}[htbp]
+  \centering
+  \begin{subfigure}{\columnwidth}
+    \centering
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2632/pm_ngc_2632.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2632/parallax_ngc_2632.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2632/hr_diagram_ngc_2632.png}
+    \end{subfigure}
+  \end{subfigure}
+  \centering
+  \begin{subfigure}{\columnwidth}
+    \centering
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2632/kmeans_pm_ngc_2632.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2632/kmeans_parallax_ngc_2632.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2632/kmeans_hr_diagram_ngc_2632.png}
+    \end{subfigure}
+  \end{subfigure}
+  \caption{NGC 2632 characterization using Clusterix+TOPCAT (top) and K-Means (bottom).
+           K-Means identifies NGC 2632 as \emph{g1}.}
+  \label{fig:app_result_ngc_2632_clusterix_kmeans}
+\end{figure}
+
+Figure \ref{fig:app_result_ngc_2632_dec} shows the groups found using
+the DEC model (first row) and the DEC model filtered (second row).
+Open cluster NGC 2632 is labeled as group \emph{g1}.
+
+\begin{figure}[htbp]
+  \centering
+  \begin{subfigure}{\columnwidth}
+    \centering
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_pm_ngc_2632.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_parallax_ngc_2632.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_hr_diagram_ngc_2632.png}
+    \end{subfigure}
+  \end{subfigure}
+  \centering
+  \begin{subfigure}{\columnwidth}
+    \centering
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_pm_filtered_ngc_2632.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_parallax_filtered_ngc_2632.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_hr_diagram_filtered_ngc_2632.png}
+    \end{subfigure}
+  \end{subfigure}
+  \caption{NGC 2632 characterization using DEC model (top) and DEC (filtered) (bottom).
+           NGC 2632 is identified as \emph{g1}.}
+  \label{fig:app_result_ngc_2632_dec}
+\end{figure}
+
+\begin{table}[htbp]
+  \begin{center}
+    \resizebox{\columnwidth}{!}{
+      \begin{tabular}{l|c|c|c|c}
+        \textbf{Method} & \emph{$\mu_{\alpha}$ $(mas/yr^{-1})$} & \emph{$\mu_{\delta}$ $(mas/yr^{-1})$} & \emph{$\varpi$ $(mas)$} & \emph{\# stars} \\
+        \hline
+        \textbf{Simbad} & -36.047 $\pm$ 0.110 & -12.917 $\pm$ 0.066 & 5.371 $\pm$ 0.003 & - \\
+        Clusterix & -36.154 $\pm$ 1.001 & -12.909 $\pm$ 0.806 & 5.327 $\pm$ 0.187 & 371 \\
+        K-Means & -26.352 $\pm$ 0.82 & -15.828 $\pm$ 0.76 & 5.394 $\pm$ 0.03 & 629 \\
+        DEC & -20.012 $\pm$ 0.69 & -14.742 $\pm$ 0.58 & 4.686 $\pm$ 0.03 & 894 \\
+        \textbf{DEC (filt.)} & -21.571 $\pm$ 0.74 & -14.234 $\pm$ 0.61 & 4.719 $\pm$ 0.03 & 714 \\
+      \end{tabular}
+    }
+    \caption{NGC 2632 results.}
+    \label{tab:app_results_ngc_2632}
+  \end{center}
+\end{table}
+
+Table \ref{tab:app_results_ngc_2632} shows a results summary for NGC 2632 analysis.
+While Table \ref{tab:app_hyperparameters_ngc_2632} shows the hyperparameters
+used for characterizing NGC 2632 with our model.
+
+\begin{table}[htbp]
+  \begin{center}
+    \begin{tabular}{l|c}
+      \textbf{Hyperparameter} & \textbf{Value} \\
+      \hline
+      Number of Clusters & 10 \\
+      Clustering Layer & $\left[ 50, 50, 40 \right]$ \\
+      Kernel Initializer Seed & 10 \\
+      Quantil Threshold & 0.1 \\
+    \end{tabular}
+    \caption{NGC 2632 DEC hyperparameters.}
+    \label{tab:app_hyperparameters_ngc_2632}
+  \end{center}
+\end{table}
+
+\subsection{NGC 2682}
+\label{sec:ngc2682}
+
+Table \ref{tab:app_results_ngc_2682} shows a results summary for NGC 2682.
+
+\begin{table}[htbp]
+  \begin{center}
+    \resizebox{\columnwidth}{!}{
+      \begin{tabular}{l|c|c|c|c}
+        \textbf{Method} & \emph{$\mu_{\alpha}$ $(mas/yr^{-1})$} & \emph{$\mu_{\delta}$ $(mas/yr^{-1})$} & \emph{$\varpi$ $(mas)$} & \emph{\# stars} \\
+        \hline
+        \textbf{Simbad} & -10.9737 $\pm$ 0.0064 & -2.9396 $\pm$ 0.0063 & 1.1325 $\pm$ 0.0011 & 1194 \\
+        Clusterix+TOPCAT & -10.970 $\pm$ 0.322 & -2.958 $\pm$ 0.327 & 1.142 $\pm$ 0.080 & 649 \\
+        K-Means & -8.616 $\pm$ 0.15 & -3.710 $\pm$ 0.16 & 1.196 $\pm$ 0.01 & 1374 \\
+        DEC & -8.926 $\pm$ 0.15 & -3.550 $\pm$ 0.15 & 1.144 $\pm$ 0.005 & 1238 \\
+        \textbf{DEC (filt.)} & -9.619 $\pm$ 0.13 & -3.317 $\pm$ 0.13 & 1.140 $\pm$ 0.003 & 990 \\
+      \end{tabular}
+    }
+    \caption{NGC 2682 results.}
+    \label{tab:app_results_ngc_2682}
+  \end{center}
+\end{table}
+
+Figure \ref{fig:app_result_ngc_2682_clusterix_kmeans} shows
+NGC 2682 characterized using Clusterix+TOPCAT tools (top row)
+and ten clusters identified by K-Means (bottom row).
+The chacterization made by K-Means labels NGC 2682 as group \emph{g4}.
+
+\begin{figure}[htbp]
+  \centering
+  \begin{subfigure}{\columnwidth}
+    \centering
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2682/pm_ngc_2682.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2682/parallax_ngc_2682.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2682/hr_diagram_ngc_2682.png}
+    \end{subfigure}
+  \end{subfigure}
+  \centering
+  \begin{subfigure}{\columnwidth}
+    \centering
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2682/kmeans_pm_ngc_2682.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2682/kmeans_parallax_ngc_2682.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2682/kmeans_hr_diagram_ngc_2682.png}
+    \end{subfigure}
+  \end{subfigure}
+  \caption{NGC 2682 K-Means characterization.
+           Top row corresponds to Clusterix+TOPCAT characterization,
+           while bottom row shows K-Means results identifying
+           NGC 2682 as cluster \emph{g4}.}
+  \label{fig:app_result_ngc_2682_clusterix_kmeans}
+\end{figure}
+
+Figure \ref{fig:app_result_ngc_2682_dec} shows the groups found using
+the DEC model (first row) and the DEC model filtered (second row).
+The cluster of interest is the group \emph{g2}.
+
+\begin{figure}[htbp]
+  \centering
+  \begin{subfigure}{\columnwidth}
+    \centering
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2682/dec_pm_ngc_2682.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2682/dec_parallax_ngc_2682.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2682/dec_hr_diagram_ngc_2682.png}
+    \end{subfigure}
+  \end{subfigure}
+  \centering
+  \begin{subfigure}{\columnwidth}
+    \centering
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2682/dec_pm_filtered_ngc_2682.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2682/dec_parallax_filtered_ngc_2682.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/ngc_2682/dec_hr_diagram_filtered_ngc_2682.png}
+    \end{subfigure}
+  \end{subfigure}
+  \caption{NGC 2682 is identified as cluster \emph{g2}.
+           Top row: DEC --- Bottom row: DEC (filtered)}
+  \label{fig:app_result_ngc_2682_dec}
+\end{figure}
+
+Table \ref{tab:app_hyperparameters_ngc_2682} shows the
+hyperparameters used for characterizing NGC 2682 with DEC method.
+
+\begin{table}[htbp]
+  \begin{center}
+    \begin{tabular}{l|c}
+      \textbf{Hyperparameter} & \textbf{Value} \\
+      \hline
+      Number of Clusters & 10 \\
+      Clustering Layer & $\left[ 50, 50, 40 \right]$ \\
+      Kernel Initializer Seed & 0 \\
+      Quantil Threshold & 0.1 \\
+    \end{tabular}
+    \caption{NGC 2682 DEC hyperparameters.}
+    \label{tab:app_hyperparameters_ngc_2682}
+  \end{center}
+\end{table}
+
+\subsection{MELOTTE 22}
+\label{sec:melotte22}
+
+\begin{figure}[htbp]
+  \centering
+  \begin{subfigure}{\columnwidth}
+    \centering
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/melotte_22/pm_melotte_22.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/melotte_22/parallax_melotte_22.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/melotte_22/hr_diagram_melotte_22.png}
+    \end{subfigure}
+  \end{subfigure}
+  \centering
+  \begin{subfigure}{\columnwidth}
+    \centering
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/melotte_22/kmeans_pm_melotte_22.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/melotte_22/kmeans_parallax_melotte_22.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/melotte_22/kmeans_hr_diagram_melotte_22.png}
+    \end{subfigure}
+  \end{subfigure}
+  \caption{Melotte 22 characterization.
+           Top row: Clusterix+TOPCAT. Bottom row: K-Means.
+           Melotte 22 is labeled as cluster \emph{g1}.}
+  \label{fig:app_result_melotte_22_clusterix_kmeans}
+\end{figure}
+
+Figure \ref{fig:app_result_melotte_22_clusterix_kmeans} shows Melotte 22
+characterized using Clusterix+TOPCAT tools (top row) and the characterization
+made by K-Means (bottom row) showing nine clusters. Melotte 22 is labeled as
+group \emph{g1}.
+
+Figure \ref{fig:app_result_melotte_22_dec} shows the groups found using
+DEC model (first row) and DEC model filtered (second row).
+Melotte 22 is labeled as group \emph{g2} by DEC model.
+
+\begin{figure}[htbp]
+  \centering
+  \begin{subfigure}{\columnwidth}
+    \centering
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/melotte_22/dec_pm_melotte_22.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/melotte_22/dec_parallax_melotte_22.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/melotte_22/dec_hr_diagram_melotte_22.png}
+    \end{subfigure}
+  \end{subfigure}
+  \centering
+  \begin{subfigure}{\columnwidth}
+    \centering
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/melotte_22/dec_pm_filtered_melotte_22.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/melotte_22/dec_parallax_filtered_melotte_22.png}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.30\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{../figures/melotte_22/dec_hr_diagram_filtered_melotte_22.png}
+    \end{subfigure}
+  \end{subfigure}
+  \caption{Melotte 22 is identified as cluster \emph{g2}.
+           Top row: DEC --- Bottom row: DEC (filtered)}
+  \label{fig:app_result_melotte_22_dec}
+\end{figure}
+
+Table \ref{tab:app_results_melotte_22} shows a results summary for Melotte 22.
+
+\begin{table}[htbp]
+  \begin{center}
+    \resizebox{\columnwidth}{!}{
+      \begin{tabular}{l|c|c|c|c}
+        \textbf{Method} & \emph{$\mu_{\alpha}$ $(mas/yr^{-1})$} & \emph{$\mu_{\delta}$ $(mas/yr^{-1})$} & \emph{$\varpi$ $(mas)$} & \emph{\# stars} \\
+        \hline
+        \textbf{Simbad} & 19.997 $\pm$ 0.127 & -45.548 $\pm$ 0.101 & 7.364 $\pm$ 0.005 & 1326 \\
+        Clusterix & 19.98 $\pm$ 1.25 & -45.47 $\pm$ 1.48 & 7.33 $\pm$ 0.21 & 634 \\
+        K-Means & 20.25 $\pm$ 0.95 & -38.01 $\pm$ 1.08 & 7.23 $\pm$ 0.06 & 1378 \\
+        DEC & 23.67 $\pm$ 1.29 & -46.23 $\pm$ 1.50 & 8.04 $\pm$ 0.09 & 878 \\
+        \textbf{DEC (filt.)} & 19.50 $\pm$ 0.41 & -44.23 $\pm$ 0.39 & 7.42 $\pm$ 0.005 & 438 \\
+      \end{tabular}
+    }
+    \caption{Melotte 22 results.}
+    \label{tab:app_results_melotte_22}
+  \end{center}
+\end{table}
+
+While Table \ref{tab:app_hyperparameters_melotte_22} shows the hyperparameters
+used for characterizing Melotte 22 with the DEC model.
+
+\begin{table}[hbtp]
+  \begin{center}
+    \begin{tabular}{l|c}
+      \textbf{Hyperparameter} & \textbf{Value} \\
+      \hline
+      Number of Clusters & 5 \\
+      Clustering Layer & $\left[ 50, 50, 200 \right]$ \\
+      Kernel Initializer Seed & 11 \\
+      Quantil Threshold & 0.1 \\
+    \end{tabular}
+    \caption{Melotte 22 DEC hyperparameters.}
+    \label{tab:app_hyperparameters_melotte_22}
+  \end{center}
+\end{table}
+
 \renewcommand{\refname}{REFERENCES}

### 2021-01-31 paper/paper.tex
paper: Add Results Discussion

@@
 
-Tras la presentación objetiva de los resultados, querrás aportar una discusión de los mismos.
+We have applied our proposed method to clusters with different typologies:
+
+\begin{itemize}
+  \item NGC 2516 (Ap. \ref{sec:ngc2516}) has it proper motion center deviated from the origin but
+        it is embedded inside a big cloud of stars with similar proper motions.
+  \item NGC 2632 (Ap. \ref{sec:ngc2632}), in addition to Melotte 22, ia s cluster whose proper motion center
+        is not located at the origin and has a well separated parallax center.
+  \item NGC 2682 (Ap. \ref{sec:ngc2682}), on the other hand, has its parallax centered within the region's gaussian,
+        which complicates its detection although its proper motion center is deviated from the origin.
+  \item Melotte 22 (Ap. \ref{sec:melotte22}), as well as Melotte 25, is a cluster closer to us than the other.
+        That makes its membership stars to be more scattered than previous clusters which are more compact.
+\end{itemize}
+
+All these clusters have a wide variety of diameters, from 25 to 330 arcmin,
+as well as the number of stars that belong to them. NGC 2682 has 3,000 stars while
+Melotte 25 is composed by more than 400,000 stars. More information about the
+studied clusters is available in Table \ref{tab:clusters_summary}.
+
+\begin{table}[htbp]
+  \begin{center}
+    \resizebox{\columnwidth}{!}{
+      \begin{tabular}{l|c|c|c|c}
+        \textbf{OC} & \emph{$\alpha$ J2000 $(degrees)$} & \emph{$\delta$ J2000 $(degrees)$} & \emph{Radius $(arcmin)$} & \emph{\# stars} \\
+        \hline
+        NGC 2516 & 119.517 & -60.753 & 15 & 12,869 \\
+        NGC 2632 & 130.1 & 19.667 & 35 & 13,167 \\
+        NGC 2682 & 132.825 & 11.8 & 12.5 & 2,839 \\
+        Melotte 22 & 56.75 & 24.117 & 60 & 61,552 \\
+        Melotte 25 & 66.725 & 15.867 & 165 & 433,996 \\
+      \end{tabular}
+    }
+    \caption{Righ ascension, declination, radius and number of stars of studied clusters.
+             The number of stars corresponds to those stars contained within a cone of center
+             $(\alpha, \delta)$ and raius the cluster's radius multiplied by a factor of 1.5.}
+    \label{tab:clusters_summary}
+  \end{center}
+\end{table}
+
+In all cases, the model has resolved properly the identity of the cluster
+and has characterized the membership stars, giving compatible results
+with the ones obtained with classic procedures and VO tools.
+
+Compared with the mentioned tools, our model can be categorized as a
+non-parameterized model since it does not depend on parameters referred
+to the cluster itself but only hyperparameters such as the initial number
+of clusters to be found, or the structure of the Clustering Layer used by DEC model.
+
+The proposed method is also non-supervised since we do not need to tell the model
+which stars belong to the cluster or assist it while its training stage.
+
+Only at the end of the characterization, a fine tuning driven by the user,
+can be applied in order to improve the selection by discarding those
+stars which fall outside the given quantiles.
+
+In any case, we do not make assumptions about the cluster profile,
+something that we have to take into account when using VO tools.
+This is another evidence that our model is non-parameterized.
 

### 2021-01-31 thesis/thesis.tex

@@
 
+\medskip
+
+This research has made use of the SIMBAD database, operated at CDS, Strasbourg, France
+\cite[2000,A\&AS,143,9]{wenger2000simbad}, "The SIMBAD astronomical database", Wenger et al.
+
 \mainmatter
@@
 
-There are several clustering algorithms: KMeans, Mean-Shift Clustering, DBSCAN \cite{ester1996density} among other.
+There are several clustering algorithms: K-Means, Mean-Shift Clustering, DBSCAN \cite{ester1996density} among other.
 While each one behaves better according to the distribution of the objects to clusterize,
-we have chosen KMeans by its simplicity and good results.
-
-KMeans requires a single parameter, the number of clusters to build (N).
+we have chosen K-Means by its simplicity and good results.
+
+K-Means requires a single parameter, the number of clusters to build (N).
 The algorithm looks for N center points which are vectors of the same dimension as the number of selected features.
-KMeans starts by making an initial groups configuration and then,
+K-Means starts by making an initial groups configuration and then,
 it reassigns objects to other groups iteratively by minimizing the distance among points
@@
 SIMBAD Astronomical Database - CDS (Strasbourg),
-Clusterix+TOPCAT method, and the results obtained for each step of this method (KMeans, DEC and DEC filtered).
-Errors associated to KMeans and DEC predictions have been calculated as the \emph{Standard Error},
+Clusterix+TOPCAT method, and the results obtained for each step of this method (K-Means, DEC and DEC filtered).
+Errors associated to K-Means and DEC predictions have been calculated as the \emph{Standard Error},
 $SE = \sigma / \sqrt{N}$, where $N$ is the number of samples in the selected group.
@@
       Clusterix+TOPCAT & 19.98 $\pm$ 1.25 & -45.47 $\pm$ 1.48 & 7.33 $\pm$ 0.21 & 634 \\
-      KMeans & 20.25 $\pm$ 0.95 & -38.01 $\pm$ 1.08 & 7.23 $\pm$ 0.06 & 1378 \\
+      K-Means & 20.25 $\pm$ 0.95 & -38.01 $\pm$ 1.08 & 7.23 $\pm$ 0.06 & 1378 \\
       DEC & 23.67 $\pm$ 1.29 & -46.23 $\pm$ 1.50 & 8.04 $\pm$ 0.09 & 878 \\
@@
       \textbf{Clusterix+TOPCAT} & 19.98 $\pm$ 1.25 & -45.47 $\pm$ 1.48 & 7.33 $\pm$ 0.21 & 634 \\
-      KMeans & 20.25 $\pm$ 0.95 & -38.01 $\pm$ 1.08 & 7.23 $\pm$ 0.06 & 1378 \\
+      K-Means & 20.25 $\pm$ 0.95 & -38.01 $\pm$ 1.08 & 7.23 $\pm$ 0.06 & 1378 \\
       DEC & 23.67 $\pm$ 1.29 & -46.23 $\pm$ 1.50 & 8.04 $\pm$ 0.09 & 878 \\
@@
 Since our aim is to get a model that fits well for a wide range of clusters,
-we will now proceed to show some results obtained for a selection of clusters with different topologies.
+we will now proceed to show some results obtained for a selection of clusters with different typologies.
 
@@
   \end{subfigure}
-  \caption{NGC 2516 KMeans characterization. Identified as cluster \emph{g3}.}
+  \caption{NGC 2516 K-Means characterization. Identified as cluster \emph{g3}.}
   \label{fig:result_ngc_2516_kmeans}
@@
 Figure \ref{fig:result_ngc_2516_clusterix} shows NGC 2516 characterized using Clusterix+TOPCAT tools.
-Figure \ref{fig:result_ngc_2516_kmeans} shows eigth clusters identified by KMeans.
+Figure \ref{fig:result_ngc_2516_kmeans} shows eight clusters identified by K-Means.
 In this case, cluster \emph{g3} is the one we are looking for.
@@
 Again, the cluster of interest is the group \emph{g3}.
-Although in general, groups between KMeans model and DEC models may not match.
+Although in general, groups between K-Means model and DEC models may not match.
 
@@
       Clusterix+TOPCAT & -4.652 $\pm$ 0.523 & 11.203 $\pm$ 0.454 & 2.409 $\pm$ 0.127 & 638 \\
-      KMeans & -4.344 $\pm$ 0.14 & 9.507 $\pm$ 0.19 & 2.268 $\pm$ 0.01 & 1542 \\
+      K-Means & -4.344 $\pm$ 0.14 & 9.507 $\pm$ 0.19 & 2.268 $\pm$ 0.01 & 1542 \\
       DEC & -4.426 $\pm$ 0.17 & 9.952 $\pm$ 0.20 & 2.436 $\pm$ 0.01 & 1532 \\
@@
   \end{subfigure}
-  \caption{NGC 2632 KMeans characterization. Identified as cluster \emph{g1}.}
+  \caption{NGC 2632 K-Means characterization. Identified as cluster \emph{g1}.}
   \label{fig:result_ngc_2632_kmeans}
@@
 Figure \ref{fig:result_ngc_2632_clusterix} shows NGC 2632 characterized by the validation method.
-Figure \ref{fig:result_ngc_2632_kmeans} shows ten clusters identified by KMeans.
+Figure \ref{fig:result_ngc_2632_kmeans} shows ten clusters identified by K-Means.
 Cluster \emph{g1} is the open cluster we are looking for.
@@
       Clusterix+TOPCAT & -36.154 $\pm$ 1.001 & -12.909 $\pm$ 0.806 & 5.327 $\pm$ 0.187 & 371 \\
-      KMeans & -26.352 $\pm$ 0.82 & -15.828 $\pm$ 0.76 & 5.394 $\pm$ 0.03 & 629 \\
+      K-Means & -26.352 $\pm$ 0.82 & -15.828 $\pm$ 0.76 & 5.394 $\pm$ 0.03 & 629 \\
       DEC & -20.012 $\pm$ 0.69 & -14.742 $\pm$ 0.58 & 4.686 $\pm$ 0.03 & 894 \\
@@
   \end{subfigure}
-  \caption{NGC 2682 KMeans characterization. Identified as cluster \emph{g4}.}
+  \caption{NGC 2682 K-Means characterization. Identified as cluster \emph{g4}.}
   \label{fig:result_ngc_2682_kmeans}
@@
 Figure \ref{fig:result_ngc_2682_clusterix} shows NGC 2682 characterized using Clusterix+TOPCAT tools.
-Figure \ref{fig:result_ngc_2682_kmeans} shows ten clusters identified by KMeans.
-KMeans labels NGC 2682 as group \emph{g4}.
+Figure \ref{fig:result_ngc_2682_kmeans} shows ten clusters identified by K-Means.
+K-Means labels NGC 2682 as group \emph{g4}.
 Figures \ref{fig:result_ngc_2682_dec} and \ref{fig:result_ngc_2682_dec_filtered}
@@
       Clusterix+TOPCAT & -10.970 $\pm$ 0.322 & -2.958 $\pm$ 0.327 & 1.142 $\pm$ 0.080 & 649 \\
-      KMeans & -8.616 $\pm$ 0.15 & -3.710 $\pm$ 0.16 & 1.196 $\pm$ 0.01 & 1374 \\
+      K-Means & -8.616 $\pm$ 0.15 & -3.710 $\pm$ 0.16 & 1.196 $\pm$ 0.01 & 1374 \\
       DEC & -8.926 $\pm$ 0.15 & -3.550 $\pm$ 0.15 & 1.144 $\pm$ 0.005 & 1238 \\
@@
   \end{subfigure}
-  \caption{Melotte 25 KMeans characterization. Identified as cluster \emph{g4}.}
+  \caption{Melotte 25 K-Means characterization. Identified as cluster \emph{g4}.}
   \label{fig:result_melotte_25_kmeans}
@@
 Figure \ref{fig:result_melotte_25_clusterix} shows Melotte 25 characterized using Clusterix+TOPCAT tools.
-Figure \ref{fig:result_melotte_25_kmeans} shows nine clusters identified by KMeans.
-KMeans labels Melotte 25 as group \emph{g3}.
+Figure \ref{fig:result_melotte_25_kmeans} shows nine clusters identified by K-Means.
+K-Means labels Melotte 25 as group \emph{g3}.
 Figures \ref{fig:result_melotte_25_dec} and \ref{fig:result_melotte_25_dec_filtered}
@@
       Clusterix+TOPCAT & 106.796 $\pm$ 6.229 & -24.870 $\pm$ 5.417 & 21.210 $\pm$ 1.115 & 109 \\
-      KMeans & 79.936 $\pm$ 3.7 & -45.362 $\pm$ 3.8 & 20.901 $\pm$ 0.31 & 374 \\
+      K-Means & 79.936 $\pm$ 3.7 & -45.362 $\pm$ 3.8 & 20.901 $\pm$ 0.31 & 374 \\
       DEC & 104.051 $\pm$ 3.2 & -33.424 $\pm$ 2.5 & 22.072 $\pm$ 0.13 & 219 \\
@@
 
-We have applied our proposed method to clusters with different topologies:
+We have applied our proposed method to clusters with different typologies:
 
@@
   \item Melotte 25 (as well as Melotte 22) is a cluster closer to us than the other.
-        That makes its memebership stars to be more scattered than previous clusters which are more compact.
+        That makes its membership stars to be more scattered than previous clusters which are more compact.
 \end{itemize}
 
-All of these clusters have a wide variety of diameters, from 25 to 330 arcmin,
+All these clusters have a wide variety of diameters, from 25 to 330 arcmin,
 as well as the number of stars that belong to them.
@@
 
-Compared with the mentioned tools, our model can be cathegorizeed as a non-parameterized
+Compared with the mentioned tools, our model can be categorized as a non-parameterized
 model since it does not depend on parameters referred to the cluster itself
@@
 
-Another possible change in our model would be using DBSCAN instead of KMeans as our initial clustering algorithm.
+Another possible change in our model would be using DBSCAN instead of K-Means as our initial clustering algorithm.
 Maybe this algorithm gives us a better starting point for the DEC model which could improve the results.
@@
 
-\includepdf[pages={-},fitpaper]{../paper/paper.pdf}
+\includepdf[pages={-},noautoscale]{../paper/paper.pdf}
 

### 2021-01-31 paper/paper.tex

@@
 \title{Machine Learning Tools for Open Cluster Characterization with Gaia DR2 Data}
-\author{CD. Álvaro, J. Álvaro, CAG. Guzmán}
+\author{CD. Álvaro, CAG. Álvarez, J. Álvaro}
 \date{24th of January, 2021}
@@
 
-Melotte 25 is located at 66.725 degrees in right ascension and 15.867 degrees in
-declination with a radius of 165 arcmin and composed by more than 400,000 stars.
+Melotte 25 is located at 66.725 degrees in right ascension and 15.867
+degrees in declination with a radius of 165 arcmin.
+This region contains more than 400,000 stars.
 
@@
 All these clusters have a wide variety of diameters, from 25 to 330 arcmin,
-as well as the number of stars that belong to them. NGC 2682 has 3,000 stars while
-Melotte 25 is composed by more than 400,000 stars. More information about the
-studied clusters is available in Table \ref{tab:clusters_summary}.
+as well as the number of stars that belong to them. NGC 2682 has 3,000 stars
+while Melotte 25 is located inside a region with more than 400,000 stars.
+More information about the studied clusters is available in Table \ref{tab:clusters_summary}.
 

### 2021-02-06 paper/paper.tex
paper: Update Abstract
Update Abstract with changes suggested by @cguz

Co-Authored-By: Cesar Guzmann <15159632+cguz@users.noreply.github.com>

@@
 \title{Machine Learning Tools for Open Cluster Characterization with Gaia DR2 Data}
-\author{CD. Álvaro, CAG. Álvarez, J. Álvaro}
+\author{CD. Álvaro, C. Guzman, J. Álvaro}
 \date{24th of January, 2021}
@@
 \resumen{
-The characterization and understanding of \emph{Open Clusters}
-allow us to better understand properties and mechanisms about the Universe
+The characterization and understanding of \emph{Open Clusters} (OCs)
+allow us to understand better properties and mechanisms about the Universe
 such as stellar formation and the regions where these events occur.
-They also provide information about stellar processes and the evolution of the galactic disk.
-
-In this paper, we develop a method to characterize open clusters by using
-\emph{machine learning} tools, such as clustering by \emph{K-Means}
-and clustering based on \emph{Artificial Neural Networks} by adapting
-the \emph{Deep Embedded Clustering} model.
-We are using \emph{Gaia DR2 database} as data source.
-
-The developed method aims to improve classic techniques of the state of the art.
-We achieve improvements not only in terms of \emph{computational efficiency},
-with lower computational requirements, but in \emph{usability}, reducing the number
-of hyperparameters to get a good characterization of the analyzed clusters.
+They also provide information about stellar processes and the evolution
+of the galactic disk.
+
+In this paper, we present a novel method to characterize OCs.
+Our method employs a model built on \emph{Artificial Neural Networks} (ANNs).
+More specifically, we adapted a state of the art model, the
+\emph{Deep Embedded Clustering} (DEC) model for our purpose.
+
+The developed method aims to improve classical state of the arts techniques.
+We improved not only in terms of computational efficiency, (with lower
+computational requirements), but in usability (reducing the number
+of hyperparameters to get a good characterization of the analyzed clusters).
+For our experiments, we used the \emph{Gaia DR2 database} as the data source,
+and compared our model with the clustering technique \emph{K-Means}.
 
 Our method achieves good results,
-becoming even better in some cases when the results are compared with current techniques.
+becoming even better (in some of the cases) than current techniques.
 }

### 2021-02-06 paper/paper.tex
paper: Update Introduction
Update Introduction with suggestions made by @cguz

Co-Authored-By: Cesar Guzmann <15159632+cguz@users.noreply.github.com>

@@
 
-\section{INTRODUCTION}
-
-Stellar open clusters (OCs) \cite{janes1982open} are groups of stars
-gravitationally bounded originated from a single molecular gas cloud.
-They share the same chemical composition and age. Moreover,
-they have similar relative positions and proper motion.
-Those astronomical objects are relevant to understand the spiral structure,
-the dynamics and the chemical evolution of our galaxy.
-
-The study of OCs has been pushed forward thanks to the huge and precise dataset
-from the Gaia mission Gaia DR2 available since 2018
+\section{Introduction}
+
+Stellar OCs~\cite{janes1982open} are groups of stars gravitationally
+bounded originated from a single molecular gas cloud.
+They share the same chemical composition and age, and that is the reason
+that their metallicity must be uniform since those stars were born from
+the same gas cloud and at the same time stage. Likewise, they have similar
+relative positions inherited from their original gas cloud, which means that
+their distances to the Earth are the same. Therefore, they have a narrow
+dispersion in their parallax value. Moreover, they share similar values
+of proper motion, both in right ascension and declination. For all these
+properties, we can see why the stellar OCs are relevant to understand the
+spiral structure, dynamics, and chemical evolution of our galaxy.
+
+The study of OCs advanced thanks to the huge and precise dataset
+from the Gaia DR2 mission, available since 2018 from
 \cite{collaboration2016description} and \cite{gaia2018gaia}.
-This dataset has helped to review already known open clusters and to find new ones.
-
-Stars that belong to the same OC share relative positions,
-inherited from their original gas cloud.
-This means that their distances to the Earth are similar for all of them and,
-therefore, they have a narrow dispersion in their parallax value.
-They also share similar values of proper motion, both in right ascension and declination.
-Another property they share is their chemical composition.
-Thus, their metallicity must be uniform,
-since these stars were born from the same gas cloud and at the same time stage.
-However, to take this last property into account, we are faced with the drawback
-that this parameter is poorly reported in Gaia DR2 database.
-
-Normally, the characterization process takes into account astrometric features
-such as proper motion in right ascension and declination, and parallax.
-Photometric features like $E(b-r) / G_{mag}$ are considered too.
-These last variables are used for generating the H-R diagram of the OC candidate stars.
-This stars should present a sharp profile corresponding to an isochrone curve derived
-from a theoretical model.
-This model does consider metallicity, mass and brightness of the stars involved.
-At the moment, a set of tools have been used for this purpose, as explained later.
-But in general, these tools require supervised and parameterized models.
-Therefore, a previous knowledge of the cluster is necessary or the process must be
-repeated iteratively in order to achieve valid results.
-
-Our aim is to build a machine learning model capable of characterizing an open
-cluster inside a stellar region with knowing nothing about the region.
-This model will take advantage of these features to train an Artificial Neural
-Network (ANN) that will clusterize all stars within the given region.
-One of these groups will be the open cluster that interests us.
-With that purpose we have adapted the Unsupervised Deep Embedding for Clustering
-Analysis \cite{xie2016unsupervised} to our problem.
-
-This model fulfills one of our main requirements.
-It has to be \emph{non-supervised} and \emph{non-parameterized} in order to make easier
-the automation process of analyzing a wide range of regions with different typologies.
-Another important requirement for our model is that it must be computationally efficient
-in order to run in common workstations.
-The model has been developed with Python using the Keras framework which takes
-advantage of modern GPUs to make its computations.
-That increases significantly the computational capacity of our model making possible
-to run it on regular workstations with one of those GPUs.
-
-Since the aim of this work is not to find new OCs but to improve the characterization
-of the already known ones, we have taken the approach of analyzing known regions.
-These regions have been taken from the OpenClust catalogue \cite{dias2002new}
-(see Figure \ref{fig:openclust_catalogue}) and downloaded from Gaia DR2 database
-increasing the radius of each region by a factor $1.5$.
-
-\begin{figure}
-    \includegraphics[width=\columnwidth]{../figures/openclust_catalogue.png}
-    \caption{OpenClust Catalogue Distribution}
-    \label{fig:openclust_catalogue}
-\end{figure}
-
-Finally, we have compared the results obtained with the model we purposed with another
-method that involves tools from the Virtual Observatory (VO) such as Clusterix \cite{balaguer2020clusterix}.
-
-As shown in Section \ref{sec:results}, the results obtained are promising.
-This allows us to consider our model as a valid tool that could be part
-of the standard set of tools for open clusters characterization.
-
-\section{STATE OF THE ART}
+Although, the Gaia DR2 database misreported the chemical composition
+of some OCs, the database has helped to review already known OCs and
+to find new ones.
+
+The characterization of an OC refers to the statistical determination
+of the stars that form it. Therefore, the characterization is based on
+the probability of a star to be a member of the cluster~\cite{sampedro2016caracterizacion}.
+
+Normally, the characterization process considers astrometic features such as
+proper motion in right ascension and declination, or parallax.
+In some cases~\cite{oliveira2013fitting}, it also considers photometric features,
+which helps to generate the H-R diagram~\cite{hypki2018gaia} of the OC candidate stars.
+Those OC candidates should present a sharp profile corresponding to an isochrone curve
+derived from a theoretical model. The model considers metallicity, mass and brightness
+of the stars involved. For all this purpose, we used a set of tools that require
+supervised and parameterized models. It means that we need a previous knowledge of the
+cluster, otherwise we must repeat the process iteratively to achieve valid results.
+
+In the present paper, we propose a machine learning model capable of characterizing
+OCs inside a stellar region with no previous knowledge about the region.
+Our model takes advantage of some features such as proper motion in right ascension
+and declination, and parallax to train an Artificial Neural Network.
+The ANN clusters all the stars within given region in groups.
+One of those groups will be the open cluster that interests us.
+
+Our model contributes with several novelties: i) it is \emph{non-supervised} and
+\emph{non-parameterized}, making easier the automation process of analyzing a
+wide range of regions with different typologies, and ii) it is computationally
+efficient to run in common workstations because it has been developed with
+Python using the Keras framework which takes advantage of modern GPUs to perform
+its computations. That increases significantly the computational capacity of
+our model doing it possible to run on regular workstations.
+
+This paper is organized as follows. Section~\ref{sec:related_works} presents
+some related works and Section~\ref{sec:aims_and_method} describes in more detail
+our model. The following Section~\ref{sec:results}, shows the results and compares
+them with a method that involves tools from the Virtual Observatory (VO) such as
+Clusterix~\cite{balaguer2020clusterix}. Finally, the last section concludes and
+outlines future research lines.
+
+\section{Related works}
+\label{sec:related_works}
 
@@
 
-\section{AIMS AND METHOD}
+\section{Open Cluster Characterization method}
+\label{sec:aims_and_method}
 

### 2021-02-07 paper/paper.tex
paper: Update Related works
Update Related works section with @cguz suggestions.

Co-Authored-By: Cesar Guzmann <15159632+cguz@users.noreply.github.com>

@@
 
-The initial approach to find open clusters is by searching for overdensities
-in the astrometric space of the galactic disk and a later identification of
-possible OCs using photometric information \cite{castro2020hunting}.
-However, although it looks simple, it is not. Here we face the problem that the
-near field around the OC is composed of two kind of populations.
-The first kind belongs to the membership stars of the OC (from tens or hundreds
-to a few thousand) while the second one is composed by a background of stars
-that do not belong to the OC (from tens to hundreds of thousands).
-To find out which stars do belong to the OC is the problem we try to solve in
-this work as this selection is critical to properly characterize the
-fundamental properties of the cluster (total mass, age, chemical composition, etc.).
-
-Since the components of the open cluster are born from the same gas cloud and they
-share both dynamic properties and relative distance, as well as age and chemical
-composition, the first step is to analyze the proper motion configuration space of
-the region of interest. This approach is the one used by many tools from the VO
-such as TOPCAT \cite{taylor2005topcat}, Clusterix 2.0 \cite{balaguer2020clusterix},
-Aladin \cite{bonnarel2000aladin} or VOSA \cite{bayo2008vosa}.
-
-TOPCAT by itself cannot find members of the open cluster. It requires to parameterize
-some of the properties of the cluster, what requires a previous knowledge of the cluster.
-Or to make a supervised and manual selection of groups based on overdensities
-in the proper motion configuration space. Clusterix 2.0, an interactive web-based tool,
-can help to this process. It takes the proper motion diagram without making any prior
-assumption about the membership of the candidate star and determines empirically
-the frequency functions. Clusterix uses normal Gaussian kernel functions, defined as:
+In~\cite{castro2020hunting}, they present one of the initial approach to detect OCs.
+They search for overdensities in the astrometric space of the galactic disk, and then,
+they identify with photometric information possible OCs. Although it looks simple,
+it is quite difficult to face the problem that the near field around the OC has two
+kind of populations. The first population belongs to the membership stars of the OC
+(from tens or hundreds to a few thousands), while the second population has a background
+of stars that do not belong to the OC (from tens to hundreds of thousands). Thus,
+we summarize our problem to find out which stars belong to the OC. This is a handicap
+work because the selection is critical to properly characterize the fundamental
+properties of the cluster (total mass, age, chemical composition, between others).
+
+Other works, such as TOPCAT~\cite{taylor2005topcat}, Clusterix 2.0~\cite{balaguer2020clusterix},
+Aladin~\cite{bonnarel2000aladin} or VOSA~\cite{bayo2008vosa}, perform as a first step
+to analyze the proper motion configuration space of the region of interest.
+
+TOPCAT cannot find open cluster, instead it requires to parameterize some of the
+properties of the cluster. This last process requires a previous knowledge of the cluster.
+In other words, they perform a supervised and manual selection of groups based on
+overdensities in the proper motion configuration space. Clusterix 2.0 is an interactive
+web-based tool that can help to the selection of groups. It takes the proper motion
+diagram without making any prior assumption about the membership of the candidate star
+and determines empirically the frequency functions. It employs normal Gaussian kernel
+functions, defined as:
 
 \begin{equation*}
-  K(a, b) = \frac{1}{2 \pi h^{2}} \exp{ \left[ - \frac{1}{2}\frac{\left( a - a_{i} \right)^{2} + \left( b - b_{i} \right)^{2}}{ h^{2}} \right]}
+  K(a, b) = \frac{1}{2 \pi h^{2}} \exp{ \left[ - \frac{1}{2}\frac{\left( a - a_{i} \right)^{2} + \left( b - b_{j} \right)^{2}}{ h^{2}} \right]}
 \end{equation*}
 
-$\left( a, b \right)$ are referred to the proper motion configuration space,
-while a point located in the center of the cell $\left( i, j \right)$
-provides the maximum contribution for calculating the local density.
-Number $h$ is called the \emph{smoothing} parameter and it is measured
-in the same units as proper motion.
-
-Clusterix is defined as an unsupervised and non-parameterized tool, but it
-critically depends on the selection of three radii in the studied region.
-The inner radius should contain stars belonging or not to the cluster.
-While the outermost radii define a ring which only contains stars that do
-not belong to the cluster. In addition, it is also sensitive to other
-parameters, such as the soften parameter \emph{h}, or other restrictions
-related with the searched proper motions.
-
-If the analysis performed by Clusterix is successful, it returns the probability
-associated to each star in the dataset to belong to the open cluster.
-These results can be imported then into TOPCAT to continue the refining process to
-get a valid characterization.
-
-Another recent approach to solve this problem based on machine learning techniques
-makes a systematic search for overdensities in the astrometric space of the galactic
-disk and a subsequent identification of open clusters using photometric information
-\cite{castro2020hunting}.
-
-This last method includes two phases: the first one uses an unsupervised clustering
-algorithm, DBSCAN, to search for overdensities $(l, b \pi, \mu_{\alpha} *, \mu_{\delta})$,
-and then applies a deep learning ANN, previously trained with magnitude diagrams,
-to identify isochrone patterns within the detected overdensities and thus proceed
-to confirm them as OC.
-
-It should be noted that for the execution of this method, MareNostrum 42
-(Barcelona Supercomputing Center) was used.
-So the neural network could handle the image recognition process with
-isochrone patterns and not applying theoretical models derived from
-values such as metallicity or masses, among other.
-
-This work concludes with the identification of 582 new open clusters
-distributed along the galactic disk for a galactic declination bellow 20 degrees.
-This result increases the number of known open clusters by 45\%.
-
-As told before, our aim in this work, linked to computational limitations compared
-to MareNostrum, is not to do a blind search for clusters, but to obtain a method,
-in the machine learning environment, that allows the characterization of open clusters.
-We require our method to be unsupervised and non-parameterized making it suitable for
-automated processes.
-
-We use the Clusterix+TOPCAT method applied to the same clusters that we have analyzed
-with our method to validate our results.
-
-In this work we make use of Gaia DR2 since DR3 has not been released in time for us to include it.
-Gaia DR2 is a multidimensional dataset obtained by ESA's Gaia mission
-(located at L2, 1.5 million kilometers from Earth) and operational since 2014.
-The catalogue has high precision and accuracy astrometric data for more than 1.7 billion stellar sources,
-and magnitudes in three photometric filters (G, BP and RP) for more than 1,300 million sources.
-
-As we will explain later, what we need to accomplish our aim is an algorithm
-that manages to make data groups based on the dynamic properties of the stars.
-
-There are several clustering algorithms: K-Means, Mean-Shift Clustering, DBSCAN \cite{ester1996density}
-among other. While each one behaves better according to the distribution of the objects
-to clusterize, we have chosen K-Means by its simplicity and good results.
-
-This algorithm works well and gives good results at first approach.
-However, we need to set a large number of clusters to find the open cluster we are looking for.
-This fact complicates the identification of the OC and many times the found cluster contains to many outliers.
-For that reason, we have search for a K-Means refinement based on an artificial neural network.
-
-The \emph{Unsupervised Deep Embedding for Clustering Analysis} model
-or \emph{DEC} \cite{xie2016unsupervised} takes K-Means as its starting point,
-but then, it trains an autoencoder to reduce the feature space
-and pass this transformed data through a Clustering Layer which refines the previous selection.
+where \(( a, b )\) are the proper motion configuration space, \(( i, j )\) is a point
+located in the center of that provides the maximum contribution for calculating the
+local density, and \(h\) is the \emph{smoothing} parameter, which it is measured in the
+same units as the proper motion.
+
+Clusterix critically depends on the selection of three radii in the studied region.
+The inner radius contains stars belonging or not to the cluster. While the outermost
+radii defines a ring which only contains stars that do not belong to the cluster.
+In addition, it is also sensitive to other parameters, such as the soften parameter
+\emph{h}, or other restrictions related with the searched proper motions.
+
+If the analysis performed by Clusterix is successful, it returns the probability of
+each star to belong to the open cluster. The results can be imported then into TOPCAT
+to continue the refining process to get a valid characterization.
+
+Another recent approach~\cite{castro2020hunting} solves the problem using machine
+learning techniques. It searches systematically for overdensities in the astrometric
+space of the galactic disk and a subsequent identification of open clusters using
+photometric information. It includes two phases:
+
+\begin{enumerate}
+  \item it employs DBSCAN, an unsupervised clustering algorithm, to search for
+        overdensities,
+  \item and it applies a deep learning ANN to identify isochrone patterns within
+        the detected overdensities and thus proceed to confirm them as OCs. They
+        trained previously the ANN with magnitude diagrams.
+\end{enumerate}
+
+They execute their experiments in the Barcelona Supercomputing Center, MareNostrum 42.
+Thereby, the neural network handles the image recognition process with isochrone
+patterns without applying theoretical models derived from values such as metallicity
+or masses, among other. Their work concludes with the identification of 582 new open
+clusters distributed along the galactic disk for a galactic declination bellow 20
+degrees, increasing the number of known open clusters by 45\%.
+
+In contrast to~\cite{castro2020hunting}, we lack of a super computer like MareNostrum
+for our experiments. For this reason, our aim is to obtain a new novel method that
+allows the characterization of open clusters without doing a blind search for clusters
+characterization. We advocate of the idea of building an unsupervised and
+non-parameterized method. So, it an be suitable for automated processes.
 

### 2021-02-09 paper/paper.tex
Update Abstract

@@
 \resumen{
-The characterization and understanding of \emph{Open Clusters} (OCs)
-allow us to understand better properties and mechanisms about the Universe
-such as stellar formation and the regions where these events occur.
-They also provide information about stellar processes and the evolution
-of the galactic disk.
-
-In this paper, we present a novel method to characterize OCs.
-Our method employs a model built on \emph{Artificial Neural Networks} (ANNs).
-More specifically, we adapted a state of the art model, the
-\emph{Deep Embedded Clustering} (DEC) model for our purpose.
-
-The developed method aims to improve classical state of the arts techniques.
-We improved not only in terms of computational efficiency, (with lower
-computational requirements), but in usability (reducing the number
-of hyperparameters to get a good characterization of the analyzed clusters).
-For our experiments, we used the \emph{Gaia DR2 database} as the data source,
-and compared our model with the clustering technique \emph{K-Means}.
-
-Our method achieves good results,
-becoming even better (in some of the cases) than current techniques.
+The characterization and understanding of \emph{Open Clusters} (OCs) allow us to
+understand better properties and mechanisms about the Universe such as stellar
+formation and the regions where these events occur. They also provide information
+about stellar processes and the evolution of the galactic disk.
+
+In this paper, we present a novel method to characterize OCs. Our method employs a
+model built on \emph{Artificial Neural Networks} (ANNs). More specifically, we adapted
+a state of the art model, the \emph{Deep Embedded Clustering} (DEC) model for our purpose.
+The developed method aims to improve classical state of the arts techniques. We improved
+not only in terms of computational efficiency (with lower computational requirements),
+but in usability (reducing the number of hyperparameters to get a good characterization
+of the analyzed clusters). For our experiments, we used the \emph{Gaia DR2 database} as
+the data source, and compared our model with the clustering technique \emph{K-Means}. Our
+method achieves good results, becoming even better (in some of the cases) than current techniques.
 }

### 2021-02-09 thesis/thesis.tex
thesis: Minor corrections
thesis: Fix appending paper
Update Abstract

@@
 
-The characterization and understanding of \emph{Open Clusters} allow us to better understand properties and mechanisms about the Universe
-such as stellar formation and the kind of regions where these events occur.
-They also provide information about stellar processes and the evolution of the galactic disk.
-
-In this work, we develop a method to characterize open clusters by using \emph{artificial intelligence} tools,
-such as clustering by \emph{K-Means} and clustering based on \emph{Artificial Neural Networks} by implementing the \emph{Deep Embedded Clustering} model.
-We are using \emph{Gaia DR2 database} as data source for testing our models.
-
-The developed method aims to improve current techniques of the state of the art.
-We achieve improvements not only in terms of \emph{computational efficiency}, with lower computational requirements, but in \emph{usability},
-reducing the number of hyperparameters to obtain a good characterization of the analyzed clusters.
-
-Our method achieves good results, becoming even better in some cases when the results are compared with current techniques.
+The characterization and understanding of \emph{Open Clusters} (OCs) allow us to
+understand better properties and mechanisms about the Universe such as stellar
+formation and the regions where these events occur. They also provide information
+about stellar processes and the evolution of the galactic disk.
+
+In this work, we present a novel method to characterize OCs. Our method employs a
+model built on \emph{Artificial Neural Networks} (ANNs). More specifically, we adapted
+a state of the art model, the \emph{Deep Embedded Clustering} (DEC) model for our purpose.
+The developed method aims to improve classical state of the arts techniques. We improved
+not only in terms of computational efficiency (with lower computational requirements),
+but in usability (reducing the number of hyperparameters to get a good characterization
+of the analyzed clusters). For our experiments, we used the \emph{Gaia DR2 database} as
+the data source, and compared our model with the clustering technique \emph{K-Means}. Our
+method achieves good results, becoming even better (in some of the cases) than current techniques.
 
@@
 
-La caracterización y conocimiento de \emph{Cúmulos Abiertos} permite conocer mejor propiedades y mecanismos del Universo tales como
-la formación de estrellas y las regiones donde se dan estos procesos.
-También permiten obtener información sobre procesos estelares y la evolución del disco galáctico.
-
-En este trabajo se desarrolla un método para caracterizar estos cúmulos mediante el uso de herramientas de \emph{inteligencia artificial}
-como agrupamiento por \emph{K-Medias} o agrupamiento basado en \emph{Redes Neuronales Artificiales} mediante la implementación del
-modelo \emph{Deep Embedded Clustering} aplicados sobre conjuntos de datos extraídos de la base de datos de \emph{Gaia DR2}.
-
-El método desarrollado pretende mejorar los ya existentes tanto en términos de \emph{eficiencia computacional}, con menores requisitos de cálculo,
-como en \emph{facilidad de uso}, reduciendo el número de hiperparámetros a configurar para obtener una buena caracterización de los cúmulos analizados.
-
-Como se detalla más adelante, nuestro método consigue buenos resultados llegando a ser incluso mejor en algunos casos cuando se comparan
-los resultados con otras técnicas actuales.
+La caracterización y conocimiento de \emph{Cúmulos Abiertos} permite conocer mejor
+propiedades y mecanismos del Universo tales como la formación de estrellas y las
+regiones donde se dan estos procesos. También permiten obtener información sobre
+procesos estelares y la evolución del disco galáctico.
+
+En este trabajo presentamos un método novedoso para caracterizar estos cúmulos.
+Nuestro método hace uso de un modelo basado en \emph{Redes Neuronales Artificiales}.
+Más concretamente, adaptamos un modelo del estado del arte, el \emph{Deep Embedded
+Clustering} (DEC), a nuestro problema. El método desarrollado tiene como objetivo
+mejorar las técnicas clásicas del estado del arte. Con nuestro método, no sólo
+mejoramos en términos de eficiencia de cálculo (consiguiendo menores requisitos
+computacionales), también mejoramos en usabilidad (reduciendo el número de
+hiperparámetros para conseguir una buena caracterización de los cúmulos analizados).
+Para nuestro experimentos, usamos la base de datos \emph{Gaia DR2} como fuente de datos,
+y comparamos nuestro modelos con el algoritmo de clustering \emph{K-Medias}. Nuestro
+método consigue buenos resultados, siendo incluso mejor (en algunos casos) que las
+técnicas actuales.
 
@@
 All these clusters have a wide variety of diameters, from 25 to 330 arcmin,
-as well as the number of stars that belong to them.
-NGC 2682 has 3,000 stars while Melotte 25 is composed by more than 400,000 stars.
+as well as the number of stars that belong to them. NGC 2682 has 3,000 stars
+while Melotte 25 is located inside a region with more than 400,000 stars.
 More information about the studied clusters is available in Table \ref{tab:clusters_summary}.
@@
 
-\includepdf[pages={-},noautoscale]{../paper/paper.pdf}
+\setboolean{@twoside}{false}
+\addtocontents{toc}{\protect\renewcommand{\protect\numberline}[1]{}}
+\includepdf[pages=-, offset=75 -75, addtotoc={1, chapter, 1, Paper, chap:paper}]{../paper/paper.pdf}
 

### 2021-02-09 paper/paper.tex
paper: Add advisor corrections
Add @cguz corrections and suggestions

Co-Authored-By: Cesar Guzmann <15159632+cguz@users.noreply.github.com>

@@
 The study of OCs advanced thanks to the huge and precise dataset
-from the Gaia DR2 mission, available since 2018 from
-\cite{collaboration2016description} and \cite{gaia2018gaia}.
-Although, the Gaia DR2 database misreported the chemical composition
-of some OCs, the database has helped to review already known OCs and
+from the Gaia DR2 mission, available since 2018 from~\cite{collaboration2016description}
+and~\cite{gaia2018gaia}. Although, the Gaia DR2 database misreported the chemical
+composition of some OCs, the database has helped to review already known OCs and
 to find new ones.
@@
 
-Normally, the characterization process considers astrometic features such as
+Normally, the characterization process considers astrometric features such as
 proper motion in right ascension and declination, or parallax.
@@
 OCs inside a stellar region with no previous knowledge about the region.
-Our model takes advantage of some features such as proper motion in right ascension
-and declination, and parallax to train an Artificial Neural Network.
-The ANN clusters all the stars within given region in groups.
+Our model takes advantage of some features (such as proper motion in right ascension
+and declination, and parallax) to train an Artificial Neural Network.
+The ANN clusters all the stars within a given region in groups.
 One of those groups will be the open cluster that interests us.
@@
 This paper is organized as follows. Section~\ref{sec:related_works} presents
-some related works and Section~\ref{sec:aims_and_method} describes in more detail
+some related works and Section~\ref{sec:aims_and_method} describes in detail
 our model. The following Section~\ref{sec:results}, shows the results and compares
@@
 
-In~\cite{castro2020hunting}, they present one of the initial approach to detect OCs.
-They search for overdensities in the astrometric space of the galactic disk, and then,
-they identify with photometric information possible OCs. Although it looks simple,
-it is quite difficult to face the problem that the near field around the OC has two
-kind of populations. The first population belongs to the membership stars of the OC
-(from tens or hundreds to a few thousands), while the second population has a background
-of stars that do not belong to the OC (from tens to hundreds of thousands). Thus,
-we summarize our problem to find out which stars belong to the OC. This is a handicap
-work because the selection is critical to properly characterize the fundamental
-properties of the cluster (total mass, age, chemical composition, between others).
-
-Other works, such as TOPCAT~\cite{taylor2005topcat}, Clusterix 2.0~\cite{balaguer2020clusterix},
-Aladin~\cite{bonnarel2000aladin} or VOSA~\cite{bayo2008vosa}, perform as a first step
-to analyze the proper motion configuration space of the region of interest.
-
-TOPCAT cannot find open cluster, instead it requires to parameterize some of the
-properties of the cluster. This last process requires a previous knowledge of the cluster.
-In other words, they perform a supervised and manual selection of groups based on
-overdensities in the proper motion configuration space. Clusterix 2.0 is an interactive
-web-based tool that can help to the selection of groups. It takes the proper motion
-diagram without making any prior assumption about the membership of the candidate star
-and determines empirically the frequency functions. It employs normal Gaussian kernel
-functions, defined as:
+In~\cite{castro2020hunting}, they present one of the initial approaches to detect OCs.
+They search for overdensities in the astrometric space of the galactic disk. Then,
+they identify with photometric information possible OCs. Although it seems easy,
+it is quite hard to face the problem that the near field around the OC has two kinds of
+populations. The first population belongs to the members of the OC stars (from tens or
+hundreds to a few thousand). The second population has a background of stars that do
+not belong to the OC (from tens to hundreds of thousands). Thus, we summarize our problem
+in finding out which stars belong to the OC.
+
+Other works (TOPCAT~\cite{taylor2005topcat}, Clusterix 2.0~\cite{balaguer2020clusterix},
+Aladin~\cite{bonnarel2000aladin}, or VOSA~\cite{bayo2008vosa}), first analyze the proper
+motion configuration space of the region of interest. TOPCAT cannot find an open cluster.
+Instead, it requires parameterizing some of the properties of the cluster. This last
+process requires previous knowledge of the cluster. In other words, they perform a supervised
+and manual selection of groups based on overdensities in the proper motion configuration space.
+Clusterix 2.0 is an interactive web-based tool that can help the choice of groups. Take the
+proper motion diagram without making any previous assumptions about the membership of the
+candidate star. And empirically determine the frequency functions. It employs normal Gaussian
+kernel functions, defined as:
 
@@
 characterization. We advocate of the idea of building an unsupervised and
-non-parameterized method. So, it an be suitable for automated processes.
+non-parameterized method. So, it can be suitable for automated processes.
 
@@
 
-Our aim is to \emph{build an unsupervised clustering model for open cluster characterization}.
-The model will be \emph{non-supervised and non-parameterized} in order to fit a wide range
-of clusters without the need for fine-tuning a high number of hyperparameters.
-
-\subsection{DOWNLOAD}
-
-We start by selecting a region from the OpenClust catalogue and downloading it from
-Gaia DR2 database into a docker database. This allows us to repeat the analysis without
-having to download the data every time. The radius of the downloaded region from Gaia
-is 1.5 times larger than the one registered in OpenClust. This way we ensure to include
-several stars that do not belong to the open cluster.
-
-\subsection{FEATURE SELECTION}
+In the literature, we have several clustering algorithms, such as K-Means, Mean-Shift
+Clustering, DBSCAN~\cite{ester1996density} among other, that can help us to achieve
+our aim. Each one behaves better according to the distribution of the objects to
+clusterize. However, we need to set a large number of clusters to find the open cluster
+we are looking for. It complicates the identification of the OC because most of the
+time falls in many outliers.
+
+The unsupervised \emph{Deep Embedding for Clustering Analysis} (\emph{DEC})
+model~\cite{xie2016unsupervised} is a refinement K-Means based on ANN. It starts with
+K-Means. And then, it trains an autoencoder to reduce the feature space and pass them
+through a Clustering Layer which refines the previous selection.
+
+With all this in mind, our method is based on DEC model. We adapted it to manage data
+groups based on the dynamic properties of the stars. Thus, we have an unsupervised
+clustering model for open cluster characterization. The model fits a wide range of
+clusters without the need for fine-tuning a high number of hyperparameters.
+
+\subsection{Selection of the catalogue}
+
+In this work we make use of Gaia DR2 since DR3 has not been released in time for us to
+include it. Gaia DR2 is a multidimensional dataset obtained by ESA's Gaia mission
+(located at L2, 1.5 million kilometers from Earth) and operational since 2014. The
+catalogue has high precision and accuracy astrometric data for more than 1.7 billion
+stellar sources, and magnitudes in three photometric filters (G, BP and RP) for more
+than 1,300 million sources.
+
+We start by selecting a region from the OpenClust~\cite{dias2002new} catalogue and
+we downloaded it from Gaia DR2 database. The radius of the downloaded region from Gaia
+is 1.5 times larger than the one registered in OpenClust, ensuring to include several
+stars that do not belong to the open cluster.
+
+\subsection{Selection of features}
 \label{sec:feature_selection}
 
-\begin{figure}[htbp]
-  \includegraphics[width=\columnwidth]{../figures/melotte_22/features_melotte_22.png}
+For the selection of features, we studied the Melotte 22 dataset. Table~\ref{tab:
+sample_features_melotte22_region} describes an example of the region Melotte 22. Each
+row represents the features of a star.
+
+\begin{table}[!htbp]
+  \begin{center}
+    \resizebox{\columnwidth}{!}{
+      \begin{tabular}{c|c|c}
+        % \textbf{source\_id} &
+        \emph{pmra $(mas \cdot yr^{-1})$} & \emph{pmdec $(mas \cdot yr^{-1})$} & \emph{parallax $(mas)$} \\
+        \hline
+        % 64035217900973056 &
+        2.848682 & -3.291204 & 0.399680 \\
+        % 64035217900973440 &
+        0.894901 & -3.445501 & 0.416639 \\
+        % 64035217901215616 &
+        7.924372 & -0.241281 & 0.397743 \\
+        % 64035351043566208 &
+        -4.433802 & -2.584965 & 0.410695 \\
+        % 64035385403304960 &
+        0.055990 & -1.760018 & 0.413813 \\
+      \end{tabular}
+    }
+    \caption{Features of Melotte 22 region.}
+    \label{tab:sample_features_melotte22_region}
+  \end{center}
+\end{table}
+
+Proper motion in right ascension (column pmra) and declination (column pmdec) seems
+like a natural choice since stars belonging to the same OC share a common motion vector.
+Parallax (column parallax) is another relevant feature. It lets us know how far stars
+are from us. Besides, since all stars within an open cluster were born from the same
+dust cloud, they must all have similar parallax.
+
+However, we are not going to use these raw features. Instead, we are taking a combination
+of them. Figure~\ref{fig:features_melotte_22} shows a pairwise relationship of our
+combination of features from Melotte 22.
+
+\begin{figure*}[!hbt]
+  \centering
+  \includegraphics[scale=0.5]{../figures/melotte_22/features_melotte_22.png}
   \caption{Pairwise relationships among variables using Melotte 22 data}
   \label{fig:features_melotte_22}
-\end{figure}
-
-As mentioned before, we want our model to characterize open clusters by looking at
-their dynamic properties. Also, we want to maintain as simple as possible our
-clustering model in order to save computing resources.
-
-Proper motion in right ascension and declination seems like a natural choice since,
-as we know, stars belonging to the same OC share a common motion vector.
-
-Parallax is another important feature. It lets us know how far stars are from us.
-In addition, since all stars within an open cluster were born from the same dust cloud,
-they must all have similar parallax.
-
-However, we are not going to use these raw features.
-Instead, we are taking a combination of them.
+\end{figure*}
 
 First, we correct proper motion in right ascension and declination by dividing them by
-the parallax. That way, we normalize these quantities and help our clustering models
-to improve their performance.
-
-The modulus of the proper motion is another computed property that we are considering.
-We use it to relate both features and therefore, to force our model to keep them tight.
-
-Figure \ref{fig:features_melotte_22} shows a pairwise relationship among some features
-available in the dataset for Melotte 22 data.
-
-\subsection{SOFT CLUSTERING}
-
-Our first approach to find the open cluster is using the K-Means algorithm.
-Since we are looking for a single cluster, it seems reasonable to use a clustering
-algorithm and set it to find two clusters. One for the desired OC and another which
-contains stars that do not belong to the open cluster.
-However, this idea is not completely right.
-
-This is due to the fact that OC's stars are surrounded by other stars with possibly similar
-properties. Therefore, setting the number of clusters to two is too low to separate them properly.
-
-\begin{figure}
-  \centering
-  \begin{subfigure}{\columnwidth}
-    \centering
-    \begin{subfigure}[t]{0.3\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n2_pm_melotte_22.png}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.3\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n5_pm_melotte_22.png}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.3\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n8_pm_melotte_22.png}
-    \end{subfigure}
-  \end{subfigure}
-  \medskip
-  \begin{subfigure}{\columnwidth}
-    \centering
-    \begin{subfigure}[t]{0.3\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n2_parallax_melotte_22.png}
-      \caption{N = 2}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.3\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n5_parallax_melotte_22.png}
-      \caption{N = 5}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.3\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n8_parallax_melotte_22.png}
-      \caption{N = 8}
-    \end{subfigure}
-  \end{subfigure}
-  \caption{K-Means comparisons with Melotte 22}
-  \label{fig:kmeans_comparisons_melotte_22}
-\end{figure}
-
-As shown in Figure \ref{fig:kmeans_comparisons_melotte_22}, larger values for the number of
-clusters allow us to isolate more accurately the resonance in parallax at $\approx 7.3mas$.
-However, we have the disadvantage that more groups are formed.
-
-This effect complicates our task of finding the desired open cluster, since we would like
-to get just two groups, one for the OC and another with the remaining stars.
-Therefore, we have to find a way to set the right value for the number of clusters
-to isolate the searched cluster without creating too many groups.
-
-To solve this issue, we will try to estimate the best number of clusters by using the
-\emph{silhouette score} \cite{rousseeuw1987silhouettes}.
-
-K-Means does a good job making an initial clustering. However, too many clusters arise
-from this characterization and the OC is still polluted with stars that do not belong to it.
-Moreover, we would like to reduce the amount of clusters too.
-
-Therefore, we would like to find a better model that improves this initial characterization
-by reducing the amount of clusters and also that removes outsiders from the OC.
-
-\subsection{DEC MODEL}
-
-Since we do not have a labeled dataset to train a supervised model,
-we have no choice but to use an unsupervised self-trained model.
-
-For that reason we have adapted the \emph{Unsupervised Deep Embedding for Clustering Analysis}
-to our work.
-
-The model is composed by a \emph{deep autoencoder}, a \emph{deep encoder} and
-a \emph{clustering layer}.
-
-Encoders are used to transform the input data into a latent space
-using a non-linear mapping function $f_{\theta} : X \rightarrow Z$.
-
-Although, as explained in section \ref{sec:feature_selection},
-the number of features we are managing is not too large,
-this latent space helps us reduce the number of features
-and avoids the \emph{"curse of dimensionality"} \cite{bellman1961curse}.
-
-These encoders are pretrained before fitting the model to generate predictions. Then,
-the encoder is used with the aim of transforming input data to the latent space $Z$. Once the
-data has been transformed, a K-Means clusterer is used in order to make an initial clustering.
+the parallax (variables pmra\_corr and pmdec\_corr, respectively). That way, we normalize
+these quantities and help our clustering models to improve their performance. The modulus
+of the proper motion (variable pmmod in Figure~\ref{fig:features_melotte_22}) is another
+computed property that we considered. We use it to relate both features and therefore
+force our model to keep them tight.
+
+\subsection{Deep Open Clustering of stars}
+\label{sec:deep_open_clustering}
+
+In this section we explain in more detail our model.
+
+In order to define our model, one of the main problems is that we do not have a labeled
+dataset to train a supervised model. Thus, we have to deal with an unsupervised self-trained
+model. We adapted the Unsupervised \emph{DEC} model~\cite{xie2016unsupervised} to our requirements.
+
+Same as the DEC model, we have two main components:
+
+\begin{itemize}
+  \item \textbf{deep autoencoder:} It is trained before passing the data through the
+        \emph{clustering layer}. It is composed by encoder layers followed by decoder
+        layers. Recent research has shown that this autoencoder provides meaningful and
+        well-separated representations on real-world datasets~\cite{vincent2010stacked}
+        and~\cite{hinton2006reducing} to the DEC model.
+  \item \textbf{clustering layer:} This layer recibes data transformed by the autoencoder
+        and it is iteratively trained until a convergence criterion is met.
+\end{itemize}
+
+The deep autoencoder is used to transform the input data into a latent space using a
+non-linear mapping function \(f_{\theta} : X \rightarrow Z\).
+
+As it was described in Section~\ref{sec:feature_selection}, the number of features we
+deal is not too large. This latent space helps us to start in a reduced number of
+features and avoids the \emph{"curse of dimensionality"}~\cite{bellman1961curse}.
+
+The autoencoder is pretrained before fitting the model to generate predictions. Then,
+the encoder layers of the autoencoder are used with the aim of transforming input data to
+the latent space \(Z\). Once the data has been transformed, a K-Means clusterer is used in
+order to make an initial clustering. K-Means cluster centers are used as the initial
+weights for the clustering layer.
 
 With that initial configuration, the model iterates alternating between computing an auxiliary
-target distribution (Soft Assignment) and minimizing the Kullback-Leiber (KL) divergence
-\cite{kullback1951information} to it. This unsupervised algorithm allows us to improve the clustering.
+target distribution (Soft Assignment) and minimizing the Kullback-Leiber (KL)
+divergence~\cite{kullback1951information} to it.
+This unsupervised algorithm allows us to improve the clustering.
 
@@
 The choice of this target distribution is crucial for DEC's performance.
-In this work we have taken the target distribution from DEC's original paper \cite{xie2016unsupervised},
-which is defined in Equation \ref{eq:student_tdistribution}.
-
-Figure \ref{fig:dec_model_setup} shows the layer setup of our DEC model.
-It is simpler than the one tested on the original paper \cite{xie2016unsupervised},
+In this work we have taken the target distribution from DEC's original paper~\cite{xie2016unsupervised},
+which is defined in Equation~\ref{eq:student_tdistribution}.
+
+\begin{figure*}[!hbt]
+  \centering
+  \includegraphics[scale=0.8]{../figures/dec_diagram.pdf}
+  \caption{DEC model layer setup}
+  \label{fig:dec_model_setup}
+\end{figure*}
+
+Figure~\ref{fig:dec_model_setup} shows the layer setup of our DEC model.
+It is simpler than the one tested on the original paper~\cite{xie2016unsupervised},
 since the number of selected features in our work is smaller than in the original one.
@@
 
-\begin{figure}
-  \includegraphics[width=\columnwidth]{../figures/dec_diagram.pdf}
-  \caption{DEC model layer setup}
-  \label{fig:dec_model_setup}
-\end{figure}
+\begin{figure*}[!htbp]
+  \centering
+  \begin{subfigure}[b]{0.3\textwidth}
+    \centering
+      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n2_pm_melotte_22.png}
+      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n2_parallax_melotte_22.png}
+      \caption{N = 2}
+  \end{subfigure}
+  \medskip
+  \begin{subfigure}[b]{0.3\textwidth}
+    \centering
+      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n5_pm_melotte_22.png}
+      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n5_parallax_melotte_22.png}
+      \caption{N = 5}
+  \end{subfigure}
+  \medskip
+  \begin{subfigure}[b]{0.3\textwidth}
+    \centering
+      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n8_pm_melotte_22.png}
+      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n8_parallax_melotte_22.png}
+      \caption{N = 8}
+  \end{subfigure}
+  \caption{K-Means with Melotte 22}
+  \label{fig:kmeans_comparisons_melotte_22}
+\end{figure*}
 
 Finally, we can refine this selection by filtering those stars which are bellow and above
-the 0.10 and 0.90 quantiles for each group, respectively.
-That way we remove the most doubtful values from the selection.
-
-\begin{figure}[htbp]
-  \centering
-  \begin{subfigure}{\columnwidth}
-    \centering
-    \begin{subfigure}[t]{0.3\textwidth}
+the 0.10 and 0.90 quantiles for each group, respectively. That way we remove the most
+doubtful values from the selection.
+
+\paragraph{Discussion.}
+
+Since we are looking for a single cluster, it seems reasonable to use a clustering algorithm
+like K-Means to find two clusters. One for the desired OC and another for stars that do not
+belong to the OC. However, the idea is not completely right. This is due to the fact that OC's
+stars are surrounded by other stars with possibly similar properties. Therefore, setting the
+number of clusters to two is too low to separate them properly.
+
+Figure \ref{fig:kmeans_comparisons_melotte_22} shows an example. It describes the results of
+applying K-Means with N=2, N=5 and N=8 in Melotte 22. As we can see, large values of the number
+of clusters allow us to isolate more accurately the resonance in parallax at \(\approx 7.3mas\).
+
+The disadvantage is that it forms more groups of stars, which complicates the task of finding the
+desired OC, since we would like to get just two groups. Therefore, with a K-Means we have to find
+a way to set the right value for the number of clusters to isolate the searched cluster without
+creating too many groups. To solve this issue, we can try to estimate the best number of clusters
+by using the \emph{silhouette score}~\cite{rousseeuw1987silhouettes}.
+
+K-Means does a good job making an initial clustering. However, too many clusters arise
+from this characterization and the OC is still polluted with stars that do not belong to it.
+Moreover, we would like to reduce the amount of clusters too.
+
+\section{Results}
+\label{sec:results}
+
+%% Comparación con K-Means, DEC y DEC filter
+\begin{figure*}[!htbp]
+  \centering
+  \begin{subfigure}[b]{0.3\textwidth}
     \centering
       \includegraphics[width=\textwidth]{../figures/melotte_22/kmeans_pm_melotte_22.png}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.3\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/melotte_22/dec_pm_melotte_22.png}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.3\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/melotte_22/dec_pm_filtered_melotte_22.png}
-    \end{subfigure}
-  \end{subfigure}
-  \medskip
-  \begin{subfigure}{\columnwidth}
-    \centering
-    \begin{subfigure}[t]{0.3\textwidth}
-      \centering
       \includegraphics[width=\textwidth]{../figures/melotte_22/kmeans_hr_diagram_melotte_22.png}
@@
   \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.3\textwidth}
-      \centering
+  \medskip
+  \begin{subfigure}[b]{0.3\textwidth}
+    \centering
+      \includegraphics[width=\textwidth]{../figures/melotte_22/dec_pm_melotte_22.png}
       \includegraphics[width=\textwidth]{../figures/melotte_22/dec_hr_diagram_melotte_22.png}
@@
   \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.3\textwidth}
-      \centering
+  \medskip
+  \begin{subfigure}[b]{0.3\textwidth}
+    \centering
+      \includegraphics[width=\textwidth]{../figures/melotte_22/dec_pm_filtered_melotte_22.png}
       \includegraphics[width=\textwidth]{../figures/melotte_22/dec_hr_diagram_filtered_melotte_22.png}
@@
   \end{subfigure}
-  \end{subfigure}
   \caption{Evolution of Melotte 22 characterization}
   \label{fig:melotte_22_characterization_evolution}
-\end{figure}
-
-Figure \ref{fig:melotte_22_characterization_evolution} shows the evolution
+\end{figure*}
+
+We have tested our model with the Melotte 25 dataset, which is one of the most difficult dataset
+to characterize OC by its distribution and extension. In order to support the discussion presented
+in Section~\ref{sec:deep_open_clustering}, we compared our model against a K-Means classification
+algorithm. In addition, we perform a second execution only in our model where we filtered the
+quantiles in the Melotte 22 dataset.
+
+Our model was implemented in Python 3.8 using the Keras 2.2 framework and Jupyter
+Notebooks~\cite{Kluyver2016jupyter}. All tests were run on an Apple Mac Pro Late 2013 with a
+2.7GHz 12-Core Intel Xeon E5-2697v2, and 64GB RAM 1866MHz DDR3. For the GPU, we use a graphic
+card AMD FirePro D700 with 6GB~\footnote{All resources developed for this project are available
+at \href{https://github.com/cdalvaro/machine-learning-master-thesis}
+{https://github.com/cdalvaro/machine-learning-master-thesis}}.
+
+% More results are available at Appendix \ref{sec:appendix}.
+
+Figure~\ref{fig:melotte_22_characterization_evolution} shows the evolution
 of the characterization of Melotte 22. The first column corresponds to the
@@
 
-All resources developed for this project are available at the GitHub repository:
-\href{https://github.com/cdalvaro/machine-learning-master-thesis}{cdalvaro/machine-learning-master-thesis}
-
-It includes the DEC model developed in Python using the Keras framework
-and Jupyter Notebooks \cite{Kluyver2016jupyter} with examples.
-
-\section{CONTRIBUTION}
-
-The developed method is successful since it meets all our established requirements:
-
-\begin{itemize}
-  \item Our method gives valid results compatible with current methods as described in Section \ref{sec:results}.
-  \item As required from the beginning, it is non-supervised and non-parameterized.
-  \item The model shows a great performance compared with current methods.
-  \item It is able to deal with large and open regions. As an example,
-        it has succeeded characterizing Melotte 25 (Hiades) OC. Also,
-        the relative proximity of this cluster, 45 parsec, makes it difficult to
-        characterize. Clusterix fails characterizing this cluster. In this case,
-        our method proposes at least two more non-catalogued clusters. The analysis
-        of these new clusters brings out an interesting stellar dynamics in the studied
-        region. It could mean that there was a recent interaction among different molecular clouds.
-\end{itemize}
-
-For all this, we consider that the presented method is good enough to be included
-in the Virtual Observatory toolset. Only a few tweaks must be done in order to make
-our model compatible with other tools of the VO.
-
-\section{RESULTS}
-\label{sec:results}
-
-We now present the results obtained after having analyzed Melotte 25.
-This is the largest cluster we have studied and one of the most difficult
-to characterize by its distribution and extension.
-
-More results are available at Appendix \ref{sec:appendix}.
+%% Comparación con K-Means, DEC y DEC filter
 
@@
 
-\begin{table}[htbp]
+\begin{table}[!htbp]
   \begin{center}
@@
       \begin{tabular}{l|c|c|c|c}
-        \textbf{Method} & \emph{$\mu_{\alpha}$ $(mas/yr^{-1})$} & \emph{$\mu_{\delta}$ $(mas/yr^{-1})$} & \emph{$\varpi$ $(mas)$} & \emph{\# stars} \\
+        \textbf{Method} & \emph{$\mu_{\alpha}$ $(mas \cdot yr^{-1})$} & \emph{$\mu_{\delta}$ $(mas \cdot yr^{-1})$} & \emph{$\varpi$ $(mas)$} & \emph{\# stars} \\
         \hline
@@
 
-\section{RESULTS DISCUSSION}
+\section{Result discussions}
 
@@
 
-\section{CONCLUSIONS}
-
-From the results shown in Section \ref{sec:results}, we can say that, in general,
+\section{Conclusions}
+
+From the results shown in Section~\ref{sec:results}, we can say that, in general,
 our model works fine identifying and characterizing open clusters.
@@
 in order to improve its accuracy and precision.
+
+Another achievement is that we have developed a model that does not require
+complex hardware. Instead, it can be run on workstations with a common GPU,
+making it accessible to be implemented in many research centers.
 
@@
 
-\subsection{FURTHER RESEARCH}
+Furthermore, our model is able to deal with large and open regions. As an example,
+it has succeeded characterizing Melotte 25 (Hiades) OC. Also, the relative proximity
+of this cluster, 45 parsec, makes it difficult to characterize. Clusterix fails
+characterizing this cluster. In this case, our method proposes at least two more
+non-catalogued clusters. The analysis of these new clusters brings out an interesting
+stellar dynamics in the studied region. It could mean that there was a recent
+interaction among different molecular clouds.
+
+For all this, we consider that the presented method is good enough to be included
+in the Virtual Observatory toolset. Only a few tweaks must be done in order to make
+our model compatible with other tools of the VO.
+
+\subsection{Future works}
 

### 2021-02-09 thesis/thesis.tex

@@
 \usepackage{minted}
+\usemintedstyle{friendly}
 \usepackage{picture}
@@
 This research has made use of the VizieR catalogue access tool, CDS, Strasbourg, France (DOI : 10.26093/cds/vizier).
-The original description of the VizieR service was published in 2000, \cite[A\&AS 143, 23]{ochsenbein2000vizier}
+The original description of the VizieR service was published in 2000~\cite[A\&AS 143, 23]{ochsenbein2000vizier}
 
@@
 
-Stellar open clusters (OCs) \cite{janes1982open} are groups of stars gravitationally bounded originated from a single molecular gas cloud.
+Stellar open clusters (OCs)~\cite{janes1982open} are groups of stars gravitationally bounded originated from a single molecular gas cloud.
 They share the same chemical composition and age. Moreover, they have similar relative positions and proper motion.
@@
 are formed in clustered environments and spend a period of time gravitationally bounded with their siblings embedded
-in their original molecular cloud \cite{clarke2000theformationofstellarclusters} \cite{portegies2010young}.
+in their original molecular cloud~\cite{clarke2000theformationofstellarclusters}~\cite{portegies2010young}.
 The evolution of those systems tends to sparse them in a few million years by interacting gravitationally with other systems.
@@
 The study of OCs has been pushed forward thanks to the huge and precise dataset from the Gaia mission
-\cite{collaboration2016description} Gaia DR2 \cite{gaia2018gaia}, available since 2018.
+\cite{collaboration2016description} Gaia DR2~\cite{gaia2018gaia}, available since 2018.
 This dataset has helped to review already known open clusters and to find new ones.
@@
 We will avoid this issue by looking at $E(b-r) / G_{mag}$ diagrams that show their connection with the isochrone curves.
-These curves are a tool used to determine how stars evolve in time \cite{bressan2012parsec}.
+These curves are a tool used to determine how stars evolve in time~\cite{bressan2012parsec}.
 The isochrones are derived from theoretical models.
@@
 They also have a direct relation with that presented by the Hertzsprung–Russell (H-R) diagram of the stars belonging to the cluster.
-Examples of these curves are shown in Figure \ref{fig:examples_of_isochrones}.
+Examples of these curves are shown in Figure~\ref{fig:examples_of_isochrones}.
 
@@
 with low scattering through the main sequence in the H-R diagram.
-Examples for this kind of diagrams are presented in Figure \ref{fig:examples_of_hr_diagrams}.
+Examples for this kind of diagrams are presented in Figure~\ref{fig:examples_of_hr_diagrams}.
 
@@
 Although the members of a cluster appear in the observational visual field as an overdensity in positions,
-as shown in Figure \ref{fig:pos_ngc_2682},
+as shown in Figure~\ref{fig:pos_ngc_2682},
 these coordinates are not useful to separate those stars that belong to the cluster from the other that do not.
 However, if we look for overdensities in the proper motion configuration spaces, it is possible, at least at first instance,
-to assume a possible membership cut (see Figure \ref{fig:pm_ngc_2682}.
+to assume a possible membership cut (see Figure~\ref{fig:pm_ngc_2682}.
 
@@
 Furthermore, it frequently requires a parametrization based on prior knowledge regarding the studied region.
-The aim of this work is to test an unsupervised and non-parametrized model based on Machine Learning (ML) tools.
+The aim of this work is to test an unsupervised and non-parameterized model based on Machine Learning (ML) tools.
 
@@
 The primary aim of this thesis is to \emph{build an unsupervised clustering model for open cluster characterization}.
-The model will be \emph{non-supervised and non-parameterized} to fit a wide range of clusters without
+The model must be \emph{non-supervised and non-parameterized} to fit a wide range of clusters without
 the need for fine-tuning a high number of hyperparameters.
@@
 This selection is crucial to properly characterize the fundamental properties of the cluster
-(dynamics, total mass, age, chemical composition, etc.).
+(dynamics, total mass, age, chemical composition, among other).
 
@@
 
-Figure \ref{fig:pm_melotte_22} shows the proper motion distribution in right ascension and declination.
+Figure~\ref{fig:pm_melotte_22} shows the proper motion distribution in right ascension and declination.
 It is easy to see that a subgroup can be located at [20, -45].
-Figure \ref{fig:pm_vec_melotte_22} shows the parallax distribution and confirms an overdensity at
-$\approx 7.3 mas$ corresponding to Mellote 22 OC (See also Figure \ref{fig:melotte_22_parallax_zoom}).
-
-However, as shown in Figures \ref{fig:pm_ngc_6494} and \ref{fig:parallax_ngc_6494},
+Figure~\ref{fig:pm_vec_melotte_22} shows the parallax distribution and confirms an overdensity at
+\(\approx 7.3 mas\) corresponding to Mellote 22 OC (See also Figure~\ref{fig:melotte_22_parallax_zoom}).
+
+However, as shown in Figures~\ref{fig:pm_ngc_6494} and~\ref{fig:parallax_ngc_6494},
 in general it is not as easy and becomes necessary to consider other parameters such as distances,
@@
 and magnitudes in three photometric filters (G, BP and RP) for more than 1,300 million sources.
-80\% of Gaia DR2 sources are weaker than $G_{mag} \approx 18$.
-For magnitudes $G \approx 21$ nominal uncertainties reach $2 mas$ for parallaxes and $5 mas \cdot years^{-1}$ in proper motions.
-But for closer sources, (at the bright end, $G < 14$), the precision is of the order of $0.02 mas$
-and $0.05 mas \cdot year^{-1}$ respectively \cite{cantat2018gaia}.
+80\% of Gaia DR2 sources are weaker than \(G_{mag} \approx 18\).
+For magnitudes \(G \approx 21\) nominal uncertainties reach \(2 mas\) for parallaxes and \(5 mas \cdot years^{-1}\) in proper motions.
+But for closer sources, (at the bright end, \(G < 14\)), the precision is of the order of \(0.02 mas\)
+and \(0.05 mas \cdot year^{-1}\) respectively~\cite{cantat2018gaia}.
 In general, open clusters, which are the main subject of this work, are at optimal data thresholds.
@@
 new approaches are being developed to improve current detection methods and automating OC characterization as much as possible.
-This is the case of tools such as Clusterix 2.0 \cite{balaguer2020clusterix}, TOPCAT \cite{taylor2005topcat}
-and VOSA \cite{bayo2008vosa}, (all of them developed by the Virtual Observatory).
+This is the case of tools such as Clusterix 2.0~\cite{balaguer2020clusterix}, TOPCAT~\cite{taylor2005topcat}
+and VOSA~\cite{bayo2008vosa}, (all of them developed by the Virtual Observatory).
 These tools work together and offer good results,
@@
 
-Another tool is Clusterix 2.0, which is an interactive web-based tool \cite{balaguer2020clusterix}.
+Another tool is Clusterix 2.0, which is an interactive web-based tool~\cite{balaguer2020clusterix}.
 It takes the proper motion diagram without making any prior assumption about the membership of the candidate star
@@
 \begin{equation}
-  K(a, b) = \frac{1}{2 \pi h^{2}} \exp{ \left[ - \frac{1}{2}\frac{\left( a - a_{i} \right)^{2} + \left( b - b_{i} \right)^{2}}{ h^{2}} \right]}
+  K(a, b) = \frac{1}{2 \pi h^{2}} \exp{ \left[ - \frac{1}{2}\frac{\left( a - a_{i} \right)^{2} + \left( b - b_{j} \right)^{2}}{ h^{2}} \right]}
 \end{equation}
 
-$\left( a, b \right)$ are referred to the proper motion configuration space,
-while a point located in the center of the cell $\left( i, j \right)$ provides the maximum
+\(\left( a, b \right)\) are referred to the proper motion configuration space,
+while a point located in the center of the cell \(\left( i, j \right)\) provides the maximum
 contribution for calculating the local density.
-Number $h$ is called the \emph{smoothing} parameter and it is measured in the same units as proper motion.
+\(h\) is called the \emph{smoothing} parameter and it is measured in the same units as proper motion.
 
 Various clustering algorithms have been studied in the past with the purpose of finding a valid non-parameterized method.
-UPMASK \cite{krone2014upmask} was developed to only use photometry and positions,
-and then it was adapted to the Gaia data \cite{cantat2018gaia} based only on proper motion and parallax.
+UPMASK~\cite{krone2014upmask} was developed to only use photometry and positions,
+and then it was adapted to the Gaia data~\cite{cantat2018gaia} based only on proper motion and parallax.
 A similar approach has been carried out using DBSCAN and machine learning algorithms to discover new clusters.
@@
 which makes it possible to distinguish two regions in the workspace:
-the only field region (label 'f'), dominated by star fields,
-and the cluster + field region (label 'c+f'), which includes both star fields and not star fields.
-\cite{balaguer2020clusterix}
+the only field region (label `f`), dominated by star fields,
+and the cluster + field region (label `c+f`), which includes both star fields and not star
+fields~\cite{balaguer2020clusterix}.
 \end{displayquote}
@@
 This fact implies defining three areas or regions with different radius.
-The first 'c+f' corresponds to the one in which the cluster members are
+The first `c+f` corresponds to the one in which the cluster members are
 presumed to be contained together with other star fields that are not part of the cluster.
@@
 
-The right choice of these radii, even having a previous estimation for the 'c+f' region,
-highly affects the execution of the algorithm and, in general, requires a considerable wide field 'f'.
+The right choice of these radii, even having a previous estimation for the `c+f` region,
+highly affects the execution of the algorithm and, in general, requires a considerable wide field `f`.
 There is no rule of thumb that defines relative proportions of these areas.
@@
 This new dataset can be exported to TOPCAT for further optimization and final processing.
-Figure \ref{fig:clusterix_probability} shows an example of the probability distribution
+Figure~\ref{fig:clusterix_probability} shows an example of the probability distribution
 in the proper motion space for NGC 2682.
@@
 
-Finally, it is necessary to refer to the work developed on \cite{castro2020hunting}.
+Finally, it is necessary to refer to the work developed on~\cite{castro2020hunting}.
 It presents a method based on machine learning techniques to make a systematic search
 for overdensities in the astrometric space of the galactic disk
-and a later identification of OC using photometric information, also from Gaia DR2.
+and a later identification of OCs using photometric information, also from Gaia DR2.
 
 The method includes two phases: the first one uses an unsupervised clustering algorithm, DBSCAN,
-to search for overdensities $(l, b \pi, \mu_{\alpha} *, \mu_{\delta})$,
+to search for overdensities \((l, b \pi, \mu_{\alpha} *, \mu_{\delta})\),
 and then applies a deep learning Artificial Neural Network (ANN),
@@
 The result of this work is the recognition of 582 new open clusters
-distributed along the galactic disk for a galactic declination $b < 20^{\circ}$,
+distributed along the galactic disk for a galactic declination \(b < 20^{\circ}\),
 which has meant an increase of 45\% over previously ones.
@@
 Our aim in this work, linked to computational limitations compared to MareNostrum,
-is not to do a blind search for clusters, but to obtain a method,
-within the machine learning environment,
-that allows the characterization of open clusters through unsupervised and non-parametrized procedures in a single step.
+is not to perform a blind search for clusters, but to obtain a method, within the machine learning environment,
+that allows the characterization of open clusters through unsupervised and non-parameterized procedures in a single step.
 This method should not require further refinements with other tools.
@@
 
-There are several clustering algorithms: K-Means, Mean-Shift Clustering, DBSCAN \cite{ester1996density} among other.
+There are several clustering algorithms: \emph{K-Means}, \emph{Mean-Shift Clustering}, \emph{DBSCAN}~\cite{ester1996density} among other.
 While each one behaves better according to the distribution of the objects to clusterize,
@@
 
-K-Means requires a single parameter, the number of clusters to build (N).
+K-Means requires a single parameter, the number of clusters to build (\(N\)).
 The algorithm looks for N center points which are vectors of the same dimension as the number of selected features.
@@
 This fact complicates the identification of the OC and many times the found cluster contains to many outliers.
-For that reason, we have search for a K-Means refinement based on an artificial neural network.
+For that reason, we searched for a K-Means refinement based on an artificial neural network.
 
 The \emph{Unsupervised Deep Embedding for Clustering Analysis} model
-or \emph{DEC} \cite{xie2016unsupervised} takes K-Means as its starting point,
+or \emph{DEC}~\cite{xie2016unsupervised} takes K-Means as its starting point,
 but then, it trains an autoencoder to reduce the feature space
@@
 
-The mechanism of this model is explained in Section \ref{sec:deep_embedding_clustering}
+The mechanism of this model is explained in Section~\ref{sec:deep_embedding_clustering}
 and its implementation is available in \verb|cdalvaro.ml.dec| package.
@@
 
-All code has been developed with the Python programming language \cite{Python3}.
-Other auxiliary tools like \emph{Docker} \cite{merkel2014docker},
-\emph{PostgreSQL} \cite{postgresql}, \emph{Jupyter Notebooks} \cite{Kluyver2016jupyter},
-and frameworks such as \emph{Astropy} \cite{astropy:2013} \cite{astropy:2018},
-\emph{Scikit-Learn} \cite{scikit-learn}, \emph{Seaborn} \cite{michael_waskom_2017_883859},
-\emph{SQLAlchemy} \cite{sqlalchemy} and \emph{Keras} \cite{chollet2015keras} have been used too.
+All code has been developed with the Python programming language~\cite{Python3}.
+Other auxiliary tools like \emph{Docker}~\cite{merkel2014docker},
+\emph{PostgreSQL}~\cite{postgresql}, \emph{Jupyter Notebooks}~\cite{Kluyver2016jupyter},
+and frameworks such as \emph{Astropy}~\cite{astropy:2013}~\cite{astropy:2018},
+\emph{Scikit-Learn}~\cite{scikit-learn}, \emph{Seaborn}~\cite{michael_waskom_2017_883859},
+\emph{SQLAlchemy}~\cite{sqlalchemy} and \emph{Keras}~\cite{chollet2015keras} have been used too.
 
@@
 
-For the sake of simplicity, we are taking \emph{Melotte 22 (Messier 45)} \cite{elsanhoury2019ppmxl} to illustrate the whole method.
+For the sake of simplicity, we are taking \emph{Melotte 22 (Messier 45)}~\cite{elsanhoury2019ppmxl} to illustrate the whole method.
 This cluster is well studied and its stars are not too mixed with other stars that do not belong to the OC
@@
 
-Figure \ref{fig:raw_pm_melotte_22} shows \emph{proper motions in right ascension and declination}
+Figure~\ref{fig:raw_pm_melotte_22} shows \emph{proper motion in right ascension and declination}
 for a sample of the downloaded dataset for Melotte 22.
@@
 
-Figure \ref{fig:raw_parallax_hr_diagram_mellote_22} shows the parallax histogram and the H-R diagram for this cluster, respectively.
-
-The figure on the left shows a resonance at $\approx 7.3mas$ belonging the OC.
+Figure~\ref{fig:raw_parallax_hr_diagram_mellote_22} shows the parallax histogram and the H-R diagram for this cluster, respectively.
+
+The figure on the left shows a resonance at \(\approx 7.3mas\) belonging the OC.
 While the figure on the right would help us to look for isochrone curves and so,
@@
 Before being able to develop and test our clustering model,
-the first step is to download the required data from Gaia repository
+the first step is to download the required data from the Gaia repository
 and store it in a custom database for later access.
@@
 In order to reduce the size of the dataset to be downloaded,
-the OpenClust catalogue \cite{dias2002new} (Figure \ref{fig:OpenClustComplete})
+the OpenClust catalogue~\cite{dias2002new} (Figure~\ref{fig:OpenClustComplete})
 has been used to restrict the sky regions to be explored.
@@
 
-As shown in Figure \ref{fig:OpenClustSelection},
+As shown in Figure~\ref{fig:OpenClustSelection},
 these constrains give us a smaller dataset but it is still a good representation of all clusters in the Milky Way
@@
 This script is prepared to load the OpenClust catalogue, connect to the Gaia DR2 database,
-download all stars contained inside each cluster (with a radius multiplier to increase the area to be downloaded)
+download all stars contained inside each cluster (with a radius factor to increase the area to be downloaded)
 and save the downloaded data inside our custom database hosted in the second container.
@@
 These containers are orchestrated with a \verb|docker-compose.yml| file which automatically launches one instance of each container.
-Figure \ref{fig:services_diagram} shows a diagram describing the services architecture for this work.
+Figure~\ref{fig:services_diagram} shows a diagram describing the services architecture for this work.
 
@@
 storing all stars related to one region inside an independent table for optimized access performance.
-Partition tables are located inside \verb|gaiadr2| schema. See Figure \ref{fig:customdb_diagram}.
+Partition tables are located inside \verb|gaiadr2| schema. See Figure~\ref{fig:customdb_diagram}.
 
@@
 We only need to know the \verb|id| for the desired region and query for its stars to the main table as follows:
+
+\pagebreak
 
@@
 
-Table \ref{tab:important_fields} shows a brief description of the fields used in this work.
+Table~\ref{tab:important_fields} shows a brief description of the fields used in this work.
 
@@
 As mentioned before, we want our model to characterize open clusters by looking at their dynamic properties.
-
 Also, we want to maintain as simple as possible our clustering model in order to save computing resources.
@@
 
-Parallax is another important feature. It lets us know how far stars are from us.
+Parallax is another important feature. It lets us know how far stars are from the Earth.
 In addition, since all stars within an open cluster were born from the same dust cloud,
@@
 
-However, we are not going to use these raw features.
-Instead, we are taking a combination of them.
+However, we are not going to use these raw features. Instead, we are taking a combination of them.
 
@@
 
-Figure \ref{fig:features_melotte_22} shows a pairwise relationship among some features available
+Figure~\ref{fig:features_melotte_22} shows a pairwise relationship among some features available
 in the dataset for Melotte 22 data.
@@
 \begin{itemize}
-  \item \emph{Proper motion in right ascension} (corrected by parallax): $\mu_{\alpha}$
-  \item \emph{Proper motion in declination} (corrected by parallax): $\mu_{\delta}$
-  \item \emph{Parallax}: $\varpi$
-  \item \emph{Proper motion modulus}: $\left\| \vec{\mu} \right\|$
+  \item \emph{Proper motion in right ascension} (corrected by parallax): \(\mu_{\alpha}\)
+  \item \emph{Proper motion in declination} (corrected by parallax): \(\mu_{\delta}\)
+  \item \emph{Parallax}: \( \varpi \)
+  \item \emph{Proper motion modulus}: \(\left\| \vec{\mu} \right\|\)
 \end{itemize}
@@
 with a fixed number of clusters and varying the feature selection in each case.
-The final selection is the one with best \emph{silhouette score} \cite{rousseeuw1987silhouettes}.
-
-The silhouette score is a metric used to determine how good a cluster is based on intra-cluster distances and nearest-cluster distances.
-The best possible value is 1, and the worst is -1.
+The final selection is the one with best \emph{silhouette score}~\cite{rousseeuw1987silhouettes}.
+
+The silhouette score is a metric used to determine how good a cluster is based on intra-cluster
+distances and nearest-cluster distances. The best possible value is 1, and the worst is -1.
 Negative values mean that some samples have been assigned to the wrong cluster.
@@
 Once we have found the set of features that best describes our problem,
-we can begin searching for the OC within the study region.
+we can begin searching for the OC within the selected region.
 
@@
 
-As shown in Figure \ref{fig:kmeans_comparisons_melotte_22},
-larger values for the number of clusters allow us to isolate more accurately the resonance in parallax at $\approx 7.3mas$.
-However, we have the disadvantage that more groups are formed.
+As shown in Figure~\ref{fig:kmeans_comparisons_melotte_22},
+larger values for the number of clusters allow us to isolate more accurately the resonance in parallax
+at \(\approx 7.3mas\). However, we have the disadvantage that more groups are formed.
 
@@
 K-Means does a good job making an initial clustering,
-as shown in Figure \ref{fig:kmeans_melotte_22}.
+as shown in Figure~\ref{fig:kmeans_melotte_22}.
 However, too many clusters arise from this characterization
@@
 
-If we take a look into the H-R diagram (Figure \ref{fig:kmeans_hr_diagram_melotte_22}),
+If we take a look into the H-R diagram (Figure~\ref{fig:kmeans_hr_diagram_melotte_22}),
 we can identify the isochrone curve for Melotte 22 cluster.
@@
 
-The model is composed by a \emph{deep autoencoder}, a \emph{deep encoder} and a \emph{clustering layer}.
-
-Encoders are used to transform the input data into a latent space
-using a non-linear mapping function $f_{\theta} : X \rightarrow Z$.
-
-Although, as explained in section \ref{sec:feature_selection},
+The model is composed by a \emph{deep autoencoder} and a \emph{clustering layer}.
+
+The autoencoder is used to transform the input data into a latent space
+using a non-linear mapping function \(f_{\theta} : X \rightarrow Z\).
+
+Although, as explained in Section~\ref{sec:feature_selection},
 the number of features we are managing is not too large,
 this latent space helps us reduce the number of features
-and avoids the \emph{"curse of dimensionality"} \cite{bellman1961curse}.
-
-These encoders are pretrained before fitting the model to generate predictions.
-Then, the encoder is used with the aim of transforming input data to the latent space $Z$.
-Once the data has been transformed, a K-Means clusterer is used in order to make an initial clustering.
+and avoids the \emph{``curse of dimensionality``}~\cite{bellman1961curse}.
+
+The autoencoder is pretrained before fitting the model to generate predictions. Then,
+the encoder layers of the autoencoder are used with the aim of transforming input data to
+the latent space \(Z\). Once the data has been transformed, a K-Means clusterer is used in
+order to make an initial clustering. K-Means cluster centers are used as the initial
+weights for the clustering layer.
 
@@
 the model iterates alternating between computing an auxiliary target distribution (Soft Assignment)
-and minimizing the Kullback-Leiber (KL) divergence \cite{kullback1951information} to it.
+and minimizing the Kullback-Leiber (KL) divergence~\cite{kullback1951information} to it.
 This unsupervised algorithm allows us to improve the clustering.
@@
 The choice of this target distribution is crucial for DEC's performance.
-In this work we have taken the target distribution from DEC's original paper \cite{xie2016unsupervised},
-which is defined in Equation \ref{eq:student_tdistribution}.
-
-Figure \ref{fig:dec_model_setup} shows the layer setup of our DEC model.
-It is simpler than the one tested on the original paper \cite{xie2016unsupervised},
+In this work we have taken the target distribution from DEC's original paper~\cite{xie2016unsupervised},
+which is defined in Equation~\ref{eq:student_tdistribution}.
+
+Figure~\ref{fig:dec_model_setup} shows the layer setup of our DEC model.
+It is simpler than the one tested on the original paper~\cite{xie2016unsupervised},
 since the number of selected features in our work is smaller than in the original one.
@@
 
-As shown in Figure \ref{fig:dec_melotte_22},
+As shown in Figure~\ref{fig:dec_melotte_22},
 the DEC model is able to move stars from one group to another,
@@
 and the isochrone curve of the open cluster is sharper than the one obtained with the K-Means algorithm
-(Figure \ref{fig:kmeans_hr_diagram_melotte_22}).
-
-Figure \ref{fig:hr_diagram_comparison} shows a comparison between the H-R diagram obtained
+(Figure~\ref{fig:kmeans_hr_diagram_melotte_22}).
+
+Figure~\ref{fig:hr_diagram_comparison} shows a comparison between the H-R diagram obtained
 by DEC clustering and the one obtained using the Clusterix + TOPCAT method.
@@
 That way we remove the most doubtful values from the selection.
-Figure \ref{fig:melotte_22_filtered} shows the result after filtering DEC clustering selection.
+Figure~\ref{fig:melotte_22_filtered} shows the result after filtering DEC clustering selection.
 
@@
 
-Table \ref{tab:results_melotte_22} shows a summary of Melotte 22 with values taken from different sources and methods:
+Table~\ref{tab:results_melotte_22} shows a summary of Melotte 22 with values taken from different sources and methods:
 SIMBAD Astronomical Database - CDS (Strasbourg),
@@
 Errors associated to K-Means and DEC predictions have been calculated as the \emph{Standard Error},
-$SE = \sigma / \sqrt{N}$, where $N$ is the number of samples in the selected group.
+\(SE = \sigma / \sqrt{N}\), where \(N\) is the number of samples in the selected group.
 
@@
     \begin{tabular}{l|c|c|c|c}
-      \textbf{Source / method} & \emph{$\mu_{\alpha}$ $(mas/yr^{-1})$} & \emph{$\mu_{\delta}$ $(mas/yr^{-1})$} & \emph{$\varpi$ $(mas)$} & \emph{\# stars} \\
+      \textbf{Source / method} & \emph{\(\mu_{\alpha}\) \((mas \cdot yr^{-1})\)} & \emph{\(\mu_{\delta}\) \((mas \cdot yr^{-1})\)}
+      & \emph{\( \varpi \) \((mas)\)} & \emph{\# stars} \\
       \hline
       \textbf{Simbad.u-strasbg.fr}\tablefootnote{Results have been taken from \protect\citeA[p.~25. Table A.3. Pleiades]{babusiaux2018gaia}} &
-        19.997 $\pm$ 0.127 & -45.548 $\pm$ 0.101 & 7.364 $\pm$ 0.005 & 1326 \\
-      Clusterix+TOPCAT & 19.98 $\pm$ 1.25 & -45.47 $\pm$ 1.48 & 7.33 $\pm$ 0.21 & 634 \\
-      K-Means & 20.25 $\pm$ 0.95 & -38.01 $\pm$ 1.08 & 7.23 $\pm$ 0.06 & 1378 \\
-      DEC & 23.67 $\pm$ 1.29 & -46.23 $\pm$ 1.50 & 8.04 $\pm$ 0.09 & 878 \\
-      \textbf{DEC (filtered)} & 19.50 $\pm$ 0.41 & -44.23 $\pm$ 0.39 & 7.42 $\pm$ 0.005 & 438 \\
+        19.997 \( \pm \) 0.127 & -45.548 \( \pm \) 0.101 & 7.364 \( \pm \) 0.005 & 1326 \\
+      Clusterix+TOPCAT & 19.98 \( \pm \) 1.25 & -45.47 \( \pm \) 1.48 & 7.33 \( \pm \) 0.21 & 634 \\
+      K-Means & 20.25 \( \pm \) 0.95 & -38.01 \( \pm \) 1.08 & 7.23 \( \pm \) 0.06 & 1378 \\
+      DEC & 23.67 \( \pm \) 1.29 & -46.23 \( \pm \) 1.50 & 8.04 \( \pm \) 0.09 & 878 \\
+      \textbf{DEC (filtered)} & 19.50 \( \pm \) 0.41 & -44.23 \( \pm \) 0.39 & 7.42 \( \pm \) 0.005 & 438 \\
     \end{tabular}
@@
 
-We have seen in the previous chapter that DEC model is capable
+We have seen in Section~\ref{sec:deep_embedding_clustering} that DEC model is capable
 of refining the clustering selection made by K-Means
@@
 We take Melotte 22 again as our study case to make easier the comparison process between
-the method presented in Chapter \ref{chap:method} and the alternative one.
+the method presented in Chapter~\ref{chap:method} and the alternative one.
 
@@
 and manual selection for the cluster members.
-Figure \ref{fig:clusterix_selection} shows a screenshot of the Clusterix 2.0 selection panel.
+Figure~\ref{fig:clusterix_selection} shows a screenshot of the Clusterix 2.0 selection panel.
 
@@
            while the radius is the value registered for this cluster in the
-           OpenClust catalogue multiplied by 1.5 (as explained in \ref{sec:data_mining})}
+           OpenClust catalogue multiplied by 1.5 (as explained in~\ref{sec:data_mining})}
   \label{fig:clusterix_selection}
@@
 Clusterix returns a position map for the downloaded stars from Gaia DR2 database
-using Aladin \cite{bonnarel2000aladin} as an atlas of the sky (See Figure \ref{fig:aladin_melotte_22}).
+using Aladin~\cite{bonnarel2000aladin} as an atlas of the sky (See Figure~\ref{fig:aladin_melotte_22}).
 At this point, we can choose sending the recovered data to TOPCAT or Aladin,
@@
 For example, in the case of Melotte 22, the proper motion in declination direction
-of the membership stars around $-45 mas \cdot year^{-1}$,
-so it is necessary to open the $\mu$ range above the estimated values (in absolute value).
+of the membership stars around \(-45 mas \cdot year^{-1}\),
+so it is necessary to open the \(\mu\) range above the estimated values (in absolute value).
 If we do not do that we lose the possibility of detection,
-since Clusterix sets this value to $30 mas \cdot year^{-1}$ by default.
+since Clusterix sets this value to \(30 mas \cdot year^{-1}\) by default.
 But even in this case, it is necessary to set a threshold for magnitude values,
-$G \leq 14$, to help getting the right result.
-Figure \ref{fig:clusterix_step2_melotte_22} shows the region selection panel on Clusterix 2.0.
+\(G \leq 14\), to help getting the right result.
+Figure~\ref{fig:clusterix_step2_melotte_22} shows the region selection panel on Clusterix 2.0.
 
@@
 Clusterix performs the proper motion frequency analysis and returns the probability field
-associated to each star of the whole region (see Figure \ref{fig:clusterix_probability_melotte_22}).
+associated to each star of the whole region (see Figure~\ref{fig:clusterix_probability_melotte_22}).
 
@@
 Once data is available in TOPCAT, we have to carefully analyze the imported data.
-At this point, we can perform a first stars selection by excluding those stars with probability value lower than 0.8.
+At this point, we can perform a first star selection by excluding those stars with probability value lower than 0.8.
 This first result can be taken as a good starting point, but we still need to make some fine tunning.
-As we can see in Figure \ref{fig:topcat_prob_selection_melotte_22},
+As we can see in Figure~\ref{fig:topcat_prob_selection_melotte_22},
 the selection is too restrictive and we have lost too many stars.
 
-Looking at Figure \ref{fig:topcat_prob_parallax_melotte_22},
-we can assume that cluster members are around $7.3mas$ in the parallax histogram.
-Also, Figure \ref{fig:topcat_prob_pm_melotte_22} reveals that cluster stars
+Looking at Figure~\ref{fig:topcat_prob_parallax_melotte_22},
+we can assume that cluster members are around \(7.3mas\) in the parallax histogram.
+Also, Figure~\ref{fig:topcat_prob_pm_melotte_22} reveals that cluster stars
 have low scattering rate in the proper motion configuration space.
@@
   \end{subfigure}
-  \caption{Melotte 22 first selection using validation method. Parallax is $7.304mas$ with 0.797 of standard deviation.}
+  \caption{Melotte 22 first selection using validation method. Parallax is \(7.304mas\) with 0.797 of standard deviation.}
   \label{fig:topcat_prob_selection_melotte_22}
@@
 for parallax and its standard deviation and selects those stars from the whole set
-which fall inside the limits $7.304mas \pm 0.797$ (see Figure \ref{fig:topcat_parallax_melotte_22}).
+which fall inside the limits \(7.304mas \pm 0.797\) (see Figure~\ref{fig:topcat_parallax_melotte_22}).
 
@@
 
-Looking at the H-R diagrams (Figure \ref{fig:clusterix_selection_hr_diagrams_melotte_22}),
+Looking at the H-R diagrams (Figure~\ref{fig:clusterix_selection_hr_diagrams_melotte_22}),
 we can compare both selections and see how we have extended the selection without
@@
 Finally, we filter the second selection by discarding those stars whose parallax relative error in absolute value
-is behind 0.05 (see Figure \ref{fig:clusterix_final_selection_melotte_22}).
+is behind 0.05 (see Figure~\ref{fig:clusterix_final_selection_melotte_22}).
 
@@
 
-For the sake of completeness, Table \ref{tab:results_melotte_22} is shown again.
+For the sake of completeness, Table~\ref{tab:results_melotte_22} is shown again.
 This table summarizes the results obtained with the validation technique and our method.
@@
     \begin{tabular}{l|c|c|c|c}
-      \textbf{Source / method} & \emph{$\mu_{\alpha}$ $(mas/yr^{-1})$} & \emph{$\mu_{\delta}$ $(mas/yr^{-1})$} & \emph{$\varpi$ $(mas)$} & \emph{\# stars} \\
+      \textbf{Source / method} & \emph{\(\mu_{\alpha}\) \((mas \cdot yr^{-1})\)} & \emph{\(\mu_{\delta}\) \((mas \cdot yr^{-1})\)}
+      & \emph{\( \varpi \) \((mas)\)} & \emph{\# stars} \\
       \hline
-      \textbf{Simbad.u-strasbg.fr} & 19.997 $\pm$ 0.127 & -45.548 $\pm$ 0.101 & 7.364 $\pm$ 0.005 & 1326 \\
-      \textbf{Clusterix+TOPCAT} & 19.98 $\pm$ 1.25 & -45.47 $\pm$ 1.48 & 7.33 $\pm$ 0.21 & 634 \\
-      K-Means & 20.25 $\pm$ 0.95 & -38.01 $\pm$ 1.08 & 7.23 $\pm$ 0.06 & 1378 \\
-      DEC & 23.67 $\pm$ 1.29 & -46.23 $\pm$ 1.50 & 8.04 $\pm$ 0.09 & 878 \\
-      DEC (filtered) & 19.50 $\pm$ 0.41 & -44.23 $\pm$ 0.39 & 7.42 $\pm$ 0.005 & 438 \\
+      \textbf{Simbad.u-strasbg.fr} & 19.997 \( \pm \) 0.127 & -45.548 \( \pm \) 0.101 & 7.364 \( \pm \) 0.005 & 1326 \\
+      \textbf{Clusterix+TOPCAT} & 19.98 \( \pm \) 1.25 & -45.47 \( \pm \) 1.48 & 7.33 \( \pm \) 0.21 & 634 \\
+      K-Means & 20.25 \( \pm \) 0.95 & -38.01 \( \pm \) 1.08 & 7.23 \( \pm \) 0.06 & 1378 \\
+      DEC & 23.67 \( \pm \) 1.29 & -46.23 \( \pm \) 1.50 & 8.04 \( \pm \) 0.09 & 878 \\
+      DEC (filtered) & 19.50 \( \pm \) 0.41 & -44.23 \( \pm \) 0.39 & 7.42 \( \pm \) 0.005 & 438 \\
     \end{tabular}
@@
 All results have been computed with an Apple Mac Pro Late 2013,
-2.7 GHz 12-Core Intel Xeon E5 with two AMD FirePro D700 6GB and 64 GB
-of memory, running macOS BigSur 11.1, Keras 2.2 and PlaidML 0.7.
-
-The majority of the runnings have taken a few minutes, while big clusters such as Melotte 25,
-have taken around an hour to perform the computations.
-This contrasts with \citeA{castro2020hunting} work,
-that used the power of the MareNostrum 42 supercomputer in order to perform its computations.
+2.7GHz 12-Core Intel Xeon E5-2697v2, 64GB RAM 1866MHz DDR3 and two AMD FirePro D700 6GB.
+The operative system is macOS BigSur 11.1, with Python 3.8, Keras 2.2 and PlaidML 0.7.
+
+Most executions took a few minutes, while large clusters such as Melotte 25,
+took around an hour to perform the calculations.
+This is in contrast to~\citeA{castro2020hunting} work,
+that used the power of the MareNostrum 42 supercomputer to accomplish its computations.
 
@@
 
-Figure \ref{fig:result_ngc_2516_clusterix} shows NGC 2516 characterized using Clusterix+TOPCAT tools.
-Figure \ref{fig:result_ngc_2516_kmeans} shows eight clusters identified by K-Means.
+Figure~\ref{fig:result_ngc_2516_clusterix} shows NGC 2516 characterized using Clusterix+TOPCAT tools.
+Figure~\ref{fig:result_ngc_2516_kmeans} shows eight clusters identified by K-Means.
 In this case, cluster \emph{g3} is the one we are looking for.
-Figures \ref{fig:result_ngc_2516_dec} and \ref{fig:result_ngc_2516_dec_filtered}
+Figures~\ref{fig:result_ngc_2516_dec} and~\ref{fig:result_ngc_2516_dec_filtered}
 show the groups found using the DEC model and the DEC model filtered, respectively.
@@
 
-Table \ref{tab:hyperparameters_ngc_2516} shows the hyperparameters used for characterizing NGC 2516
-with the DEC model. Table \ref{tab:results_ngc_2516} shows a results summary for NGC 2516.
+Table~\ref{tab:hyperparameters_ngc_2516} shows the hyperparameters used for characterizing NGC 2516
+with the DEC model. Table~\ref{tab:results_ngc_2516} shows a results summary for NGC 2516.
 
@@
       Number of Clusters & 8 \\
-      Clustering Layer & $\left[ 50, 50, 60 \right]$ \\
+      Clustering Layer & \(\left[ 50, 50, 60 \right]\) \\
       Kernel Initializer Seed & 2 \\
@@
     \begin{tabular}{l|c|c|c|c}
-      \textbf{Source / method} & \emph{$\mu_{\alpha}$ $(mas/yr^{-1})$} & \emph{$\mu_{\delta}$ $(mas/yr^{-1})$} & \emph{$\varpi$ $(mas)$} & \emph{\# stars} \\
+      \textbf{Source / method} & \emph{\(\mu_{\alpha}\) \((mas \cdot yr^{-1})\)} & \emph{\(\mu_{\delta}\) \((mas \cdot yr^{-1})\)}
+      & \emph{\( \varpi \) \((mas)\)} & \emph{\# stars} \\
       \hline
-      \textbf{Simbad.u-strasbg.fr} & -4.6579 $\pm$ 0.0075 & 11.1517 $\pm$ 0.0075 & 2.4118 $\pm$ 0.0006 & 1727 \\
-      Clusterix+TOPCAT & -4.652 $\pm$ 0.523 & 11.203 $\pm$ 0.454 & 2.409 $\pm$ 0.127 & 638 \\
-      K-Means & -4.344 $\pm$ 0.14 & 9.507 $\pm$ 0.19 & 2.268 $\pm$ 0.01 & 1542 \\
-      DEC & -4.426 $\pm$ 0.17 & 9.952 $\pm$ 0.20 & 2.436 $\pm$ 0.01 & 1532 \\
-      \textbf{DEC (filtered)} & -4.502 $\pm$ 0.14 & 10.114 $\pm$ 0.17 & 2.392 $\pm$ 0.004 & 1072 \\
+      \textbf{Simbad.u-strasbg.fr} & -4.6579 \( \pm \) 0.0075 & 11.1517 \( \pm \) 0.0075 & 2.4118 \( \pm \) 0.0006 & 1727 \\
+      Clusterix+TOPCAT & -4.652 \( \pm \) 0.523 & 11.203 \( \pm \) 0.454 & 2.409 \( \pm \) 0.127 & 638 \\
+      K-Means & -4.344 \( \pm \) 0.14 & 9.507 \( \pm \) 0.19 & 2.268 \( \pm \) 0.01 & 1542 \\
+      DEC & -4.426 \( \pm \) 0.17 & 9.952 \( \pm \) 0.20 & 2.436 \( \pm \) 0.01 & 1532 \\
+      \textbf{DEC (filtered)} & -4.502 \( \pm \) 0.14 & 10.114 \( \pm \) 0.17 & 2.392 \( \pm \) 0.004 & 1072 \\
     \end{tabular}
@@
 
-Figure \ref{fig:result_ngc_2632_clusterix} shows NGC 2632 characterized by the validation method.
-Figure \ref{fig:result_ngc_2632_kmeans} shows ten clusters identified by K-Means.
+Figure~\ref{fig:result_ngc_2632_clusterix} shows NGC 2632 characterized by the validation method.
+Figure~\ref{fig:result_ngc_2632_kmeans} shows ten clusters identified by K-Means.
 Cluster \emph{g1} is the open cluster we are looking for.
-Figures \ref{fig:result_ngc_2632_dec} and \ref{fig:result_ngc_2632_dec_filtered}
+Figures~\ref{fig:result_ngc_2632_dec} and~\ref{fig:result_ngc_2632_dec_filtered}
 show the groups found using the DEC model and the DEC model filtered, respectively.
@@
 
-Table \ref{tab:hyperparameters_ngc_2632} shows the hyperparameters used for characterizing NGC 2632
-with our model. Table \ref{tab:results_ngc_2632} shows a results summary for NGC 2632 analysis.
+Table~\ref{tab:hyperparameters_ngc_2632} shows the hyperparameters used for characterizing NGC 2632
+with our model. Table~\ref{tab:results_ngc_2632} shows a results summary for NGC 2632 analysis.
 
@@
       Number of Clusters & 10 \\
-      Clustering Layer & $\left[ 50, 50, 40 \right]$ \\
+      Clustering Layer & \(\left[ 50, 50, 40 \right]\) \\
       Kernel Initializer Seed & 10 \\
@@
     \begin{tabular}{l|c|c|c|c}
-      \textbf{Source / method} & \emph{$\mu_{\alpha}$ $(mas/yr^{-1})$} & \emph{$\mu_{\delta}$ $(mas/yr^{-1})$} & \emph{$\varpi$ $(mas)$} & \emph{\# stars} \\
+      \textbf{Source / method} & \emph{\(\mu_{\alpha}\) \((mas \cdot yr^{-1})\)} & \emph{\(\mu_{\delta}\) \((mas \cdot yr^{-1})\)}
+      & \emph{\( \varpi \) \((mas)\)} & \emph{\# stars} \\
       \hline
-      \textbf{Simbad.u-strasbg.fr} & -36.047 $\pm$ 0.110 & -12.917 $\pm$ 0.066 & 5.371 $\pm$ 0.003 & - \\
-      Clusterix+TOPCAT & -36.154 $\pm$ 1.001 & -12.909 $\pm$ 0.806 & 5.327 $\pm$ 0.187 & 371 \\
-      K-Means & -26.352 $\pm$ 0.82 & -15.828 $\pm$ 0.76 & 5.394 $\pm$ 0.03 & 629 \\
-      DEC & -20.012 $\pm$ 0.69 & -14.742 $\pm$ 0.58 & 4.686 $\pm$ 0.03 & 894 \\
-      \textbf{DEC (filtered)} & -21.571 $\pm$ 0.74 & -14.234 $\pm$ 0.61 & 4.719 $\pm$ 0.03 & 714 \\
+      \textbf{Simbad.u-strasbg.fr} & -36.047 \( \pm \) 0.110 & -12.917 \( \pm \) 0.066 & 5.371 \( \pm \) 0.003 & - \\
+      Clusterix+TOPCAT & -36.154 \( \pm \) 1.001 & -12.909 \( \pm \) 0.806 & 5.327 \( \pm \) 0.187 & 371 \\
+      K-Means & -26.352 \( \pm \) 0.82 & -15.828 \( \pm \) 0.76 & 5.394 \( \pm \) 0.03 & 629 \\
+      DEC & -20.012 \( \pm \) 0.69 & -14.742 \( \pm \) 0.58 & 4.686 \( \pm \) 0.03 & 894 \\
+      \textbf{DEC (filtered)} & -21.571 \( \pm \) 0.74 & -14.234 \( \pm \) 0.61 & 4.719 \( \pm \) 0.03 & 714 \\
     \end{tabular}
@@
 
-Figure \ref{fig:result_ngc_2682_clusterix} shows NGC 2682 characterized using Clusterix+TOPCAT tools.
-Figure \ref{fig:result_ngc_2682_kmeans} shows ten clusters identified by K-Means.
+Figure~\ref{fig:result_ngc_2682_clusterix} shows NGC 2682 characterized using Clusterix+TOPCAT tools.
+Figure~\ref{fig:result_ngc_2682_kmeans} shows ten clusters identified by K-Means.
 K-Means labels NGC 2682 as group \emph{g4}.
-Figures \ref{fig:result_ngc_2682_dec} and \ref{fig:result_ngc_2682_dec_filtered}
+Figures~\ref{fig:result_ngc_2682_dec} and~\ref{fig:result_ngc_2682_dec_filtered}
 show the groups found using the DEC model and the DEC model filtered, respectively.
@@
 
-Table \ref{tab:hyperparameters_ngc_2682} shows the hyperparameters used for characterizing NGC 2682
-with the DEC model. Table \ref{tab:results_ngc_2682} shows a results summary for NGC 2682.
+Table~\ref{tab:hyperparameters_ngc_2682} shows the hyperparameters used for characterizing NGC 2682
+with the DEC model. Table~\ref{tab:results_ngc_2682} shows a results summary for NGC 2682.
 
@@
       Number of Clusters & 10 \\
-      Clustering Layer & $\left[ 50, 50, 40 \right]$ \\
+      Clustering Layer & \(\left[ 50, 50, 40 \right]\) \\
       Kernel Initializer Seed & 0 \\
@@
     \begin{tabular}{l|c|c|c|c}
-      \textbf{Source / method} & \emph{$\mu_{\alpha}$ $(mas/yr^{-1})$} & \emph{$\mu_{\delta}$ $(mas/yr^{-1})$} & \emph{$\varpi$ $(mas)$} & \emph{\# stars} \\
+      \textbf{Source / method} & \emph{\(\mu_{\alpha}\) \((mas \cdot yr^{-1})\)} & \emph{\(\mu_{\delta}\) \((mas \cdot yr^{-1})\)}
+      & \emph{\( \varpi \) \((mas)\)} & \emph{\# stars} \\
       \hline
-      \textbf{Simbad.u-strasbg.fr} & -10.9737 $\pm$ 0.0064 & -2.9396 $\pm$ 0.0063 & 1.1325 $\pm$ 0.0011 & 1194 \\
-      Clusterix+TOPCAT & -10.970 $\pm$ 0.322 & -2.958 $\pm$ 0.327 & 1.142 $\pm$ 0.080 & 649 \\
-      K-Means & -8.616 $\pm$ 0.15 & -3.710 $\pm$ 0.16 & 1.196 $\pm$ 0.01 & 1374 \\
-      DEC & -8.926 $\pm$ 0.15 & -3.550 $\pm$ 0.15 & 1.144 $\pm$ 0.005 & 1238 \\
-      \textbf{DEC (filtered)} & -9.619 $\pm$ 0.13 & -3.317 $\pm$ 0.13 & 1.140 $\pm$ 0.003 & 990 \\
+      \textbf{Simbad.u-strasbg.fr} & -10.9737 \( \pm \) 0.0064 & -2.9396 \( \pm \) 0.0063 & 1.1325 \( \pm \) 0.0011 & 1194 \\
+      Clusterix+TOPCAT & -10.970 \( \pm \) 0.322 & -2.958 \( \pm \) 0.327 & 1.142 \( \pm \) 0.080 & 649 \\
+      K-Means & -8.616 \( \pm \) 0.15 & -3.710 \( \pm \) 0.16 & 1.196 \( \pm \) 0.01 & 1374 \\
+      DEC & -8.926 \( \pm \) 0.15 & -3.550 \( \pm \) 0.15 & 1.144 \( \pm \) 0.005 & 1238 \\
+      \textbf{DEC (filtered)} & -9.619 \( \pm \) 0.13 & -3.317 \( \pm \) 0.13 & 1.140 \( \pm \) 0.003 & 990 \\
     \end{tabular}
@@
 
-Figure \ref{fig:result_melotte_25_clusterix} shows Melotte 25 characterized using Clusterix+TOPCAT tools.
-Figure \ref{fig:result_melotte_25_kmeans} shows nine clusters identified by K-Means.
+Figure~\ref{fig:result_melotte_25_clusterix} shows Melotte 25 characterized using Clusterix+TOPCAT tools.
+Figure~\ref{fig:result_melotte_25_kmeans} shows nine clusters identified by K-Means.
 K-Means labels Melotte 25 as group \emph{g3}.
-Figures \ref{fig:result_melotte_25_dec} and \ref{fig:result_melotte_25_dec_filtered}
+Figures~\ref{fig:result_melotte_25_dec} and~\ref{fig:result_melotte_25_dec_filtered}
 show the groups found using the DEC model and the DEC model filtered, respectively.
@@
 
-Table \ref{tab:hyperparameters_melotte_25} shows the hyperparameters used for characterizing Melotte 25
-with the DEC model. Table \ref{tab:results_melotte_25} shows a results summary for Melotte 25.
+Table~\ref{tab:hyperparameters_melotte_25} shows the hyperparameters used for characterizing Melotte 25
+with the DEC model. Table~\ref{tab:results_melotte_25} shows a results summary for Melotte 25.
 
@@
       Number of Clusters & 9 \\
-      Clustering Layer & $\left[ 50, 50, 40 \right]$ \\
+      Clustering Layer & \(\left[ 50, 50, 40 \right]\) \\
       Kernel Initializer Seed & 11 \\
@@
     \begin{tabular}{l|c|c|c|c}
-      \textbf{Source / method} & \emph{$\mu_{\alpha}$ $(mas/yr^{-1})$} & \emph{$\mu_{\delta}$ $(mas/yr^{-1})$} & \emph{$\varpi$ $(mas)$} & \emph{\# stars} \\
+      \textbf{Source / method} & \emph{\(\mu_{\alpha}\) \((mas \cdot yr^{-1})\)} & \emph{\(\mu_{\delta}\) \((mas \cdot yr^{-1})\)}
+      & \emph{\( \varpi \) \((mas)\)} & \emph{\# stars} \\
       \hline
-      \textbf{Simbad.u-strasbg.fr} & 104.92 $\pm$ 0.12 & -28.00 $\pm$ 0.09 & 21.052 $\pm$ 0.065 & - \\
-      Clusterix+TOPCAT & 106.796 $\pm$ 6.229 & -24.870 $\pm$ 5.417 & 21.210 $\pm$ 1.115 & 109 \\
-      K-Means & 79.936 $\pm$ 3.7 & -45.362 $\pm$ 3.8 & 20.901 $\pm$ 0.31 & 374 \\
-      DEC & 104.051 $\pm$ 3.2 & -33.424 $\pm$ 2.5 & 22.072 $\pm$ 0.13 & 219 \\
-      \textbf{DEC (filtered)} & 105.96 $\pm$ 3.5 & -30.00 $\pm$ 2.4 & 21.74 $\pm$ 0.07 & 175 \\
+      \textbf{Simbad.u-strasbg.fr} & 104.92 \( \pm \) 0.12 & -28.00 \( \pm \) 0.09 & 21.052 \( \pm \) 0.065 & - \\
+      Clusterix+TOPCAT & 106.796 \( \pm \) 6.229 & -24.870 \( \pm \) 5.417 & 21.210 \( \pm \) 1.115 & 109 \\
+      K-Means & 79.936 \( \pm \) 3.7 & -45.362 \( \pm \) 3.8 & 20.901 \( \pm \) 0.31 & 374 \\
+      DEC & 104.051 \( \pm \) 3.2 & -33.424 \( \pm \) 2.5 & 22.072 \( \pm \) 0.13 & 219 \\
+      \textbf{DEC (filtered)} & 105.96 \( \pm \) 3.5 & -30.00 \( \pm \) 2.4 & 21.74 \( \pm \) 0.07 & 175 \\
     \end{tabular}
@@
 while Melotte 25 is located inside a region with more than 400,000 stars.
-More information about the studied clusters is available in Table \ref{tab:clusters_summary}.
+More information about the studied clusters is available in Table~\ref{tab:clusters_summary}.
 
@@
 
-From the results shown in Chapter \ref{sec:comparing_and_validating},
+From the results shown in Chapter~\ref{sec:comparing_and_validating},
 we can say that, in general, our model works fine identifying and characterizing open clusters.
@@
 Nevertheless some tweaks can be done regarding the cluster,
-like changing the initial number of clusters or modifying the clustering layer structure
+like changing the initial number of clusters or modifying the DEC's layer structure
 in order to improve the final characterization.
@@
 This kernel prepares data before passing it to the ANN
-and the results may vary significantly depending on this kernel.
+and results may vary significantly depending on this kernel.
 For example, simply varying the kernel seed can make a huge difference on the result.
@@
 We could determine with higher accuracy the typologies that our model works well with,
-and which ones the model does not resolve properly.
+and which ones the model fails to resolve properly.
 Furthermore, we could stablish different sets of hyperparameters regarding the typology of the cluster.

### 2021-02-09 paper/paper.tex
paper: Minor corrections

@@
 
-\section{Related works}
+\section{Related work}
 \label{sec:related_works}
@@
 
-For the selection of features, we studied the Melotte 22 dataset. Table~\ref{tab:
-sample_features_melotte22_region} describes an example of the region Melotte 22. Each
-row represents the features of a star.
+For the selection of features, we studied the Melotte 22 dataset.
+Table~\ref{tab:sample_features_melotte22_region} describes an example of the region Melotte 22.
+Each row represents the features of a star.
 
@@
         % \textbf{source\_id} &
-        \emph{pmra $(mas \cdot yr^{-1})$} & \emph{pmdec $(mas \cdot yr^{-1})$} & \emph{parallax $(mas)$} \\
+        \emph{pmra \((mas \cdot yr^{-1})\)} & \emph{pmdec \((mas \cdot yr^{-1})\)} & \emph{parallax \((mas)\)} \\
         \hline
@@
 deal is not too large. This latent space helps us to start in a reduced number of
-features and avoids the \emph{"curse of dimensionality"}~\cite{bellman1961curse}.
+features and avoids the \emph{``curse of dimensionality``}~\cite{bellman1961curse}.
 
@@
 
-Figure \ref{fig:kmeans_comparisons_melotte_22} shows an example. It describes the results of
+Figure~\ref{fig:kmeans_comparisons_melotte_22} shows an example. It describes the results of
 applying K-Means with N=2, N=5 and N=8 in Melotte 22. As we can see, large values of the number
@@
 
-Table \ref{tab:results_melotte_25} shows a results summary for Melotte 25.
+Table~\ref{tab:results_melotte_25} shows a results summary for Melotte 25.
 
@@
       \begin{tabular}{l|c|c|c|c}
-        \textbf{Method} & \emph{$\mu_{\alpha}$ $(mas \cdot yr^{-1})$} & \emph{$\mu_{\delta}$ $(mas \cdot yr^{-1})$} & \emph{$\varpi$ $(mas)$} & \emph{\# stars} \\
+        \textbf{Method} & \emph{\(\mu_{\alpha}\) \((mas \cdot yr^{-1})\)} & \emph{\(\mu_{\delta}\) \((mas \cdot yr^{-1})\)}
+        & \emph{\( \varpi \) \((mas)\)} & \emph{\# stars} \\
         \hline
-        \textbf{Simbad}\tablefootnote{Results have been taken from the SIMBAD astronomical database\cite{wenger2000simbad}} & 104.92 $\pm$ 0.12 & -28.00 $\pm$ 0.09 & 21.052 $\pm$ 0.065 & - \\
-        Clusterix & 106.796 $\pm$ 6.229 & -24.870 $\pm$ 5.417 & 21.210 $\pm$ 1.115 & 109 \\
-        K-Means & 79.936 $\pm$ 3.7 & -45.362 $\pm$ 3.8 & 20.901 $\pm$ 0.31 & 374 \\
-        DEC & 104.051 $\pm$ 3.2 & -33.424 $\pm$ 2.5 & 22.072 $\pm$ 0.13 & 219 \\
-        \textbf{DEC (filt.)} & 105.96 $\pm$ 3.5 & -30.00 $\pm$ 2.4 & 21.74 $\pm$ 0.07 & 175 \\
+        \textbf{Simbad}\tablefootnote{Results have been taken from the SIMBAD astronomical database\cite{wenger2000simbad}}
+        & 104.92 \( \pm \) 0.12 & -28.00 \( \pm \) 0.09 & 21.052 \( \pm \) 0.065 & - \\
+        Clusterix & 106.796 \( \pm \) 6.229 & -24.870 \( \pm \) 5.417 & 21.210 \( \pm \) 1.115 & 109 \\
+        K-Means & 79.936 \( \pm \) 3.7 & -45.362 \( \pm \) 3.8 & 20.901 \( \pm \) 0.31 & 374 \\
+        DEC & 104.051 \( \pm \) 3.2 & -33.424 \( \pm \) 2.5 & 22.072 \( \pm \) 0.13 & 219 \\
+        \textbf{DEC (filt.)} & 105.96 \( \pm \) 3.5 & -30.00 \( \pm \) 2.4 & 21.74 \( \pm \) 0.07 & 175 \\
       \end{tabular}
@@
 
-Figure \ref{fig:result_melotte_25_dec} shows the groups found using the DEC model.
+Figure~\ref{fig:result_melotte_25_dec} shows the groups found using the DEC model.
 This model labels Melotte 25 as group \emph{g5}.
@@
 
-Table \ref{tab:hyperparameters_melotte_25} shows the hyperparameters
+Table~\ref{tab:hyperparameters_melotte_25} shows the hyperparameters
 used for the characterization.
@@
 \begin{itemize}
-  \item NGC 2516 (Ap. \ref{sec:ngc2516}) has it proper motion center deviated from the origin but
+  \item NGC 2516 (Ap.~\ref{sec:ngc2516}) has it proper motion center deviated from the origin but
         it is embedded inside a big cloud of stars with similar proper motions.
-  \item NGC 2632 (Ap. \ref{sec:ngc2632}), in addition to Melotte 22, ia s cluster whose proper motion center
+  \item NGC 2632 (Ap.~\ref{sec:ngc2632}), in addition to Melotte 22, ia s cluster whose proper motion center
         is not located at the origin and has a well separated parallax center.
-  \item NGC 2682 (Ap. \ref{sec:ngc2682}), on the other hand, has its parallax centered within the region's gaussian,
+  \item NGC 2682 (Ap.~\ref{sec:ngc2682}), on the other hand, has its parallax centered within the region's gaussian,
         which complicates its detection although its proper motion center is deviated from the origin.
-  \item Melotte 22 (Ap. \ref{sec:melotte22}), as well as Melotte 25, is a cluster closer to us than the other.
+  \item Melotte 22 (Ap.~\ref{sec:melotte22}), as well as Melotte 25, is a cluster closer to us than the other.
         That makes its membership stars to be more scattered than previous clusters which are more compact.
@@
       \begin{tabular}{l|c|c|c|c}
-        \textbf{OC} & \emph{$\alpha$ J2000 $(degrees)$} & \emph{$\delta$ J2000 $(degrees)$} & \emph{Radius $(arcmin)$} & \emph{\# stars} \\
+        \textbf{OC} & \emph{\( \alpha \) J2000 \((degrees)\)} & \emph{\( \delta \) J2000 \((degrees)\)}
+        & \emph{Radius \((arcmin)\)} & \emph{\# stars} \\
         \hline
@@
     }
-    \caption{Righ ascension, declination, radius and number of stars of studied clusters.
+    \caption{Right ascension, declination, radius and number of stars of studied clusters.
              The number of stars corresponds to those stars contained within a cone of center
-             $(\alpha, \delta)$ and raius the cluster's radius multiplied by a factor of 1.5.}
+             \((\alpha, \delta)\) and radius the cluster's radius multiplied by a factor of 1.5.}
     \label{tab:clusters_summary}
@@
 
-\subsection{Future works}
+\subsection{Future work}
 
@@
 \appendix
-\section{APPENDIX}
+\section{Appendix}
 \label{sec:appendix}
@@
       \begin{tabular}{l|c|c|c|c}
-        \textbf{Method} & \emph{$\mu_{\alpha}$ $(mas \cdot yr^{-1})$} & \emph{$\mu_{\delta}$ $(mas \cdot yr^{-1})$} & \emph{$\varpi$ $(mas)$} & \emph{\# stars} \\
+        \textbf{Method} & \emph{\(\mu_{\alpha}\) \((mas \cdot yr^{-1})\)} & \emph{\(\mu_{\delta}\) \((mas \cdot yr^{-1})\)}
+        & \emph{\( \varpi \) \((mas)\)} & \emph{\# stars} \\
         \hline
-        \textbf{Simbad} & -4.6579 $\pm$ 0.0075 & 11.1517 $\pm$ 0.0075 & 2.4118 $\pm$ 0.0006 & 1727 \\
-        Clusterix & -4.652 $\pm$ 0.523 & 11.203 $\pm$ 0.454 & 2.409 $\pm$ 0.127 & 638 \\
-        K-Means & -4.344 $\pm$ 0.14 & 9.507 $\pm$ 0.19 & 2.268 $\pm$ 0.01 & 1542 \\
-        DEC & -4.426 $\pm$ 0.17 & 9.952 $\pm$ 0.20 & 2.436 $\pm$ 0.01 & 1532 \\
-        \textbf{DEC (filt.)} & -4.502 $\pm$ 0.14 & 10.114 $\pm$ 0.17 & 2.392 $\pm$ 0.004 & 1072 \\
+        \textbf{Simbad} & -4.6579 \( \pm \) 0.0075 & 11.1517 \( \pm \) 0.0075 & 2.4118 \( \pm \) 0.0006 & 1727 \\
+        Clusterix & -4.652 \( \pm \) 0.523 & 11.203 \( \pm \) 0.454 & 2.409 \( \pm \) 0.127 & 638 \\
+        K-Means & -4.344 \( \pm \) 0.14 & 9.507 \( \pm \) 0.19 & 2.268 \( \pm \) 0.01 & 1542 \\
+        DEC & -4.426 \( \pm \) 0.17 & 9.952 \( \pm \) 0.20 & 2.436 \( \pm \) 0.01 & 1532 \\
+        \textbf{DEC (filt.)} & -4.502 \( \pm \) 0.14 & 10.114 \( \pm \) 0.17 & 2.392 \( \pm \) 0.004 & 1072 \\
       \end{tabular}
@@
 
-Table \ref{tab:app_results_ngc_2516} shows a results summary for NGC 2516.
-The top row of Figure \ref{fig:app_result_ngc_2516_clusterix_kmeans} shows
+Table~\ref{tab:app_results_ngc_2516} shows a results summary for NGC 2516.
+The top row of Figure~\ref{fig:app_result_ngc_2516_clusterix_kmeans} shows
 the characterization using Clusterix+TOPCAT tools while the bottom row
@@
 
-Figure \ref{fig:app_result_ngc_2516_dec} shows the groups found using the
+Figure~\ref{fig:app_result_ngc_2516_dec} shows the groups found using the
 DEC model (top row) and the DEC model filtered (bottom row).
@@
 groups between K-Means and DEC models may not match.
-Hyperparameters used in this characterization are listed in Table
-\ref{tab:app_hyperparameters_ngc_2516}.
+Hyperparameters used in this characterization are listed in
+Table~\ref{tab:app_hyperparameters_ngc_2516}.
 
@@
 
-Figure \ref{fig:app_result_ngc_2632_clusterix_kmeans} shows NGC 2632 characterized
+Figure~\ref{fig:app_result_ngc_2632_clusterix_kmeans} shows NGC 2632 characterized
 by the validation method (first row) and the characterization made by K-Means
@@
 
-Figure \ref{fig:app_result_ngc_2632_dec} shows the groups found using
+Figure~\ref{fig:app_result_ngc_2632_dec} shows the groups found using
 the DEC model (first row) and the DEC model filtered (second row).
@@
       \begin{tabular}{l|c|c|c|c}
-        \textbf{Method} & \emph{$\mu_{\alpha}$ $(mas \cdot yr^{-1})$} & \emph{$\mu_{\delta}$ $(mas \cdot yr^{-1})$} & \emph{$\varpi$ $(mas)$} & \emph{\# stars} \\
+        \textbf{Method} & \emph{\(\mu_{\alpha}\) \((mas \cdot yr^{-1})\)} & \emph{\(\mu_{\delta}\) \((mas \cdot yr^{-1})\)}
+        & \emph{\( \varpi \) \((mas)\)} & \emph{\# stars} \\
         \hline
-        \textbf{Simbad} & -36.047 $\pm$ 0.110 & -12.917 $\pm$ 0.066 & 5.371 $\pm$ 0.003 & - \\
-        Clusterix & -36.154 $\pm$ 1.001 & -12.909 $\pm$ 0.806 & 5.327 $\pm$ 0.187 & 371 \\
-        K-Means & -26.352 $\pm$ 0.82 & -15.828 $\pm$ 0.76 & 5.394 $\pm$ 0.03 & 629 \\
-        DEC & -20.012 $\pm$ 0.69 & -14.742 $\pm$ 0.58 & 4.686 $\pm$ 0.03 & 894 \\
-        \textbf{DEC (filt.)} & -21.571 $\pm$ 0.74 & -14.234 $\pm$ 0.61 & 4.719 $\pm$ 0.03 & 714 \\
+        \textbf{Simbad} & -36.047 \( \pm \) 0.110 & -12.917 \( \pm \) 0.066 & 5.371 \( \pm \) 0.003 & - \\
+        Clusterix & -36.154 \( \pm \) 1.001 & -12.909 \( \pm \) 0.806 & 5.327 \( \pm \) 0.187 & 371 \\
+        K-Means & -26.352 \( \pm \) 0.82 & -15.828 \( \pm \) 0.76 & 5.394 \( \pm \) 0.03 & 629 \\
+        DEC & -20.012 \( \pm \) 0.69 & -14.742 \( \pm \) 0.58 & 4.686 \( \pm \) 0.03 & 894 \\
+        \textbf{DEC (filt.)} & -21.571 \( \pm \) 0.74 & -14.234 \( \pm \) 0.61 & 4.719 \( \pm \) 0.03 & 714 \\
       \end{tabular}
@@
 
-Table \ref{tab:app_results_ngc_2632} shows a results summary for NGC 2632 analysis.
-While Table \ref{tab:app_hyperparameters_ngc_2632} shows the hyperparameters
+Table~\ref{tab:app_results_ngc_2632} shows a results summary for NGC 2632 analysis.
+While Table~\ref{tab:app_hyperparameters_ngc_2632} shows the hyperparameters
 used for characterizing NGC 2632 with our model.
@@
 
-Table \ref{tab:app_results_ngc_2682} shows a results summary for NGC 2682.
+Table~\ref{tab:app_results_ngc_2682} shows a results summary for NGC 2682.
 
@@
       \begin{tabular}{l|c|c|c|c}
-        \textbf{Method} & \emph{$\mu_{\alpha}$ $(mas \cdot yr^{-1})$} & \emph{$\mu_{\delta}$ $(mas \cdot yr^{-1})$} & \emph{$\varpi$ $(mas)$} & \emph{\# stars} \\
+        \textbf{Method} & \emph{\(\mu_{\alpha}\) \((mas \cdot yr^{-1})\)} & \emph{\(\mu_{\delta}\) \((mas \cdot yr^{-1})\)}
+        & \emph{\( \varpi \) \((mas)\)} & \emph{\# stars} \\
         \hline
-        \textbf{Simbad} & -10.9737 $\pm$ 0.0064 & -2.9396 $\pm$ 0.0063 & 1.1325 $\pm$ 0.0011 & 1194 \\
-        Clusterix+TOPCAT & -10.970 $\pm$ 0.322 & -2.958 $\pm$ 0.327 & 1.142 $\pm$ 0.080 & 649 \\
-        K-Means & -8.616 $\pm$ 0.15 & -3.710 $\pm$ 0.16 & 1.196 $\pm$ 0.01 & 1374 \\
-        DEC & -8.926 $\pm$ 0.15 & -3.550 $\pm$ 0.15 & 1.144 $\pm$ 0.005 & 1238 \\
-        \textbf{DEC (filt.)} & -9.619 $\pm$ 0.13 & -3.317 $\pm$ 0.13 & 1.140 $\pm$ 0.003 & 990 \\
+        \textbf{Simbad} & -10.9737 \( \pm \) 0.0064 & -2.9396 \( \pm \) 0.0063 & 1.1325 \( \pm \) 0.0011 & 1194 \\
+        Clusterix+TOPCAT & -10.970 \( \pm \) 0.322 & -2.958 \( \pm \) 0.327 & 1.142 \( \pm \) 0.080 & 649 \\
+        K-Means & -8.616 \( \pm \) 0.15 & -3.710 \( \pm \) 0.16 & 1.196 \( \pm \) 0.01 & 1374 \\
+        DEC & -8.926 \( \pm \) 0.15 & -3.550 \( \pm \) 0.15 & 1.144 \( \pm \) 0.005 & 1238 \\
+        \textbf{DEC (filt.)} & -9.619 \( \pm \) 0.13 & -3.317 \( \pm \) 0.13 & 1.140 \( \pm \) 0.003 & 990 \\
       \end{tabular}
@@
 
-Figure \ref{fig:app_result_ngc_2682_clusterix_kmeans} shows
+Figure~\ref{fig:app_result_ngc_2682_clusterix_kmeans} shows
 NGC 2682 characterized using Clusterix+TOPCAT tools (top row)
 and ten clusters identified by K-Means (bottom row).
-The chacterization made by K-Means labels NGC 2682 as group \emph{g4}.
+The characterization made by K-Means labels NGC 2682 as group \emph{g4}.
 
@@
 
-Figure \ref{fig:app_result_ngc_2682_dec} shows the groups found using
+Figure~\ref{fig:app_result_ngc_2682_dec} shows the groups found using
 the DEC model (first row) and the DEC model filtered (second row).
@@
 
-Table \ref{tab:app_hyperparameters_ngc_2682} shows the
+Table~\ref{tab:app_hyperparameters_ngc_2682} shows the
 hyperparameters used for characterizing NGC 2682 with DEC method.
@@
 
-\subsection{MELOTTE 22}
+\subsection{Melotte 22}
 \label{sec:melotte22}
@@
 
-Figure \ref{fig:app_result_melotte_22_clusterix_kmeans} shows Melotte 22
+Figure~\ref{fig:app_result_melotte_22_clusterix_kmeans} shows Melotte 22
 characterized using Clusterix+TOPCAT tools (top row) and the characterization
@@
 
-Figure \ref{fig:app_result_melotte_22_dec} shows the groups found using
+Figure~\ref{fig:app_result_melotte_22_dec} shows the groups found using
 DEC model (first row) and DEC model filtered (second row).
@@
 
-Table \ref{tab:app_results_melotte_22} shows a results summary for Melotte 22.
+Table~\ref{tab:app_results_melotte_22} shows a results summary for Melotte 22.
 
@@
       \begin{tabular}{l|c|c|c|c}
-        \textbf{Method} & \emph{$\mu_{\alpha}$ $(mas \cdot yr^{-1})$} & \emph{$\mu_{\delta}$ $(mas \cdot yr^{-1})$} & \emph{$\varpi$ $(mas)$} & \emph{\# stars} \\
+        \textbf{Method} & \emph{\(\mu_{\alpha}\) \((mas \cdot yr^{-1})\)} & \emph{\(\mu_{\delta}\) \((mas \cdot yr^{-1})\)}
+        & \emph{\( \varpi \) \((mas)\)} & \emph{\# stars} \\
         \hline
-        \textbf{Simbad} & 19.997 $\pm$ 0.127 & -45.548 $\pm$ 0.101 & 7.364 $\pm$ 0.005 & 1326 \\
-        Clusterix & 19.98 $\pm$ 1.25 & -45.47 $\pm$ 1.48 & 7.33 $\pm$ 0.21 & 634 \\
-        K-Means & 20.25 $\pm$ 0.95 & -38.01 $\pm$ 1.08 & 7.23 $\pm$ 0.06 & 1378 \\
-        DEC & 23.67 $\pm$ 1.29 & -46.23 $\pm$ 1.50 & 8.04 $\pm$ 0.09 & 878 \\
-        \textbf{DEC (filt.)} & 19.50 $\pm$ 0.41 & -44.23 $\pm$ 0.39 & 7.42 $\pm$ 0.005 & 438 \\
+        \textbf{Simbad} & 19.997 \( \pm \) 0.127 & -45.548 \( \pm \) 0.101 & 7.364 \( \pm \) 0.005 & 1326 \\
+        Clusterix & 19.98 \( \pm \) 1.25 & -45.47 \( \pm \) 1.48 & 7.33 \( \pm \) 0.21 & 634 \\
+        K-Means & 20.25 \( \pm \) 0.95 & -38.01 \( \pm \) 1.08 & 7.23 \( \pm \) 0.06 & 1378 \\
+        DEC & 23.67 \( \pm \) 1.29 & -46.23 \( \pm \) 1.50 & 8.04 \( \pm \) 0.09 & 878 \\
+        \textbf{DEC (filt.)} & 19.50 \( \pm \) 0.41 & -44.23 \( \pm \) 0.39 & 7.42 \( \pm \) 0.005 & 438 \\
       \end{tabular}
@@
 
-While Table \ref{tab:app_hyperparameters_melotte_22} shows the hyperparameters
+While Table~\ref{tab:app_hyperparameters_melotte_22} shows the hyperparameters
 used for characterizing Melotte 22 with the DEC model.
@@
       Number of Clusters & 5 \\
-      Clustering Layer & $\left[ 50, 50, 200 \right]$ \\
+      Clustering Layer & \(\left[ 50, 50, 200 \right]\) \\
       Kernel Initializer Seed & 11 \\
@@
 
-\renewcommand{\refname}{REFERENCES}
+\renewcommand{\refname}{References}
 \bibliographystyle{unsrt}

### 2021-02-10 thesis/thesis.tex

@@
     \end{tabular}
-    \caption{Righ ascension, declination, radius and number of stars of studied clusters.
+    \caption{Right ascension, declination, radius and number of stars of studied clusters.
              The number of stars corresponds to those stars contained within a cone of center
-             $(\alpha, \delta)$ and raius the cluster's radius multiplied by a factor of 1.5.}
+             \((\alpha, \delta)\) and radius the cluster's radius multiplied by a factor of 1.5.}
     \label{tab:clusters_summary}

### 2021-02-10 thesis/thesis.tex

@@
 However, if we look for overdensities in the proper motion configuration spaces, it is possible, at least at first instance,
-to assume a possible membership cut (see Figure~\ref{fig:pm_ngc_2682}.
+to assume a possible membership cut (see Figure~\ref{fig:pm_ngc_2682}).
 
@@
 which makes it possible to distinguish two regions in the workspace:
-the only field region (label `f`), dominated by star fields,
-and the cluster + field region (label `c+f`), which includes both star fields and not star
+the only field region (label `f'), dominated by star fields,
+and the cluster + field region (label `c+f'), which includes both star fields and not star
 fields~\cite{balaguer2020clusterix}.
@@
 This fact implies defining three areas or regions with different radius.
-The first `c+f` corresponds to the one in which the cluster members are
+The first `c+f' corresponds to the one in which the cluster members are
 presumed to be contained together with other star fields that are not part of the cluster.
@@
 
-The right choice of these radii, even having a previous estimation for the `c+f` region,
-highly affects the execution of the algorithm and, in general, requires a considerable wide field `f`.
+The right choice of these radii, even having a previous estimation for the `c+f' region,
+highly affects the execution of the algorithm and, in general, requires a considerable wide field `f'.
 There is no rule of thumb that defines relative proportions of these areas.
@@
 The method includes two phases: the first one uses an unsupervised clustering algorithm, DBSCAN,
-to search for overdensities \((l, b \pi, \mu_{\alpha} *, \mu_{\delta})\),
+to search for overdensities \((l, b, \pi, \mu_{\alpha} *, \mu_{\delta})\),
 and then applies a deep learning Artificial Neural Network (ANN),
@@
 for a sample of the downloaded dataset for Melotte 22.
-At first sight, two main clusters can be distinguished, one of them centered nearly at (0, 0)
-and the second one with center at (20, -45). This second cluster is the one we are looking for.
+At first sight, two main clusters can be distinguished, one of them centered nearly at [0, 0]
+and the second one with center at [20, -45]. This second cluster is the one we are looking for.
 
@@
 this latent space helps us reduce the number of features
-and avoids the \emph{``curse of dimensionality``}~\cite{bellman1961curse}.
+and avoids the \emph{``curse of dimensionality''}~\cite{bellman1961curse}.
 

### 2021-02-21 paper/paper.tex
paper: Fix NGC 2682 Clusterix method name
Add final corrections
paper: Final changes
Co-Authored-By: Cesar Guzmann <15159632+cguz@users.noreply.github.com>

@@
 \resumen{
-The characterization and understanding of \emph{Open Clusters} (OCs) allow us to
+Characterize and understand \emph{Open Clusters} (OCs) allow us to
 understand better properties and mechanisms about the Universe such as stellar
@@
 
-The characterization of an OC refers to the statistical determination
-of the stars that form it. Therefore, the characterization is based on
-the probability of a star to be a member of the cluster~\cite{sampedro2016caracterizacion}.
+Characterize an OC refers to the statistical determination of the stars that
+form it. Therefore, the characterization is based on the probability of a star
+being a member of the cluster~\cite{sampedro2016caracterizacion}.
 
@@
 Our model takes advantage of some features (such as proper motion in right ascension
-and declination, and parallax) to train an Artificial Neural Network.
+and declination, and parallax) to train an ANN.
 The ANN clusters all the stars within a given region in groups.
@@
 
-\section{Related work}
+\section{Related Works}
 \label{sec:related_works}
@@
 
-Clusterix critically depends on the selection of three radii in the studied region.
-The inner radius contains stars belonging or not to the cluster. While the outermost
-radii defines a ring which only contains stars that do not belong to the cluster.
-In addition, it is also sensitive to other parameters, such as the soften parameter
-\emph{h}, or other restrictions related with the searched proper motions.
-
-If the analysis performed by Clusterix is successful, it returns the probability of
-each star to belong to the open cluster. The results can be imported then into TOPCAT
-to continue the refining process to get a valid characterization.
+Clusterix depends fundamentally on the selection of three radii in the studied region.
+The inner radius contains stars that belong or not to the cluster. While the outermost
+radii define a ring that only contains stars that do not belong to the cluster.
+Furthermore, it is also sensitive to other parameters, such as the soften parameter
+\(h\), or other restrictions related to the searched proper motions.
+
+If the analysis performed by Clusterix is successful, it returns the probability that
+each star belongs to the open cluster. The results can then be imported into TOPCAT
+to continue the refinement process and obtain a valid characterization.
 
@@
 learning techniques. It searches systematically for overdensities in the astrometric
-space of the galactic disk and a subsequent identification of open clusters using
+space of the galactic disk and subsequent identification of open clusters using
 photometric information. It includes two phases:
@@
         overdensities,
-  \item and it applies a deep learning ANN to identify isochrone patterns within
+  \item and applies a deep learning ANN to identify isochrone patterns within
         the detected overdensities and thus proceed to confirm them as OCs. They
-        trained previously the ANN with magnitude diagrams.
+        previously trained the ANN with magnitude diagrams.
 \end{enumerate}
 
-They execute their experiments in the Barcelona Supercomputing Center, MareNostrum 42.
+They carry out their experiments at the Barcelona Supercomputing Center, MareNostrum 42.
 Thereby, the neural network handles the image recognition process with isochrone
 patterns without applying theoretical models derived from values such as metallicity
-or masses, among other. Their work concludes with the identification of 582 new open
-clusters distributed along the galactic disk for a galactic declination bellow 20
+or masses, among others. Their work concludes with the identification of 582 new open
+clusters distributed along with the galactic disk for a galactic declination below 20
 degrees, increasing the number of known open clusters by 45\%.
 
-In contrast to~\cite{castro2020hunting}, we lack of a super computer like MareNostrum
-for our experiments. For this reason, our aim is to obtain a new novel method that
-allows the characterization of open clusters without doing a blind search for clusters
-characterization. We advocate of the idea of building an unsupervised and
-non-parameterized method. So, it can be suitable for automated processes.
+In contrast to~\cite{castro2020hunting}, we lack a supercomputer like MareNostrum
+for our experiments. For this reason, our goal is to obtain a new novel method that
+allows the characterization of open clusters without doing a blind search for
+characterization clusters. We advocate the idea of building an unsupervised and
+non-parameterized method. Therefore, it can be suitable for automated processes.
 
@@
 In the literature, we have several clustering algorithms, such as K-Means, Mean-Shift
-Clustering, DBSCAN~\cite{ester1996density} among other, that can help us to achieve
+Clustering, DBSCAN~\cite{ester1996density} among others, that can help us achieve
 our aim. Each one behaves better according to the distribution of the objects to
-clusterize. However, we need to set a large number of clusters to find the open cluster
+be grouped. However, we need to establish a large number of clusters to find the open cluster
 we are looking for. It complicates the identification of the OC because most of the
@@
 
-With all this in mind, our method is based on DEC model. We adapted it to manage data
+With all this in mind, our method is based on the DEC model. We adapted it to manage data
 groups based on the dynamic properties of the stars. Thus, we have an unsupervised
@@
 
-\subsection{Selection of the catalogue}
-
-In this work we make use of Gaia DR2 since DR3 has not been released in time for us to
+\subsection{Selection of the catalog}
+
+In this work, we make use of Gaia DR2 as DR3 has not been released in time for us to
 include it. Gaia DR2 is a multidimensional dataset obtained by ESA's Gaia mission
 (located at L2, 1.5 million kilometers from Earth) and operational since 2014. The
-catalogue has high precision and accuracy astrometric data for more than 1.7 billion
-stellar sources, and magnitudes in three photometric filters (G, BP and RP) for more
+catalog has high precision and accuracy astrometric data for more than 1.7 billion
+stellar sources and magnitudes in three photometric filters (G, BP, and RP) for more
 than 1,300 million sources.
 
-We start by selecting a region from the OpenClust~\cite{dias2002new} catalogue and
-we downloaded it from Gaia DR2 database. The radius of the downloaded region from Gaia
-is 1.5 times larger than the one registered in OpenClust, ensuring to include several
+We start by selecting a region from the OpenClust~\cite{dias2002new} catalog and
+downloading it from the Gaia DR2 database. The radius of the downloaded region from Gaia
+is 1.5 times greater than that recorded in OpenClust, ensuring that it includes several
 stars that do not belong to the open cluster.
@@
 
-However, we are not going to use these raw features. Instead, we are taking a combination
+However, we are not going to use these raw features. Instead, we take a combination
 of them. Figure~\ref{fig:features_melotte_22} shows a pairwise relationship of our
@@
 
-\subsection{Deep Open Clustering of stars}
+\subsection{Deep Open Clustering of Stars}
 \label{sec:deep_open_clustering}
 
-In this section we explain in more detail our model.
-
-In order to define our model, one of the main problems is that we do not have a labeled
-dataset to train a supervised model. Thus, we have to deal with an unsupervised self-trained
-model. We adapted the Unsupervised \emph{DEC} model~\cite{xie2016unsupervised} to our requirements.
-
-Same as the DEC model, we have two main components:
+In this section, we explain our model in more detail.
+
+One of the main problems to define our model is that we do not have a labeled
+dataset to train a supervised model. To solve it, we employ an unsupervised self-trained
+model. Thus, we adapted the Unsupervised \emph{DEC} model~\cite{xie2016unsupervised}
+to our requirements.
+
+Like the DEC model, we have two main components:
 
 \begin{itemize}
-  \item \textbf{deep autoencoder:} It is trained before passing the data through the
-        \emph{clustering layer}. It is composed by encoder layers followed by decoder
-        layers. Recent research has shown that this autoencoder provides meaningful and
-        well-separated representations on real-world datasets~\cite{vincent2010stacked}
-        and~\cite{hinton2006reducing} to the DEC model.
-  \item \textbf{clustering layer:} This layer recibes data transformed by the autoencoder
-        and it is iteratively trained until a convergence criterion is met.
+  \item \textbf{deep autoencoder:} it is a neural network that learns to copy its input
+        to its output. It is composed of encoder layers followed by decoder layers.
+        We train it before passing the data through the next component, the
+        \emph{clustering layer}. Recent research~\cite{vincent2010stacked, hinton2006reducing}
+         showed that the autoencoder provides meaningful and well-separated representations
+         on real-world datasets.
+  \item \textbf{clustering layer:} it receives data transformed by the autoencoder
+        and iteratively is trained until a convergence criterion is met.
 \end{itemize}
@@
 
+We trained the autoencoder before fitting the model to generate predictions.
+Then, we use the encoder layers of the autoencoder with the aim of transforming
+input data to the latent space \(Z\). Once the data has been transformed, a K-Means
+clusterer is used to make an initial clustering. K-Means cluster centers are used
+as the initial weights for the clustering layer.
+
+With that initial configuration, the model iterates alternating between computing an
+auxiliary target distribution (soft assignment) and minimizing the Kullback-Leiber (KL)
+divergence~\cite{kullback1951information} to it. This unsupervised algorithm allows us
+to improve the clustering.
+
 As it was described in Section~\ref{sec:feature_selection}, the number of features we
-deal is not too large. This latent space helps us to start in a reduced number of
-features and avoids the \emph{``curse of dimensionality``}~\cite{bellman1961curse}.
-
-The autoencoder is pretrained before fitting the model to generate predictions. Then,
-the encoder layers of the autoencoder are used with the aim of transforming input data to
-the latent space \(Z\). Once the data has been transformed, a K-Means clusterer is used in
-order to make an initial clustering. K-Means cluster centers are used as the initial
-weights for the clustering layer.
-
-With that initial configuration, the model iterates alternating between computing an auxiliary
-target distribution (Soft Assignment) and minimizing the Kullback-Leiber (KL)
-divergence~\cite{kullback1951information} to it.
-This unsupervised algorithm allows us to improve the clustering.
+deal is not too large. However, this latent space helps us to start with a reduced number
+of features and avoids the \emph{``curse of dimensionality''}~\cite{bellman1961curse}.
+
+In the soft assignment stage, we used \emph{Student's t-distribution} as a kernel to
+measure the similarity between the embedded points and the cluster centroid.
+While in the KL divergence minimization the algorithm iteratively refines clusters by learning
+from their high confidence assignments with the help of an auxiliary target distribution.
+The model is trained by matching the soft assignment to the target distribution.
+The choice of this target distribution is crucial for model's performance.
+
+In this work, we took the target distribution~\cite{xie2016unsupervised}
+defined in Equation~\ref{eq:student_tdistribution}.
 
@@
 
-In the soft assignment stage,
-the \emph{Student's t-distribution} is used as a kernel to measure the similarity
-between the embedded points and the cluster centroid.
-While in the KL divergence minimization the algorithm iteratively refines clusters by learning
-from their high confidence assignments with the help of an auxiliary target distribution.
-The model is trained by matching the soft assignment to the target distribution.
-The choice of this target distribution is crucial for DEC's performance.
-In this work we have taken the target distribution from DEC's original paper~\cite{xie2016unsupervised},
-which is defined in Equation~\ref{eq:student_tdistribution}.
+This target distribution has the following properties:
+
+\begin{itemize}
+  \item strengthen predictions (i.e., improve cluster purity),
+  \item put more emphasis on data points assigned with high confidence,
+  \item normalize loss contribution of each centroid to prevent large clusters
+        from distorting the hidden feature space.
+\end{itemize}
 
@@
   \includegraphics[scale=0.8]{../figures/dec_diagram.pdf}
-  \caption{DEC model layer setup}
+  \caption{Layer setup of our model}
   \label{fig:dec_model_setup}
@@
 
-Figure~\ref{fig:dec_model_setup} shows the layer setup of our DEC model.
-It is simpler than the one tested on the original paper~\cite{xie2016unsupervised},
-since the number of selected features in our work is smaller than in the original one.
-Therefore, using the same configuration would result in a model so
-powerful that would incur in overfitting issues unable to make right predictions.
+Figure~\ref{fig:dec_model_setup} shows the layer setup of our model.
+It is simpler than the used in~\cite{xie2016unsupervised} because we selected a smaller
+number of features. Therefore, using the same configuration would result in a model so
+powerful that it would incur overfitting issues that cannot make the correct predictions.
 
@@
 
-Finally, we can refine this selection by filtering those stars which are bellow and above
+Finally, we can refine this selection by filtering those stars that are below and above
 the 0.10 and 0.90 quantiles for each group, respectively. That way we remove the most
@@
 Figure~\ref{fig:kmeans_comparisons_melotte_22} shows an example. It describes the results of
-applying K-Means with N=2, N=5 and N=8 in Melotte 22. As we can see, large values of the number
-of clusters allow us to isolate more accurately the resonance in parallax at \(\approx 7.3mas\).
-
-The disadvantage is that it forms more groups of stars, which complicates the task of finding the
-desired OC, since we would like to get just two groups. Therefore, with a K-Means we have to find
-a way to set the right value for the number of clusters to isolate the searched cluster without
-creating too many groups. To solve this issue, we can try to estimate the best number of clusters
-by using the \emph{silhouette score}~\cite{rousseeuw1987silhouettes}.
+applying different configurations of K-Means in Melotte 22, N=2, N=5 and N=8. For each configuration
+of K-Means, we show the proper motion in right ascension and declination (figures in the top position),
+and the parallax distribution (figures in the down position).
+
+As we can see in the parallax distribution, lower values of the number of clusters (e.g., N=2) assign
+many stars to the group \emph{0} (purple) that should belong to group \emph{g1}. These badly assigned stars
+are the ones that have the parallax far from the 7.3mas resonance. In contrast, large values of the
+number of clusters allow us to isolate more accurately the resonance in parallax at \(\approx 7.3mas\).
+For instance, at N=8, more groups are better compartmentalized. Group \emph{g2} would be the one corresponding
+to the OC that we are looking for, and all the other groups are those that we would like to be a single
+group. That is the stars that do not belong to the OC. The disadvantage is that K-Means with N=8 forms more
+groups of stars, which complicates the task of finding the desired OC, since we would like to get just two
+groups. Therefore, with a K-Means we have to find  a way to set the right value for the number of clusters
+to isolate the searched cluster without creating too many groups. To solve this issue, we can try to
+estimate the best number of clusters by using the \emph{silhouette score}~\cite{rousseeuw1987silhouettes}.
 
@@
         it is embedded inside a big cloud of stars with similar proper motions.
-  \item NGC 2632 (Ap.~\ref{sec:ngc2632}), in addition to Melotte 22, ia s cluster whose proper motion center
+  \item NGC 2632 (Ap.~\ref{sec:ngc2632}), in addition to Melotte 22, is a cluster whose proper motion center
         is not located at the origin and has a well separated parallax center.
@@
 
-\subsection{Future work}
+\subsection{Future Works}
 
@@
         \textbf{Simbad} & -10.9737 \( \pm \) 0.0064 & -2.9396 \( \pm \) 0.0063 & 1.1325 \( \pm \) 0.0011 & 1194 \\
-        Clusterix+TOPCAT & -10.970 \( \pm \) 0.322 & -2.958 \( \pm \) 0.327 & 1.142 \( \pm \) 0.080 & 649 \\
+        Clusterix & -10.970 \( \pm \) 0.322 & -2.958 \( \pm \) 0.327 & 1.142 \( \pm \) 0.080 & 649 \\
         K-Means & -8.616 \( \pm \) 0.15 & -3.710 \( \pm \) 0.16 & 1.196 \( \pm \) 0.01 & 1374 \\

### 2021-02-25 paper/paper.tex
paper: New corrections
Co-Authored-By: Cesar Guzmann <15159632+cguz@users.noreply.github.com>

@@
 
-\section{Open Cluster Characterization method}
+\section{Open Cluster Characterization Method}
 \label{sec:aims_and_method}
@@
 
-\subsection{Selection of the catalog}
+\subsection{Selection of the Catalog}
 
@@
         to its output. It is composed of encoder layers followed by decoder layers.
-        We train it before passing the data through the next component, the
-        \emph{clustering layer}. Recent research~\cite{vincent2010stacked, hinton2006reducing}
+        We trained it before passing the data through the next component, the
+        \emph{clustering layer}. Recent research~\cite{vincent2010stacked,hinton2006reducing}
         showed that the autoencoder provides meaningful and well-separated representations
@@
 As we can see in the parallax distribution, lower values of the number of clusters (e.g., N=2) assign
-many stars to the group \emph{0} (purple) that should belong to group \emph{g1}. These badly assigned stars
+many stars to the group \emph{g0} (purple) that should belong to group \emph{g1}. These badly assigned stars
 are the ones that have the parallax far from the 7.3mas resonance. In contrast, large values of the
@@
     \centering
-      \includegraphics[width=\textwidth]{../figures/melotte_22/kmeans_pm_melotte_22.png}
-      \includegraphics[width=\textwidth]{../figures/melotte_22/kmeans_hr_diagram_melotte_22.png}
+      \includegraphics[width=\textwidth]{../figures/melotte_25/kmeans_pm_melotte_25.png}
+      \includegraphics[width=\textwidth]{../figures/melotte_25/kmeans_hr_diagram_melotte_25.png}
       \caption{K-Means}
@@
     \centering
-      \includegraphics[width=\textwidth]{../figures/melotte_22/dec_pm_melotte_22.png}
-      \includegraphics[width=\textwidth]{../figures/melotte_22/dec_hr_diagram_melotte_22.png}
+      \includegraphics[width=\textwidth]{../figures/melotte_25/dec_pm_melotte_25.png}
+      \includegraphics[width=\textwidth]{../figures/melotte_25/dec_hr_diagram_melotte_25.png}
       \caption{DEC}
@@
     \centering
-      \includegraphics[width=\textwidth]{../figures/melotte_22/dec_pm_filtered_melotte_22.png}
-      \includegraphics[width=\textwidth]{../figures/melotte_22/dec_hr_diagram_filtered_melotte_22.png}
+      \includegraphics[width=\textwidth]{../figures/melotte_25/dec_pm_filtered_melotte_25.png}
+      \includegraphics[width=\textwidth]{../figures/melotte_25/dec_hr_diagram_filtered_melotte_25.png}
       \caption{DEC (filt.)}
   \end{subfigure}
-  \caption{Evolution of Melotte 22 characterization}
-  \label{fig:melotte_22_characterization_evolution}
+  \caption{Evolution of Melotte 25 characterization}
+  \label{fig:melotte_25_characterization_evolution}
 \end{figure*}
@@
 We have tested our model with the Melotte 25 dataset, which is one of the most difficult dataset
-to characterize OC by its distribution and extension. In order to support the discussion presented
-in Section~\ref{sec:deep_open_clustering}, we compared our model against a K-Means classification
-algorithm. In addition, we perform a second execution only in our model where we filtered the
-quantiles in the Melotte 22 dataset.
+to characterize OC by its distribution and extension, because it is located in a region with
+400,000 stars.
+
+In order to support the discussion presented in Section~\ref{sec:deep_open_clustering}, we compared
+our model against a K-Means classification algorithm. In addition, we perform a second execution
+only in our model where we filtered the quantiles in the Melotte 25 dataset.
 
@@
 
-Figure~\ref{fig:melotte_22_characterization_evolution} shows the evolution
-of the characterization of Melotte 22. The first column corresponds to the
-characterization obtained by applying K-Means alone to the dataset. The middle
-column is the characterization made by the DEC model. And the column on the right
-shows the characterization made by the DEC and having filtered the quantiles
-lower than 0.10 and higher than 0.90. The top row shows the proper motion
-configuration space while the bottom one shows the Hertzsprung–Russell diagrams.
+Figure~\ref{fig:melotte_25_characterization_evolution} shows the evolution of the characterization
+of Melotte 25. The first column corresponds to the characterization obtained by applying K-Means
+(with N=10) alone to the dataset. The middle column is the characterization made by our model.
+And the column on the right shows the characterization made by our model but filtering the
+quantiles lower than 0.10 and higher than 0.90. The top row shows the proper motion configuration
+space, while the bottom one shows the Hertzsprung–Russell diagrams.
 
@@
 
-Melotte 25 is located at 66.725 degrees in right ascension and 15.867
-degrees in declination with a radius of 165 arcmin.
-This region contains more than 400,000 stars.
-
-Table~\ref{tab:results_melotte_25} shows a results summary for Melotte 25.
+The clustering made by K-Means arises 9 groups, but most of them contain stars that do not belong
+to the Melotte 25 OC. However, after applying our model, most of the stars are grouped into a single
+cluster. One of the other groups is the OC we are looking for, and the other seven could be new OCs
+inside the studied region.
+
+The dataset Melotte 25 is located at 66.725 degrees in right ascension and 15.867 degrees in
+declination with a radius of 165 arcmin. As it was mentioned before, this region contains more than
+400,000 stars making it difficult to group stars of the same region.
+
+Table~\ref{tab:results_melotte_25} summarizes the results for the Melotte 25 dataset. The first column
+shows the name of the method. Second and third columns describe the proper motion in right ascension and
+declination, respectively. The fourth column shows the parallax. And the last column presents the number
+of stars corresponding to the OC. Columns from second to fourth have their respective deviations computed
+as the standard error of the predicted group.
+
+By comparing the results obtained with our model, we can see that they are in range with those extracted
+from SIMBAD astronomical database~\cite{wenger2000simbad}. While the results provided by K-Means model are
+far from the official data. This comparison demonstrates that our model improves the results obtained with
+a raw K-Means.
 
@@
         \hline
-        \textbf{Simbad}\tablefootnote{Results have been taken from the SIMBAD astronomical database\cite{wenger2000simbad}}
+        \textbf{Simbad}\tablefootnote{Results have been taken from the SIMBAD astronomical database~\cite{wenger2000simbad}}
         & 104.92 \( \pm \) 0.12 & -28.00 \( \pm \) 0.09 & 21.052 \( \pm \) 0.065 & - \\
-        Clusterix & 106.796 \( \pm \) 6.229 & -24.870 \( \pm \) 5.417 & 21.210 \( \pm \) 1.115 & 109 \\
+        %Clusterix & 106.796 \( \pm \) 6.229 & -24.870 \( \pm \) 5.417 & 21.210 \( \pm \) 1.115 & 109 \\
         K-Means & 79.936 \( \pm \) 3.7 & -45.362 \( \pm \) 3.8 & 20.901 \( \pm \) 0.31 & 374 \\
-        DEC & 104.051 \( \pm \) 3.2 & -33.424 \( \pm \) 2.5 & 22.072 \( \pm \) 0.13 & 219 \\
-        \textbf{DEC (filt.)} & 105.96 \( \pm \) 3.5 & -30.00 \( \pm \) 2.4 & 21.74 \( \pm \) 0.07 & 175 \\
+        DECOCC & 104.051 \( \pm \) 3.2 & -33.424 \( \pm \) 2.5 & 22.072 \( \pm \) 0.13 & 219 \\
+        \textbf{DECOCC (filt.)} & 105.96 \( \pm \) 3.5 & -30.00 \( \pm \) 2.4 & 21.74 \( \pm \) 0.07 & 175 \\
       \end{tabular}
     }
-    \caption{Melotte 25 results.}
+    \caption{Summary for the Melotte 25 results.}
     \label{tab:results_melotte_25}
@@
 
-Parameters shown are proper motion in right ascension and declination, parallax
-with their respective deviations and number of stars corresponding to the OC.
-
-Figure~\ref{fig:result_melotte_25_dec} shows the groups found using the DEC model.
-This model labels Melotte 25 as group \emph{g5}.
-
-\begin{figure}[htbp]
-  \centering
-  \begin{subfigure}{\columnwidth}
+Figure~\ref{fig:result_melotte_25_dec} shows the groups found using our model.
+Our model labels Melotte 25 as group \emph{g5}.
+
+\begin{figure*}[!hbt]
   \centering
@@
   \end{subfigure}
-  \end{subfigure}
-  \caption{Melotte 25 characterization using the DEC model.
-           From the left to the right: proper motion configuration
-           space, parallax histogram and H-R diagram.}
+  \caption{Melotte 25 characterization.
+           From left to right: proper motion configuration space,
+           parallax histogram and H-R diagram.}
   \label{fig:result_melotte_25_dec}
-\end{figure}
+\end{figure*}
 
@@
     \end{tabular}
-    \caption{Melotte 25 DEC hyperparameters.}
+    \caption{Hyperparameters for Melotte 25.}
     \label{tab:hyperparameters_melotte_25}
@@
 
-\section{Result discussions}
+These hyperparameters are used to fine-tuning our model and improve its accuracy. Although they depend on
+the studied region, no one of them is related with physical properties of the cluster. That makes easier
+to automate the selection of hyperparameters.
+
+\paragraph{Final Discussion.}
 

### 2021-03-06 paper/paper.tex
paper: Update title
paper: Improve results section
Co-Authored-By: Cesar Guzmann <15159632+cguz@users.noreply.github.com>

@@
 \usepackage{subcaption}
-\usepackage{tablefootnote}
+\usepackage{tablefootnote, multirow}
 
@@
 %---------------------------
-\title{Machine Learning Tools for Open Cluster Characterization with Gaia DR2 Data}
+\title{Deep Embedded Clustering for Open Cluster Characterization with Gaia DR2 Data}
 \author{CD. Álvaro, C. Guzman, J. Álvaro}
@@
 
-\subsection{Selection of features}
+\subsection{Selection of Features}
 \label{sec:feature_selection}
@@
 
+%% Comparación con K-Means, DEC y DEC filter
+\begin{figure*}[!htp]
+  \centering
+  \begin{subfigure}[b]{0.3\textwidth}
+    \centering
+      \includegraphics[width=\textwidth]{../figures/melotte_25/kmeans_pm_melotte_25.png}
+      \includegraphics[width=\textwidth]{../figures/melotte_25/kmeans_hr_diagram_melotte_25.png}
+      \caption{K-Means}
+  \end{subfigure}
+  \medskip
+  \begin{subfigure}[b]{0.3\textwidth}
+    \centering
+      \includegraphics[width=\textwidth]{../figures/melotte_25/dec_pm_melotte_25.png}
+      \includegraphics[width=\textwidth]{../figures/melotte_25/dec_hr_diagram_melotte_25.png}
+      \caption{DECOCC}
+  \end{subfigure}
+  \medskip
+  \begin{subfigure}[b]{0.3\textwidth}
+    \centering
+      \includegraphics[width=\textwidth]{../figures/melotte_25/dec_pm_filtered_melotte_25.png}
+      \includegraphics[width=\textwidth]{../figures/melotte_25/dec_hr_diagram_filtered_melotte_25.png}
+      \caption{DECOCC (filt.)}
+  \end{subfigure}
+  \caption{Evolution of Melotte 25 characterization}
+  \label{fig:melotte_25_characterization_evolution}
+\end{figure*}
+
 Since we are looking for a single cluster, it seems reasonable to use a clustering algorithm
@@
 
+\newpage
+
 \section{Results}
@@
 
-%% Comparación con K-Means, DEC y DEC filter
-\begin{figure*}[!htbp]
-  \centering
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-      \includegraphics[width=\textwidth]{../figures/melotte_25/kmeans_pm_melotte_25.png}
-      \includegraphics[width=\textwidth]{../figures/melotte_25/kmeans_hr_diagram_melotte_25.png}
-      \caption{K-Means}
-  \end{subfigure}
-  \medskip
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-      \includegraphics[width=\textwidth]{../figures/melotte_25/dec_pm_melotte_25.png}
-      \includegraphics[width=\textwidth]{../figures/melotte_25/dec_hr_diagram_melotte_25.png}
-      \caption{DEC}
-  \end{subfigure}
-  \medskip
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-      \includegraphics[width=\textwidth]{../figures/melotte_25/dec_pm_filtered_melotte_25.png}
-      \includegraphics[width=\textwidth]{../figures/melotte_25/dec_hr_diagram_filtered_melotte_25.png}
-      \caption{DEC (filt.)}
-  \end{subfigure}
-  \caption{Evolution of Melotte 25 characterization}
-  \label{fig:melotte_25_characterization_evolution}
-\end{figure*}
-
-We have tested our model with the Melotte 25 dataset, which is one of the most difficult dataset
-to characterize OC by its distribution and extension, because it is located in a region with
-400,000 stars.
-
-In order to support the discussion presented in Section~\ref{sec:deep_open_clustering}, we compared
-our model against a K-Means classification algorithm. In addition, we perform a second execution
-only in our model where we filtered the quantiles in the Melotte 25 dataset.
-
-Our model was implemented in Python 3.8 using the Keras 2.2 framework and Jupyter
+We performed two tested of our model:
+
+\begin{enumerate}
+  \item with the Melotte 25 dataset, which is one of the most difficult dataset to characterize OC
+        by its distribution and extension, because it is located in a region with 400,000 stars.
+  \item with a variety of clusters of different typologies.
+\end{enumerate}
+
+In order to support the discussion presented in Section~\ref{sec:deep_open_clustering}, in the both
+tests we compared our model against a K-Means classification algorithm. In addition, we perform a
+second execution only in our model where we filtered the quantiles in the Melotte 25 dataset.
+
+Our DECOCC model was implemented in Python 3.8 using the Keras 2.2 framework and Jupyter
 Notebooks~\cite{Kluyver2016jupyter}. All tests were run on an Apple Mac Pro Late 2013 with a
@@
 
-% More results are available at Appendix \ref{sec:appendix}.
+\subsection{Melotte 25 dataset}
 
 Figure~\ref{fig:melotte_25_characterization_evolution} shows the evolution of the characterization
-of Melotte 25. The first column corresponds to the characterization obtained by applying K-Means
-(with N=10) alone to the dataset. The middle column is the characterization made by our model.
-And the column on the right shows the characterization made by our model but filtering the
-quantiles lower than 0.10 and higher than 0.90. The top row shows the proper motion configuration
-space, while the bottom one shows the Hertzsprung–Russell diagrams.
+of Melotte 25. The first column corresponds to the characterization obtained by applying only
+K-Means (with N=10). The middle column shows the result of our model. And the last column describes
+the characterization made by our model but filtering the quantiles lower than 0.10 and higher than
+0.90. The top row shows the proper motion configuration space, while the bottom one shows the
+Hertzsprung–Russell diagrams.
 
@@
 The clustering made by K-Means arises 9 groups, but most of them contain stars that do not belong
-to the Melotte 25 OC. However, after applying our model, most of the stars are grouped into a single
-cluster. One of the other groups is the OC we are looking for, and the other seven could be new OCs
-inside the studied region.
+to the Melotte 25 OC. However, in our model, the stars regroup and although we have the same 9
+groups of K-Means, almost all of the stars are assigned to a group that would contain the stars
+that would not belong to the OC. The remaining groups contain either the OC we are looking for or
+other possible new OCs inside the studied region. Moreover, if we exclude those stars that remain
+below 0.10 and above 0.90 percentiles, and execute again this refinement in our model (third column
+in Figure~\ref{fig:melotte_25_characterization_evolution}), we observe an improvement in the
+results of our DECOCC model.
+
+\begin{figure}[!hbt]
+  % \centering
+  % \begin{subfigure}[t]{0.30\textwidth}
+  %   \centering
+  %   \includegraphics[width=\textwidth]{../figures/melotte_25/dec_pm_melotte_25.png}
+  % \end{subfigure}
+  % \hfill
+  % \begin{subfigure}[t]{0.30\textwidth}
+    \centering
+    \includegraphics[width=0.36\textwidth]{../figures/melotte_25/dec_parallax_melotte_25.png}
+  % \end{subfigure}
+  % \hfill
+  % \begin{subfigure}[t]{0.30\textwidth}
+  %   \centering
+  %   \includegraphics[width=\textwidth]{../figures/melotte_25/dec_hr_diagram_melotte_25.png}
+  % \end{subfigure}
+  % \caption{Melotte 25 characterization.
+  %          From left to right: proper motion configuration space,
+  %          parallax histogram and H-R diagram.}
+  \caption{Parallax histogram of Melotte 25 characterization.}
+  \label{fig:result_melotte_25_dec}
+\end{figure}
 
@@
 declination with a radius of 165 arcmin. As it was mentioned before, this region contains more than
-400,000 stars making it difficult to group stars of the same region.
-
-Table~\ref{tab:results_melotte_25} summarizes the results for the Melotte 25 dataset. The first column
-shows the name of the method. Second and third columns describe the proper motion in right ascension and
-declination, respectively. The fourth column shows the parallax. And the last column presents the number
-of stars corresponding to the OC. Columns from second to fourth have their respective deviations computed
-as the standard error of the predicted group.
-
-By comparing the results obtained with our model, we can see that they are in range with those extracted
-from SIMBAD astronomical database~\cite{wenger2000simbad}. While the results provided by K-Means model are
-far from the official data. This comparison demonstrates that our model improves the results obtained with
-a raw K-Means.
+400,000 stars making it difficult to group stars of the same region. Our model labels Melotte 25 as
+group \emph{g5}. Figure~\ref{fig:result_melotte_25_dec} shows the parallax histogram.
 
@@
         \hline
-        \textbf{Simbad}\tablefootnote{Results have been taken from the SIMBAD astronomical database~\cite{wenger2000simbad}}
-        & 104.92 \( \pm \) 0.12 & -28.00 \( \pm \) 0.09 & 21.052 \( \pm \) 0.065 & - \\
-        %Clusterix & 106.796 \( \pm \) 6.229 & -24.870 \( \pm \) 5.417 & 21.210 \( \pm \) 1.115 & 109 \\
+        Simbad & 104.92 \( \pm \) 0.12 & -28.00 \( \pm \) 0.09 & 21.052 \( \pm \) 0.065 & - \\
+        % Clusterix & 106.796 \( \pm \) 6.229 & -24.870 \( \pm \) 5.417 & 21.210 \( \pm \) 1.115 & 109 \\
         K-Means & 79.936 \( \pm \) 3.7 & -45.362 \( \pm \) 3.8 & 20.901 \( \pm \) 0.31 & 374 \\
         DECOCC & 104.051 \( \pm \) 3.2 & -33.424 \( \pm \) 2.5 & 22.072 \( \pm \) 0.13 & 219 \\
-        \textbf{DECOCC (filt.)} & 105.96 \( \pm \) 3.5 & -30.00 \( \pm \) 2.4 & 21.74 \( \pm \) 0.07 & 175 \\
+        DECOCC (filt.) & 105.96 \( \pm \) 3.5 & -30.00 \( \pm \) 2.4 & 21.74 \( \pm \) 0.07 & 175 \\
       \end{tabular}
@@
 
-Figure~\ref{fig:result_melotte_25_dec} shows the groups found using our model.
-Our model labels Melotte 25 as group \emph{g5}.
-
-\begin{figure*}[!hbt]
-  \centering
-  \begin{subfigure}[t]{0.30\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{../figures/melotte_25/dec_pm_melotte_25.png}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[t]{0.30\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{../figures/melotte_25/dec_parallax_melotte_25.png}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[t]{0.30\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{../figures/melotte_25/dec_hr_diagram_melotte_25.png}
-  \end{subfigure}
-  \caption{Melotte 25 characterization.
-           From left to right: proper motion configuration space,
-           parallax histogram and H-R diagram.}
-  \label{fig:result_melotte_25_dec}
-\end{figure*}
-
-Table~\ref{tab:hyperparameters_melotte_25} shows the hyperparameters
-used for the characterization.
-
-\begin{table}[h]
+Table~\ref{tab:results_melotte_25} summarizes the results for the Melotte 25 dataset. The first column
+shows the name of the method. Second and third columns describe the proper motion in right ascension and
+declination, respectively. The fourth column shows the parallax. And the last column presents the number
+of stars corresponding to the OC. Columns from second to fourth have their respective deviations computed
+as the standard error of the predicted group.
+
+By comparing the results obtained with our DECOCC model, we can see that they are in range with those
+extracted from Simbad astronomical database~\cite{wenger2000simbad}. While the results provided by K-Means
+model are far from the official data. This comparison demonstrates that our model improves the results
+obtained with a raw K-Means.
+
+\begin{table}[!htb]
   \begin{center}
@@
 
-These hyperparameters are used to fine-tuning our model and improve its accuracy. Although they depend on
-the studied region, no one of them is related with physical properties of the cluster. That makes easier
-to automate the selection of hyperparameters.
-
-\paragraph{Final Discussion.}
-
-We have applied our proposed method to clusters with different typologies:
+Table~\ref{tab:hyperparameters_melotte_25} shows the hyperparameters used for the characterization. We used
+them to fine-tuning our model and improve its accuracy. Although they depend on the studied region, none of
+them are related with physical properties of the cluster, which makes it easy to automate the selection of
+hyperparameters.
+
+\subsection{Clusters with different typologies}
+
+We perform now more exhaustive tests of our DECOCC model using open clusters with different typologies (e.g.,
+large, small, more extensive, or more concentrated within the region to be analyzed):
 
 \begin{itemize}
-  \item NGC 2516 (Ap.~\ref{sec:ngc2516}) has it proper motion center deviated from the origin but
-        it is embedded inside a big cloud of stars with similar proper motions.
-  \item NGC 2632 (Ap.~\ref{sec:ngc2632}), in addition to Melotte 22, is a cluster whose proper motion center
-        is not located at the origin and has a well separated parallax center.
-  \item NGC 2682 (Ap.~\ref{sec:ngc2682}), on the other hand, has its parallax centered within the region's gaussian,
-        which complicates its detection although its proper motion center is deviated from the origin.
-  \item Melotte 22 (Ap.~\ref{sec:melotte22}), as well as Melotte 25, is a cluster closer to us than the other.
-        That makes its membership stars to be more scattered than previous clusters which are more compact.
+  \item NGC 2516, which has the proper motion center deviated from the origin but embedded within a large cloud
+        of stars with similar proper motions.
+  \item NGC 2632, which, in addition to Melotte 22, is a cluster whose proper motion center is not at the origin
+        and has a well-separated parallax center.
+  \item NGC 2682, which has the parallax centered within the region's Gaussian, making it difficult its detection.
+        Moreover, same as NGC 2516, the proper motion center is deviated from the origin.
+  \item Melotte 22, as well as Melotte 25, is a cluster closer to our galaxy. That causes its membership stars to
+        be more scattered than previous clusters which are more compact.
 \end{itemize}
 
-All these clusters have a wide variety of diameters, from 25 to 330 arcmin,
-as well as the number of stars that belong to them. NGC 2682 has 3,000 stars
-while Melotte 25 is located inside a region with more than 400,000 stars.
-More information about the studied clusters is available in Table~\ref{tab:clusters_summary}.
+Table~\ref{tab:clusters_summary} describes more specific information about the previous clusters (we also include
+information about the Melotte 25 open cluster, which was previously analyzed). The first column shows the name of
+the cluster. The remaining columns are the right ascension, declination, radius, and the number of stars of studied
+clusters. The column \# stars correspond to those stars contained within a cone of center \((\alpha, \delta)\) and
+radius the cluster's radius multiplied by a factor of 1.5.
 
@@
     }
-    \caption{Right ascension, declination, radius and number of stars of studied clusters.
-             The number of stars corresponds to those stars contained within a cone of center
-             \((\alpha, \delta)\) and radius the cluster's radius multiplied by a factor of 1.5.}
+    \caption{Properties of clusters for further testing of our model.}
     \label{tab:clusters_summary}
@@
 
-In all cases, the model has resolved properly the identity of the cluster
-and has characterized the membership stars, giving compatible results
-with the ones obtained with classic procedures and VO tools.
-
-Compared with the mentioned tools, our model can be categorized as a
-non-parameterized model since it does not depend on parameters referred
-to the cluster itself but only hyperparameters such as the initial number
-of clusters to be found, or the structure of the Clustering Layer used by DEC model.
-
-The proposed method is also non-supervised since we do not need to tell the model
-which stars belong to the cluster or assist it while its training stage.
-
-Only at the end of the characterization, a fine tuning driven by the user,
-can be applied in order to improve the selection by discarding those
-stars which fall outside the given quantiles.
-
-In any case, we do not make assumptions about the cluster profile,
-something that we have to take into account when using VO tools.
-This is another evidence that our model is non-parameterized.
+%% Tabla comparativa de todos los open clusters
+\begin{table*}[!htb]
+  \begin{center}
+    % \resizebox{\columnwidth}{!}{
+      \begin{tabular}{c|l|c|c|c|c}
+         & Method & \emph{\(\mu_{\alpha}\) \((mas \cdot yr^{-1})\)} & \emph{\(\mu_{\delta}\) \((mas \cdot yr^{-1})\)}
+        & \emph{\( \varpi \) \((mas)\)} & \emph{\# stars} \\
+        \hline
+        \multirow{4}{*}{\rotatebox[origin=c]{90}{NGC 2516}}
+        & Simbad & -4.6579 \( \pm \) 0.0075 & 11.1517 \( \pm \) 0.0075 & 2.4118 \( \pm \) 0.0006 & 1727 \\
+        & K-Means & -4.344 \( \pm \) 0.14 & 9.507 \( \pm \) 0.19 & 2.268 \( \pm \) 0.01 & 1542 \\
+        & DECOCC & -4.426 \( \pm \) 0.17 & 9.952 \( \pm \) 0.20 & 2.436 \( \pm \) 0.01 & 1532 \\
+        & DECOCC (filt.) & -4.502 \( \pm \) 0.14 & 10.114 \( \pm \) 0.17 & 2.392 \( \pm \) 0.004 & 1072 \\\\
+
+        % \caption{NGC 2632 results.}
+        \multirow{4}{*}{\rotatebox[origin=c]{90}{NGC 2632}} & Simbad & -36.047 \( \pm \) 0.110 & -12.917 \( \pm \) 0.066 & 5.371 \( \pm \) 0.003 & - \\
+        & K-Means & -26.352 \( \pm \) 0.82 & -15.828 \( \pm \) 0.76 & 5.394 \( \pm \) 0.03 & 629 \\
+        & DECOCC & -20.012 \( \pm \) 0.69 & -14.742 \( \pm \) 0.58 & 4.686 \( \pm \) 0.03 & 894 \\
+        & DECOCC (filt.) & -21.571 \( \pm \) 0.74 & -14.234 \( \pm \) 0.61 & 4.719 \( \pm \) 0.03 & 714 \\\\
+
+        % \caption{NGC 2682 results.}
+         \multirow{4}{*}{\rotatebox[origin=c]{90}{NGC 2682}} & Simbad & -10.9737 \( \pm \) 0.0064 & -2.9396 \( \pm \) 0.0063 & 1.1325 \( \pm \) 0.0011 & 1194 \\
+        & K-Means & -8.616 \( \pm \) 0.15 & -3.710 \( \pm \) 0.16 & 1.196 \( \pm \) 0.01 & 1374 \\
+        & DECOCC & -8.926 \( \pm \) 0.15 & -3.550 \( \pm \) 0.15 & 1.144 \( \pm \) 0.005 & 1238 \\
+        & DECOCC (filt.) & -9.619 \( \pm \) 0.13 & -3.317 \( \pm \) 0.13 & 1.140 \( \pm \) 0.003 & 990 \\\\
+
+        % \caption{Melotte 22 results.}
+        \multirow{4}{*}{\rotatebox[origin=c]{90}{Melotte 22}} & Simbad & 19.997 \( \pm \) 0.127 & -45.548 \( \pm \) 0.101 & 7.364 \( \pm \) 0.005 & 1326 \\
+        & K-Means & 20.25 \( \pm \) 0.95 & -38.01 \( \pm \) 1.08 & 7.23 \( \pm \) 0.06 & 1378 \\
+        & DECOCC & 23.67 \( \pm \) 1.29 & -46.23 \( \pm \) 1.50 & 8.04 \( \pm \) 0.09 & 878 \\
+        & DECOCC (filt.) & 19.50 \( \pm \) 0.41 & -44.23 \( \pm \) 0.39 & 7.42 \( \pm \) 0.005 & 438 \\
+
+      \end{tabular}
+    % }
+    \caption{Summary of the results for the clusters with different typologies.}
+    \label{tab:app_results_ngc_2516}
+  \end{center}
+\end{table*}
+
+As we can see, all those clusters have a wide variety of diameters, from 25 to 330 arcmin, as well as the number
+of stars that belong to them. For instance, NGC 2682 has 3,000 stars while Melotte 25 is inside a region with more
+than 400,000 stars.
+
+Table~\ref{tab:app_results_ngc_2516} describes a summary of the results for the clusters with different typologies.
+Again, our DECOCC model presents similar results from those extracted from the astronomical database Simbad. Except
+for the open cluster NGC 2632. In this cluster, the K-Means presents better results than our model. And this is so
+because the center of the proper motion is not at the origin and generates a more separate center of parallax, which
+makes it difficult to characterize the open cluster in our model. However, our model presents more robust results in
+more difficult open clusters, such as the NGC 2682 that has the parallax centered within the region's Gaussian.
+
+%% comparativa de las figuras de los open clusters
+\begin{figure*}[!hbt]
+  \rotatebox[origin=c]{90}{{\bfseries NGC 2516}\strut}
+  \rotatebox[origin=c]{90}{parallax \strut}
+  \begin{subfigure}{0.3\textwidth}
+    \includegraphics[width=\textwidth]{../figures/ngc_2516/kmeans_parallax_ngc_2516.png}
+  \end{subfigure}
+  \begin{subfigure}{0.3\textwidth}
+    \includegraphics[width=\textwidth]{../figures/ngc_2516/dec_parallax_ngc_2516.png}
+  \end{subfigure}
+  \begin{subfigure}{0.3\textwidth}
+    \includegraphics[width=\textwidth]{../figures/ngc_2516/dec_parallax_filtered_ngc_2516.png}
+  \end{subfigure}
+
+  \rotatebox[origin=c]{90}{{\bfseries NGC 2516} \strut}
+  \rotatebox[origin=c]{90}{hr diagram\strut}
+  \begin{subfigure}{0.3\textwidth}
+    \includegraphics[width=\textwidth]{../figures/ngc_2516/kmeans_hr_diagram_ngc_2516.png}
+  \end{subfigure}
+  \begin{subfigure}{0.3\textwidth}
+    \includegraphics[width=\textwidth]{../figures/ngc_2516/dec_hr_diagram_ngc_2516.png}
+  \end{subfigure}
+  \begin{subfigure}{0.3\textwidth}
+    \includegraphics[width=\textwidth]{../figures/ngc_2516/dec_hr_diagram_filtered_ngc_2516.png}
+  \end{subfigure}
+
+  \rotatebox[origin=c]{90}{{\bfseries NGC 2632}\strut}
+  \rotatebox[origin=c]{90}{parallax \strut}
+  \begin{subfigure}{0.3\textwidth}
+    \includegraphics[width=\textwidth]{../figures/ngc_2632/kmeans_parallax_ngc_2632.png}
+  \end{subfigure}
+  \begin{subfigure}{0.3\textwidth}
+    \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_parallax_ngc_2632.png}
+  \end{subfigure}
+  \begin{subfigure}{0.3\textwidth}
+    \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_parallax_filtered_ngc_2632.png}
+  \end{subfigure}
+
+  \rotatebox[origin=c]{90}{{\bfseries NGC 2632} \strut}
+  \rotatebox[origin=c]{90}{hr diagram\strut}
+  \begin{subfigure}{0.3\textwidth}
+    \includegraphics[width=\textwidth]{../figures/ngc_2632/kmeans_hr_diagram_ngc_2632.png}
+    \caption{K-Means}
+  \end{subfigure}
+  \begin{subfigure}{0.3\textwidth}
+    \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_hr_diagram_ngc_2632.png}
+    \caption{DECOCC}
+  \end{subfigure}%
+  \begin{subfigure}{0.3\textwidth}
+    \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_hr_diagram_filtered_ngc_2632.png}
+    \caption{DECOCC (filt.)}
+  \end{subfigure}
+  \caption{Summary of the characterization for the clusters NGC 2516 and NGC 2632.}
+  \label{fig:ngc_2516_ngc_2632_characterization_evolution}
+\end{figure*}
+
+\begin{figure*}[!hbt]
+  \rotatebox[origin=c]{90}{{\bfseries NGC 2682}\strut}
+  \rotatebox[origin=c]{90}{parallax \strut}
+  \begin{subfigure}{0.3\textwidth}
+    \includegraphics[width=\textwidth]{../figures/ngc_2682/kmeans_parallax_ngc_2682.png}
+  \end{subfigure}
+  \begin{subfigure}{0.3\textwidth}
+    \includegraphics[width=\textwidth]{../figures/ngc_2682/dec_parallax_ngc_2682.png}
+  \end{subfigure}
+  \begin{subfigure}{0.3\textwidth}
+    \includegraphics[width=\textwidth]{../figures/ngc_2682/dec_parallax_filtered_ngc_2682.png}
+  \end{subfigure}
+
+  \rotatebox[origin=c]{90}{{\bfseries NGC 2682} \strut}
+  \rotatebox[origin=c]{90}{hr diagram\strut}
+  \begin{subfigure}{0.3\textwidth}
+    \includegraphics[width=\textwidth]{../figures/ngc_2682/kmeans_hr_diagram_ngc_2682.png}
+  \end{subfigure}
+  \begin{subfigure}{0.3\textwidth}
+    \includegraphics[width=\textwidth]{../figures/ngc_2682/dec_hr_diagram_ngc_2682.png}
+  \end{subfigure}
+  \begin{subfigure}{0.3\textwidth}
+    \includegraphics[width=\textwidth]{../figures/ngc_2682/dec_hr_diagram_filtered_ngc_2682.png}
+  \end{subfigure}
+
+  \rotatebox[origin=c]{90}{{\bfseries Melotte 22}\strut}
+  \rotatebox[origin=c]{90}{parallax \strut}
+  \begin{subfigure}{0.3\textwidth}
+    \includegraphics[width=\textwidth]{../figures/melotte_22/kmeans_parallax_melotte_22.png}
+  \end{subfigure}
+  \begin{subfigure}{0.3\textwidth}
+    \includegraphics[width=\textwidth]{../figures/melotte_22/dec_parallax_melotte_22.png}
+  \end{subfigure}
+  \begin{subfigure}{0.3\textwidth}
+    \includegraphics[width=\textwidth]{../figures/melotte_22/dec_parallax_filtered_melotte_22.png}
+  \end{subfigure}
+
+  \rotatebox[origin=c]{90}{{\bfseries Melotte 22} \strut}
+  \rotatebox[origin=c]{90}{hr diagram\strut}
+  \begin{subfigure}{0.3\textwidth}
+    \includegraphics[width=\textwidth]{../figures/melotte_22/kmeans_hr_diagram_melotte_22.png}
+    \caption{K-Means}
+  \end{subfigure}
+  \begin{subfigure}{0.3\textwidth}
+    \includegraphics[width=\textwidth]{../figures/melotte_22/dec_hr_diagram_melotte_22.png}
+    \caption{DECOCC}
+  \end{subfigure}
+  \begin{subfigure}{0.3\textwidth}
+    \includegraphics[width=\textwidth]{../figures/melotte_22/dec_hr_diagram_filtered_melotte_22.png}
+    \caption{DECOCC (filt.)}
+  \end{subfigure}
+  \caption{Summary of the characterization for the clusters NGC 2682 and Melotte 22.}
+  \label{fig:melotte_22_ngc_2682_characterization_evolution}
+\end{figure*}
+
+Figures~\ref{fig:ngc_2516_ngc_2632_characterization_evolution} and~\ref{fig:melotte_22_ngc_2682_characterization_evolution}
+show the groups found for the three models. For each open cluster, we describe the parallax and hr diagram.
+
+K-Means and DECOCC have labeled NGC 2516 as \emph{g3}. Although in general, groups between K-Means and our
+model may not match.
+
+%% tabla comparativa de los hyperparametros
+\begin{table}[htbp]
+  \begin{center}
+
+    % \resizebox{\columnwidth}{!}{
+        \begin{tabular}{l|l|c}
+          & \textbf{Hyperparameter} & \textbf{Value} \\
+          \hline
+          \multirow{4}{*}{\rotatebox[origin=c]{90}{NGC 2516}} & Number of Clusters & 8 \\
+          & Clustering Layer & \(\left[ 50, 50, 60 \right]\) \\
+          & Kernel Initializer Seed & 2 \\
+          & Quantil Threshold & 0.15 \\\\
+
+          % \caption{NGC 2632 DEC hyperparameters.}
+          \multirow{4}{*}{\rotatebox[origin=c]{90}{NGC 2632}} & Number of Clusters & 10 \\
+          & Clustering Layer & \(\left[ 50, 50, 40 \right]\) \\
+          & Kernel Initializer Seed & 10 \\
+          & Quantil Threshold & 0.1 \\\\
+
+         % \caption{NGC 2682 DEC hyperparameters.}
+          \multirow{4}{*}{\rotatebox[origin=c]{90}{NGC 2682}} & Number of Clusters & 10 \\
+          & Clustering Layer & \(\left[ 50, 50, 40 \right]\) \\
+          & Kernel Initializer Seed & 0 \\
+          & Quantil Threshold & 0.1 \\\\
+
+          % \caption{Melotte 22 DEC hyperparameters.}
+          \multirow{4}{*}{\rotatebox[origin=c]{90}{Melotte 22}} & Number of Clusters & 5 \\
+          & Clustering Layer & \(\left[ 50, 50, 200 \right]\) \\
+          & Kernel Initializer Seed & 11 \\
+          & Quantil Threshold & 0.1 \\
+
+        \end{tabular}
+    % }
+    \caption{Hyperparameters used by DECOCC model for the open clusters with different typologies.}
+    \label{tab:app_hyperparameters_ngc_2516}
+  \end{center}
+\end{table}
+
+Table~\ref{tab:app_hyperparameters_ngc_2516} shows the hyperparameters used for characterizing the
+open clusters with our method.
+
+In all cases, the DECOCC model has resolved properly the identity of the cluster and has characterized
+the membership stars, giving compatible results with the ones obtained with classic procedures and VO
+tools.
+
+Compared with the mentioned tools, our model can be categorized as a non-parameterized model since it
+does not depend on parameters referred to the cluster itself but only hyperparameters such as the
+initial number of clusters to be found, or the structure of the Clustering Layer used by DEC model.
+
+The proposed method is also non-supervised since we do not need to tell the model which stars belong
+to the cluster or assist it while its training stage.
+
+Only at the end of the characterization, a fine-tuning driven by the user, can be applied in order to
+improve the selection by discarding those stars which fall outside the given quantiles.
+
+In any case, we do not make assumptions about the cluster profile, something that we have to take into
+account when using VO tools. This is another evidence that our model is non-parameterized.
 
@@
 
-\appendix
-\section{Appendix}
-\label{sec:appendix}
-
-\subsection{NGC 2516}
-\label{sec:ngc2516}
-
-\begin{table}[htbp]
-  \begin{center}
-    \resizebox{\columnwidth}{!}{
-      \begin{tabular}{l|c|c|c|c}
-        \textbf{Method} & \emph{\(\mu_{\alpha}\) \((mas \cdot yr^{-1})\)} & \emph{\(\mu_{\delta}\) \((mas \cdot yr^{-1})\)}
-        & \emph{\( \varpi \) \((mas)\)} & \emph{\# stars} \\
-        \hline
-        \textbf{Simbad} & -4.6579 \( \pm \) 0.0075 & 11.1517 \( \pm \) 0.0075 & 2.4118 \( \pm \) 0.0006 & 1727 \\
-        Clusterix & -4.652 \( \pm \) 0.523 & 11.203 \( \pm \) 0.454 & 2.409 \( \pm \) 0.127 & 638 \\
-        K-Means & -4.344 \( \pm \) 0.14 & 9.507 \( \pm \) 0.19 & 2.268 \( \pm \) 0.01 & 1542 \\
-        DEC & -4.426 \( \pm \) 0.17 & 9.952 \( \pm \) 0.20 & 2.436 \( \pm \) 0.01 & 1532 \\
-        \textbf{DEC (filt.)} & -4.502 \( \pm \) 0.14 & 10.114 \( \pm \) 0.17 & 2.392 \( \pm \) 0.004 & 1072 \\
-      \end{tabular}
-    }
-    \caption{NGC 2516 results.}
-    \label{tab:app_results_ngc_2516}
-  \end{center}
-\end{table}
-
-Table~\ref{tab:app_results_ngc_2516} shows a results summary for NGC 2516.
-The top row of Figure~\ref{fig:app_result_ngc_2516_clusterix_kmeans} shows
-the characterization using Clusterix+TOPCAT tools while the bottom row
-shows eight clusters found by K-Means.
-
-\begin{figure}[htbp]
-  \centering
-  \begin{subfigure}{\columnwidth}
-    \centering
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/ngc_2516/pm_ngc_2516.png}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/ngc_2516/parallax_ngc_2516.png}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/ngc_2516/hr_diagram_ngc_2516.png}
-    \end{subfigure}
-  \end{subfigure}
-  \centering
-  \begin{subfigure}{\columnwidth}
-    \centering
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/ngc_2516/kmeans_pm_ngc_2516.png}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/ngc_2516/kmeans_parallax_ngc_2516.png}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/ngc_2516/kmeans_hr_diagram_ngc_2516.png}
-    \end{subfigure}
-  \end{subfigure}
-  \caption{NGC 2516 characterization.
-           Top: Clusterix+TOPCAT. Bottom: K-Means (\emph{g3}).}
-  \label{fig:app_result_ngc_2516_clusterix_kmeans}
-\end{figure}
-
-Figure~\ref{fig:app_result_ngc_2516_dec} shows the groups found using the
-DEC model (top row) and the DEC model filtered (bottom row).
-K-Means and DEC have labeled NGC 2516 as \emph{g3}. Although in general,
-groups between K-Means and DEC models may not match.
-Hyperparameters used in this characterization are listed in
-Table~\ref{tab:app_hyperparameters_ngc_2516}.
-
-\begin{figure}[htbp]
-  \centering
-  \begin{subfigure}{\columnwidth}
-    \centering
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/ngc_2516/dec_pm_ngc_2516.png}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/ngc_2516/dec_parallax_ngc_2516.png}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/ngc_2516/dec_hr_diagram_ngc_2516.png}
-    \end{subfigure}
-  \end{subfigure}
-  \centering
-  \begin{subfigure}{\columnwidth}
-    \centering
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/ngc_2516/dec_pm_filtered_ngc_2516.png}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/ngc_2516/dec_parallax_filtered_ngc_2516.png}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/ngc_2516/dec_hr_diagram_filtered_ngc_2516.png}
-    \end{subfigure}
-  \end{subfigure}
-  \caption{NGC 2516 characterization (labeled as \emph{g3}).
-           DEC (top). DEC filtered (bottom).}
-  \label{fig:app_result_ngc_2516_dec}
-\end{figure}
-
-\begin{table}[htbp]
-  \begin{center}
-    \begin{tabular}{l|c}
-      \textbf{Hyperparameter} & \textbf{Value} \\
-      \hline
-      Number of Clusters & 8 \\
-      Clustering Layer & \(\left[ 50, 50, 60 \right]\) \\
-      Kernel Initializer Seed & 2 \\
-      Quantil Threshold & 0.15 \\
-    \end{tabular}
-    \caption{NGC 2516 DEC hyperparameters.}
-    \label{tab:app_hyperparameters_ngc_2516}
-  \end{center}
-\end{table}
-
-\subsection{NGC 2632}
-\label{sec:ngc2632}
-
-Figure~\ref{fig:app_result_ngc_2632_clusterix_kmeans} shows NGC 2632 characterized
-by the validation method (first row) and the characterization made by K-Means
-showing ten clusters (second row). K-Means labels NGC 2632 as \emph{g1}.
-
-\begin{figure}[htbp]
-  \centering
-  \begin{subfigure}{\columnwidth}
-    \centering
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/ngc_2632/pm_ngc_2632.png}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/ngc_2632/parallax_ngc_2632.png}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/ngc_2632/hr_diagram_ngc_2632.png}
-    \end{subfigure}
-  \end{subfigure}
-  \centering
-  \begin{subfigure}{\columnwidth}
-    \centering
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/ngc_2632/kmeans_pm_ngc_2632.png}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/ngc_2632/kmeans_parallax_ngc_2632.png}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/ngc_2632/kmeans_hr_diagram_ngc_2632.png}
-    \end{subfigure}
-  \end{subfigure}
-  \caption{NGC 2632 characterization using Clusterix+TOPCAT (top) and K-Means (bottom).
-           K-Means identifies NGC 2632 as \emph{g1}.}
-  \label{fig:app_result_ngc_2632_clusterix_kmeans}
-\end{figure}
-
-Figure~\ref{fig:app_result_ngc_2632_dec} shows the groups found using
-the DEC model (first row) and the DEC model filtered (second row).
-Open cluster NGC 2632 is labeled as group \emph{g1}.
-
-\begin{figure}[htbp]
-  \centering
-  \begin{subfigure}{\columnwidth}
-    \centering
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_pm_ngc_2632.png}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_parallax_ngc_2632.png}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_hr_diagram_ngc_2632.png}
-    \end{subfigure}
-  \end{subfigure}
-  \centering
-  \begin{subfigure}{\columnwidth}
-    \centering
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_pm_filtered_ngc_2632.png}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_parallax_filtered_ngc_2632.png}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_hr_diagram_filtered_ngc_2632.png}
-    \end{subfigure}
-  \end{subfigure}
-  \caption{NGC 2632 characterization using DEC model (top) and DEC (filtered) (bottom).
-           NGC 2632 is identified as \emph{g1}.}
-  \label{fig:app_result_ngc_2632_dec}
-\end{figure}
-
-\begin{table}[htbp]
-  \begin{center}
-    \resizebox{\columnwidth}{!}{
-      \begin{tabular}{l|c|c|c|c}
-        \textbf{Method} & \emph{\(\mu_{\alpha}\) \((mas \cdot yr^{-1})\)} & \emph{\(\mu_{\delta}\) \((mas \cdot yr^{-1})\)}
-        & \emph{\( \varpi \) \((mas)\)} & \emph{\# stars} \\
-        \hline
-        \textbf{Simbad} & -36.047 \( \pm \) 0.110 & -12.917 \( \pm \) 0.066 & 5.371 \( \pm \) 0.003 & - \\
-        Clusterix & -36.154 \( \pm \) 1.001 & -12.909 \( \pm \) 0.806 & 5.327 \( \pm \) 0.187 & 371 \\
-        K-Means & -26.352 \( \pm \) 0.82 & -15.828 \( \pm \) 0.76 & 5.394 \( \pm \) 0.03 & 629 \\
-        DEC & -20.012 \( \pm \) 0.69 & -14.742 \( \pm \) 0.58 & 4.686 \( \pm \) 0.03 & 894 \\
-        \textbf{DEC (filt.)} & -21.571 \( \pm \) 0.74 & -14.234 \( \pm \) 0.61 & 4.719 \( \pm \) 0.03 & 714 \\
-      \end{tabular}
-    }
-    \caption{NGC 2632 results.}
-    \label{tab:app_results_ngc_2632}
-  \end{center}
-\end{table}
-
-Table~\ref{tab:app_results_ngc_2632} shows a results summary for NGC 2632 analysis.
-While Table~\ref{tab:app_hyperparameters_ngc_2632} shows the hyperparameters
-used for characterizing NGC 2632 with our model.
-
-\begin{table}[htbp]
-  \begin{center}
-    \begin{tabular}{l|c}
-      \textbf{Hyperparameter} & \textbf{Value} \\
-      \hline
-      Number of Clusters & 10 \\
-      Clustering Layer & \(\left[ 50, 50, 40 \right]\) \\
-      Kernel Initializer Seed & 10 \\
-      Quantil Threshold & 0.1 \\
-    \end{tabular}
-    \caption{NGC 2632 DEC hyperparameters.}
-    \label{tab:app_hyperparameters_ngc_2632}
-  \end{center}
-\end{table}
-
-\subsection{NGC 2682}
-\label{sec:ngc2682}
-
-Table~\ref{tab:app_results_ngc_2682} shows a results summary for NGC 2682.
-
-\begin{table}[htbp]
-  \begin{center}
-    \resizebox{\columnwidth}{!}{
-      \begin{tabular}{l|c|c|c|c}
-        \textbf{Method} & \emph{\(\mu_{\alpha}\) \((mas \cdot yr^{-1})\)} & \emph{\(\mu_{\delta}\) \((mas \cdot yr^{-1})\)}
-        & \emph{\( \varpi \) \((mas)\)} & \emph{\# stars} \\
-        \hline
-        \textbf{Simbad} & -10.9737 \( \pm \) 0.0064 & -2.9396 \( \pm \) 0.0063 & 1.1325 \( \pm \) 0.0011 & 1194 \\
-        Clusterix & -10.970 \( \pm \) 0.322 & -2.958 \( \pm \) 0.327 & 1.142 \( \pm \) 0.080 & 649 \\
-        K-Means & -8.616 \( \pm \) 0.15 & -3.710 \( \pm \) 0.16 & 1.196 \( \pm \) 0.01 & 1374 \\
-        DEC & -8.926 \( \pm \) 0.15 & -3.550 \( \pm \) 0.15 & 1.144 \( \pm \) 0.005 & 1238 \\
-        \textbf{DEC (filt.)} & -9.619 \( \pm \) 0.13 & -3.317 \( \pm \) 0.13 & 1.140 \( \pm \) 0.003 & 990 \\
-      \end{tabular}
-    }
-    \caption{NGC 2682 results.}
-    \label{tab:app_results_ngc_2682}
-  \end{center}
-\end{table}
-
-Figure~\ref{fig:app_result_ngc_2682_clusterix_kmeans} shows
-NGC 2682 characterized using Clusterix+TOPCAT tools (top row)
-and ten clusters identified by K-Means (bottom row).
-The characterization made by K-Means labels NGC 2682 as group \emph{g4}.
-
-\begin{figure}[htbp]
-  \centering
-  \begin{subfigure}{\columnwidth}
-    \centering
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/ngc_2682/pm_ngc_2682.png}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/ngc_2682/parallax_ngc_2682.png}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/ngc_2682/hr_diagram_ngc_2682.png}
-    \end{subfigure}
-  \end{subfigure}
-  \centering
-  \begin{subfigure}{\columnwidth}
-    \centering
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/ngc_2682/kmeans_pm_ngc_2682.png}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/ngc_2682/kmeans_parallax_ngc_2682.png}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/ngc_2682/kmeans_hr_diagram_ngc_2682.png}
-    \end{subfigure}
-  \end{subfigure}
-  \caption{NGC 2682 K-Means characterization.
-           Top row corresponds to Clusterix+TOPCAT characterization,
-           while bottom row shows K-Means results identifying
-           NGC 2682 as cluster \emph{g4}.}
-  \label{fig:app_result_ngc_2682_clusterix_kmeans}
-\end{figure}
-
-Figure~\ref{fig:app_result_ngc_2682_dec} shows the groups found using
-the DEC model (first row) and the DEC model filtered (second row).
-The cluster of interest is the group \emph{g2}.
-
-\begin{figure}[htbp]
-  \centering
-  \begin{subfigure}{\columnwidth}
-    \centering
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/ngc_2682/dec_pm_ngc_2682.png}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/ngc_2682/dec_parallax_ngc_2682.png}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/ngc_2682/dec_hr_diagram_ngc_2682.png}
-    \end{subfigure}
-  \end{subfigure}
-  \centering
-  \begin{subfigure}{\columnwidth}
-    \centering
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/ngc_2682/dec_pm_filtered_ngc_2682.png}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/ngc_2682/dec_parallax_filtered_ngc_2682.png}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/ngc_2682/dec_hr_diagram_filtered_ngc_2682.png}
-    \end{subfigure}
-  \end{subfigure}
-  \caption{NGC 2682 is identified as cluster \emph{g2}.
-           Top row: DEC --- Bottom row: DEC (filtered)}
-  \label{fig:app_result_ngc_2682_dec}
-\end{figure}
-
-Table~\ref{tab:app_hyperparameters_ngc_2682} shows the
-hyperparameters used for characterizing NGC 2682 with DEC method.
-
-\begin{table}[htbp]
-  \begin{center}
-    \begin{tabular}{l|c}
-      \textbf{Hyperparameter} & \textbf{Value} \\
-      \hline
-      Number of Clusters & 10 \\
-      Clustering Layer & \(\left[ 50, 50, 40 \right]\) \\
-      Kernel Initializer Seed & 0 \\
-      Quantil Threshold & 0.1 \\
-    \end{tabular}
-    \caption{NGC 2682 DEC hyperparameters.}
-    \label{tab:app_hyperparameters_ngc_2682}
-  \end{center}
-\end{table}
-
-\subsection{Melotte 22}
-\label{sec:melotte22}
-
-\begin{figure}[htbp]
-  \centering
-  \begin{subfigure}{\columnwidth}
-    \centering
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/melotte_22/pm_melotte_22.png}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/melotte_22/parallax_melotte_22.png}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/melotte_22/hr_diagram_melotte_22.png}
-    \end{subfigure}
-  \end{subfigure}
-  \centering
-  \begin{subfigure}{\columnwidth}
-    \centering
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/melotte_22/kmeans_pm_melotte_22.png}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/melotte_22/kmeans_parallax_melotte_22.png}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/melotte_22/kmeans_hr_diagram_melotte_22.png}
-    \end{subfigure}
-  \end{subfigure}
-  \caption{Melotte 22 characterization.
-           Top row: Clusterix+TOPCAT. Bottom row: K-Means.
-           Melotte 22 is labeled as cluster \emph{g1}.}
-  \label{fig:app_result_melotte_22_clusterix_kmeans}
-\end{figure}
-
-Figure~\ref{fig:app_result_melotte_22_clusterix_kmeans} shows Melotte 22
-characterized using Clusterix+TOPCAT tools (top row) and the characterization
-made by K-Means (bottom row) showing nine clusters. Melotte 22 is labeled as
-group \emph{g1}.
-
-Figure~\ref{fig:app_result_melotte_22_dec} shows the groups found using
-DEC model (first row) and DEC model filtered (second row).
-Melotte 22 is labeled as group \emph{g2} by DEC model.
-
-\begin{figure}[htbp]
-  \centering
-  \begin{subfigure}{\columnwidth}
-    \centering
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/melotte_22/dec_pm_melotte_22.png}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/melotte_22/dec_parallax_melotte_22.png}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/melotte_22/dec_hr_diagram_melotte_22.png}
-    \end{subfigure}
-  \end{subfigure}
-  \centering
-  \begin{subfigure}{\columnwidth}
-    \centering
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/melotte_22/dec_pm_filtered_melotte_22.png}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/melotte_22/dec_parallax_filtered_melotte_22.png}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.30\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{../figures/melotte_22/dec_hr_diagram_filtered_melotte_22.png}
-    \end{subfigure}
-  \end{subfigure}
-  \caption{Melotte 22 is identified as cluster \emph{g2}.
-           Top row: DEC --- Bottom row: DEC (filtered)}
-  \label{fig:app_result_melotte_22_dec}
-\end{figure}
-
-Table~\ref{tab:app_results_melotte_22} shows a results summary for Melotte 22.
-
-\begin{table}[htbp]
-  \begin{center}
-    \resizebox{\columnwidth}{!}{
-      \begin{tabular}{l|c|c|c|c}
-        \textbf{Method} & \emph{\(\mu_{\alpha}\) \((mas \cdot yr^{-1})\)} & \emph{\(\mu_{\delta}\) \((mas \cdot yr^{-1})\)}
-        & \emph{\( \varpi \) \((mas)\)} & \emph{\# stars} \\
-        \hline
-        \textbf{Simbad} & 19.997 \( \pm \) 0.127 & -45.548 \( \pm \) 0.101 & 7.364 \( \pm \) 0.005 & 1326 \\
-        Clusterix & 19.98 \( \pm \) 1.25 & -45.47 \( \pm \) 1.48 & 7.33 \( \pm \) 0.21 & 634 \\
-        K-Means & 20.25 \( \pm \) 0.95 & -38.01 \( \pm \) 1.08 & 7.23 \( \pm \) 0.06 & 1378 \\
-        DEC & 23.67 \( \pm \) 1.29 & -46.23 \( \pm \) 1.50 & 8.04 \( \pm \) 0.09 & 878 \\
-        \textbf{DEC (filt.)} & 19.50 \( \pm \) 0.41 & -44.23 \( \pm \) 0.39 & 7.42 \( \pm \) 0.005 & 438 \\
-      \end{tabular}
-    }
-    \caption{Melotte 22 results.}
-    \label{tab:app_results_melotte_22}
-  \end{center}
-\end{table}
-
-While Table~\ref{tab:app_hyperparameters_melotte_22} shows the hyperparameters
-used for characterizing Melotte 22 with the DEC model.
-
-\begin{table}[hbtp]
-  \begin{center}
-    \begin{tabular}{l|c}
-      \textbf{Hyperparameter} & \textbf{Value} \\
-      \hline
-      Number of Clusters & 5 \\
-      Clustering Layer & \(\left[ 50, 50, 200 \right]\) \\
-      Kernel Initializer Seed & 11 \\
-      Quantil Threshold & 0.1 \\
-    \end{tabular}
-    \caption{Melotte 22 DEC hyperparameters.}
-    \label{tab:app_hyperparameters_melotte_22}
-  \end{center}
-\end{table}
-
 \renewcommand{\refname}{References}

### 2021-03-07 paper/paper.tex
paper: Last changes
Co-Authored-By: Cesar Guzmann <15159632+cguz@users.noreply.github.com>

@@
 
-\subsection{Melotte 25 dataset}
+\subsection{Melotte 25 Dataset}
 
@@
 By comparing the results obtained with our DECOCC model, we can see that they are in range with those
-extracted from Simbad astronomical database~\cite{wenger2000simbad}. While the results provided by K-Means
-model are far from the official data. This comparison demonstrates that our model improves the results
-obtained with a raw K-Means.
+extracted from Simbad astronomical database~\cite{wenger2000simbad} (first row in
+Table~\ref{tab:results_melotte_25}). While the results provided by K-Means model are far from the official
+data. This comparison demonstrates that our model improves the results obtained with a raw K-Means.
 
@@
 
-\subsection{Clusters with different typologies}
+\subsection{Clusters With Different Typologies}
 
@@
         of stars with similar proper motions.
-  \item NGC 2632, which, in addition to Melotte 22, is a cluster whose proper motion center is not at the origin
-        and has a well-separated parallax center.
+  \item NGC 2632, which, in addition to Melotte 22 (explained below), is a cluster whose proper motion center is
+        not at the origin and has a well-separated parallax center.
   \item NGC 2682, which has the parallax centered within the region's Gaussian, making it difficult its detection.
@@
 the cluster. The remaining columns are the right ascension, declination, radius, and the number of stars of studied
-clusters. The column \# stars correspond to those stars contained within a cone of center \((\alpha, \delta)\) and
-radius the cluster's radius multiplied by a factor of 1.5.
+clusters. The last column (\# stars) corresponds to those stars contained within a cone of center \((\alpha, \delta)\)
+and radius the cluster's radius multiplied by a factor of 1.5.
 
@@
 Table~\ref{tab:app_results_ngc_2516} describes a summary of the results for the clusters with different typologies.
-Again, our DECOCC model presents similar results from those extracted from the astronomical database Simbad. Except
-for the open cluster NGC 2632. In this cluster, the K-Means presents better results than our model. And this is so
-because the center of the proper motion is not at the origin and generates a more separate center of parallax, which
-makes it difficult to characterize the open cluster in our model. However, our model presents more robust results in
-more difficult open clusters, such as the NGC 2682 that has the parallax centered within the region's Gaussian.
+Again, our DECOCC model presents similar results from those extracted from the astronomical database Simbad. Except for
+the open cluster NGC 2632. This is so because the center of the proper motion is not at the origin and generates a more
+separate center of parallax, which makes it difficult to characterize the open cluster. Compared to K-Means, our model
+presents more robust results in more difficult open clusters, such as the NGC 2682 that has the parallax centered within
+the region's Gaussian.
 
@@
 
-  \rotatebox[origin=c]{90}{{\bfseries NGC 2516} \strut}
+  \rotatebox[origin=c]{90}{{\bfseries} \strut}
   \rotatebox[origin=c]{90}{hr diagram\strut}
@@
 
-  \rotatebox[origin=c]{90}{{\bfseries NGC 2632} \strut}
+  \rotatebox[origin=c]{90}{{\bfseries} \strut}
   \rotatebox[origin=c]{90}{hr diagram\strut}
@@
 
-  \rotatebox[origin=c]{90}{{\bfseries NGC 2682} \strut}
+  \rotatebox[origin=c]{90}{{\bfseries} \strut}
   \rotatebox[origin=c]{90}{hr diagram\strut}
@@
 
-  \rotatebox[origin=c]{90}{{\bfseries Melotte 22} \strut}
+  \rotatebox[origin=c]{90}{{\bfseries} \strut}
   \rotatebox[origin=c]{90}{hr diagram\strut}
@@
 
-Figures~\ref{fig:ngc_2516_ngc_2632_characterization_evolution} and~\ref{fig:melotte_22_ngc_2682_characterization_evolution}
-show the groups found for the three models. For each open cluster, we describe the parallax and hr diagram.
+Figures~\ref{fig:ngc_2516_ngc_2632_characterization_evolution} (describes NGC 2516 and NGC 2632)
+and~\ref{fig:melotte_22_ngc_2682_characterization_evolution} (describes NGC 2682 and Melotte 22) show the
+groups found for the three models. For each open cluster, we describe the parallax and hr diagram.
 
-K-Means and DECOCC have labeled NGC 2516 as \emph{g3}. Although in general, groups between K-Means and our
-model may not match.
+In each open cluster, the models K-Means and DECOCC label the open cluster in the following groups:
+
+\begin{itemize}
+  \item NGC 2516: K-Means and our model label as \emph{g3}.
+  \item NGC 2632: K-Means and our model label as \emph{g1}.
+  \item NGC 2682: K-Means labels as \emph{g4}, and our model labels as \emph{g2}.
+  \item Melotte 22: K-Means labels as \emph{g1}, and our model labels as \emph{g2}.
+\end{itemize}
+
+As we can observe, labels between K-Means and DECOCC do not match (except for NGC 2632), since we tested these two
+models independently. Thus, each model can label each group differently from the other. Moreover, running the same model
+twice can label the resulting groups differently between the first and the second run (as long as the same seed is not
+assigned). However, regardless whether the labels are different, the stars within the generated clusters should be very
+similar.
+
+In all cases, the DECOCC model resolved properly the identity of the cluster and characterized the membership stars,
+giving compatible results with the ones obtained with classic procedures and VO tools~\cite{valdivielso2009iphas}.
+
+Table~\ref{tab:app_hyperparameters_ngc_2516} describes the hyperparameters used for characterizing the open clusters
+with our method. Compared with the VO tools, we categorize our model as a non-parameterized model since it does not
+depend on parameters referred to as the cluster itself but only on hyperparameters such as the initial number of
+clusters to be found, or the structure of the Clustering Layer.
+
+The proposed method is also non-supervised since we do not need to tell the model which stars belong to the cluster or
+assist it while its training stage.
+
+Only at the end of the characterization, a fine-tuning driven by the user is applied to improve the selection by
+discarding those stars which fall outside the given quantiles. Nevertheless, we do not conceive assumptions about the
+cluster profile, something that we consider in VO tools. This is another evidence that our model is non-parameterized.
 
@@
     % }
-    \caption{Hyperparameters used by DECOCC model for the open clusters with different typologies.}
+    \caption{Hyperparameters of DECOCC model for the open clusters with different typologies.}
     \label{tab:app_hyperparameters_ngc_2516}
@@
 
-Table~\ref{tab:app_hyperparameters_ngc_2516} shows the hyperparameters used for characterizing the
-open clusters with our method.
-
-In all cases, the DECOCC model has resolved properly the identity of the cluster and has characterized
-the membership stars, giving compatible results with the ones obtained with classic procedures and VO
-tools.
-
-Compared with the mentioned tools, our model can be categorized as a non-parameterized model since it
-does not depend on parameters referred to the cluster itself but only hyperparameters such as the
-initial number of clusters to be found, or the structure of the Clustering Layer used by DEC model.
-
-The proposed method is also non-supervised since we do not need to tell the model which stars belong
-to the cluster or assist it while its training stage.
-
-Only at the end of the characterization, a fine-tuning driven by the user, can be applied in order to
-improve the selection by discarding those stars which fall outside the given quantiles.
-
-In any case, we do not make assumptions about the cluster profile, something that we have to take into
-account when using VO tools. This is another evidence that our model is non-parameterized.
+In summary, we can say that our results improve K-Means. We based this judgment by comparing right ascension,
+declination, parallax, and the number of stars of the cluster found by the two models with the Simbad database. In other
+words, we statistically analyze the results not individualized by stars belonging to the cluster found. We arrange it
+with Simbad because as far as we are concerned, there is no Gaia dataset with classified open clusters.
 
@@
 
-From the results shown in Section~\ref{sec:results}, we can say that, in general,
-our model works fine identifying and characterizing open clusters.
-So we have succeeded building a model non-parameterized and non-supervised
-for open cluster characterization. However, some improvements should be done
-in order to improve its accuracy and precision.
+This paper presents a model to characterize open clusters, comprising an Artificial Neural Network, which is neither
+parameterized nor supervised. The model does not require complex hardware. Instead, it can be run on workstations with a
+common GPU, making it accessible for deployment in many research centers.
 
-Another achievement is that we have developed a model that does not require
-complex hardware. Instead, it can be run on workstations with a common GPU,
-making it accessible to be implemented in many research centers.
+The model works well in identifying and characterizing open clusters. However, some improvements must be made to improve
+its accuracy and precision.
 
-We have built a non-parameterized model since the user is not required to provide
-any information about the cluster profile or its properties.
-Nevertheless some tweaks can be done regarding the cluster, like changing the initial
-number of clusters or modifying the clustering layer structure in order to improve
-the final characterization.
+The number of hyperparameters in our model is small and we do not need previous knowledge about the cluster studied,
+which makes it easier to test different settings and automate the process.
 
-Another key point in our model is the initializer kernel.
-This kernel prepares data before passing it to the ANN
-and the results may vary significantly depending on this kernel.
-This is something to avoid, so an improvement to solve this issue is necessary
-in order to have a reliable model which does not depend on the dataset order.
+Building a non-parameterized model is a difficult and sometimes and unfeasible task due to every machine learning model
+requires initial definitions of the hyperparameters to work properly. However, in our model, the user is not required to
+provide any information about the cluster profile or its properties. Making it a non-parameterized model. Nevertheless,
+some tweaks can be done regarding the cluster and the studied region, like changing the initial number of clusters or
+modifying the clustering layer structure to improve the final characterization.
 
-Despite that, comparing our method with Clusterix, we can still say that the
-number of parameters (or hyperparameters) in our model is smaller and that
-we do not need previous knowledge about the studied cluster.
-That makes easier to test different configurations and automate the process.
+The initializer kernel prepares data before passing them to the ANN. This is one of the reasons because the results may
+vary significantly. In future works, we want to avoid it to have a reliable model which does not depend on the dataset
+order.
 
-Furthermore, our model is able to deal with large and open regions. As an example,
-it has succeeded characterizing Melotte 25 (Hiades) OC. Also, the relative proximity
-of this cluster, 45 parsec, makes it difficult to characterize. Clusterix fails
-characterizing this cluster. In this case, our method proposes at least two more
-non-catalogued clusters. The analysis of these new clusters brings out an interesting
-stellar dynamics in the studied region. It could mean that there was a recent
-interaction among different molecular clouds.
+In future, we also want to test our model with a wider range of clusters. In general, we could think of applying it to
+the entire VizieR catalog. So, we could get a better idea about the current limitations of our model and improve it. For
+instance, we could determine with higher accuracy the typologies that our model works well with. Furthermore, we could
+establish different sets of hyperparameters regarding the typology of the cluster.
 
-For all this, we consider that the presented method is good enough to be included
-in the Virtual Observatory toolset. Only a few tweaks must be done in order to make
-our model compatible with other tools of the VO.
+An important aspect we noticed is that with our model new uncatalogued clusters arise from the data apart of the main
+ones. Making it of special interest to study these individual new clusters.
 
-\subsection{Future Works}
+Another possible improvements are either to use DBSCAN instead of K-Means as our initial clustering algorithm, and
+employ DR3 dataset instead of DR2 dataset of Gaia.
 
-One open point to do from now is testing the model with a wider range of clusters.
-In general we could think of applying it to the whole VizieR catalogue.
-That way we could have a better idea about the current limitations of our model
-and we could make some adjustments to improve it. We could determine with higher
-accuracy the typologies that our model works well with, and which ones the model
-does not resolve properly. Furthermore, we could stablish different sets of
-hyperparameters regarding the typology of the cluster.
-
-With our model, new uncatalogued clusters arise from the data apart of the main ones.
-Hence, it is also interesting the individual study of these new clusters.
-
-Another possible change in our model would be using DBSCAN instead of K-Means
-as our initial clustering algorithm. Maybe this algorithm gives us a better starting
-point for the DEC model which could improve the results.
-
-Finally, we have used the Gaia DR2 database as our data source for this work,
-but recently the DR3 dataset has been released.
-Therefore, it is evident the interest on testing our model with this new data.
+Furthermore, our model can deal with large and open regions. As an example, it has succeeded in characterizing Melotte
+25 (Hiades) OC. Also, the relative proximity of this cluster, 45 parsecs, makes it difficult to characterize. Our method
+proposes at least two more non-catalogued clusters. The analysis of these new clusters brings out interesting stellar
+dynamics in the studied region. It could mean that there was a recent interaction among different molecular clouds.
 

### 2021-03-09 paper/paper.tex

@@
 
-%---------------------------
-%marges
-%---------------------------
-%\usepackage[margin=1.9cm]{geometry}
-%---------------------------
-%---------------------------
-%---------------------------
-%---------------------------
+\resumen{
+Characterize and understand \emph{Open Clusters} (OCs) allow us to understand better properties and mechanisms about the Universe such as stellar formation and the regions where these events occur.
+They also provide information about stellar processes and the evolution of the galactic disk.
 
-\resumen{
-Characterize and understand \emph{Open Clusters} (OCs) allow us to
-understand better properties and mechanisms about the Universe such as stellar
-formation and the regions where these events occur. They also provide information
-about stellar processes and the evolution of the galactic disk.
-
-In this paper, we present a novel method to characterize OCs. Our method employs a
-model built on \emph{Artificial Neural Networks} (ANNs). More specifically, we adapted
-a state of the art model, the \emph{Deep Embedded Clustering} (DEC) model for our purpose.
-The developed method aims to improve classical state of the arts techniques. We improved
-not only in terms of computational efficiency (with lower computational requirements),
-but in usability (reducing the number of hyperparameters to get a good characterization
-of the analyzed clusters). For our experiments, we used the \emph{Gaia DR2 database} as
-the data source, and compared our model with the clustering technique \emph{K-Means}. Our
-method achieves good results, becoming even better (in some of the cases) than current techniques.
+In this paper, we present a novel method to characterize OCs.
+Our method employs a model built on \emph{Artificial Neural Networks} (ANNs).
+More specifically, we adapted a state of the art model, the \emph{Deep Embedded Clustering} (DEC) model for our purpose.
+The developed method aims to improve classical state of the arts techniques.
+We improved not only in terms of computational efficiency (with lower computational requirements), but in usability (reducing the number of hyperparameters to get a good characterization of the analyzed clusters).
+For our experiments, we used the \emph{Gaia DR2 database} as the data source, and compared our model with the clustering technique \emph{K-Means}.
+Our method achieves good results, becoming even better (in some of the cases) than current techniques.
 }
@@
 
-%\renewcommand{\listfigurename}{Índice de Ilustraciones}
-%\renewcommand{\listtablename}{Índice de Tablas}
-%\renewcommand{\contentsname}{Índice de Contenidos}
-%\renewcommand{\figurename}{Figura}
-%\renewcommand{\tablename}{Tabla}
-%\twocolumn
-
-%\frontmatter
-%\tableofcontents
-%\listoffigures
-%\listoftables
-
 \section{Introduction}
 
-Stellar OCs~\cite{janes1982open} are groups of stars gravitationally
-bounded originated from a single molecular gas cloud.
-They share the same chemical composition and age, and that is the reason
-that their metallicity must be uniform since those stars were born from
-the same gas cloud and at the same time stage. Likewise, they have similar
-relative positions inherited from their original gas cloud, which means that
-their distances to the Earth are the same. Therefore, they have a narrow
-dispersion in their parallax value. Moreover, they share similar values
-of proper motion, both in right ascension and declination. For all these
-properties, we can see why the stellar OCs are relevant to understand the
-spiral structure, dynamics, and chemical evolution of our galaxy.
+Stellar OCs~\cite{janes1982open} are groups of stars gravitationally bounded originated from a single molecular gas cloud.
+They share the same chemical composition and age, and that is the reason that their metallicity must be uniform since those stars were born from the same gas cloud and at the same time stage.
+Likewise, they have similar relative positions inherited from their original gas cloud, which means that their distances to the Earth are the same.
+Therefore, they have a narrow dispersion in their parallax value.
+Moreover, they share similar values of proper motion, both in right ascension and declination.
+For all these properties, we can see why the stellar OCs are relevant to understand the spiral structure, dynamics, and chemical evolution of our galaxy.
 
-The study of OCs advanced thanks to the huge and precise dataset
-from the Gaia DR2 mission, available since 2018 from~\cite{collaboration2016description}
-and~\cite{gaia2018gaia}. Although, the Gaia DR2 database misreported the chemical
-composition of some OCs, the database has helped to review already known OCs and
-to find new ones.
+The study of OCs advanced thanks to the huge and precise dataset from the Gaia DR2 mission, available since 2018 from~\cite{collaboration2016description} and~\cite{gaia2018gaia}.
+Although, the Gaia DR2 database misreported the chemical composition of some OCs, the database has helped to review already known OCs and to find new ones.
 
-Characterize an OC refers to the statistical determination of the stars that
-form it. Therefore, the characterization is based on the probability of a star
-being a member of the cluster~\cite{sampedro2016caracterizacion}.
+Characterize an OC refers to the statistical determination of the stars that form it.
+Therefore, the characterization is based on the probability of a star being a member of the cluster~\cite{sampedro2016caracterizacion}.
 
-Normally, the characterization process considers astrometric features such as
-proper motion in right ascension and declination, or parallax.
-In some cases~\cite{oliveira2013fitting}, it also considers photometric features,
-which helps to generate the H-R diagram~\cite{hypki2018gaia} of the OC candidate stars.
-Those OC candidates should present a sharp profile corresponding to an isochrone curve
-derived from a theoretical model. The model considers metallicity, mass and brightness
-of the stars involved. For all this purpose, we used a set of tools that require
-supervised and parameterized models. It means that we need a previous knowledge of the
-cluster, otherwise we must repeat the process iteratively to achieve valid results.
+Normally, the characterization process considers astrometric features such as proper motion in right ascension and declination, or parallax.
+In some cases~\cite{oliveira2013fitting}, it also considers photometric features, which helps to generate the H-R diagram~\cite{hypki2018gaia} of the OC candidate stars.
+Those OC candidates should present a sharp profile corresponding to an isochrone curve derived from a theoretical model.
+The model considers metallicity, mass and brightness of the stars involved.
+For all this purpose, we used a set of tools that require supervised and parameterized models.
+It means that we need a previous knowledge of the cluster, otherwise we must repeat the process iteratively to achieve valid results.
 
-In the present paper, we propose a machine learning model capable of characterizing
-OCs inside a stellar region with no previous knowledge about the region.
-Our model takes advantage of some features (such as proper motion in right ascension
-and declination, and parallax) to train an ANN.
+In the present paper, we propose a machine learning model capable of characterizing OCs inside a stellar region with no previous knowledge about the region.
+Our model takes advantage of some features (such as proper motion in right ascension and declination, and parallax) to train an ANN.
 The ANN clusters all the stars within a given region in groups.
@@
 
-Our model contributes with several novelties: i) it is \emph{non-supervised} and
-\emph{non-parameterized}, making easier the automation process of analyzing a
-wide range of regions with different typologies, and ii) it is computationally
-efficient to run in common workstations because it has been developed with
-Python using the Keras framework which takes advantage of modern GPUs to perform
-its computations. That increases significantly the computational capacity of
-our model doing it possible to run on regular workstations.
+Our model contributes with several novelties: i) it is \emph{non-supervised} and \emph{non-parameterized}, making easier the automation process of analyzing a wide range of regions with different typologies, and ii) it is computationally efficient to run in common workstations because it has been developed with Python using the Keras framework which takes advantage of modern GPUs to perform its computations.
+That increases significantly the computational capacity of our model doing it possible to run on regular workstations.
 
-This paper is organized as follows. Section~\ref{sec:related_works} presents
-some related works and Section~\ref{sec:aims_and_method} describes in detail
-our model. The following Section~\ref{sec:results}, shows the results and compares
-them with a method that involves tools from the Virtual Observatory (VO) such as
-Clusterix~\cite{balaguer2020clusterix}. Finally, the last section concludes and
-outlines future research lines.
+This paper is organized as follows.
+Section~\ref{sec:related_works} presents some related works and Section~\ref{sec:aims_and_method} describes in detail our model.
+The following Section~\ref{sec:results}, shows the results and compares them with a method that involves tools from the Virtual Observatory (VO) such as Clusterix~\cite{balaguer2020clusterix}.
+Finally, the last section concludes and outlines future research lines.
 
@@
 In~\cite{castro2020hunting}, they present one of the initial approaches to detect OCs.
-They search for overdensities in the astrometric space of the galactic disk. Then,
-they identify with photometric information possible OCs. Although it seems easy,
-it is quite hard to face the problem that the near field around the OC has two kinds of
-populations. The first population belongs to the members of the OC stars (from tens or
-hundreds to a few thousand). The second population has a background of stars that do
-not belong to the OC (from tens to hundreds of thousands). Thus, we summarize our problem
-in finding out which stars belong to the OC.
+They search for overdensities in the astrometric space of the galactic disk.
+Then, they identify with photometric information possible OCs.
+Although it seems easy, it is quite hard to face the problem that the near field around the OC has two kinds of populations.
+The first population belongs to the members of the OC stars (from tens or hundreds to a few thousand).
+The second population has a background of stars that do not belong to the OC (from tens to hundreds of thousands).
+Thus, we summarize our problem in finding out which stars belong to the OC.
 
-Other works (TOPCAT~\cite{taylor2005topcat}, Clusterix 2.0~\cite{balaguer2020clusterix},
-Aladin~\cite{bonnarel2000aladin}, or VOSA~\cite{bayo2008vosa}), first analyze the proper
-motion configuration space of the region of interest. TOPCAT cannot find an open cluster.
-Instead, it requires parameterizing some of the properties of the cluster. This last
-process requires previous knowledge of the cluster. In other words, they perform a supervised
-and manual selection of groups based on overdensities in the proper motion configuration space.
-Clusterix 2.0 is an interactive web-based tool that can help the choice of groups. Take the
-proper motion diagram without making any previous assumptions about the membership of the
-candidate star. And empirically determine the frequency functions. It employs normal Gaussian
-kernel functions, defined as:
+Other works (TOPCAT~\cite{taylor2005topcat}, Clusterix 2.0~\cite{balaguer2020clusterix}, Aladin~\cite{bonnarel2000aladin}, or VOSA~\cite{bayo2008vosa}), first analyze the proper motion configuration space of the region of interest.
+TOPCAT cannot find an open cluster.
+Instead, it requires parameterizing some of the properties of the cluster.
+This last process requires previous knowledge of the cluster.
+In other words, they perform a supervised and manual selection of groups based on overdensities in the proper motion configuration space.
+Clusterix 2.0 is an interactive web-based tool that can help the choice of groups.
+Take the proper motion diagram without making any previous assumptions about the membership of the candidate star.
+And empirically determine the frequency functions.
+It employs normal Gaussian kernel functions, defined as:
 
@@
 
-where \(( a, b )\) are the proper motion configuration space, \(( i, j )\) is a point
-located in the center of that provides the maximum contribution for calculating the
-local density, and \(h\) is the \emph{smoothing} parameter, which it is measured in the
-same units as the proper motion.
+where \(( a, b )\) are the proper motion configuration space, \(( i, j )\) is a point located in the center of that provides the maximum contribution for calculating the local density, and \(h\) is the \emph{smoothing} parameter, which it is measured in the same units as the proper motion.
 
 Clusterix depends fundamentally on the selection of three radii in the studied region.
-The inner radius contains stars that belong or not to the cluster. While the outermost
-radii define a ring that only contains stars that do not belong to the cluster.
-Furthermore, it is also sensitive to other parameters, such as the soften parameter
-\(h\), or other restrictions related to the searched proper motions.
+The inner radius contains stars that belong or not to the cluster.
+While the outermost radii define a ring that only contains stars that do not belong to the cluster.
+Furthermore, it is also sensitive to other parameters, such as the soften parameter \(h\), or other restrictions related to the searched proper motions.
 
-If the analysis performed by Clusterix is successful, it returns the probability that
-each star belongs to the open cluster. The results can then be imported into TOPCAT
-to continue the refinement process and obtain a valid characterization.
+If the analysis performed by Clusterix is successful, it returns the probability that each star belongs to the open cluster.
+The results can then be imported into TOPCAT to continue the refinement process and obtain a valid characterization.
 
-Another recent approach~\cite{castro2020hunting} solves the problem using machine
-learning techniques. It searches systematically for overdensities in the astrometric
-space of the galactic disk and subsequent identification of open clusters using
-photometric information. It includes two phases:
+Another recent approach~\cite{castro2020hunting} solves the problem using machine learning techniques.
+It searches systematically for overdensities in the astrometric space of the galactic disk and subsequent identification of open clusters using photometric information.
+It includes two phases:
 
 \begin{enumerate}
-  \item it employs DBSCAN, an unsupervised clustering algorithm, to search for
-        overdensities,
-  \item and applies a deep learning ANN to identify isochrone patterns within
-        the detected overdensities and thus proceed to confirm them as OCs. They
-        previously trained the ANN with magnitude diagrams.
+  \item it employs DBSCAN, an unsupervised clustering algorithm, to search for overdensities,
+  \item and applies a deep learning ANN to identify isochrone patterns within the detected overdensities and thus proceed to confirm them as OCs.
+They previously trained the ANN with magnitude diagrams.
 \end{enumerate}
@@
 They carry out their experiments at the Barcelona Supercomputing Center, MareNostrum 42.
-Thereby, the neural network handles the image recognition process with isochrone
-patterns without applying theoretical models derived from values such as metallicity
-or masses, among others. Their work concludes with the identification of 582 new open
-clusters distributed along with the galactic disk for a galactic declination below 20
-degrees, increasing the number of known open clusters by 45\%.
+Thereby, the neural network handles the image recognition process with isochrone patterns without applying theoretical models derived from values such as metallicity or masses, among others.
+Their work concludes with the identification of 582 new open clusters distributed along with the galactic disk for a galactic declination below 20 degrees, increasing the number of known open clusters by 45\%.
 
-In contrast to~\cite{castro2020hunting}, we lack a supercomputer like MareNostrum
-for our experiments. For this reason, our goal is to obtain a new novel method that
-allows the characterization of open clusters without doing a blind search for
-characterization clusters. We advocate the idea of building an unsupervised and
-non-parameterized method. Therefore, it can be suitable for automated processes.
+In contrast to~\cite{castro2020hunting}, we lack a supercomputer like MareNostrum for our experiments.
+For this reason, our goal is to obtain a new novel method that allows the characterization of open clusters without doing a blind search for characterization clusters.
+We advocate the idea of building an unsupervised and non-parameterized method.
+Therefore, it can be suitable for automated processes.
 
@@
 
-In the literature, we have several clustering algorithms, such as K-Means, Mean-Shift
-Clustering, DBSCAN~\cite{ester1996density} among others, that can help us achieve
-our aim. Each one behaves better according to the distribution of the objects to
-be grouped. However, we need to establish a large number of clusters to find the open cluster
-we are looking for. It complicates the identification of the OC because most of the
-time falls in many outliers.
+In the literature, we have several clustering algorithms, such as K-Means, Mean-Shift Clustering, DBSCAN~\cite{ester1996density} among others, that can help us achieve our aim.
+Each one behaves better according to the distribution of the objects to be grouped.
+However, we need to establish a large number of clusters to find the open cluster we are looking for.
+It complicates the identification of the OC because most of the time falls in many outliers.
 
-The unsupervised \emph{Deep Embedding for Clustering Analysis} (\emph{DEC})
-model~\cite{xie2016unsupervised} is a refinement K-Means based on ANN. It starts with
-K-Means. And then, it trains an autoencoder to reduce the feature space and pass them
-through a Clustering Layer which refines the previous selection.
+The unsupervised \emph{Deep Embedding for Clustering Analysis} (\emph{DEC}) model~\cite{xie2016unsupervised} is a refinement K-Means based on ANN.
+It starts with K-Means.
+And then, it trains an autoencoder to reduce the feature space and pass them through a Clustering Layer which refines the previous selection.
 
-With all this in mind, our method is based on the DEC model. We adapted it to manage data
-groups based on the dynamic properties of the stars. Thus, we have an unsupervised
-clustering model for open cluster characterization. The model fits a wide range of
-clusters without the need for fine-tuning a high number of hyperparameters.
+With all this in mind, our method is based on the DEC model.
+We adapted it to manage data groups based on the dynamic properties of the stars.
+Thus, we have an unsupervised clustering model for open cluster characterization.
+The model fits a wide range of clusters without the need for fine-tuning a high number of hyperparameters.
 
@@
 
-In this work, we make use of Gaia DR2 as DR3 has not been released in time for us to
-include it. Gaia DR2 is a multidimensional dataset obtained by ESA's Gaia mission
-(located at L2, 1.5 million kilometers from Earth) and operational since 2014. The
-catalog has high precision and accuracy astrometric data for more than 1.7 billion
-stellar sources and magnitudes in three photometric filters (G, BP, and RP) for more
-than 1,300 million sources.
+In this work, we make use of Gaia DR2 as DR3 has not been released in time for us to include it.
+Gaia DR2 is a multidimensional dataset obtained by ESA's Gaia mission (located at L2, 1.5 million kilometers from Earth) and operational since 2014.
+The catalog has high precision and accuracy astrometric data for more than 1.7 billion stellar sources and magnitudes in three photometric filters (G, BP, and RP) for more than 1,300 million sources.
 
-We start by selecting a region from the OpenClust~\cite{dias2002new} catalog and
-downloading it from the Gaia DR2 database. The radius of the downloaded region from Gaia
-is 1.5 times greater than that recorded in OpenClust, ensuring that it includes several
-stars that do not belong to the open cluster.
+We start by selecting a region from the OpenClust~\cite{dias2002new} catalog and downloading it from the Gaia DR2 database.
+The radius of the downloaded region from Gaia is 1.5 times greater than that recorded in OpenClust, ensuring that it includes several stars that do not belong to the open cluster.
 
@@
 
-Proper motion in right ascension (column pmra) and declination (column pmdec) seems
-like a natural choice since stars belonging to the same OC share a common motion vector.
-Parallax (column parallax) is another relevant feature. It lets us know how far stars
-are from us. Besides, since all stars within an open cluster were born from the same
-dust cloud, they must all have similar parallax.
+Proper motion in right ascension (column pmra) and declination (column pmdec) seems like a natural choice since stars belonging to the same OC share a common motion vector.
+Parallax (column parallax) is another relevant feature.
+It lets us know how far stars are from us.
+Besides, since all stars within an open cluster were born from the same dust cloud, they must all have similar parallax.
 
-However, we are not going to use these raw features. Instead, we take a combination
-of them. Figure~\ref{fig:features_melotte_22} shows a pairwise relationship of our
-combination of features from Melotte 22.
+However, we are not going to use these raw features.
+Instead, we take a combination of them.
+Figure~\ref{fig:features_melotte_22} shows a pairwise relationship of our combination of features from Melotte 22.
 
@@
 
-First, we correct proper motion in right ascension and declination by dividing them by
-the parallax (variables pmra\_corr and pmdec\_corr, respectively). That way, we normalize
-these quantities and help our clustering models to improve their performance. The modulus
-of the proper motion (variable pmmod in Figure~\ref{fig:features_melotte_22}) is another
-computed property that we considered. We use it to relate both features and therefore
-force our model to keep them tight.
+First, we correct proper motion in right ascension and declination by dividing them by the parallax (variables pmra\_corr and pmdec\_corr, respectively).
+That way, we normalize these quantities and help our clustering models to improve their performance.
+The modulus of the proper motion (variable pmmod in Figure~\ref{fig:features_melotte_22}) is another computed property that we considered.
+We use it to relate both features and therefore force our model to keep them tight.
 
@@
 
-One of the main problems to define our model is that we do not have a labeled
-dataset to train a supervised model. To solve it, we employ an unsupervised self-trained
-model. Thus, we adapted the Unsupervised \emph{DEC} model~\cite{xie2016unsupervised}
-to our requirements.
+One of the main problems to define our model is that we do not have a labeled dataset to train a supervised model.
+To solve it, we employ an unsupervised self-trained model.
+Thus, we adapted the Unsupervised \emph{DEC} model~\cite{xie2016unsupervised} to our requirements.
 
@@
 \begin{itemize}
-  \item \textbf{deep autoencoder:} it is a neural network that learns to copy its input
-        to its output. It is composed of encoder layers followed by decoder layers.
-        We trained it before passing the data through the next component, the
-        \emph{clustering layer}. Recent research~\cite{vincent2010stacked,hinton2006reducing}
-        showed that the autoencoder provides meaningful and well-separated representations
-        on real-world datasets.
-  \item \textbf{clustering layer:} it receives data transformed by the autoencoder
-        and iteratively is trained until a convergence criterion is met.
+  \item \textbf{deep autoencoder:} it is a neural network that learns to copy its input to its output.
+It is composed of encoder layers followed by decoder layers.
+We trained it before passing the data through the next component, the \emph{clustering layer}.
+Recent research~\cite{vincent2010stacked,hinton2006reducing} showed that the autoencoder provides meaningful and well-separated representations on real-world datasets.
+  \item \textbf{clustering layer:} it receives data transformed by the autoencoder and iteratively is trained until a convergence criterion is met.
 \end{itemize}
 
-The deep autoencoder is used to transform the input data into a latent space using a
-non-linear mapping function \(f_{\theta} : X \rightarrow Z\).
+The deep autoencoder is used to transform the input data into a latent space using a non-linear mapping function \(f_{\theta} : X \rightarrow Z\).
 
 We trained the autoencoder before fitting the model to generate predictions.
-Then, we use the encoder layers of the autoencoder with the aim of transforming
-input data to the latent space \(Z\). Once the data has been transformed, a K-Means
-clusterer is used to make an initial clustering. K-Means cluster centers are used
-as the initial weights for the clustering layer.
+Then, we use the encoder layers of the autoencoder with the aim of transforming input data to the latent space \(Z\).
+Once the data has been transformed, a K-Means clusterer is used to make an initial clustering.
+K-Means cluster centers are used as the initial weights for the clustering layer.
 
-With that initial configuration, the model iterates alternating between computing an
-auxiliary target distribution (soft assignment) and minimizing the Kullback-Leiber (KL)
-divergence~\cite{kullback1951information} to it. This unsupervised algorithm allows us
-to improve the clustering.
+With that initial configuration, the model iterates alternating between computing an auxiliary target distribution (soft assignment) and minimizing the Kullback-Leiber (KL) divergence~\cite{kullback1951information} to it.
+This unsupervised algorithm allows us to improve the clustering.
 
-As it was described in Section~\ref{sec:feature_selection}, the number of features we
-deal is not too large. However, this latent space helps us to start with a reduced number
-of features and avoids the \emph{``curse of dimensionality''}~\cite{bellman1961curse}.
+As it was described in Section~\ref{sec:feature_selection}, the number of features we deal is not too large.
+However, this latent space helps us to start with a reduced number of features and avoids the \emph{``curse of dimensionality''}~\cite{bellman1961curse}.
 
-In the soft assignment stage, we used \emph{Student's t-distribution} as a kernel to
-measure the similarity between the embedded points and the cluster centroid.
-While in the KL divergence minimization the algorithm iteratively refines clusters by learning
-from their high confidence assignments with the help of an auxiliary target distribution.
+In the soft assignment stage, we used \emph{Student's t-distribution} as a kernel to measure the similarity between the embedded points and the cluster centroid.
+While in the KL divergence minimization the algorithm iteratively refines clusters by learning from their high confidence assignments with the help of an auxiliary target distribution.
 The model is trained by matching the soft assignment to the target distribution.
@@
 
-In this work, we took the target distribution~\cite{xie2016unsupervised}
-defined in Equation~\ref{eq:student_tdistribution}.
+In this work, we took the target distribution~\cite{xie2016unsupervised} defined in Equation~\ref{eq:student_tdistribution}.
 
@@
   \item put more emphasis on data points assigned with high confidence,
-  \item normalize loss contribution of each centroid to prevent large clusters
-        from distorting the hidden feature space.
+  \item normalize loss contribution of each centroid to prevent large clusters from distorting the hidden feature space.
 \end{itemize}
@@
 Figure~\ref{fig:dec_model_setup} shows the layer setup of our model.
-It is simpler than the used in~\cite{xie2016unsupervised} because we selected a smaller
-number of features. Therefore, using the same configuration would result in a model so
-powerful that it would incur overfitting issues that cannot make the correct predictions.
+It is simpler than the used in~\cite{xie2016unsupervised} because we selected a smaller number of features.
+Therefore, using the same configuration would result in a model so powerful that it would incur overfitting issues that cannot make the correct predictions.
 
@@
 
-Finally, we can refine this selection by filtering those stars that are below and above
-the 0.10 and 0.90 quantiles for each group, respectively. That way we remove the most
-doubtful values from the selection.
+Finally, we can refine this selection by filtering those stars that are below and above the 0.10 and 0.90 quantiles for each group, respectively.
+That way we remove the most doubtful values from the selection.
 
@@
 
-Since we are looking for a single cluster, it seems reasonable to use a clustering algorithm
-like K-Means to find two clusters. One for the desired OC and another for stars that do not
-belong to the OC. However, the idea is not completely right. This is due to the fact that OC's
-stars are surrounded by other stars with possibly similar properties. Therefore, setting the
-number of clusters to two is too low to separate them properly.
+Since we are looking for a single cluster, it seems reasonable to use a clustering algorithm like K-Means to find two clusters.
+One for the desired OC and another for stars that do not belong to the OC.
+However, the idea is not completely right.
+This is due to the fact that OC's stars are surrounded by other stars with possibly similar properties.
+Therefore, setting the number of clusters to two is too low to separate them properly.
 
-Figure~\ref{fig:kmeans_comparisons_melotte_22} shows an example. It describes the results of
-applying different configurations of K-Means in Melotte 22, N=2, N=5 and N=8. For each configuration
-of K-Means, we show the proper motion in right ascension and declination (figures in the top position),
-and the parallax distribution (figures in the down position).
+Figure~\ref{fig:kmeans_comparisons_melotte_22} shows an example.
+It describes the results of applying different configurations of K-Means in Melotte 22, N=2, N=5 and N=8.
+For each configuration of K-Means, we show the proper motion in right ascension and declination (figures in the top position), and the parallax distribution (figures in the down position).
 
-As we can see in the parallax distribution, lower values of the number of clusters (e.g., N=2) assign
-many stars to the group \emph{g0} (purple) that should belong to group \emph{g1}. These badly assigned stars
-are the ones that have the parallax far from the 7.3mas resonance. In contrast, large values of the
-number of clusters allow us to isolate more accurately the resonance in parallax at \(\approx 7.3mas\).
-For instance, at N=8, more groups are better compartmentalized. Group \emph{g2} would be the one corresponding
-to the OC that we are looking for, and all the other groups are those that we would like to be a single
-group. That is the stars that do not belong to the OC. The disadvantage is that K-Means with N=8 forms more
-groups of stars, which complicates the task of finding the desired OC, since we would like to get just two
-groups. Therefore, with a K-Means we have to find a way to set the right value for the number of clusters
-to isolate the searched cluster without creating too many groups. To solve this issue, we can try to
-estimate the best number of clusters by using the \emph{silhouette score}~\cite{rousseeuw1987silhouettes}.
+As we can see in the parallax distribution, lower values of the number of clusters (e.g., N=2) assign many stars to the group \emph{g0} (purple) that should belong to group \emph{g1}.
+These badly assigned stars are the ones that have the parallax far from the 7.3mas resonance.
+In contrast, large values of the number of clusters allow us to isolate more accurately the resonance in parallax at \(\approx 7.3mas\).
+For instance, at N=8, more groups are better compartmentalized.
+Group \emph{g2} would be the one corresponding to the OC that we are looking for, and all the other groups are those that we would like to be a single group.
+That is the stars that do not belong to the OC.
+The disadvantage is that K-Means with N=8 forms more groups of stars, which complicates the task of finding the desired OC, since we would like to get just two groups.
+Therefore, with a K-Means we have to find a way to set the right value for the number of clusters to isolate the searched cluster without creating too many groups.
+To solve this issue, we can try to estimate the best number of clusters by using the \emph{silhouette score}~\cite{rousseeuw1987silhouettes}.
 
-K-Means does a good job making an initial clustering. However, too many clusters arise
-from this characterization and the OC is still polluted with stars that do not belong to it.
+K-Means does a good job making an initial clustering.
+However, too many clusters arise from this characterization and the OC is still polluted with stars that do not belong to it.
 Moreover, we would like to reduce the amount of clusters too.
@@
 \begin{enumerate}
-  \item with the Melotte 25 dataset, which is one of the most difficult dataset to characterize OC
-        by its distribution and extension, because it is located in a region with 400,000 stars.
+  \item with the Melotte 25 dataset, which is one of the most difficult dataset to characterize OC by its distribution and extension, because it is located in a region with 400,000 stars.
   \item with a variety of clusters of different typologies.
@@
 
-In order to support the discussion presented in Section~\ref{sec:deep_open_clustering}, in the both
-tests we compared our model against a K-Means classification algorithm. In addition, we perform a
-second execution only in our model where we filtered the quantiles in the Melotte 25 dataset.
+In order to support the discussion presented in Section~\ref{sec:deep_open_clustering}, in the both tests we compared our model against a K-Means classification algorithm.
+In addition, we perform a second execution only in our model where we filtered the quantiles in the Melotte 25 dataset.
 
-Our DECOCC model was implemented in Python 3.8 using the Keras 2.2 framework and Jupyter
-Notebooks~\cite{Kluyver2016jupyter}. All tests were run on an Apple Mac Pro Late 2013 with a
-2.7GHz 12-Core Intel Xeon E5-2697v2, and 64GB RAM 1866MHz DDR3. For the GPU, we use a graphic
-card AMD FirePro D700 with 6GB~\footnote{All resources developed for this project are available
-at \href{https://github.com/cdalvaro/machine-learning-master-thesis}
-{https://github.com/cdalvaro/machine-learning-master-thesis}}.
+Our DECOCC model was implemented in Python 3.8 using the Keras 2.2 framework and Jupyter Notebooks~\cite{Kluyver2016jupyter}.
+All tests were run on an Apple Mac Pro Late 2013 with a 2.7GHz 12-Core Intel Xeon E5-2697v2, and 64GB RAM 1866MHz DDR3.
+For the GPU, we use a graphic card AMD FirePro D700 with 6GB~\footnote{All resources developed for this project are available at \href{https://github.com/cdalvaro/machine-learning-master-thesis}{https://github.com/cdalvaro/machine-learning-master-thesis}}.
 
@@
 
-Figure~\ref{fig:melotte_25_characterization_evolution} shows the evolution of the characterization
-of Melotte 25. The first column corresponds to the characterization obtained by applying only
-K-Means (with N=10). The middle column shows the result of our model. And the last column describes
-the characterization made by our model but filtering the quantiles lower than 0.10 and higher than
-0.90. The top row shows the proper motion configuration space, while the bottom one shows the
-Hertzsprung–Russell diagrams.
+Figure~\ref{fig:melotte_25_characterization_evolution} shows the evolution of the characterization of Melotte 25.
+The first column corresponds to the characterization obtained by applying only K-Means (with N=10).
+The middle column shows the result of our model.
+And the last column describes the characterization made by our model but filtering the quantiles lower than 0.10 and higher than 0.90.
+The top row shows the proper motion configuration space, while the bottom one shows the Hertzsprung–Russell diagrams.
 
@@
 
-The clustering made by K-Means arises 9 groups, but most of them contain stars that do not belong
-to the Melotte 25 OC. However, in our model, the stars regroup and although we have the same 9
-groups of K-Means, almost all of the stars are assigned to a group that would contain the stars
-that would not belong to the OC. The remaining groups contain either the OC we are looking for or
-other possible new OCs inside the studied region. Moreover, if we exclude those stars that remain
-below 0.10 and above 0.90 percentiles, and execute again this refinement in our model (third column
-in Figure~\ref{fig:melotte_25_characterization_evolution}), we observe an improvement in the
-results of our DECOCC model.
+The clustering made by K-Means arises 9 groups, but most of them contain stars that do not belong to the Melotte 25 OC.
+However, in our model, the stars regroup and although we have the same 9 groups of K-Means, almost all of the stars are assigned to a group that would contain the stars that would not belong to the OC.
+The remaining groups contain either the OC we are looking for or other possible new OCs inside the studied region.
+Moreover, if we exclude those stars that remain below 0.10 and above 0.90 percentiles, and execute again this refinement in our model (third column in Figure~\ref{fig:melotte_25_characterization_evolution}), we observe an improvement in the results of our DECOCC model.
 
@@
 
-The dataset Melotte 25 is located at 66.725 degrees in right ascension and 15.867 degrees in
-declination with a radius of 165 arcmin. As it was mentioned before, this region contains more than
-400,000 stars making it difficult to group stars of the same region. Our model labels Melotte 25 as
-group \emph{g5}. Figure~\ref{fig:result_melotte_25_dec} shows the parallax histogram.
+The dataset Melotte 25 is located at 66.725 degrees in right ascension and 15.867 degrees in declination with a radius of 165 arcmin.
+As it was mentioned before, this region contains more than 400,000 stars making it difficult to group stars of the same region.
+Our model labels Melotte 25 as group \emph{g5}.
+Figure~\ref{fig:result_melotte_25_dec} shows the parallax histogram.
 
@@
 
-Table~\ref{tab:results_melotte_25} summarizes the results for the Melotte 25 dataset. The first column
-shows the name of the method. Second and third columns describe the proper motion in right ascension and
-declination, respectively. The fourth column shows the parallax. And the last column presents the number
-of stars corresponding to the OC. Columns from second to fourth have their respective deviations computed
-as the standard error of the predicted group.
+Table~\ref{tab:results_melotte_25} summarizes the results for the Melotte 25 dataset.
+The first column shows the name of the method.
+Second and third columns describe the proper motion in right ascension and declination, respectively.
+The fourth column shows the parallax.
+And the last column presents the number of stars corresponding to the OC.
+Columns from second to fourth have their respective deviations computed as the standard error of the predicted group.
 
-By comparing the results obtained with our DECOCC model, we can see that they are in range with those
-extracted from Simbad astronomical database~\cite{wenger2000simbad} (first row in
-Table~\ref{tab:results_melotte_25}). While the results provided by K-Means model are far from the official
-data. This comparison demonstrates that our model improves the results obtained with a raw K-Means.
+By comparing the results obtained with our DECOCC model, we can see that they are in range with those extracted from Simbad astronomical database~\cite{wenger2000simbad} (first row in Table~\ref{tab:results_melotte_25}).
+While the results provided by K-Means model are far from the official data.
+This comparison demonstrates that our model improves the results obtained with a raw K-Means.
 
@@
 
-Table~\ref{tab:hyperparameters_melotte_25} shows the hyperparameters used for the characterization. We used
-them to fine-tuning our model and improve its accuracy. Although they depend on the studied region, none of
-them are related with physical properties of the cluster, which makes it easy to automate the selection of
-hyperparameters.
+Table~\ref{tab:hyperparameters_melotte_25} shows the hyperparameters used for the characterization.
+We used them to fine-tuning our model and improve its accuracy.
+Although they depend on the studied region, none of them are related with physical properties of the cluster, which makes it easy to automate the selection of hyperparameters.
 
@@
 
-We perform now more exhaustive tests of our DECOCC model using open clusters with different typologies (e.g.,
-large, small, more extensive, or more concentrated within the region to be analyzed):
+We perform now more exhaustive tests of our DECOCC model using open clusters with different typologies (e.g., large, small, more extensive, or more concentrated within the region to be analyzed):
 
 \begin{itemize}
-  \item NGC 2516, which has the proper motion center deviated from the origin but embedded within a large cloud
-        of stars with similar proper motions.
-  \item NGC 2632, which, in addition to Melotte 22 (explained below), is a cluster whose proper motion center is
-        not at the origin and has a well-separated parallax center.
-  \item NGC 2682, which has the parallax centered within the region's Gaussian, making it difficult its detection.
-        Moreover, same as NGC 2516, the proper motion center is deviated from the origin.
-  \item Melotte 22, as well as Melotte 25, is a cluster closer to our galaxy. That causes its membership stars to
-        be more scattered than previous clusters which are more compact.
+  \item NGC 2516, which has the proper motion center deviated from the origin but embedded within a large cloud of stars with similar proper motions.
+  \item NGC 2632, which, in addition to Melotte 22 (explained below), is a cluster whose proper motion center is not at the origin and has a well-separated parallax center.
+  \item NGC 2682, which has the parallax centered within the region's Gaussian, making it difficult its detection. Moreover, same as NGC 2516, the proper motion center is deviated from the origin.
+  \item Melotte 22, as well as Melotte 25, is a cluster closer to our galaxy. That causes its membership stars to be more scattered than previous clusters which are more compact.
 \end{itemize}
 
-Table~\ref{tab:clusters_summary} describes more specific information about the previous clusters (we also include
-information about the Melotte 25 open cluster, which was previously analyzed). The first column shows the name of
-the cluster. The remaining columns are the right ascension, declination, radius, and the number of stars of studied
-clusters. The last column (\# stars) corresponds to those stars contained within a cone of center \((\alpha, \delta)\)
-and radius the cluster's radius multiplied by a factor of 1.5.
+Table~\ref{tab:clusters_summary} describes more specific information about the previous clusters (we also include information about the Melotte 25 open cluster, which was previously analyzed).
+The first column shows the name of the cluster.
+The remaining columns are the right ascension, declination, radius, and the number of stars of studied clusters.
+The last column (\# stars) corresponds to those stars contained within a cone of center \((\alpha, \delta)\) and radius the cluster's radius multiplied by a factor of 1.5.
 
@@
 
-As we can see, all those clusters have a wide variety of diameters, from 25 to 330 arcmin, as well as the number
-of stars that belong to them. For instance, NGC 2682 has 3,000 stars while Melotte 25 is inside a region with more
-than 400,000 stars.
+As we can see, all those clusters have a wide variety of diameters, from 25 to 330 arcmin, as well as the number of stars that belong to them.
+For instance, NGC 2682 has 3,000 stars while Melotte 25 is inside a region with more than 400,000 stars.
 
 Table~\ref{tab:app_results_ngc_2516} describes a summary of the results for the clusters with different typologies.
-Again, our DECOCC model presents similar results from those extracted from the astronomical database Simbad. Except for
-the open cluster NGC 2632. This is so because the center of the proper motion is not at the origin and generates a more
-separate center of parallax, which makes it difficult to characterize the open cluster. Compared to K-Means, our model
-presents more robust results in more difficult open clusters, such as the NGC 2682 that has the parallax centered within
-the region's Gaussian.
+Again, our DECOCC model presents similar results from those extracted from the astronomical database Simbad.
+Except for the open cluster NGC 2632.
+This is so because the center of the proper motion is not at the origin and generates a more separate center of parallax, which makes it difficult to characterize the open cluster.
+Compared to K-Means, our model presents more robust results in more difficult open clusters, such as the NGC 2682 that has the parallax centered within the region's Gaussian.
 
@@
 
-Figures~\ref{fig:ngc_2516_ngc_2632_characterization_evolution} (describes NGC 2516 and NGC 2632)
-and~\ref{fig:melotte_22_ngc_2682_characterization_evolution} (describes NGC 2682 and Melotte 22) show the
-groups found for the three models. For each open cluster, we describe the parallax and hr diagram.
+Figures~\ref{fig:ngc_2516_ngc_2632_characterization_evolution} (describes NGC 2516 and NGC 2632) and~\ref{fig:melotte_22_ngc_2682_characterization_evolution} (describes NGC 2682 and Melotte 22) show the groups found for the three models.
+For each open cluster, we describe the parallax and hr diagram.
 
@@
 
-As we can observe, labels between K-Means and DECOCC do not match (except for NGC 2632), since we tested these two
-models independently. Thus, each model can label each group differently from the other. Moreover, running the same model
-twice can label the resulting groups differently between the first and the second run (as long as the same seed is not
-assigned). However, regardless whether the labels are different, the stars within the generated clusters should be very
-similar.
+As we can observe, labels between K-Means and DECOCC do not match (except for NGC 2632), since we tested these two models independently.
+Thus, each model can label each group differently from the other.
+Moreover, running the same model twice can label the resulting groups differently between the first and the second run (as long as the same seed is not assigned).
+However, regardless whether the labels are different, the stars within the generated clusters should be very similar.
 
-In all cases, the DECOCC model resolved properly the identity of the cluster and characterized the membership stars,
-giving compatible results with the ones obtained with classic procedures and VO tools~\cite{valdivielso2009iphas}.
+In all cases, the DECOCC model resolved properly the identity of the cluster and characterized the membership stars, giving compatible results with the ones obtained with classic procedures and VO tools~\cite{valdivielso2009iphas}.
 
-Table~\ref{tab:app_hyperparameters_ngc_2516} describes the hyperparameters used for characterizing the open clusters
-with our method. Compared with the VO tools, we categorize our model as a non-parameterized model since it does not
-depend on parameters referred to as the cluster itself but only on hyperparameters such as the initial number of
-clusters to be found, or the structure of the Clustering Layer.
+Table~\ref{tab:app_hyperparameters_ngc_2516} describes the hyperparameters used for characterizing the open clusters with our method.
+Compared with the VO tools, we categorize our model as a non-parameterized model since it does not depend on parameters referred to as the cluster itself but only on hyperparameters such as the initial number of clusters to be found, or the structure of the Clustering Layer.
 
-The proposed method is also non-supervised since we do not need to tell the model which stars belong to the cluster or
-assist it while its training stage.
+The proposed method is also non-supervised since we do not need to tell the model which stars belong to the cluster or assist it while its training stage.
 
-Only at the end of the characterization, a fine-tuning driven by the user is applied to improve the selection by
-discarding those stars which fall outside the given quantiles. Nevertheless, we do not conceive assumptions about the
-cluster profile, something that we consider in VO tools. This is another evidence that our model is non-parameterized.
+Only at the end of the characterization, a fine-tuning driven by the user is applied to improve the selection by discarding those stars which fall outside the given quantiles.
+Nevertheless, we do not conceive assumptions about the cluster profile, something that we consider in VO tools.
+This is another evidence that our model is non-parameterized.
 
@@
 
-In summary, we can say that our results improve K-Means. We based this judgment by comparing right ascension,
-declination, parallax, and the number of stars of the cluster found by the two models with the Simbad database. In other
-words, we statistically analyze the results not individualized by stars belonging to the cluster found. We arrange it
-with Simbad because as far as we are concerned, there is no Gaia dataset with classified open clusters.
+In summary, we can say that our results improve K-Means.
+We based this judgment by comparing right ascension, declination, parallax, and the number of stars of the cluster found by the two models with the Simbad database.
+In other words, we statistically analyze the results not individualized by stars belonging to the cluster found.
+We arrange it with Simbad because as far as we are concerned, there is no Gaia dataset with classified open clusters.
 
@@
 
-This paper presents a model to characterize open clusters, comprising an Artificial Neural Network, which is neither
-parameterized nor supervised. The model does not require complex hardware. Instead, it can be run on workstations with a
-common GPU, making it accessible for deployment in many research centers.
+This paper presents a model to characterize open clusters, comprising an Artificial Neural Network, which is neither parameterized nor supervised.
+The model does not require complex hardware.
+Instead, it can be run on workstations with a common GPU, making it accessible for deployment in many research centers.
 
-The model works well in identifying and characterizing open clusters. However, some improvements must be made to improve
-its accuracy and precision.
+The model works well in identifying and characterizing open clusters. However, some improvements must be made to improve its accuracy and precision.
 
-The number of hyperparameters in our model is small and we do not need previous knowledge about the cluster studied,
-which makes it easier to test different settings and automate the process.
+The number of hyperparameters in our model is small and we do not need previous knowledge about the cluster studied, which makes it easier to test different settings and automate the process.
 
-Building a non-parameterized model is a difficult and sometimes and unfeasible task due to every machine learning model
-requires initial definitions of the hyperparameters to work properly. However, in our model, the user is not required to
-provide any information about the cluster profile or its properties. Making it a non-parameterized model. Nevertheless,
-some tweaks can be done regarding the cluster and the studied region, like changing the initial number of clusters or
-modifying the clustering layer structure to improve the final characterization.
+Building a non-parameterized model is a difficult and sometimes and unfeasible task due to every machine learning model requires initial definitions of the hyperparameters to work properly.
+However, in our model, the user is not required to provide any information about the cluster profile or its properties.
+Making it a non-parameterized model.
+Nevertheless, some tweaks can be done regarding the cluster and the studied region, like changing the initial number of clusters or modifying the clustering layer structure to improve the final characterization.
 
-The initializer kernel prepares data before passing them to the ANN. This is one of the reasons because the results may
-vary significantly. In future works, we want to avoid it to have a reliable model which does not depend on the dataset
-order.
+The initializer kernel prepares data before passing them to the ANN.
+This is one of the reasons because the results may vary significantly.
+In future works, we want to avoid it to have a reliable model which does not depend on the dataset order.
 
-In future, we also want to test our model with a wider range of clusters. In general, we could think of applying it to
-the entire VizieR catalog. So, we could get a better idea about the current limitations of our model and improve it. For
-instance, we could determine with higher accuracy the typologies that our model works well with. Furthermore, we could
-establish different sets of hyperparameters regarding the typology of the cluster.
+In future, we also want to test our model with a wider range of clusters.
+In general, we could think of applying it to the entire VizieR catalog.
+So, we could get a better idea about the current limitations of our model and improve it.
+For instance, we could determine with higher accuracy the typologies that our model works well with.
+Furthermore, we could establish different sets of hyperparameters regarding the typology of the cluster.
 
-An important aspect we noticed is that with our model new uncatalogued clusters arise from the data apart of the main
-ones. Making it of special interest to study these individual new clusters.
+An important aspect we noticed is that with our model new uncatalogued clusters arise from the data apart of the main ones. Making it of special interest to study these individual new clusters.
 
-Another possible improvements are either to use DBSCAN instead of K-Means as our initial clustering algorithm, and
-employ DR3 dataset instead of DR2 dataset of Gaia.
+Another possible improvements are either to use DBSCAN instead of K-Means as our initial clustering algorithm, and employ DR3 dataset instead of DR2 dataset of Gaia.
 
-Furthermore, our model can deal with large and open regions. As an example, it has succeeded in characterizing Melotte
-25 (Hiades) OC. Also, the relative proximity of this cluster, 45 parsecs, makes it difficult to characterize. Our method
-proposes at least two more non-catalogued clusters. The analysis of these new clusters brings out interesting stellar
-dynamics in the studied region. It could mean that there was a recent interaction among different molecular clouds.
+Furthermore, our model can deal with large and open regions.
+As an example, it has succeeded in characterizing Melotte 25 (Hiades) OC.
+Also, the relative proximity of this cluster, 45 parsecs, makes it difficult to characterize.
+Our method proposes at least two more non-catalogued clusters.
+The analysis of these new clusters brings out interesting stellar dynamics in the studied region.
+It could mean that there was a recent interaction among different molecular clouds.
 

### 2021-04-07 paper/paper.tex

@@
-\documentclass[11pt,a4paper,english,twocolumn]{article}
-\usepackage{unir_paper}
+\documentclass[preprint,12pt,authoryear]{elsarticle}
+
 \usepackage{subcaption}
 \usepackage{tablefootnote, multirow}
+\usepackage{amssymb}
+\usepackage{amsmath}
+\usepackage{hyperref}
 
-%---------------------------
-%título del trabajo y autor
-%---------------------------
+\journal{Astronomy and Computing}
+
+\begin{document}
+\begin{frontmatter}
+
+%% Title, authors and addresses
+
+%% use the tnoteref command within \title for footnotes;
+%% use the tnotetext command for theassociated footnote;
+%% use the fnref command within \author or \address for footnotes;
+%% use the fntext command for theassociated footnote;
+%% use the corref command within \author for corresponding author footnotes;
+%% use the cortext command for theassociated footnote;
+%% use the ead command for the email address,
+%% and the form \ead[url] for the home page:
+%% \title{Title\tnoteref{label1}}
+%% \tnotetext[label1]{}
+%% \author{Name\corref{cor1}\fnref{label2}}
+%% \ead{email address}
+%% \ead[url]{home page}
+%% \fntext[label2]{}
+%% \cortext[cor1]{}
+%% \affiliation{organization={},
+%%             addressline={},
+%%             city={},
+%%             postcode={},
+%%             state={},
+%%             country={}}
+%% \fntext[label3]{}
+
 \title{Deep Embedded Clustering for Open Cluster Characterization with Gaia DR2 Data}
-\author{CD. Álvaro, C. Guzman, J. Álvaro}
-\date{24th of January, 2021}
 
-\resumen{
+%\author[MET]{CD. Álvaro}
+
+%\affiliation[MET]{organization={Meteologica},
+             %addressline={Calle Costa Brava, 10},
+             %city={Madrid},
+             %postcode={28034},
+             %country={Spain}}
+
+%\author[UNAD,IAC]{C. Guzman}
+
+%% use optional labels to link authors explicitly to addresses:
+%% \author[label1,label2]{}
+%\affiliation[IAC]{organization={Instituto de Astrofisíca de Canarias},
+%             addressline={Calle Vía Láctea, s/n},
+%             city={ San Cristóbal de La Laguna},
+%             postcode={ 38205},
+%             country={Spain}}
+%\affiliation[UNAD]{organization={Universidad Nacional Abierta y a Distancia},
+%             addressline={Sede JCM Calle 14 Sur \# 14-23},
+%             city={Bogotá},
+%             country={Colombia}}
+
+%\author[NIE,SEA,FAAE]{J. Álvaro}
+
+%\affiliation[SEA]{organization={SEA Astronomia},
+%             country={Spain}}
+
+%\affiliation[NIE]{organization={Nodo Ibérico de Europlanet, Spain \& Portugal},
+%             country={Spain / Portugal}}
+
+%\affiliation[FAAE]{organization={Federación de Asociaciones Astronómicas de España},
+%             addressline={C/ Serrano, n. 117},
+%             city={Madrid},
+%             postcode={28006},
+%             country={Spain}}
+
+% https://ascl.net/code/v/2823
+
+% \date{24th of January, 2021}
+
+\begin{abstract}
 Characterize and understand \emph{Open Clusters} (OCs) allow us to understand better properties and mechanisms about the Universe such as stellar formation and the regions where these events occur.
@@
 Our method achieves good results, becoming even better (in some of the cases) than current techniques.
-}
+\end{abstract}
 
-\palabrasclave{
-    characterization,
-    data analysis,
-    deep embedded clustering,
-    gaia,
-    machine learning,
+%%Graphical abstract
+% \begin{graphicalabstract}
+% \includegraphics{grabs}
+% \end{graphicalabstract}
+
+%%Research highlights
+\begin{highlights}
+\item We present a novel method to characterize Open Clusters
+\item The method employs Artificial Neural Networks.
+\item We adapted the Deep Embedded Clustering model.
+\item Compare the results with K-Means using the Gaia DR2 database.
+\item The code is freely available.
+\end{highlights}
+
+\begin{keyword}
+%% keywords here, in the form: keyword \sep keyword
+ characterization \sep
+    data analysis  \sep
+    deep embedded clustering  \sep
+    gaia  \sep
+    machine learning  \sep
     open cluster
-}
+%% PACS codes here, in the form: \PACS code \sep code
+% \PACS 0000 \sep 1111
+%% MSC codes here, in the form: \MSC code \sep code
+%% or \MSC[2008] code \sep code (2000 is the default)
+% \MSC 0000 \sep 1111
+\end{keyword}
 
-\begin{document}
-\twocolumn[
-\begin{@twocolumnfalse}
-\maketitle
-\end{@twocolumnfalse}
-]
+\end{frontmatter}
 
@@
 Section~\ref{sec:related_works} presents some related works and Section~\ref{sec:aims_and_method} describes in detail our model.
-The following Section~\ref{sec:results}, shows the results and compares them with a method that involves tools from the Virtual Observatory (VO) such as Clusterix~\cite{balaguer2020clusterix}.
+The following Section~\ref{sec:results}, shows the results using either the Melotte 25 dataset and a variety of clusters of different typologies.
+We also compare our model against a K-Means classification algorithm.
 Finally, the last section concludes and outlines future research lines.
@@
 %% Comparación con K-Means, DEC y DEC filter
-\begin{figure*}[!htp]
+\begin{figure*}[!htbp]
   \centering
@@
 
-\newpage
-
 \section{Results}
@@
 All tests were run on an Apple Mac Pro Late 2013 with a 2.7GHz 12-Core Intel Xeon E5-2697v2, and 64GB RAM 1866MHz DDR3.
-For the GPU, we use a graphic card AMD FirePro D700 with 6GB~\footnote{All resources developed for this project are available at \href{https://github.com/cdalvaro/machine-learning-master-thesis}{https://github.com/cdalvaro/machine-learning-master-thesis}}.
+For the GPU, we use a graphic card AMD FirePro D700 with 6GB~\footnote{All resources developed for this project are available at [LINK]}.
+%\href{https://github.com/cdalvaro/machine-learning-master-thesis}{https://github.com/cdalvaro/machine-learning-master-thesis}}.
 
@@
   % \end{subfigure}
-  % \caption{Melotte 25 characterization.
-  %          From left to right: proper motion configuration space,
-  %          parallax histogram and H-R diagram.}
+  % \caption{Melotte 25 characterization. From left to right: proper motion configuration space, parallax histogram and H-R diagram.}
   \caption{Parallax histogram of Melotte 25 characterization.}
@@
   \rotatebox[origin=c]{90}{parallax \strut}
-  \begin{subfigure}{0.3\textwidth}
+  \begin{subfigure}{0.29\textwidth}
     \includegraphics[width=\textwidth]{../figures/ngc_2516/kmeans_parallax_ngc_2516.png}
   \end{subfigure}
-  \begin{subfigure}{0.3\textwidth}
+  \begin{subfigure}{0.29\textwidth}
     \includegraphics[width=\textwidth]{../figures/ngc_2516/dec_parallax_ngc_2516.png}
   \end{subfigure}
-  \begin{subfigure}{0.3\textwidth}
+  \begin{subfigure}{0.29\textwidth}
     \includegraphics[width=\textwidth]{../figures/ngc_2516/dec_parallax_filtered_ngc_2516.png}
@@
 
-  \rotatebox[origin=c]{90}{{\bfseries} \strut}
+  \rotatebox[origin=c]{90}{{\bfseries } \strut}
   \rotatebox[origin=c]{90}{hr diagram\strut}
-  \begin{subfigure}{0.3\textwidth}
+  \begin{subfigure}{0.29\textwidth}
     \includegraphics[width=\textwidth]{../figures/ngc_2516/kmeans_hr_diagram_ngc_2516.png}
   \end{subfigure}
-  \begin{subfigure}{0.3\textwidth}
+  \begin{subfigure}{0.29\textwidth}
     \includegraphics[width=\textwidth]{../figures/ngc_2516/dec_hr_diagram_ngc_2516.png}
   \end{subfigure}
-  \begin{subfigure}{0.3\textwidth}
+  \begin{subfigure}{0.29\textwidth}
     \includegraphics[width=\textwidth]{../figures/ngc_2516/dec_hr_diagram_filtered_ngc_2516.png}
@@
   \rotatebox[origin=c]{90}{parallax \strut}
-  \begin{subfigure}{0.3\textwidth}
+  \begin{subfigure}{0.29\textwidth}
     \includegraphics[width=\textwidth]{../figures/ngc_2632/kmeans_parallax_ngc_2632.png}
   \end{subfigure}
-  \begin{subfigure}{0.3\textwidth}
+  \begin{subfigure}{0.29\textwidth}
     \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_parallax_ngc_2632.png}
   \end{subfigure}
-  \begin{subfigure}{0.3\textwidth}
+  \begin{subfigure}{0.29\textwidth}
     \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_parallax_filtered_ngc_2632.png}
@@
 
-  \rotatebox[origin=c]{90}{{\bfseries} \strut}
+  \rotatebox[origin=c]{90}{{\bfseries } \strut}
   \rotatebox[origin=c]{90}{hr diagram\strut}
-  \begin{subfigure}{0.3\textwidth}
+  \begin{subfigure}{0.29\textwidth}
     \includegraphics[width=\textwidth]{../figures/ngc_2632/kmeans_hr_diagram_ngc_2632.png}
@@
   \end{subfigure}
-  \begin{subfigure}{0.3\textwidth}
+  \begin{subfigure}{0.29\textwidth}
     \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_hr_diagram_ngc_2632.png}
@@
   \end{subfigure}%
-  \begin{subfigure}{0.3\textwidth}
+  \begin{subfigure}{0.29\textwidth}
     \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_hr_diagram_filtered_ngc_2632.png}
@@
   \rotatebox[origin=c]{90}{parallax \strut}
-  \begin{subfigure}{0.3\textwidth}
+  \begin{subfigure}{0.29\textwidth}
     \includegraphics[width=\textwidth]{../figures/ngc_2682/kmeans_parallax_ngc_2682.png}
   \end{subfigure}
-  \begin{subfigure}{0.3\textwidth}
+  \begin{subfigure}{0.29\textwidth}
     \includegraphics[width=\textwidth]{../figures/ngc_2682/dec_parallax_ngc_2682.png}
   \end{subfigure}
-  \begin{subfigure}{0.3\textwidth}
+  \begin{subfigure}{0.29\textwidth}
     \includegraphics[width=\textwidth]{../figures/ngc_2682/dec_parallax_filtered_ngc_2682.png}
@@
 
-  \rotatebox[origin=c]{90}{{\bfseries} \strut}
+  \rotatebox[origin=c]{90}{{\bfseries } \strut}
   \rotatebox[origin=c]{90}{hr diagram\strut}
-  \begin{subfigure}{0.3\textwidth}
+  \begin{subfigure}{0.29\textwidth}
     \includegraphics[width=\textwidth]{../figures/ngc_2682/kmeans_hr_diagram_ngc_2682.png}
   \end{subfigure}
-  \begin{subfigure}{0.3\textwidth}
+  \begin{subfigure}{0.29\textwidth}
     \includegraphics[width=\textwidth]{../figures/ngc_2682/dec_hr_diagram_ngc_2682.png}
   \end{subfigure}
-  \begin{subfigure}{0.3\textwidth}
+  \begin{subfigure}{0.29\textwidth}
     \includegraphics[width=\textwidth]{../figures/ngc_2682/dec_hr_diagram_filtered_ngc_2682.png}
@@
   \rotatebox[origin=c]{90}{parallax \strut}
-  \begin{subfigure}{0.3\textwidth}
+  \begin{subfigure}{0.29\textwidth}
     \includegraphics[width=\textwidth]{../figures/melotte_22/kmeans_parallax_melotte_22.png}
   \end{subfigure}
-  \begin{subfigure}{0.3\textwidth}
+  \begin{subfigure}{0.29\textwidth}
     \includegraphics[width=\textwidth]{../figures/melotte_22/dec_parallax_melotte_22.png}
   \end{subfigure}
-  \begin{subfigure}{0.3\textwidth}
+  \begin{subfigure}{0.29\textwidth}
     \includegraphics[width=\textwidth]{../figures/melotte_22/dec_parallax_filtered_melotte_22.png}
@@
 
-  \rotatebox[origin=c]{90}{{\bfseries} \strut}
+  \rotatebox[origin=c]{90}{{\bfseries } \strut}
   \rotatebox[origin=c]{90}{hr diagram\strut}
-  \begin{subfigure}{0.3\textwidth}
+  \begin{subfigure}{0.29\textwidth}
     \includegraphics[width=\textwidth]{../figures/melotte_22/kmeans_hr_diagram_melotte_22.png}
@@
   \end{subfigure}
-  \begin{subfigure}{0.3\textwidth}
+  \begin{subfigure}{0.29\textwidth}
     \includegraphics[width=\textwidth]{../figures/melotte_22/dec_hr_diagram_melotte_22.png}
@@
   \end{subfigure}
-  \begin{subfigure}{0.3\textwidth}
+  \begin{subfigure}{0.29\textwidth}
     \includegraphics[width=\textwidth]{../figures/melotte_22/dec_hr_diagram_filtered_melotte_22.png}
@@
 
-\renewcommand{\refname}{References}
-\bibliographystyle{unsrt}
+\section{Acknowledgements}
+
+This work has made use of data from the European Space Agency (ESA) mission
+ Gaia (\url{https://www.cosmos.esa.int/gaia}), processed by the Gaia
+Data Processing and Analysis Consortium (DPAC\footnote{ \url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}}). Funding for the DPAC
+has been provided by national institutions participating in the {\it Gaia} Multilateral Agreement.
+
+% \renewcommand{\refname}{References}
+\bibliographystyle{elsarticle-harv}
 \bibliography{references}
